### 2012-10-18 large_number_squeezing_with_atoms.tex
Initial creation of paper files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\usepackage{graphicx}

\begin{document}

\title{Iconic representations of molluscs through the ages}

\author{Mattias T Johnsson$^1$, Graham R Dennis$^2$ and Joseph J Hope$^1$}

\address{$^1$Department of Quantum Science, The Australian National University, Canberra ACT 0200, Australia}
\address{$^2$Research School of Physics and Engineering, The Australian National University, Canberra ACT 0200, Australia}
\ead{mattias.johnsson@anu.edu.au}

\begin{abstract}
More squeezing than you can shake a stick at.

\end{abstract}

%Uncomment for PACS numbers title message
%\pacs{00.00, 20.00, 42.10}
% Keywords required only for MST, PB, PMB, PM, JOA, JOB? 
%\vspace{2pc}
%\noindent{\it Keywords}: Article preparation, IOP journals
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
% Comment out if separate title page not required
\maketitle

\section{Introduction}
Wibble.

\subsection{Subsection}
This is a subsection

\section{Model and single mode solutions}
Mollusc mollusc.
\begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/pulse_scheme.eps}
    \caption{Mollusc mollusc}
    \label{fig:PulseScheme}
\end{figure}

\section{Conclusion}

\clearpage

\appendix
\section{Derivation of single mode squeezing formula}
\label{appendixSMderivation}

\section*{References}
\begin{thebibliography}{10}
\bibitem{gross2010} Gross C, Zibold T, Est{\`{e}}ve J and Oberthaler M K 2010 Nonlinear atom interferometer surpasses classical
precision limit {\it Nature} {\bf 464} 1165--1169
\bibitem{dennis2013} Dennis G R, Hope J J and Johnsson M T 2013 XMDS2: Fast, scalable simulation of coupled stochastic partial
differential equations {\it Comput. Phys. Commun.} {\bf 184} 201--208

\end{thebibliography}

\end{document}

### 2012-10-18 large_number_squeezing_with_atoms.tex
Add some bullet points to intro as a test
Address adjustment as commit test
Trivial change to try using github for the first time.

@@
 
-\address{$^1$Department of Quantum Science, The Australian National University, Canberra ACT 0200, Australia}
-\address{$^2$Research School of Physics and Engineering, The Australian National University, Canberra ACT 0200, Australia}
+\address{$^1$Department of Quantum Science, Research School of Physics and Engineering, The Australian National University, Canberra ACT 0200, Australia}
+\address{$^2$Plasma Research Laboratory, Research School of Physics and Engineering, The Australian National University, Canberra ACT 0200, Australia}
 \ead{mattias.johnsson@anu.edu.au}
@@
 \section{Introduction}
-Wibble.
+\begin{itemize}
+  \item General non-classical state generation with atoms
+  \item Talk about number squeezing with atoms
+  \item Link to interferometry
+  \item Motivate why large number squeezing
+  \item Point out no one has done it before, only low number squeeze
+  \item Lay out what sections we have and what we've done
+\end{itemize}
 

### 2012-10-18 large_number_squeezing_with_atoms.tex
Added skeleton of paper, sections, main point to include etc

--- 2012-10-18 large_number_squeezing_with_atoms.tex
@@
 \section{Introduction}
-Wibble.
-
-\subsection{Subsection}
-This is a subsection
+\label{sectionIntroduction}
+\begin{itemize}
+  \item General non-classical state generation with atoms
+  \item Talk about number squeezing with atoms
+  \item Link to interferometry
+  \item Motivate why large number squeezing
+  \item Point out no one has done it before, only low number squeeze
+  \item Lay out what sections we have and what we've done
+\end{itemize}
 
 \section{Model and single mode solutions}
-Mollusc mollusc.
+\begin{itemize}
+  \item Lay out the scheme with diagram, define all variable etc
+  \item Present multimode analytic solution, details go into an appendix
+  \item Plot a few solutions to give an idea of what things look like, plausible times and so on
+  \item Maybe point out it agrees with Oberthaler, or save that for discussion on MM effects?
+  \item Point out that number *difference* squeezing can often be considerably higher than absolute number squeezing in a single state
+\end{itemize}
+
 \begin{figure}
@@
 
+\section{Multimode background section}
+\begin{itemize}
+  \item Motivate why MM analysis is necessary
+  \item Brief recap why MM QFT is hard
+  \item Say how great phase space methods are, quick overview of how they work, say only wigner works here, can probably crib some explanation and formulae from our fully quantum multimode atom laser paper
+  \item Talk about all the checks I've done to make sure Wigner holds over the parameters we've simulated despite the fact that TWA is an uncontrolled approximation
+  \item Cite XMDS, and say we're amazing at doing QFT sims, send money
+  Derive link between mode shape, nonlinearity U, and the SM equivalent nonlinearity, so we can always compare a given high-dimensional multimode sim against the equivalent ideal analytic single mode case
+\end{itemize}
+
+\section{Meanfield multimode problems}
+\begin{itemize}
+  \item Point out that different densities have phases that evolve at different rates
+  \item This leads to poor mode matching when recombining if two states have different nonlinearity*number values 
+  \item give a plot of the mode in |a> and |b> at the beginning and end of the hold time to show mismatch
+  \item This means constant density is good since mode shape doesn't change
+  \item Plot of squeezing for given set of params for gaussian, TF, constant density, show it get steadily better
+  \item Say we can use a pi pulse to swap the populations half way through, so if the states have different nonlinearities we still end up with the same mode shapes when we recombine.
+The fact that this works will be clear from the SM equations for a(t), b(t) when the beam splitter is on
+  \item show some plots of squeezing with and without the pi pulse
+\end{itemize}
+
+\section{Mean field quantum statistical results}
+  \begin{itemize}
+  \item Two issues. 1) as we add dimensions results get worse (maybe point out how cool we are for being able to do multimode 3D simulations). 2) As our volume increases results get worse.
+  \item To gain insight into this we do a bog analysis
+  \item Do a comparision plot between SM analytic model and Bog to show everything agrees
+  \item show that the number of mode available degrades squeezing
+  \item Plot some bog results v MM sims for increasing box sizes, and increasing dimensions
+  \item Can we put some numbers on how bad 3D is compared to 2D, in order to motivate tight trapping?
+  \end{itemize}
+
+\section{Mitigating multimode quantum statistical effects}
+  \begin{itemize}
+  \item One way to approximate constant density is to look at core of BEC, since if we only look at number in the central 1 sigma portion of gaussian it's close to flat; show results compared to single mode equivalent
+
+  \item Another improvement is to put the BEC in a box since then we get constant density, discuss feasibility
+  \item 3D still sucks even with box, so point out we can win by freezing out dimensions
+  \item Show derivation for what physical situations are needed to freeze out 1, 2 and 3 dimensions.
+In general we need small number, which makes it hard since we want large number squeezing.
+Large number requirement leads to implausible geometry in 1D (i.e.
+a tube with a ridiculous aspect ratio), but going from 3D down to 2D is probably do-able
+  \end{itemize}
+
 \section{Conclusion}
+\label{sectionConclusion}
+\begin{itemize}
+  \item Kerr nonlinearities can give arbitrarily good squeezing in BECs with high number
+  \item In a straightforward approach we can't get good squeezing with high number due to MM effects
+  \item Mode matching issues can be solved with equal scattering lengths or a pi pulse or a box
+  \item Another MM effect scales with dimension and physical size
+  \item Can solve that problem with tight trapping and box
+\end{itemize}
+
+\ack
+Acknowledgements in here; funding sources etc
 

### 2012-10-18 large_number_squeezing_with_atoms.tex
Start fleshing out two-mode section

--- 2012-10-18 large_number_squeezing_with_atoms.tex
@@
 
+As our goal is to describe the number squeezing possibilities of a two component BEC, we begin by considering a simplified version of the problem, consisting of a two-mode system, with the modes corresponding to the two states of the condensate.
+In the full problem, each of the components will include many spacial modes; our two mode system makes the assumption that only one of these spacial modes is relevant for each of the condensate's components.
+
+We allow the atoms to interact via a standard Kerr-type nonlinearity arising from atom-atom interactions, as well as allowing a linear coupling between the two modes.
+Denoting the two modes by $|a\rangle$ and $|b\rangle$, the Hamiltonian governing this system is given by
+\begin{equation}
+\hat{H} = \hbar \omega_a \hat{a}^{\dagger} \hat{a} +  \hbar \omega_b \hat{b}^{\dagger} \hat{b} 
+          + \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
+          + \chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
+          + \frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}
+          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{a}^{\dagger} \hat{b} )
+\label{eqTwoModeHamiltonian}
+\end{equation}
+where the $\chi_{ij}$ describe the various inter- and intra-mode nonlinearities, and $\Omega$ couples the two modes and allows for population transfer between them.
+
+The system is prepared with all the population in $|a\rangle$, and vacuum in $|b\rangle$.
+$|a\rangle$ is well-described by a coherent state {\bf *** Justify this.
+***} with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number in our system.
+
+At time $t=0$ the coupling $\Omega$ is switched on and applied until time $t=t_1$, at which point it is turned off, resulting in a portion of the population being transferred into mode $|a\rangle$.
+As the initial state was a coherent state in both modes, after the population transfer both $|a\rangle$ and $|b\rangle$ are described by coherent states, with mean number $\langle \hat{a}^{\dagger} \hat{a} \rangle = n_a$ and $\langle \hat{b}^{\dagger} \hat{b} \rangle = n_b$ respectively.
+
+The coupling $\Omega$ is then switched off until time $t_2$.
+During this period the atoms interact solely through the nonlinear terms, with no population being transferred.
+Due to the form of the Hamiltonian, the number variance remains unchanged, but the quadrature variances in the two modes change over time, which can lead to quadrature squeezing \cite{xxx}.
+
+After this hold time $\tau_{\mathrm{hold}} = t_2 - t_1 $, the coupling $\Omega$ is switched back on until time $t_3$, with a phase shift $\phi$ compared to it's first application.
+During this last stage, the two modes exchange population and the quadrature variance is converted into number variance, in the same way homodyne measurements are used in quantum optics to convert quadrature squeezing (which cannot be directly measured) into number squeezing (which can).
+In this interpretation, $\phi$ is the relative phase of the strong local oscillator, which allows specific phase angles of quadrature squeezing to be examined.
+
+The entire sequence of pulses and the resulting populations are shown schematically in Figure~\ref{figPulseScheme}.
+
+This system can be solved analytically, and expressions for quadrature, absolute and relative number squeezing can be derived.
+We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}.
+If we define
+\begin{eqnarray}
+s &=& \sin(\Omega (t_3 - t_2)) \\
+c &=& \cos(\Omega (t_3 - t_2)) \\
+\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} / \hbar \\
+A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
+A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
+B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
+B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )]
+\end{eqnarray}
+then the expectation values for number and number squared at time $t=t_3$ are given by
+\begin{eqnarray}
+\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \rangle &=& n_a c^2 + n_b s^2 + i c s (A B^* - A^* B) \\
+%
+\langle \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \rangle &=& n_a s^2 + n_b c^2 - i c s (A B^* - A^* B) \\
+%
+\langle \hat{a}^{\dagger}(t_3)  \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}(t_3) \rangle &=& (n_a^2 + n_a) c^4 + (n_b^2 + n_b) s^4 \nonumber \\
+&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2] \nonumber \\
+&& + i c^3 s [2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})}] A B^* - 2 n_a e^{i(\lambda_{aa}-\lambda_{ab})} A^* B + A B^* - A^* B] \nonumber \\
+&& + i c s^3 [2 n_b e^{i(\lambda_{bb}-\lambda_{ab})}] A B^* - 2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})} A^* B + A B^* - A^* B] \\
+%
+\langle \hat{b}^{\dagger}(t_3)  \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \hat{b}(t_3) \rangle &=& (n_b^2 + n_b) c^4 + (n_a^2 + n_a) s^4 \nonumber \\
+&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2 - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^*] \nonumber \\
+&& + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* + A^* B - A B^*] \nonumber \\
+&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* + A^* B - A B^*] \\
+\end{eqnarray}
+
 \begin{figure}
@@
     \caption{Mollusc mollusc}
-    \label{fig:PulseScheme}
+    \label{figPulseScheme}
 \end{figure}
@@
 \section{Derivation of single mode squeezing formula}
-\label{appendixSMderivation}
+\label{appendixTwoModeDerivation}
 

### 2012-10-24 large_number_squeezing_with_atoms.tex
Gave formulae for variances, tweaked figure

@@
 \begin{equation}
-\hat{H} = \hbar \omega_a \hat{a}^{\dagger} \hat{a} +  \hbar \omega_b \hat{b}^{\dagger} \hat{b} 
+\hat{H} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b} 
           + \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
@@
 B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
-B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )]
+B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
+D &=& AB^* - A^*B
 \end{eqnarray}
-then the expectation values for number and number squared at time $t=t_3$ are given by
+then the expectation values for number, number variance, and number difference variance after appling the final coupling field for a time $\tau = t_3-t_2$ are given by
 \begin{eqnarray}
-\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \rangle &=& n_a c^2 + n_b s^2 + i c s (A B^* - A^* B) \\
+N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D \\
 %
-\langle \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \rangle &=& n_a s^2 + n_b c^2 - i c s (A B^* - A^* B) \\
+N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D \\
 %
-\langle \hat{a}^{\dagger}(t_3)  \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}(t_3) \rangle &=& (n_a^2 + n_a) c^4 + (n_b^2 + n_b) s^4 \nonumber \\
-&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2] \nonumber \\
-&& + i c^3 s [2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})}] A B^* - 2 n_a e^{i(\lambda_{aa}-\lambda_{ab})} A^* B + A B^* - A^* B] \nonumber \\
-&& + i c s^3 [2 n_b e^{i(\lambda_{bb}-\lambda_{ab})}] A B^* - 2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})} A^* B + A B^* - A^* B] \\
+{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \\
+       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2) \nonumber \\
+       && + i c s D (1-2 n_a c^2 -2 n_b s^2)  \nonumber\\
+       && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
+       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ) \nonumber\\
 %
-\langle \hat{b}^{\dagger}(t_3)  \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \hat{b}(t_3) \rangle &=& (n_b^2 + n_b) c^4 + (n_a^2 + n_a) s^4 \nonumber \\
-&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2 - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^*] \nonumber \\
-&& + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* + A^* B - A B^*] \nonumber \\
-&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* + A^* B - A B^*] \\
+{\mathrm{Var}} [ N_b](t_3) &=&  n_a s^2 + n_b c^2 + 2 n_a n_b c^2 s^2 \nonumber \\
+       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2 ) \nonumber \\
+       && + i c s D (-1+2 n_a s^2 + 2 n_b c^2)  \nonumber \\
+       && + 2 i c^3 s n_b (e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - e^{i(\lambda_{bb} - \lambda_{ab})} A B^* ) \nonumber \\
+       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ) \\
+%
+{\mathrm{Var}} [ N_a - N_b](t_3) &=& n_a + n_b + 4 i c s D (n_a - n_b)(s^2 - c^2) \nonumber \\
+       && + 4 c^2 s^2 (2 n_a n_b +D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} A_2^* B_2 - e^{-i(\lambda_{aa} - \lambda_{bb})} A_2 B_2^* ) \nonumber \\
+       && + 4 i c s (c^2 - s^2) (n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B 
+                     + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^*)
 \end{eqnarray}
@@
 
+\begin{eqnarray}
+s &=& \sin(\Omega (t_3 - t_2)) \\
+c &=& \cos(\Omega (t_3 - t_2)) \\
+\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} / \hbar \\
+A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
+A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
+B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
+B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )]
+\end{eqnarray}
+then the expectation values for number and number squared at time $t=t_3$ are given by
+\begin{eqnarray}
+\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \rangle &=& n_a c^2 + n_b s^2 + i c s (A B^* - A^* B) \\
+%
+\langle \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \rangle &=& n_a s^2 + n_b c^2 - i c s (A B^* - A^* B) \\
+%
+\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle &=& (n_a^2 + n_a) c^4 + (n_b^2 + n_b) s^4 \nonumber \\
+&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2] \nonumber \\
+&& + i c^3 s [2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})}] A B^* - 2 n_a e^{i(\lambda_{aa}-\lambda_{ab})} A^* B + A B^* - A^* B] \nonumber \\
+&& + i c s^3 [2 n_b e^{i(\lambda_{bb}-\lambda_{ab})}] A B^* - 2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})} A^* B + A B^* - A^* B] \\
+%
+\langle \hat{b}^{\dagger}(t_3)  \hat{b}(t_3) \hat{b}^{\dagger}(t_3)  \hat{b}(t_3) \rangle &=& (n_b^2 + n_b) c^4 + (n_a^2 + n_a) s^4 \nonumber \\
+&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2 - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^*] \nonumber \\
+&& + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* - A^* B + A B^*] \nonumber \\
+&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* - A^* B + A B^*] 
+\end{eqnarray}
+
+\begin{eqnarray}
+\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle &-& \langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \rangle^2 \nonumber \\
+%
+&=& \langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle \\
+&-& n_a^2 c^4 + n_b^2 s^4 - c^2 s^2 (A B^* - A^* B)^2 + 2 i n_a c^3 s (A B^* - A^* B) + 2 i n_b c s^3 (A B^* - A^* B) + n_a n_b c^2 s^2 \\
+%
+\langle \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \rangle &=& n_a s^2 + n_b c^2 - i c s (A B^* - A^* B) \\
+%
+\langle \hat{a}^{\dagger}(t_3)  \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}(t_3) \rangle &=& (n_a^2 + n_a) c^4 + (n_b^2 + n_b) s^4 \nonumber \\
+&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2] \nonumber \\
+&& + i c^3 s [2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})}] A B^* - 2 n_a e^{i(\lambda_{aa}-\lambda_{ab})} A^* B + A B^* - A^* B] \nonumber \\
+&& + i c s^3 [2 n_b e^{i(\lambda_{bb}-\lambda_{ab})}] A B^* - 2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})} A^* B + A B^* - A^* B] \\
+%
+\langle \hat{b}^{\dagger}(t_3)  \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \hat{b}(t_3) \rangle &=& (n_b^2 + n_b) c^4 + (n_a^2 + n_a) s^4 \nonumber \\
+&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2 - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^*] \nonumber \\
+&& + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* + A^* B - A B^*] \nonumber \\
+&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* + A^* B - A B^*] 
+\end{eqnarray}
+
 \section*{References}

### 2012-10-25 large_number_squeezing_with_atoms.tex
Add more detail to two-mode section

@@
 
+{\bf{Maybe expand on the link between a full BEC in a trap and a the following hamiltonian?}}
+
 We allow the atoms to interact via a standard Kerr-type nonlinearity arising from atom-atom interactions, as well as allowing a linear coupling between the two modes.
@@
 In this interpretation, $\phi$ is the relative phase of the strong local oscillator, which allows specific phase angles of quadrature squeezing to be examined.
+
+In another interpretation, the sequence of pulses results in a Ramsey interferometer, with a final beam splitter phase of $\phi$.
 
@@
 \end{eqnarray}
+In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, e.g.
+${\mathrm{Var}}[N_a]=N_a$, an no number squeezing is possible.
+
+There are three independent nonlinearities $\chi_{ij}$ in the Hamiltonian (\ref{eqTwoModeHamiltonian}).
+In the expressions for the number variances, however, the nonlinearities only appear in the cyclic combinations $\chi_{aa}-\chi_{ab}$, $\chi_{bb}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$, meaning that number squeezing is parameterized by only two independent quantities which will be linear combinations of the nonlinearities.
+For example, since $\chi_{bb}-\chi_{ab} = (\chi_{aa}-\chi_{ab}) - (\chi_{aa}-\chi_{bb})$ it is possible to parameterize the number squeezing entirely in terms of the nonlinear differences $\chi_{aa}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$.
+
+{\bf{There seems to be this bogus rule of thumb in the literature that the amount of squeezing you can get is related to $\chi_{11}+\chi_{22}-2\chi_{12}$.
+Is it worth pointing out that this is wrong?}}
+
+While the analytic expressions for the number variances are not particularly transparent, it is easy to evaluate and plot them for any given set of parameters.
+As an example, in Figure \ref{figxxx} we plot the solution for the normalized number variance in mode $|a\rangle$ for a particular set of nonlinearities, choosing a number of different values of $\phi$.
+The plots illustrate the basic features of the problem: During the final coupling period, the number variance undergoes cyclic variation, with a series of minima.
+It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
+
+%Can Uaa + Ubb - 2Uab be zero and still get squeezing? Yep, Uaa=0.03, Ubb=0.01, Uab=0.2 gives squeezing for all quantities of interest
 
@@
   \item This leads to poor mode matching when recombining if two states have different nonlinearity*number values 
-  \item give a plot of the mode in |a> and |b> at the beginning and end of the hold time to show mismatch
+  \item give a plot of the mode in $|a\rangle$ and $|b\rangle$ at the beginning and end of the hold time to show mismatch
   \item This means constant density is good since mode shape doesn't change

### 2012-10-25 large_number_squeezing_with_atoms.tex
Do brain dump of dimensional reduction from 3D to 0D, 1D, and 2D

@@
 
+In order to reduce the number of modes available, we can attempt to make our system single mode in one or more dimensions.
+To do this we need to ensure that that any excitation along that dimension is energetically impossible, so we always stay in the ground state.
+This in turn means we need the harmonic oscillator energy per particle to be greater than the nonlinear energy per particle.
+{\bf{*** Not happy with this explanation.
+Make it better ***}}.
+
+The criteria we need to accomplish this depends on the number of dimensions we want to freeze out, so investigate the cases where we reduce a 3D system to an effective 0D, 1D, and 2D system seperately.
+In all cases we assume we have a box potential with length $L$ in the dimension(s) that aren't frozen out, and tight harmonic trapping potential of frequency $\omega$ in the frozen dimensions.
+
+Reduction from 3D to 0D: Here we have tight trapping along all three dimensions, so we have a harmonic oscillator wavefunction along all dimensions.
+With $N$ particles in the system, the density is given by
+\begin{equation}
+|\psi|^2 = N \left( \frac{m \omega}{\pi \hbar} \right)^{3/2} \exp \left[ -\frac{m \omega} {\hbar} (x^2 +y^2+z^2) \right] .
+\end{equation}
+In this system, the nonlinear energy per particle is
+\begin{eqnarray}
+E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
+       &=& \frac{U N}{4 \sqrt{2}} \left( \frac{m \omega}{\pi \hbar} \right)^{3/2}
+\end{eqnarray}
+where $U = \pi \hbar^2 a/m$ and $a$ is the s-wave scattering length of the particles. The criterion $\hbar \omega > E_{NL}$ thus becomes
+\begin{equation}
+%N < \frac{1}{a} \sqrt{2 \pi \hbar}{m \omega}.
+\sqrt{\frac{\hbar}{m \omega}} > \frac{aN}{\sqrt{2 \pi}}.
+\end{equation}
+Assuming Rubidium atoms, typical scattering lengths ($a\sim$10nm) and large particle numbers (say $N\sim 10^6$), we require a harmonic oscillator length $\sqrt{\hbar/ m \omega}>4$mm.
+A BEC that's more than a centimetre across isn't particularly feasible.
+
+Reduction from 3D to 1D: Here we have tight trapping along $x$ and $y$, with a box potential along $z$. With $N$ particles in the system, the density is given by
+\begin{equation}
+|\psi|^2 = \frac{N}{L} \frac{m \omega}{\pi \hbar} \exp \left[ -\frac{m \omega} {\hbar}  (x^2 +y^2) \right] .
+\end{equation}
+In this system, the nonlinear energy per particle is
+\begin{eqnarray}
+E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
+       &=& \frac{U N m \omega}{4 \pi \hbar L}
+\end{eqnarray}
+The criterion $\hbar \omega > E_{NL}$ thus becomes
+\begin{equation}
+L > a N / \pi.
+\end{equation}
+Assuming Rubidium atoms, typical scattering lengths ($a\sim$10nm) and large particle numbers (say $N\sim 10^6$) we need a BEC that's a few millimeters long, which isn't particularly feasible.
+
+Reduction from 3D to 2D: Here we have tight trapping along $z$, with a box potential along $x$ and $y$. With $N$ particles in the system, the density is given by
+\begin{equation}
+|\psi|^2 = \frac{N}{L^2} \left( {\frac{m \omega}{\pi \hbar}} \right) ^{1/2} \exp \left[ -\frac{m \omega} {\hbar} z^2  \right]
+\end{equation}
+In this system, the nonlinear energy per particle is
+\begin{eqnarray}
+E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
+       &=& \frac{m \omega U N}{4 \pi \hbar L^2}
+\end{eqnarray}
+The criterion $\hbar \omega > E_{NL}$ thus becomes
+\begin{equation}
+\omega > \frac{8 \pi a^2 N^2 \hbar}{m L^4}.
+\end{equation}
+Once again taking a Rubidium atom and considering typical parameters, say $a\sim$10nm, $N\sim 10^6$, $L\sim 200\mu$m, we see that we need $\omega > 2\pi\times 190$Hz, which is feasible.
+
+The takeaway point is that dimensional reduction to 0D and 1D is possible in the case where we have low particle numbers, which is how Oberthaler got his 6dB (or 8dB depending how you count) of squeezing.
+He only had 400 atoms, so could reduce his problem to 0D and freeze out any multimode dynamics.
+But if we're interested in large number squeezing, the best we can really do is reduction from 3D to 2D.
+
 \section{Conclusion}

### 2012-10-25 large_number_squeezing_with_atoms.tex
Begin fleshing out the appendix

@@
 
-\section{Model and single mode solutions}
+\section{Model and two mode solutions}
+\label{secTwoModeAnalytic}
 \begin{itemize}
@@
           + \frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}
-          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{a}^{\dagger} \hat{b} )
+          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger}  \hat{a} )
 \label{eqTwoModeHamiltonian}
@@
 \appendix
-\section{Derivation of single mode squeezing formula}
+\section{Derivation of two mode squeezing formula}
 \label{appendixTwoModeDerivation}
+In this Appendix we derive the equations for the number variances which were earlier stated without proof.
+We use the model described in Section \ref{secTwoModeAnalytic}, where we consider a two mode system with $N$ particles, initially prepared as a coherent state with all the population in mode $|a\rangle$ at time $t=0$.
+A coupling pulse is then applied, so that the system evolves under the Hamiltonian
+\begin{equation}
+\hat{H_1} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b} 
+          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger} \hat{a} )
+\end{equation}
+until time $t=t_1$. From time $t_1$ until time $t=t_2$ the coupling is turned off, and the system evolves under the Hamiltonian
+\begin{equation}
+\hat{H_2} = \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
+          + \chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
+          + \frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}.
+\end{equation}
+Finally, from time $t_2$ until time $t=t_3$, the coupling is once again applied with a phase offset of $\phi$ relative to the first time, and the system evolves via
+\begin{equation}
+\hat{H_3} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b}
+          + \Omega (e^{i\phi} \hat{a}^{\dagger} \hat{b} + e^{-i\phi} \hat{b}^{\dagger} \hat{a} )
+\end{equation}
+It is clear that the linear self-energy terms proportional to $\omega$ result in a constant phase rotation of the fields $\hat{a}$ and $\hat{b}$.
+As this system is self-contained and global phase is unmeasurable, all phase measurements can only rely on the difference in phase between the two modes.
+This phase rotation is therefore irrelevent and will be ignored in the following analysis.
+
+We will work in the Heisenberg picture, and derive expressions for $\hat{a}(t_3)$ and $\hat{b}(t_3)$ in terms of $\hat{a}(t_2)$ and $\hat{b}(t_2)$, which in turn can be expressed in terms of $\hat{a}(t_1)$ and $\hat{b}(t_1)$.
+We will show that the state of the system at time $t_1$ consists of seperate coherent states in modes $|a\rangle$ and $|b\rangle$, enabling us to calculate expectation values at time $t_3$ in terms of the known state at $t_1$.
+
+In order to have more compact expressions we will use the notation
+\begin{eqnarray}
+\hat{a}(t_j) &=& \hat{a}_j \\
+\hat{b}(t_j) &=& \hat{b}_j.
+\end{eqnarray}
+
+Beginning with the Hamiltonian $H_3$, during the period $(t_2, t_3)$ mode $|a\rangle$ will evolve as
+\begin{equation}
+\hat{a}_3 = e^{ i H_3 t/ \hbar} \, \hat{a}_2 \, e^{-i H_3 t/ \hbar}.
+\end{equation}
+Utilizing the Hadamard lemma
+\begin{equation}
+e^X Y e^{-X} = Y + [X,Y] + \frac{1}{2!}[X,[X,Y]] + \frac{1}{3!}[X,[X,[X,Y]]] + \ldots
+\end{equation}
+we obtain
+\begin{eqnarray}
+\hat{a}_3 &=& \hat{a}_2 - i e^{i\phi} \Omega \tau \hat{b}_2 + \frac{(i \Omega \tau)^2}{2!} \hat{a}_2 - \frac{(i \Omega \tau)^3}{3!}e^{i\phi} \hat{b}_2 + \frac{(i \Omega \tau)^4}{4!} \hat{a}_2 - \ldots \nonumber \\
+          &=& \cos (\Omega \tau) \hat{a}_2 -i e^{i\phi} \sin (\Omega\tau) \hat{b}_2
+\end{eqnarray}
+where $\tau=t_3-t_2$ is the period for which the coupling is applied. Similarly, we find
+\begin{equation}
+\hat{b}_3 = \cos (\Omega \tau) \hat{b}_2 -i e^{-i\phi} \sin (\Omega\tau) \hat{a}_2.
+\end{equation}
+
+(1) Find $\hat{a}_2$, $\hat{a}_2$ in terms of $\hat{a}_1$, $\hat{b}_1$.
+
+(2) Derive 1631.5 and explain we need expectation values of the bilinear, trilinear and quartic terms. We can use result of previous step to do them, provided we normally order everything
+
+(3) Take each set of terms from 1631.5 individually, normal order them, and evaluate
+
+(4) Plug list of expectation values back into 1631.5
 

### 2012-10-26 large_number_squeezing_with_atoms.tex
More appendix. Everyone likes appendix.

@@
 
-Beginning with the Hamiltonian $H_3$, during the period $(t_2, t_3)$ mode $|a\rangle$ will evolve as
+Beginning with the Hamiltonian $H_1$, during the period $(0, t_1)$ mode $\hat{a}$ will evolve as
 \begin{equation}
-\hat{a}_3 = e^{ i H_3 t/ \hbar} \, \hat{a}_2 \, e^{-i H_3 t/ \hbar}.
+\hat{a}_1 = e^{ i H_1 t/ \hbar} \, \hat{a}_0 \, e^{-i H_1 t/ \hbar}.
 \end{equation}
@@
 e^X Y e^{-X} = Y + [X,Y] + \frac{1}{2!}[X,[X,Y]] + \frac{1}{3!}[X,[X,[X,Y]]] + \ldots
+\label{eqHadamard}
 \end{equation}
@@
 \begin{eqnarray}
-\hat{a}_3 &=& \hat{a}_2 - i e^{i\phi} \Omega \tau \hat{b}_2 + \frac{(i \Omega \tau)^2}{2!} \hat{a}_2 - \frac{(i \Omega \tau)^3}{3!}e^{i\phi} \hat{b}_2 + \frac{(i \Omega \tau)^4}{4!} \hat{a}_2 - \ldots \nonumber \\
-          &=& \cos (\Omega \tau) \hat{a}_2 -i e^{i\phi} \sin (\Omega\tau) \hat{b}_2
+\hat{a}_1 &=& \hat{a}_0 - i \Omega t_1 \hat{b}_0 + \frac{(i \Omega t_1)^2}{2!} \hat{a}_0 - \frac{(i \Omega t_1)^3}{3!} \hat{b}_0 + \frac{(i \Omega t_1)^4}{4!} \hat{a}_2 - \ldots \nonumber \\
+          &=& \cos (\Omega t_1) \hat{a}_0 -i \sin (\Omega t_1) \hat{b}_0
+\label{eqa1}
 \end{eqnarray}
-where $\tau=t_3-t_2$ is the period for which the coupling is applied. Similarly, we find
+Similarly, we find
 \begin{equation}
-\hat{b}_3 = \cos (\Omega \tau) \hat{b}_2 -i e^{-i\phi} \sin (\Omega\tau) \hat{a}_2.
+\hat{b}_1 = \cos (\Omega t_1) \hat{b}_0 - i \sin (\Omega t_1) \hat{a}_0.
+\label{eqb1}
+\end{equation}
+From the form of (\ref{eqa1}) and (\ref{eqb1}) it is clear that as both modes $\hat{a}$ and $\hat{b}$ began in coherent states, they will remain in coherent states, with only their amplitudes changing.
+Specifically, at time $t_1$ the system is in a product of coherent states, which we denote $|\alpha, \beta\rangle$ with $\alpha=\sqrt{n_a}$ and $\beta=-i\sqrt{n_b}$.
+
+Next we consider evolution under $H_2$.
+Taking mode $\hat{a}$ first, we note that as it commutes with the $\hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}$ term, and since $\hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}$ commutes with $\hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}$ we have
+\begin{equation}
+\hat{a}_2 = e^{i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1 \hat{b}_1^{\dagger} \hat{b}_1 } 
+          e^{ i \frac{\lambda_{aa}} {2} \hat{a}_1^{\dagger} \hat{a}_1^{\dagger} \hat{a}_1 \hat{a}_1}\, \hat{a}_1 \,  
+          e^{ -i \frac{\lambda_{aa}}{2} \hat{a}_1^{\dagger} \hat {a}_1^{\dagger} \hat{a}_1 \hat{a}_1 }
+          e^{-i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1 \hat{b}_1^{\dagger} \hat{b}_1}.
+\label{eqa2evolution}
+\end{equation}
+where we have defined $\lambda_{ij} = \chi_{ij} (t_2-t_1)/\hbar$. Utilizing (\ref{eqHadamard}) to move $\hat{a}_1$ through the exponential, with some algebra one can show that
+\begin{eqnarray}
+e^{ i \frac{\lambda_{aa}} {2} \hat{a}_1^{\dagger} \hat{a}_1^{\dagger} \hat{a}_1 \hat{a}_1} \hat{a}_1 
+         e^{ -i \frac{\lambda_{aa}} {2} \hat{a}_1^{\dagger} \hat{a}_1^{\dagger} \hat{a}_1 \hat{a}_1} &=& \left[ \sum_{n=0} (-i \lambda_{aa})^n (\hat{a}_1^{\dagger} \hat{a}_1)^n \right] \hat{a}_1 \nonumber \nonumber \\
+   &=& \exp[-i \lambda_{aa} \hat{a}_1^{\dagger} \hat{a}_1] \hat{a}_1
+\end{eqnarray}
+To handle the cross-nonlinearity term in (\ref{eqa2evolution}) we use the identity \cite{louisell}
+\begin{equation}
+e^{x \hat{a}^{\dagger} \hat{a}} f(\hat{a}, \hat{a}^{\dagger}) e^{-x \hat{a}^{\dagger} \hat{a}} = f(\hat{a}e^{-x}, \hat{a}^{\dagger} e^{x})
+\label{eqefeidentity}
+\end{equation}
+to obtain
+\begin{eqnarray}
+\hat{a}_2 &=& e^{i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1 \hat{b}_1^{\dagger} \hat{b}_1 } 
+          e^{-i \lambda_{aa} \hat{a}_1^{\dagger} \hat{a}_1} \, \hat{a}_1 \,
+          e^{-i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1 \hat{b}_1^{\dagger} \hat{b}_1} \nonumber \\
+%
+          &=& e^{-i \lambda_{aa} \hat{a}_1^{\dagger} \hat{a}_1} 
+              e^{i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1 \hat{b}_1^{\dagger} \hat{b}_1 } \, \hat{a}_1 \,
+          e^{-i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1 \hat{b}_1^{\dagger} \hat{b}_1} \nonumber \\
+%
+          &=& e^{-i \lambda_{aa} \hat{a}_1^{\dagger} \hat{a}_1} e^{-i \lambda_{ab} \hat{b}_1^{\dagger} \hat{b}_1} \, \hat{a}_1
+\label{eqa2}
+\end{eqnarray}
+Similarly, from permutation symmetry, we have
+\begin{equation}
+\hat{b}_2 = e^{-i \lambda_{bb} \hat{b}_1^{\dagger} \hat{b}_1} e^{-i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1} \, \hat{b}_1.     
+\label{eqb2}
+\end{equation}
+Finally, to obtain $\hat{a}_3$ and $\hat{b}_3$ we utilize (\ref{eqa1}) and (\ref{eqb1}) with the phase factor attached to the $\hat{b}$ operator to obtain
+\begin{eqnarray}
+\hat{a}_3 &=& \cos (\Omega (t_3-t_2)) \hat{a}_2 -i e^{i\phi} \sin (\Omega (t_3-t_2)) \hat{b}_2 \label{eqa3} \\
+\hat{b}_3 &=& \cos (\Omega (t_3-t_2)) \hat{b}_2 - i e^{-i\phi} \sin (\Omega (t_3-t_2)) \hat{a}_2 \label{eqb3}
+\end{eqnarray} 
+We are now in a position to evaluate the number variance of the system throughout period of the final coupling. We first consider the number variance of the $\hat{a}$ field, which is defined as
+\begin{equation}
+{\mathrm{Var}}[N_a] = \langle \hat{a}^{\dagger}_3 \hat{a}_3 \hat{a}^{\dagger}_3 \hat{a}_3 \rangle - \langle \hat{a}^{\dagger}_3 \hat{a}_3 \rangle ^2.
+\label{eqNavariance}
+\end{equation}
+Making use of (\ref{eqa3}) and using the shorthand notation 
+\begin{eqnarray}
+s &=& \sin[\Omega (t_3 - t_2)] \\
+c &=& \cos[\Omega (t_3 - t_2)]
+\end{eqnarray}
+we have
+\begin{eqnarray}
+\hat{a}^{\dagger}_3 \hat{a}_3 &=& \hat{a}^{\dagger}_2 \hat{a}_2 c^2 +  \hat{b}^{\dagger}_2 \hat{b}_2 s^2 + i c s e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 - i c s e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 \\
+%
+\hat{a}^{\dagger}_3 \hat{a}_3 \hat{a}^{\dagger}_3 \hat{a}_3 &=& \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 c^4 + \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 s^4 \nonumber \\
+%
+&& + c^2 s^2 [ \hat{a}^{\dagger}_2 \hat{a}_2 + \hat{b}^{\dagger}_2 \hat{b}_2 + 4 \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 -e^{-2 i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 -e^{-2 i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 ] \nonumber \\
+%
+&& + i c^3 s [2 e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}_2 - 2 e^{i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}_2 + e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 - e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 ] \nonumber \\
+%
+&& + i c s^3 [2 e^{-i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{b}_2 - 2 e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 + e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 - e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 ]
+\end{eqnarray}
+Clearly, to calculate (\ref{eqNavariance}) we will need the expectation values of terms quartic and quadratic in $\hat{a}_2$ and ${\hat{b}_2}$.
+Specifically, we require $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \rangle$, $\langle \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$,  $\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \rangle$, $\langle \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$ , $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 \rangle$, and $\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$, as well as their Hermitian conjugates.
+
+As $\hat{a}^{\dagger}_1 \hat{a}_1$ and $\hat{b}^{\dagger}_1 \hat{b}_1$ commute with the Hamiltonian $H_2$, they are constants of motion so we have
+\begin{equation}
+\langle \hat{a}^{\dagger}_2 \hat{a}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}_1 |\alpha, \beta \rangle = |\alpha|^2 = n_a
+\end{equation}
+%
+\begin{equation}
+\langle \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = |\beta|^2 = n_b
+\end{equation}
+%
+\begin{equation}
+\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}_1 \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = |\alpha|^2 |\beta|^2 = n_a n_b
+\end{equation}
+Furthermore, $\hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2$ and $\hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2$ also commute with $H_2$, giving
+\begin{equation}
+\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \hat{a}^{\dagger}_1 \hat{a}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}^{\dagger}_1 \hat{a}_1 \hat{a}_1 + \hat{a}^{\dagger}_1 \hat{a}_1 |\alpha, \beta \rangle = n_a^2 + n_a
+\end{equation}
+%
+\begin{equation}
+\langle \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{b}^{\dagger}_1 \hat{b}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{b}^{\dagger}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \hat{b}_1 + \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = n_b^2 + n_b
+\end{equation}
+The remaining terms are not constants of motion, so we proceed by making use of (\ref{eqa2}) and (\ref{eqb2}). For $\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle$ we have
+\begin{equation}
+\hat{a}^{\dagger}_2 \hat{b}_2 = \hat{a}^{\dagger}_1 e^{-i (\lambda_{bb} - \lambda_{ab}) \hat{b}^{\dagger}_1 \hat{b}_1} e^{i (\lambda_{aa} - \lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1} \hat{b}_1 
+\end{equation}
+To calculate the expectation value of this expression, the creation and annihilation operators must be normally ordered. Making use of the identity \cite{louisell}
+\begin{equation}
+e^{x \hat{a}^{\dagger} \hat{a}} = \sum_{n=0}^{\infty} \frac{(e^x - 1)^n} {n!} (\hat{a}^{\dagger})^n (\hat{a})^n
+\end{equation}
+allows us to compute expectation values of coherent states of the form
+\begin{eqnarray}
+\langle \alpha | e^{x \hat{a}^{\dagger} \hat{a}} |\alpha \rangle &=& \sum_{n=0}^{\infty} \frac{(e^x - 1)^n} {n!} |\alpha|^{2n} \nonumber \\
+%
+&=& \exp[|\alpha|^2(e^x -1)] \label{eqExpectationValueOfExponential}
+\end{eqnarray}
+This gives
+\begin{eqnarray}
+\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle &=& \langle \alpha, \beta | \hat{a}^{\dagger}_1 e^{-i (\lambda_{bb} - \lambda_{ab}) \hat{b}^{\dagger}_1 \hat{b}_1} e^{i (\lambda_{aa} - \lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1} \hat{b}_1 | \alpha, \beta \rangle \nonumber \\
+%
+&=& \alpha^* \beta \exp[|\alpha|^2(e^{i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2 (e^{-i (\lambda_{bb} - \lambda_{ab})} -1)]
+\end{eqnarray}
+
+Next we tackle $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$. Using (\ref{eqa2}) and (\ref{eqb2}) and the fact that $\hat{a}$ and $\hat{b}$ commute, we have
+\begin{eqnarray}
+\hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 &=& \hat{a}^{\dagger}_1 
+  e^{i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
+  \hat{b}_1 \hat{a}^{\dagger_1}
+  e^{i (\lambda_{ab}-\lambda_{bb}) \hat{b}^{\dagger}_1 \hat{b}_1}
+  e^{i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
+  \hat{b}_1 \\
+%
+&=& \hat{a}^{\dagger}_1 \hat{a}^{\dagger}_1 
+    e^{2i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
+    e^{2i (\lambda_{ab}-\lambda_{bb}) \hat{b}^{\dagger}_1 \hat{b}_1}
+    \hat{b}_1 \hat{b}_1 e^{i(\lambda_{aa}-\lambda_{bb})}
+\end{eqnarray}
+where we have made multiple uses of the identity 
+\begin{equation}
+\hat{a} e^{i x \hat{a}^{\dagger} \hat{a}} = e^{i x \hat{a}^{\dagger} \hat{a}} \hat{a} e^{ix}
+\end{equation}
+which follows from (\ref{eqefeidentity}). Finally, applying (\ref{eqExpectationValueOfExponential}) we obtain
+\begin{equation}
+\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle = \alpha^{*2} \beta^2 e^{i(\lambda_{aa}-\lambda_{bb})} \exp[|\alpha|^2(e^{2i (\lambda_{aa} - \lambda_{bb})} -1)] \exp[|\beta|^2(e^{-2i (\lambda_{bb} - \lambda_{bb})} -1)]
 \end{equation}
 
-(1) Find $\hat{a}_2$, $\hat{a}_2$ in terms of $\hat{a}_1$, $\hat{b}_1$.
-
-(2) Derive 1631.5 and explain we need expectation values of the bilinear, trilinear and quartic terms. We can use result of previous step to do them, provided we normally order everything
-
-(3) Take each set of terms from 1631.5 individually, normal order them, and evaluate
-
-(4) Plug list of expectation values back into 1631.5
-
-\begin{eqnarray}
-s &=& \sin(\Omega (t_3 - t_2)) \\
-c &=& \cos(\Omega (t_3 - t_2)) \\
-\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} / \hbar \\
-A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
-A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
-B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
-B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )]
-\end{eqnarray}
 then the expectation values for number and number squared at time $t=t_3$ are given by
@@
 differential equations {\it Comput. Phys. Commun.} {\bf 184} 201--208
-
+\bibitem{louisell} Louisell Quantum Statistical Properties of Radiation
 \end{thebibliography}

### 2012-10-29 large_number_squeezing_with_atoms.tex
Finish the appendix

@@
 \begin{eqnarray}
-s &=& \sin(\Omega (t_3 - t_2)) \\
+s &=& \sin(\Omega (t_3 - t_2)) \label{eqsDef} \\
 c &=& \cos(\Omega (t_3 - t_2)) \\
@@
 B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
-D &=& AB^* - A^*B
+D &=& AB^* - A^*B \label{eqDDef}
 \end{eqnarray}
@@
        && + 4 i c s (c^2 - s^2) (n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B 
-                     + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^*)
+                     + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^*) \label{eqNumDiffVariance}
 \end{eqnarray}
@@
 \end{eqnarray} 
-We are now in a position to evaluate the number variance of the system throughout period of the final coupling. We first consider the number variance of the $\hat{a}$ field, which is defined as
-\begin{equation}
-{\mathrm{Var}}[N_a] = \langle \hat{a}^{\dagger}_3 \hat{a}_3 \hat{a}^{\dagger}_3 \hat{a}_3 \rangle - \langle \hat{a}^{\dagger}_3 \hat{a}_3 \rangle ^2.
-\label{eqNavariance}
-\end{equation}
+We are now in a position to evaluate the number variance of the system throughout the period of the final coupling. The number variances of the fields are defined as
+\begin{eqnarray}
+{\mathrm{Var}}[N_a] &=& \langle \hat{a}^{\dagger}_3 \hat{a}_3 \hat{a}^{\dagger}_3 \hat{a}_3 \rangle - \langle \hat{a}^{\dagger}_3 \hat{a}_3 \rangle ^2 \label{eqNavariance} \\
+%
+{\mathrm{Var}}[N_b] &=& \langle \hat{b}^{\dagger}_3 \hat{b}_3 \hat{b}^{\dagger}_3 \hat{b}_3 \rangle - \langle \hat{b}^{\dagger}_3 \hat{b}_3 \rangle ^2
+\label{eqNbvariance}
+\end{eqnarray}
 Making use of (\ref{eqa3}) and using the shorthand notation 
@@
 %
+%
 \hat{a}^{\dagger}_3 \hat{a}_3 \hat{a}^{\dagger}_3 \hat{a}_3 &=& \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 c^4 + \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 s^4 \nonumber \\
@@
 \end{eqnarray}
-Clearly, to calculate (\ref{eqNavariance}) we will need the expectation values of terms quartic and quadratic in $\hat{a}_2$ and ${\hat{b}_2}$.
+\begin{eqnarray}
+\hat{b}^{\dagger}_3 \hat{b}_3 &=& \hat{b}^{\dagger}_2 \hat{b}_2 c^2 +  \hat{a}^{\dagger}_2 \hat{a}_2 s^2 + i c s e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 - i c s e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 \\
+%
+%
+\hat{b}^{\dagger}_3 \hat{b}_3 \hat{b}^{\dagger}_3 \hat{b}_3 &=& \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 c^4 + \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 s^4 \nonumber \\
+%
+&& + c^2 s^2 [ \hat{b}^{\dagger}_2 \hat{b}_2 + \hat{a}^{\dagger}_2 \hat{a}_2 + 4 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{a}^{\dagger}_2 \hat{a}_2 - e^{2 i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 -e^{2 i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 ] \nonumber \\
+%
+&& + i c^3 s [2 e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 - 2 e^{-i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{a}_2 + e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 - e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 ] \nonumber \\
+%
+&& + i c s^3 [2 e^{i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{a}_2 - 2 e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}_2 + e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 - e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 ]
+\end{eqnarray}
+Clearly, to calculate (\ref{eqNavariance}) and (\ref{eqNbvariance}) we will need the expectation values of terms quartic and quadratic in $\hat{a}_2$ and ${\hat{b}_2}$.
 Specifically, we require $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \rangle$, $\langle \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$,  $\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \rangle$, $\langle \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$ , $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 \rangle$, and $\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$, as well as their Hermitian conjugates.
@@
 \end{equation}
-
-then the expectation values for number and number squared at time $t=t_3$ are given by
+Using the same techniques one can show that
+\begin{eqnarray}
+\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 \rangle &=& |\alpha|^2 \alpha \beta^* e^{-i(\lambda_{aa}-\lambda_{ab})} \exp[|\alpha|^2(e^{-i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2(e^{i (\lambda_{bb} - \lambda_{ab})} -1)] \\
+%
+\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle &=& |\beta|^2 \alpha^{*} \beta e^{-i(\lambda_{bb}-\lambda_{ab})} \exp[|\alpha|^2(e^{i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2(e^{-i (\lambda_{bb} - \lambda_{ab})} -1)].
+\end{eqnarray}
+Employing the definitions (\ref{eqsDef})--(\ref{eqDDef}) we obtain
 \begin{eqnarray}
@@
 \end{eqnarray}
-
-\begin{eqnarray}
-\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle &-& \langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \rangle^2 \nonumber \\
-%
-&=& \langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle \\
-&-& n_a^2 c^4 + n_b^2 s^4 - c^2 s^2 (A B^* - A^* B)^2 + 2 i n_a c^3 s (A B^* - A^* B) + 2 i n_b c s^3 (A B^* - A^* B) + n_a n_b c^2 s^2 \\
-%
-\langle \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \rangle &=& n_a s^2 + n_b c^2 - i c s (A B^* - A^* B) \\
-%
-\langle \hat{a}^{\dagger}(t_3)  \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}(t_3) \rangle &=& (n_a^2 + n_a) c^4 + (n_b^2 + n_b) s^4 \nonumber \\
-&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2] \nonumber \\
-&& + i c^3 s [2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})}] A B^* - 2 n_a e^{i(\lambda_{aa}-\lambda_{ab})} A^* B + A B^* - A^* B] \nonumber \\
-&& + i c s^3 [2 n_b e^{i(\lambda_{bb}-\lambda_{ab})}] A B^* - 2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})} A^* B + A B^* - A^* B] \\
-%
-\langle \hat{b}^{\dagger}(t_3)  \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \hat{b}(t_3) \rangle &=& (n_b^2 + n_b) c^4 + (n_a^2 + n_a) s^4 \nonumber \\
-&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2 - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^*] \nonumber \\
-&& + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* + A^* B - A B^*] \nonumber \\
-&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* + A^* B - A B^*] 
-\end{eqnarray}
+When substituted into (\ref{eqNbvariance}) and (\ref{eqNbvariance}) these expecation values give the absolute number variances stated in Section \ref{secTwoModeAnalytic}.
+To calculate the number difference variance, we note
+\begin{equation}
+{\mathrm{Var}}[N_a-N_b] = {\mathrm{Var}}[N_a] + {\mathrm{Var}}[N_b] - 2\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2\rangle + 2\langle \hat{a}^{\dagger}_2  \hat{a}_2 \rangle \langle \hat{b}^{\dagger}_2 \hat{b}_2\rangle \label{eqNumDiffVarianceAppendix}.
+\end{equation}
+Substituting the expectation values calculated earlier in this Appendix into (\ref{eqNumDiffVarianceAppendix}) and simplifying yields the expression (\ref{eqNumDiffVariance}) given in Section \ref{secTwoModeAnalytic}.
 

### 2012-10-31 large_number_squeezing_with_atoms.tex
Minor appendix fixes

@@
 \hat{H_3} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b}
-          + \Omega (e^{i\phi} \hat{a}^{\dagger} \hat{b} + e^{-i\phi} \hat{b}^{\dagger} \hat{a} )
+          + \Omega (e^{i\phi} \hat{a}^{\dagger} \hat{b} + e^{-i\phi} \hat{b}^{\dagger} \hat{a} ).
 \end{equation}
@@
 \hat{a}_1 &=& \hat{a}_0 - i \Omega t_1 \hat{b}_0 + \frac{(i \Omega t_1)^2}{2!} \hat{a}_0 - \frac{(i \Omega t_1)^3}{3!} \hat{b}_0 + \frac{(i \Omega t_1)^4}{4!} \hat{a}_2 - \ldots \nonumber \\
-          &=& \cos (\Omega t_1) \hat{a}_0 -i \sin (\Omega t_1) \hat{b}_0
+          &=& \cos (\Omega t_1) \hat{a}_0 -i \sin (\Omega t_1) \hat{b}_0.
 \label{eqa1}
@@
           e^{ -i \frac{\lambda_{aa}}{2} \hat{a}_1^{\dagger} \hat {a}_1^{\dagger} \hat{a}_1 \hat{a}_1 }
-          e^{-i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1 \hat{b}_1^{\dagger} \hat{b}_1}.
+          e^{-i \lambda_{ab} \hat{a}_1^{\dagger} \hat{a}_1 \hat{b}_1^{\dagger} \hat{b}_1}
 \label{eqa2evolution}
@@
          e^{ -i \frac{\lambda_{aa}} {2} \hat{a}_1^{\dagger} \hat{a}_1^{\dagger} \hat{a}_1 \hat{a}_1} &=& \left[ \sum_{n=0} (-i \lambda_{aa})^n (\hat{a}_1^{\dagger} \hat{a}_1)^n \right] \hat{a}_1 \nonumber \nonumber \\
-   &=& \exp[-i \lambda_{aa} \hat{a}_1^{\dagger} \hat{a}_1] \hat{a}_1
+   &=& \exp[-i \lambda_{aa} \hat{a}_1^{\dagger} \hat{a}_1] \hat{a}_1.
 \end{eqnarray}
@@
 %
-{\mathrm{Var}}[N_b] &=& \langle \hat{b}^{\dagger}_3 \hat{b}_3 \hat{b}^{\dagger}_3 \hat{b}_3 \rangle - \langle \hat{b}^{\dagger}_3 \hat{b}_3 \rangle ^2
+{\mathrm{Var}}[N_b] &=& \langle \hat{b}^{\dagger}_3 \hat{b}_3 \hat{b}^{\dagger}_3 \hat{b}_3 \rangle - \langle \hat{b}^{\dagger}_3 \hat{b}_3 \rangle ^2.
 \label{eqNbvariance}
@@
 \begin{eqnarray}
-\hat{a}^{\dagger}_3 \hat{a}_3 &=& \hat{a}^{\dagger}_2 \hat{a}_2 c^2 +  \hat{b}^{\dagger}_2 \hat{b}_2 s^2 + i c s e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 - i c s e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 \\
+\hat{a}^{\dagger}_3 \hat{a}_3 &=& \hat{a}^{\dagger}_2 \hat{a}_2 c^2 +  \hat{b}^{\dagger}_2 \hat{b}_2 s^2 + i c s (e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 - e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2) \\
 %
@@
 \begin{eqnarray}
-\hat{b}^{\dagger}_3 \hat{b}_3 &=& \hat{b}^{\dagger}_2 \hat{b}_2 c^2 +  \hat{a}^{\dagger}_2 \hat{a}_2 s^2 + i c s e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 - i c s e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 \\
+\hat{b}^{\dagger}_3 \hat{b}_3 &=& \hat{b}^{\dagger}_2 \hat{b}_2 c^2 +  \hat{a}^{\dagger}_2 \hat{a}_2 s^2 + i c s ( e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 - e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2) \\
 %
@@
 \begin{equation}
-\hat{a}^{\dagger}_2 \hat{b}_2 = \hat{a}^{\dagger}_1 e^{-i (\lambda_{bb} - \lambda_{ab}) \hat{b}^{\dagger}_1 \hat{b}_1} e^{i (\lambda_{aa} - \lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1} \hat{b}_1 
+\hat{a}^{\dagger}_2 \hat{b}_2 = \hat{a}^{\dagger}_1 e^{-i (\lambda_{bb} - \lambda_{ab}) \hat{b}^{\dagger}_1 \hat{b}_1} e^{i (\lambda_{aa} - \lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1} \hat{b}_1. 
 \end{equation}
@@
 %
-&=& \exp[|\alpha|^2(e^x -1)] \label{eqExpectationValueOfExponential}
+&=& \exp[|\alpha|^2(e^x -1)]. \label{eqExpectationValueOfExponential}
 \end{eqnarray}
@@
 %
-&=& \alpha^* \beta \exp[|\alpha|^2(e^{i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2 (e^{-i (\lambda_{bb} - \lambda_{ab})} -1)]
+&=& \alpha^* \beta \exp[|\alpha|^2(e^{i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2 (e^{-i (\lambda_{bb} - \lambda_{ab})} -1)].
 \end{eqnarray}
@@
 \begin{equation}
-\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle = \alpha^{*2} \beta^2 e^{i(\lambda_{aa}-\lambda_{bb})} \exp[|\alpha|^2(e^{2i (\lambda_{aa} - \lambda_{bb})} -1)] \exp[|\beta|^2(e^{-2i (\lambda_{bb} - \lambda_{bb})} -1)]
+\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle = \alpha^{*2} \beta^2 e^{i(\lambda_{aa}-\lambda_{bb})} \exp[|\alpha|^2(e^{2i (\lambda_{aa} - \lambda_{bb})} -1)] \exp[|\beta|^2(e^{-2i (\lambda_{bb} - \lambda_{bb})} -1)].
 \end{equation}
@@
 && + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* - A^* B + A B^*] \nonumber \\
-&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* - A^* B + A B^*] 
+&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* - A^* B + A B^*]. 
 \end{eqnarray}
-When substituted into (\ref{eqNbvariance}) and (\ref{eqNbvariance}) these expecation values give the absolute number variances stated in Section \ref{secTwoModeAnalytic}.
+When substituted into (\ref{eqNavariance}) and (\ref{eqNbvariance}) these expecation values give the absolute number variances stated in Section \ref{secTwoModeAnalytic}.
 To calculate the number difference variance, we note
 \begin{equation}
-{\mathrm{Var}}[N_a-N_b] = {\mathrm{Var}}[N_a] + {\mathrm{Var}}[N_b] - 2\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2\rangle + 2\langle \hat{a}^{\dagger}_2  \hat{a}_2 \rangle \langle \hat{b}^{\dagger}_2 \hat{b}_2\rangle \label{eqNumDiffVarianceAppendix}.
+{\mathrm{Var}}[N_a-N_b] = {\mathrm{Var}}[N_a] + {\mathrm{Var}}[N_b] - 2\langle \hat{a}^{\dagger}_3 \hat{a}_3 \hat{b}^{\dagger}_3 \hat{b}_3\rangle + 2\langle \hat{a}^{\dagger}_3  \hat{a}_3 \rangle \langle \hat{b}^{\dagger}_3 \hat{b}_3\rangle \label{eqNumDiffVarianceAppendix}.
 \end{equation}
-Substituting the expectation values calculated earlier in this Appendix into (\ref{eqNumDiffVarianceAppendix}) and simplifying yields the expression (\ref{eqNumDiffVariance}) given in Section \ref{secTwoModeAnalytic}.
+We write (\ref{eqNumDiffVarianceAppendix}) in terms of $\hat{a}_2$ and $\hat{b}_2$ by using using (\ref{eqa3}) and (\ref{eqb3}), and then use the expectation values calculated earlier in this Appendix.
+After some algebra we obtain the expression (\ref{eqNumDiffVariance}) given in Section \ref{secTwoModeAnalytic}.
 

### 2012-11-14 large_number_squeezing_with_atoms.tex
Multimode discussion and abstract

@@
 \usepackage{graphicx}
+\usepackage{ulem}
 
@@
 
-\title{Iconic representations of molluscs through the ages}
+\title{Large squeezing in multicomponent trapped Bose-Einstein condensates}
 
@@
 \begin{abstract}
-More squeezing than you can shake a stick at.
-
+We examine the feasibility of creating and measuring large relative number squeezing in multicomponent trapped Bose-Einstein condensates.
+In the absence of multimode effects, this squeezing can be arbitrarily large, but a range of processes limit the measurable squeezing in realistic trap configurations.
+We conclude that high levels of squeezing with large numbers of atoms can only be achieved in one-dimensional traps with very even density profiles.
 \end{abstract}
@@
 
-There are three independent nonlinearities $\chi_{ij}$ in the Hamiltonian (\ref{eqTwoModeHamiltonian}).
-In the expressions for the number variances, however, the nonlinearities only appear in the cyclic combinations $\chi_{aa}-\chi_{ab}$, $\chi_{bb}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$, meaning that number squeezing is parameterized by only two independent quantities which will be linear combinations of the nonlinearities.
-For example, since $\chi_{bb}-\chi_{ab} = (\chi_{aa}-\chi_{ab}) - (\chi_{aa}-\chi_{bb})$ it is possible to parameterize the number squeezing entirely in terms of the nonlinear differences $\chi_{aa}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$.
+While there are three independent nonlinearities $\chi_{ij}$ in the Hamiltonian (\ref{eqTwoModeHamiltonian}), the expressions for the number variances only depend on differences.
+This means that number squeezing is parameterised by only two independent quantities, for example $\chi_{aa}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$.
 
 {\bf{There seems to be this bogus rule of thumb in the literature that the amount of squeezing you can get is related to $\chi_{11}+\chi_{22}-2\chi_{12}$.
-Is it worth pointing out that this is wrong?}}
+Is it worth pointing out that this is wrong?
+Joe:  Absolutely, but this might be true in some kind of weak squeezing limit, so we should check for some reference to this in the literature.}}
+%Can Uaa + Ubb - 2Uab be zero and still get squeezing? Yep, Uaa=0.03, Ubb=0.01, Uab=0.2 gives squeezing for all quantities of interest
 
-While the analytic expressions for the number variances are not particularly transparent, it is easy to evaluate and plot them for any given set of parameters.
+While the analytic expressions for the number variances are not particularly edifying, it is easy to evaluate and plot them for any given set of parameters.
 As an example, in Figure \ref{figxxx} we plot the solution for the normalized number variance in mode $|a\rangle$ for a particular set of nonlinearities, choosing a number of different values of $\phi$.
@@
 It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
-
-%Can Uaa + Ubb - 2Uab be zero and still get squeezing? Yep, Uaa=0.03, Ubb=0.01, Uab=0.2 gives squeezing for all quantities of interest
 
@@
 
-\section{Multimode background section}
+\section{Multimode analysis of squeezing}
+
+The two-mode model suggests that arbitrarily high squeezing can be achieved using the intrinsic nonlinearities in a BEC, but it ignores the existence of multiple spatial modes.
+The non-zero spatial dimensionality causes issues even when the modes do not interact.
+A brief inspection of the two-mode model shows that the wait time for optimal squeezing is highly number dependent.
+In a multimode environment, this means that the optimal wait time will be density-dependent.
+Furthermore, the phases of areas with different density will evolve at different rates, so there will be poor mode matching when recombining a field with significant density variation.
+
+\textbf{give a plot of the mode in $|a\rangle$ and $|b\rangle$ at the beginning and end of the hold time to show mismatch}
+
+We might speculate therefore that trap geometries that lead to near-constant density profiles for the BEC would enable higher squeezing.
+We cannot make strong conclusions without also considering the effects of coupling between spatial modes, however.
+An analytic solution to the higher dimensional problem is intractable, so we must use a numerical solution.
+Semiclassical approximations Without a semiclassical, umerical
+
 \begin{itemize}
-  \item Motivate why MM analysis is necessary
-  \item Brief recap why MM QFT is hard
+  \item \sout{Brief recap why MM QFT is hard}
   \item Say how great phase space methods are, quick overview of how they work, say only wigner works here, can probably crib some explanation and formulae from our fully quantum multimode atom laser paper
@@
   Derive link between mode shape, nonlinearity U, and the SM equivalent nonlinearity, so we can always compare a given high-dimensional multimode sim against the equivalent ideal analytic single mode case
-\end{itemize}
-
-\section{Meanfield multimode problems}
-\begin{itemize}
-  \item Point out that different densities have phases that evolve at different rates
-  \item This leads to poor mode matching when recombining if two states have different nonlinearity*number values 
-  \item give a plot of the mode in $|a\rangle$ and $|b\rangle$ at the beginning and end of the hold time to show mismatch
-  \item This means constant density is good since mode shape doesn't change
   \item Plot of squeezing for given set of params for gaussian, TF, constant density, show it get steadily better
@@
 
-\section{Mean field quantum statistical results}
+\section{The effects of dimensionality on squeezing}
   \begin{itemize}

### 2012-11-16 large_number_squeezing_with_atoms.tex
Add figure showing some analytic two-mode solutions
Also remove the [12pt] option so we're using the publication
layout rather than the proof layout, so we can get a better
idea of how long the paper will be.

@@
 %
-\documentclass[12pt]{iopart}
+\documentclass{iopart}
 %\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
@@
 While the analytic expressions for the number variances are not particularly edifying, it is easy to evaluate and plot them for any given set of parameters.
-As an example, in Figure \ref{figxxx} we plot the solution for the normalized number variance in mode $|a\rangle$ for a particular set of nonlinearities, choosing a number of different values of $\phi$.
-The plots illustrate the basic features of the problem: During the final coupling period, the number variance undergoes cyclic variation, with a series of minima.
+As an example, in Figure \ref{figTwoModeAnalyticExamples} we plot the solution for the normalized number variance in mode $|a\rangle$ for a particular set of nonlinearities, choosing a number of different values of $\phi$.
+The plots illustrate the basic features of the problem: During the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=0.5 \Omega$, and has a series of minima.
 It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
@@
     \label{figPulseScheme}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=10cm]{figures/analytic_two_mode_examples.eps}
+    \caption{Normalized number variance Var$[N_a]/N_a$ in mode $\hat{a}$ as a function of final recombination time plotted on a logarithmic scale, using four different mixing angles $\phi$.
+Parameters: $n_a = n_b =5 \times 10^5$, $\chi_{aa}=0.04 \hbar, \chi_{ab}=0, \chi_{bb}=0.01\hbar, \tau_{\mathrm{hold}}=4\times 10^{-4}$s.
+Solid line, $\phi=0.10$; dash-dotted line, $\phi=1.42$; dashed line, $\phi=3.24$; dotted line, $\phi=4.71$.}
+    \label{figTwoModeAnalyticExamples}
 \end{figure}

### 2012-11-16 large_number_squeezing_with_atoms.tex
Beginning of stochastic methods description

@@
 An analytic solution to the higher dimensional problem is intractable, so we must use a numerical solution.
-Semiclassical approximations Without a semiclassical, umerical
+
+We use stochastic methods based on phase space representations to simulate the dynamics of the multimode quantum fields.
+These methods alleviate the problem of directly simulating elements of the Hilbert space, which increases exponentially with the number of spatial modes \cite{stochasticRefs}.
+They do this by finding a sufficiently well-behaved quasi-probability representation for the density matrix, which can then be simulated as the average behaviour of a number of low-dimensional samples.
+These sample objects are the same size as the equivalent classical field.
+In our case, the only tractable phase space representation in widespread usage is the functional Wigner representation, which is well behaved when the quantum field is approximately Gaussian, as it is for coherent and squeezed states.
+Stochastic methods based on the functional Wigner representation have been used to model atom laser behaviour \cite{atomLaserWigner}, example2 \cite{example2Wigner}, and example3 \cite{example3Wigner}.
+
+The Hamiltonian for a pair of coupled atomic internal modes is the integral of the Hamiltonian density, which is given by:
+\begin{equation}
+\hat{\mathcal{H}} = \sum_j \hat{\psi}_j^{\dagger}\left(-\frac{\hbar}{2 m}\nabla^2+V_j(\mathbf{x})\right)\hat{\psi}_j 
+          + \sum_{i j}\frac{U_{i j}}{2} \hat{\psi}_i^{\dagger} \hat{\psi}_j^{\dagger} \hat{\psi}_j \hat{\psi}_i
+          + \Omega (\hat{\psi}_a^{\dagger} \hat{\psi}_b + \hat{\psi}_b^{\dagger}  \hat{\psi}_a )
+\label{eqFieldHamiltonian}
+\end{equation}
+where $\psi_j(\mathbf{x})$ is the annihilation field operator for a particle at position $\mathbf{x}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{x})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-mode nonlinearities, and $\Omega$ couples the two modes and allows for population transfer between them.
 
@@
   \item \sout{Brief recap why MM QFT is hard}
-  \item Say how great phase space methods are, quick overview of how they work, say only wigner works here, can probably crib some explanation and formulae from our fully quantum multimode atom laser paper
+  \item \sout{Say how great phase space methods are, quick overview of how they work, say only wigner works here, can probably crib some explanation and formulae from our fully quantum multimode atom laser paper}
   \item Talk about all the checks I've done to make sure Wigner holds over the parameters we've simulated despite the fact that TWA is an uncontrolled approximation

### 2012-11-19 large_number_squeezing_with_atoms.tex
More explanation of methods
Fairly loose, but how much detail will we need?  None of this is new.

--- 2012-11-16 large_number_squeezing_with_atoms.tex
@@
 \usepackage{ulem}
+\usepackage{amssymb}
 
@@
 An analytic solution to the higher dimensional problem is intractable, so we must use a numerical solution.
-Semiclassical approximations Without a semiclassical, umerical
 
+We use stochastic methods based on phase space representations to simulate the dynamics of the multimode quantum fields.
+These methods alleviate the problem of directly simulating elements of the Hilbert space, the size of which increases exponentially with the number of spatial modes \cite{stochasticRefs}.
+Stochastic methods achieve this by finding a sufficiently well-behaved quasi-probability representation for the density matrix, which can then be simulated as the average behaviour of a number of low-dimensional samples.
+These sample objects are the same size as the equivalent classical field.
+In our case, the only tractable phase space representation is the functional Wigner representation, which is well behaved when the quantum field is approximately Gaussian, as it is for coherent and squeezed states.
+Stochastic methods based on the functional Wigner representation have been used to model atom laser behaviour \cite{atomLaserWigner}, example2 \cite{example2Wigner}, and example3 \cite{example3Wigner}.
+
+The Hamiltonian for a pair of coupled atomic internal modes is the integral of the Hamiltonian density, which is given by:
+\begin{equation}
+\hat{\mathcal{H}} = \sum_{j} \hat{\psi}_j^{\dagger}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_j(\mathbf{x})\right)\hat{\psi}_j 
+          + \sum_{i j}\frac{U_{i j}}{2} \hat{\psi}_i^{\dagger} \hat{\psi}_j^{\dagger} \hat{\psi}_j \hat{\psi}_i
+          + \Omega \hat{\psi}_a^{\dagger} \hat{\psi}_b + \Omega^* \hat{\psi}_b^{\dagger}  \hat{\psi}_a
+\label{eqFieldHamiltonian}
+\end{equation}
+where $\hat{\psi}_j(\mathbf{x})$ is the annihilation field operator for a particle at position $\mathbf{x}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{x})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-mode nonlinearities, and $\Omega$ couples the two modes and allows for population transfer between them.
+This Hamiltonian defines the equation of motion for the density matrix, and thereby the functional Wigner distribution of the system, with which it has a one-to-one correspondence.
+The equation of motion for the functional Wigner distribution contains derivatives of third order which we assume to be negligible.
+This uncontrolled approximation is called the Truncated Wigner Approximation (TWA), and while it has been used widely on ultracold gases \cite{atomlaserWigner,orOtherUCGWigner}, care must be taken that the simulation remains valid.
+\textbf{Here we insert "all the checks I've done to make sure Wigner holds over the parameters we've simulated"}
+
+Under the TWA, the equation of motion for the functional Wigner distribution is a Fokker-Planck equation, and can therefore be sampled by a set of stochastic equations, as described in \cite{GardinerQuantumNoise}.
+These stochastic fields $\phi(\mathbf{x})$ are related to field operator expectation values by
+\begin{equation}
+\langle :\hat{\psi}^\dagger(\mathbf{x_1})\cdots\hat{\psi}^\dagger(\mathbf{x_n})\hat{\psi}(\mathbf{y_1})\cdots\hat{\psi}(\mathbf{y_m}):_{\mbox{sym}}\rangle = \mathbb{E}\left[\phi^*(\mathbf{x_1})\cdots\phi^*(\mathbf{x_n})\phi(\mathbf{y_1})\cdots\phi(\mathbf{y_m})\right]
+\label{eqExpectationRelations}
+\end{equation}
+%
+where $:\cdot:_{\mbox{sym}}$ denotes symmetric ordering of the operators, and $\mathbb{E}$ denotes a stochastic average over the variables $\phi(\mathbf{x})$.
+While we see below that the evolution of these fields is deterministic, the initial state still requires a random element, so multiple realisations are required.
+When approximating these equations on a discrete grid, the magnitude of initial noise, the relationship between stochastic averages and the expectation values, and the equations of motion all become grid-dependent.
+This is less concerning as it may at first appear, as the Hamiltonian in Eq.~(\ref{eqFieldHamiltonian}) assumes a contact potential between the atoms, and this is not correct below a length scale that can probe the details of the true potentials.
+Furthermore, well before that regime is reached, the different elements of grid dependence cancel out to give grid-independent predictions for physical observables.
+
+These stochastic equations take the form:
+\begin{eqnarray}
+\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} \left|\phi_{a}\right|^2 + \frac{U_{a b}}{2} \left|\phi_b\right|^2\right)\phi_a   + \Omega \phi_b  \nonumber \\
+\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} \left|\phi_{b}\right|^2 + \frac{U_{a b}}{2} \left|\phi_a\right|^2\right)\phi_b   + \Omega^* \phi_a
+  \label{eqStochasticEquations}
+\end{eqnarray}
+\textbf{I believe we have some rotation terms to include and explain.  Explanation can link to paragraph above}.
+
+\textbf{Note the initial state we use, and the equations.}
+
+We integrate equation (\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS \cite{Dennis2012}.
+The dimensional reduction is achieved by \textbf{Derive link between mode shape, nonlinearity U, and the SM equivalent nonlinearity, so we can always compare a given high-dimensional multimode sim against the equivalent ideal analytic single mode case}
+
+\section{Effects of trap geometries on squeezing in 1D}
 \begin{itemize}
-  \item \sout{Brief recap why MM QFT is hard}
-  \item Say how great phase space methods are, quick overview of how they work, say only wigner works here, can probably crib some explanation and formulae from our fully quantum multimode atom laser paper
-  \item Talk about all the checks I've done to make sure Wigner holds over the parameters we've simulated despite the fact that TWA is an uncontrolled approximation
-  \item Cite XMDS, and say we're amazing at doing QFT sims, send money
-  Derive link between mode shape, nonlinearity U, and the SM equivalent nonlinearity, so we can always compare a given high-dimensional multimode sim against the equivalent ideal analytic single mode case
   \item Plot of squeezing for given set of params for gaussian, TF, constant density, show it get steadily better
+\end{itemize}
+
+\section{Using a $\pi$ pulse to improve mode-matching
+\begin{itemize}
   \item Say we can use a pi pulse to swap the populations half way through, so if the states have different nonlinearities we still end up with the same mode shapes when we recombine.
@@
 
-\section{The effects of dimensionality on squeezing}
+\section{Squeezing in 2D and 3D}
   \begin{itemize}

### 2012-11-19 large_number_squeezing_with_atoms.tex
Making it compile

@@
 We conclude that high levels of squeezing with large numbers of atoms can only be achieved in one-dimensional traps with very even density profiles.
+We introduce a method of maximising the measurable squeezing by using a pi-pulse during the process to improve spatial mode-matching.
 \end{abstract}
@@
 In a multimode environment, this means that the optimal wait time will be density-dependent.
-Furthermore, the phases of areas with different density will evolve at different rates, so there will be poor mode matching when recombining a field with significant density variation.
+Furthermore, the phases of areas with different density will evolve at different rates, so there will be poor mode-matching when recombining a field with significant density variation.
 
@@
 
-\section{Using a $\pi$ pulse to improve mode-matching
+\section{Using a $\pi$ pulse to improve mode-matching}
 \begin{itemize}

### 2012-11-20 large_number_squeezing_with_atoms.tex
Expand on multimode problems

@@
 Parameters: $n_a = n_b =5 \times 10^5$, $\chi_{aa}=0.04 \hbar, \chi_{ab}=0, \chi_{bb}=0.01\hbar, \tau_{\mathrm{hold}}=4\times 10^{-4}$s.
-Solid line, $\phi=0.10$; dash-dotted line, $\phi=1.42$; dashed line, $\phi=3.24$; dotted line, $\phi=4.71$.}
+Solid line, $\phi=0.10$; dash-dotted line, $\phi=1.42$; dashed line, $\phi=3.24$; dotted line, $\phi=4.71$.
+These particular parameters yield 21dB of number squeezing for the $\phi=0.1$ and $\phi=1.42$ cases.}
     \label{figTwoModeAnalyticExamples}
@@
 
-The two-mode model suggests that arbitrarily high squeezing can be achieved using the intrinsic nonlinearities in a BEC, but it ignores the existence of multiple spatial modes.
-The non-zero spatial dimensionality causes issues even when the modes do not interact.
-A brief inspection of the two-mode model shows that the wait time for optimal squeezing is highly number dependent.
-In a multimode environment, this means that the optimal wait time will be density-dependent.
-Furthermore, the phases of areas with different density will evolve at different rates, so there will be poor mode-matching when recombining a field with significant density variation.
+The zero-dimensional, two-mode model described in the previous section shows that arbitrarily high squeezing can be achieved using the intrinsic nonlinearities in a BEC, but it ignores the existence of multiple spatial modes.
+
+If we allow the problem to have multiple spacial modes, there are a number of effects that prevent us from achieving the levels of number squeezing predicted by the zero-dimensional, two-mode analytic model.
+In this section we consider the two problems caused by the fact that the density in a spatially multimode problem will not be constant.
+
+The source of the first problem can be seen in the two-mode model.
+It shows that the parameters required to obtain best squeezing are number-dependent.
+That is, the optimum hold time $\tau_{\mathrm{hold}}$ depends on the number of atoms in the mode, as does the recombination phase $\phi$.
+In a multimode environment, this means that the optimal value of these parameters, rather than being number dependent, will be density-dependent, or, more precisely, dependent on the product of the local density and the strength of the nonlinearity in that mode.
+If the particle density is not constant across the spatial mode, different regions will squeeze at different rates (each region has a different optimum $\tau_{\mathrm{hold}}$) and each region will have a different optimum recombination phase $\phi$.
+This means that best we can do is choose values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across the range of densities, result in an overall lowest possible total number variance.
+
+The second problem arises from the fact that, due to the nonlinear term in the Hamiltonian, the local phase evolution of the atomic field is density dependent.
+If we write the order parameter of the atomic field as $\psi_i = \sqrt{\rho({\mathbf{r}})} \exp (i\varphi)$, where $j$ denotes one of the two states of our atomic field, the local phase evolution is given by
+\begin{equation}
+\frac{d \varphi}{dt} = \frac{U_{ii} \rho({\mathbf{r}})} {2 \hbar}.
+\end{equation}
+Consequently the mode shape of the atomic field changes depending on the nonlinearity-density product, which means that unless $U_{aa}|\psi_a({\mathbf{r}})|^2 = U_{bb}|\psi_b({\mathbf{r}})|^2$ for all ${\mathbf{r}}$, the mode shapes of will not overlap perfectly during the final coupling pulse.
+This degraded mode-matching will result in lower efficiency in converting the quadrature squeezing to number squeezing.
+We show an example of this effect in Figure~\ref{figXXX}.
 

### 2012-11-21 large_number_squeezing_with_atoms.tex
Add figure to paper
And add a preliminary caption so I don't forget the details

@@
 \end{itemize}
+This is shown in Figure~\ref{figModalSqueezingEffects1D}
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{figures/modal_squeezing_effects_1D.eps}
+    \caption{Effects of mode shape on squeezing.
+The plot shows the normalized number variance in mode $\hat{a}$ as a function of the time the final coupling pulse is applied.
+Thick red line, analytic two-mode solution; solid blue line, constant density mode; dashed line, Thomas-Fermi mode; dash-dotted line, gaussian mode.
+The nonlinearities of the various 1D simulations have been adjusted for their mode shape so that they are equivalent to that of the 0D, two-mode mode nonlinearity.
+Parameters: $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \hbar, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$.
+}
+    \label{figModalSqueezingEffects1D}
+\end{figure}
 

### 2012-11-21 large_number_squeezing_with_atoms.tex
spatial

@@
 As our goal is to describe the number squeezing possibilities of a two component BEC, we begin by considering a simplified version of the problem, consisting of a two-mode system, with the modes corresponding to the two states of the condensate.
-In the full problem, each of the components will include many spacial modes; our two mode system makes the assumption that only one of these spacial modes is relevant for each of the condensate's components.
+In the full problem, each of the components will include many spatial modes; our two mode system makes the assumption that only one of these spatial modes is relevant for each of the condensate's components.
 
@@
 
-If we allow the problem to have multiple spacial modes, there are a number of effects that prevent us from achieving the levels of number squeezing predicted by the zero-dimensional, two-mode analytic model.
+If we allow the problem to have multiple spatial modes, there are a number of effects that prevent us from achieving the levels of number squeezing predicted by the zero-dimensional, two-mode analytic model.
 In this section we consider the two problems caused by the fact that the density in a spatially multimode problem will not be constant.

### 2012-11-21 large_number_squeezing_with_atoms.tex
Brain dump of two-mode to multimode correspondence

@@
 \end{itemize}
-This is shown in Figure~\ref{figModalSqueezingEffects1D}
+To compare the effects of mode shape on squeezing, and to compare how close the a multimode system can get to the ideal two-mode system, it is necessary to have some method of mapping the multimode Hamiltonian to the two-mode Hamiltonian.
+We can do this by explicitly writing the multimode atomic field operator $\hat{\Psi} (x,t)$ in terms of a set of basis functions
+\begin{equation}
+\hat{\Psi}({\mathbf{r}},t) = \sum_n u_i({\mathbf{r}}) \hat{a}_i(t)
+\end{equation}
+where the $u_i$ are normalized mode functions. With this expansion the nonlinear terms in the multimode Hamiltonian becomes
+\begin{equation}
+\hat{H} = U \int  \sum_{ijkl} u_i^* u_j^* u_k u_l \, \hat{a}^{\dagger}_i \hat{a}^{\dagger}_j \hat{a}_k \hat{a}_l \, dV
+\end{equation}
+Ideally the system will remain predominantly in one mode, i.e.
+only one of the $u_i$ is significantly greater than zero.
+In this case we can make an equivalence between the multimode Hamiltonian and the two-mode Hamiltonian $\hat{H} = \chi \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}$ via
+\begin{equation}
+\chi = \int dV |u_i|^4  = \frac{1}{N^2} \int |\Psi|^4 \, dV
+\label{eqChiUequivalence}
+\end{equation}
+where $N$ is the number particles in the mode.
+
+Some common mode shapes are the groundstates of a harmonic trap with negligible nonlinearity (a gaussian), a harmonic trap with strong nonlinearity (Thomas-Fermi), and a constant potential (constant density).
+Using (\ref{eqChiUequivalence}) we find the following equivalences
+\begin{eqnarray}
+\chi &=& U \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}} \,\,\,\, ({\textrm{Gaussian}}) \\
+%
+\chi &=& \frac{4}{7} \left( \frac{15 U \omega_x \omega_y \omega_z}{16 \pi \sqrt{2}} \right)^{2/5} \left( \frac{m}{N} \right)^{3/5} \,\,\,\, ({\textrm{Thomas-Fermi}}) \\
+%
+\chi &=& \frac{U}{V} \,\,\,\, {\textrm{(constant density)}}
+\end{eqnarray}
+Choosing parameters where the mode shape of the multimode system doesn't change much during the evolution allows us to compare the effect of mode shape on the best squeezing that can be attained relative to the two mode system.
+An example in shown in Figure~\ref{figModalSqueezingEffects1D}.
+Clearly the closer the mode shape is to constant density, the better.
+
 \begin{figure}
@@
 \end{figure}
+
+As it can be difficult to engineer specific mode shapes, another alternative is to post-select a mode.
+For example, rather than considering the variance in the total number of atoms in a mode, it may be possible to only count, say the atoms in the central portion of the atomic cloud.
+For example, suppose one had a gaussian atomic distribution, but one considers a mode that only includes the atoms within $p$ standard deviations of the centre.
+For low $p$, this mode shape is considerably closer to the constant density case than a standard gaussian.
+The equivalence between the single mode case and this mode is given by
+\begin{equation}
+\chi = U \frac{{\mathrm{Erf}} (\sqrt{2p})^3}{{\mathrm{Erf}}(p)^6} \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}}
+\end{equation}
+And example comparing the best attainable squeezing for this mode compared to a single gaussian is shown in Figure~\ref{XXX}.
 

### 2012-11-22 large_number_squeezing_with_atoms.tex
Revised multimode motivation

@@
 The system is prepared with all the population in $|a\rangle$, and vacuum in $|b\rangle$.
-$|a\rangle$ is well-described by a coherent state {\bf *** Justify this.
-***} with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number in our system.
-
+We choose state $|a\rangle$ to be occupied as a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number in our system.
+As our simulation and experiment are insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, which is also equivalent to mixture of Poissonian-distributed number states.
+This state is consistent with BEC coherence experiments \cite{Hadzibabic2004}.
+%@article{Hadzibabic2004,
+%  title = {Interference of an Array of Independent Bose-Einstein Condensates},
+%  author = {Hadzibabic, Zoran and Stock, Sabine and Battelier, Baptiste and Bretin, Vincent and Dalibard, Jean},
+%  journal = {Phys. Rev. Lett.},
+%  volume = {93},
+%  issue = {18},
+%  pages = {180403},
+%  numpages = {4},
+%  year = {2004},
+%  month = {Oct},
+%  doi = {10.1103/PhysRevLett.93.180403},
+%  url = {http://link.aps.org/doi/10.1103/PhysRevLett.93.180403},
+%  publisher = {American Physical Society}
+%}
 At time $t=0$ the coupling $\Omega$ is switched on and applied until time $t=t_1$, at which point it is turned off, resulting in a portion of the population being transferred into mode $|a\rangle$.
@@
 The zero-dimensional, two-mode model described in the previous section shows that arbitrarily high squeezing can be achieved using the intrinsic nonlinearities in a BEC, but it ignores the existence of multiple spatial modes.
+There are three problems caused by the existence of multiple modes, and these will limit the achievable squeezing.
 
-If we allow the problem to have multiple spatial modes, there are a number of effects that prevent us from achieving the levels of number squeezing predicted by the zero-dimensional, two-mode analytic model.
-In this section we consider the two problems caused by the fact that the density in a spatially multimode problem will not be constant.
+The first problem is present even if the multiple spatial modes are uncoupled.
+The two-mode model shows that the parameters required to obtain best squeezing are dependent on the strength of the nonlinearity.
+This nonlinearity is proportional to the product of the mode shape and the number of atoms in that mode.
+The optimum hold time $\tau_{\mathrm{hold}}$ depends on this product, as does the recombination phase $\phi$.
+Different modes in a multimode environment will typically varying products of nonlinearity and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.
+This means that best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across the range of densities, result in an overall lowest possible total number variance.
 
-The source of the first problem can be seen in the two-mode model.
-It shows that the parameters required to obtain best squeezing are number-dependent.
-That is, the optimum hold time $\tau_{\mathrm{hold}}$ depends on the number of atoms in the mode, as does the recombination phase $\phi$.
-In a multimode environment, this means that the optimal value of these parameters, rather than being number dependent, will be density-dependent, or, more precisely, dependent on the product of the local density and the strength of the nonlinearity in that mode.
-If the particle density is not constant across the spatial mode, different regions will squeeze at different rates (each region has a different optimum $\tau_{\mathrm{hold}}$) and each region will have a different optimum recombination phase $\phi$.
-This means that best we can do is choose values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across the range of densities, result in an overall lowest possible total number variance.
-
-The second problem arises from the fact that, due to the nonlinear term in the Hamiltonian, the local phase evolution of the atomic field is density dependent.
+The second problem is one of the number-dependent dynamics of the spatial modes, which is observable even in a semiclassical simulation.
+Due to the nonlinear term in the Hamiltonian, the local phase evolution of the atomic field is density dependent.
 If we write the order parameter of the atomic field as $\psi_i = \sqrt{\rho({\mathbf{r}})} \exp (i\varphi)$, where $j$ denotes one of the two states of our atomic field, the local phase evolution is given by
 \begin{equation}
-\frac{d \varphi}{dt} = \frac{U_{ii} \rho({\mathbf{r}})} {2 \hbar}.
+\frac{d \varphi({\mathbf{r}})}{dt} = \frac{U_{ii} \rho({\mathbf{r}})} {2 \hbar}.
 \end{equation}
@@
 
-We might speculate therefore that trap geometries that lead to near-constant density profiles for the BEC would enable higher squeezing.
+We might speculate that the effects of these first two problems would be lessened by trap geometries that lead to near-constant density profiles for the BEC.
+This should enable higher squeezing.
 We cannot make strong conclusions without also considering the effects of coupling between spatial modes, however.
+The third potential problem is that coupling between modes is inevitable in the presence of a nonlinearity.
+Coupling between modes will disturb mode-matching, and mix fields with different phase evolution, which can rapidly destroy squeezing.
+
 An analytic solution to the higher dimensional problem is intractable, so we must use a numerical solution.
-
 We use stochastic methods based on phase space representations to simulate the dynamics of the multimode quantum fields.

### 2012-11-24 large_number_squeezing_with_atoms.tex
Fixed multimode discussion

@@
 
-The first problem is present even if the multiple spatial modes are uncoupled.
+The first problem is present even if the spatial modes are uncoupled.
 The two-mode model shows that the parameters required to obtain best squeezing are dependent on the strength of the nonlinearity.
-This nonlinearity is proportional to the product of the mode shape and the number of atoms in that mode.
+This nonlinearity is a function of the mode shape and the number of atoms in that mode.
 The optimum hold time $\tau_{\mathrm{hold}}$ depends on this product, as does the recombination phase $\phi$.
-Different modes in a multimode environment will typically varying products of nonlinearity and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.
-This means that best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across the range of densities, result in an overall lowest possible total number variance.
+Different modes in a multimode environment will typically have varying products of nonlinearity and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.
+This means that the best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across the range of densities, result in an overall lowest possible total number variance.
 
-The second problem is one of the number-dependent dynamics of the spatial modes, which is observable even in a semiclassical simulation.
+The second problem is due to the number-dependent dynamics of the spatial modes, which is observable even in a semiclassical simulation.
 Due to the nonlinear term in the Hamiltonian, the local phase evolution of the atomic field is density dependent.
-If we write the order parameter of the atomic field as $\psi_i = \sqrt{\rho({\mathbf{r}})} \exp (i\varphi)$, where $j$ denotes one of the two states of our atomic field, the local phase evolution is given by
-\begin{equation}
-\frac{d \varphi({\mathbf{r}})}{dt} = \frac{U_{ii} \rho({\mathbf{r}})} {2 \hbar}.
-\end{equation}
-Consequently the mode shape of the atomic field changes depending on the nonlinearity-density product, which means that unless $U_{aa}|\psi_a({\mathbf{r}})|^2 = U_{bb}|\psi_b({\mathbf{r}})|^2$ for all ${\mathbf{r}}$, the mode shapes of will not overlap perfectly during the final coupling pulse.
+Consequently the mode shape of the atomic field changes depending on the nonlinearity-density product, which means that unless $U_{aa}|\psi_a({\mathbf{r}})|^2 = U_{bb}|\psi_b({\mathbf{r}})|^2$ for all ${\mathbf{r}}$, the mode shapes will not overlap perfectly during the final coupling pulse.
 This degraded mode-matching will result in lower efficiency in converting the quadrature squeezing to number squeezing.

### 2012-11-27 large_number_squeezing_with_atoms.tex
Add figure showing dimensional effects on squeezing
Moving full Hamiltonian and dimensional reduction to first section
Reduces the extra redundancy.  We repeat things less this way.

@@
 \end{itemize}
+Our goal is to describe the number squeezing possibilities of a BEC comprising two relevant internal states denoted by $|a\rangle$ and $|b\rangle$.
+The Hamiltonian for a pair of coupled atomic internal modes is the spatial integral of the Hamiltonian density, which is given by:
+\begin{equation}
+\hat{\mathcal{H}} = \sum_{j} \hat{\psi}_j^{\dagger}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_j\right)\hat{\psi}_j 
+          + \sum_{i j}\frac{U_{i j}}{2} \hat{\psi}_i^{\dagger} \hat{\psi}_j^{\dagger} \hat{\psi}_j \hat{\psi}_i
+          + \kappa \hat{\psi}_a^{\dagger} \hat{\psi}_b + \kappa^* \hat{\psi}_b^{\dagger}  \hat{\psi}_a
+\label{eqFieldHamiltonian}
+\end{equation}
+where $\hat{\psi}_j(\mathbf{x})$ is the annihilation field operator for a particle at position $\mathbf{x}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{x})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-mode nonlinearities, and $\kappa$ describes the induced coupling between the two modes.
 
-As our goal is to describe the number squeezing possibilities of a two component BEC, we begin by considering a simplified version of the problem, consisting of a two-mode system, with the modes corresponding to the two states of the condensate.
-In the full problem, each of the components will include many spatial modes; our two mode system makes the assumption that only one of these spatial modes is relevant for each of the condensate's components.
-
-{\bf{Maybe expand on the link between a full BEC in a trap and a the following hamiltonian?}}
-
-We allow the atoms to interact via a standard Kerr-type nonlinearity arising from atom-atom interactions, as well as allowing a linear coupling between the two modes.
-Denoting the two modes by $|a\rangle$ and $|b\rangle$, the Hamiltonian governing this system is given by
+We begin by considering a simplified version of the problem, assuming that only one of these spatial modes is relevant for each of the condensate's components.
+This is done by writing the multimode atomic field operators in terms of a set of basis functions $\hat{\psi}_a({\mathbf{r}},t) = \sum_n u_{n j}({\mathbf{r}}) \hat{a}_{n}(t)$ and $\hat{\psi}_b({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{b}_{n}(t)$, where $\hat{a}_{n}$ and $\hat{b}_{n}$ annihilate a particle with normalised spatial mode $u_n$ and internal state $|a\rangle$ or $|b\rangle$ respectively.
+With this expansion, the first nonlinear term in the multimode Hamiltonian becomes
 \begin{equation}
-\hat{H} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b} 
-          + \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
+\hat{H} = \frac{U_{aa}}{2} \int dV \sum_{nmpq} u_n^* u_m^* u_p u_q \, \hat{a}^{\dagger}_{n} \hat{a}^{\dagger}_{m} \hat{a}_{p} \hat{a}_{q} 
+\end{equation}
+When the system remains predominantly in one mode, we can approximate this sum as the single term $\hat{H} = \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}$ where
+\begin{equation}
+\chi_{aa} = U_{aa} \int |u_i|^4 \, dV.
+\label{eqChiUequivalence}
+\end{equation}
+We proceed similarly for the other terms.
+In the parameter regimes we are interested in, the detuning between the two spatial modes will be irrelevant, so setting our zero of energy appropriately, we are left with only a standard Kerr-type nonlinearity arising from atom-atom interactions, as well as a linear coupling between the two modes.
+The Hamiltonian governing this reduced system is given by
+\begin{equation}
+\hat{H} = \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
           + \chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
@@
 After this hold time $\tau_{\mathrm{hold}} = t_2 - t_1 $, the coupling $\Omega$ is switched back on until time $t_3$, with a phase shift $\phi$ compared to it's first application.
-During this last stage, the two modes exchange population and the quadrature variance is converted into number variance, in the same way homodyne measurements are used in quantum optics to convert quadrature squeezing (which cannot be directly measured) into number squeezing (which can).
+During this last stage, the two modes exchange population and the quadrature variance is converted into number variance, in the same way homodyne measurements are used in quantum optics to convert quadrature squeezing into number squeezing, which can be directly measured.
 In this interpretation, $\phi$ is the relative phase of the strong local oscillator, which allows specific phase angles of quadrature squeezing to be examined.
-
-In another interpretation, the sequence of pulses results in a Ramsey interferometer, with a final beam splitter phase of $\phi$.
+This experiment can also be interpreted as a Ramsey interferometer with a final beam splitter phase of $\phi$.
 
 The entire sequence of pulses and the resulting populations are shown schematically in Figure~\ref{figPulseScheme}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{figures/pulse_scheme.eps}
+    \caption{Schematic of the pulse sequence and atomic populations as a function of time.
+Typically the coupling sequences at the beginning and end of the sequence are much shorter than the hold time, $\tau_{\mathrm{hold}} = t_2 - t_1 $, which is where the majority of the squeezing occurs.
+\textbf{Unless I just lied}.}
+    \label{figPulseScheme}
+\end{figure}
 
@@
 In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, e.g.
-${\mathrm{Var}}[N_a]=N_a$, an no number squeezing is possible.
+${\mathrm{Var}}[N_a]=N_a$, and no number squeezing is possible.
 
@@
 It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=8cm]{figures/pulse_scheme.eps}
-    \caption{Mollusc mollusc}
-    \label{figPulseScheme}
-\end{figure}
 
@@
 
-The Hamiltonian for a pair of coupled atomic internal modes is the integral of the Hamiltonian density, which is given by:
-\begin{equation}
-\hat{\mathcal{H}} = \sum_{j} \hat{\psi}_j^{\dagger}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_j(\mathbf{x})\right)\hat{\psi}_j 
-          + \sum_{i j}\frac{U_{i j}}{2} \hat{\psi}_i^{\dagger} \hat{\psi}_j^{\dagger} \hat{\psi}_j \hat{\psi}_i
-          + \Omega \hat{\psi}_a^{\dagger} \hat{\psi}_b + \Omega^* \hat{\psi}_b^{\dagger}  \hat{\psi}_a
-\label{eqFieldHamiltonian}
-\end{equation}
-where $\hat{\psi}_j(\mathbf{x})$ is the annihilation field operator for a particle at position $\mathbf{x}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{x})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-mode nonlinearities, and $\Omega$ couples the two modes and allows for population transfer between them.
-This Hamiltonian defines the equation of motion for the density matrix, and thereby the functional Wigner distribution of the system, with which it has a one-to-one correspondence.
+The equation of motion for the density matrix is defined by the Hamiltonian in Eq.~(\ref{eqFieldHamiltonian}).
+This defines the evolution of the functional Wigner distribution of the system, which has a one-to-one correspondence with the density matrix.
 The equation of motion for the functional Wigner distribution contains derivatives of third order which we assume to be negligible.
@@
 %
-where $:\cdot:_{\mbox{sym}}$ denotes symmetric ordering of the operators, and $\mathbb{E}$ denotes a stochastic average over the variables $\phi(\mathbf{x})$.
+where $:\star:_{\mbox{sym}}$ denotes symmetric ordering of the operators, and $\mathbb{E}$ denotes a stochastic average over the variables $\phi(\mathbf{x})$.
 While we see below that the evolution of these fields is deterministic, the initial state still requires a random element, so multiple realisations are required.
@@
 \begin{eqnarray}
-\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} \left|\phi_{a}\right|^2 + \frac{U_{a b}}{2} \left|\phi_b\right|^2\right)\phi_a   + \Omega \phi_b  \nonumber \\
-\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} \left|\phi_{b}\right|^2 + \frac{U_{a b}}{2} \left|\phi_a\right|^2\right)\phi_b   + \Omega^* \phi_a
+\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} \left|\phi_{a}\right|^2 + \frac{U_{a b}}{2} \left|\phi_b\right|^2\right)\phi_a   + \kappa \phi_b  \nonumber \\
+\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} \left|\phi_{b}\right|^2 + \frac{U_{a b}}{2} \left|\phi_a\right|^2\right)\phi_b   + \kappa^* \phi_a
   \label{eqStochasticEquations}
@@
 \end{itemize}
-To compare the effects of mode shape on squeezing, and to compare how close the a multimode system can get to the ideal two-mode system, it is necessary to have some method of mapping the multimode Hamiltonian to the two-mode Hamiltonian.
-We can do this by explicitly writing the multimode atomic field operator $\hat{\Psi} (x,t)$ in terms of a set of basis functions
-\begin{equation}
-\hat{\Psi}({\mathbf{r}},t) = \sum_n u_i({\mathbf{r}}) \hat{a}_i(t)
-\end{equation}
-where the $u_i$ are normalized mode functions. With this expansion the nonlinear terms in the multimode Hamiltonian becomes
-\begin{equation}
-\hat{H} = U \int  \sum_{ijkl} u_i^* u_j^* u_k u_l \, \hat{a}^{\dagger}_i \hat{a}^{\dagger}_j \hat{a}_k \hat{a}_l \, dV
-\end{equation}
-Ideally the system will remain predominantly in one mode, i.e.
-only one of the $u_i$ is significantly greater than zero.
-In this case we can make an equivalence between the multimode Hamiltonian and the two-mode Hamiltonian $\hat{H} = \chi \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}$ via
-\begin{equation}
-\chi = \int dV |u_i|^4  = \frac{1}{N^2} \int |\Psi|^4 \, dV
-\label{eqChiUequivalence}
-\end{equation}
-where $N$ is the number particles in the mode.
 
@@
   \end{itemize}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing.eps}
+    \label{figDimensionalSqueezingEffects:sub1}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing.eps}
+    \label{figDimensionalSqueezingEffects:sub2}
+  \end{subfigure}
+\caption{Effects of dimension on squeezing in two different nonlinear regimes.
+If two nonlinearities are non-zero, the deleterious effects of higher dimensions on squeezing are exacerbated.
+Thick red line, analytic two-mode solution; solid blue line, 1D; dashed line, 2D; dash-dotted line, 3D.
+(a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.
+(b) $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{bb}=\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.}
+  \label{figDimensionalSqueezingEffects}
+\end{figure}
 

### 2012-11-27 large_number_squeezing_with_atoms.tex
Add packages so it actually compiles

@@
 \usepackage{amssymb}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 We integrate equation (\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS \cite{Dennis2012}.
-The dimensional reduction is achieved by \textbf{Derive link between mode shape, nonlinearity U, and the SM equivalent nonlinearity, so we can always compare a given high-dimensional multimode sim against the equivalent ideal analytic single mode case}
+The dimensional reduction is achieved by estimating the mode shape in three dimensions, and using dimensionally reduced values for $U_{ij}$ that match the zero-dimensional reduction given by Eq.~(\ref{eqChiUequivalence}).
+An equivalent process is to match the chemical potentials of the initial states.
 
 \section{Effects of trap geometries on squeezing in 1D}
-\begin{itemize}
-  \item Plot of squeezing for given set of params for gaussian, TF, constant density, show it get steadily better
-\end{itemize}
 
-Some common mode shapes are the groundstates of a harmonic trap with negligible nonlinearity (a gaussian), a harmonic trap with strong nonlinearity (Thomas-Fermi), and a constant potential (constant density).
+A one-dimensional simulation is sufficient to examine the hypothesis that the squeezing will be higher for a BEC where the spatial mode has a more constant density profile.
+We compare squeezing for fields starting in the ground states of a harmonic trap with negligible nonlinearity (a Gaussian), a harmonic trap with strong nonlinearity (Thomas-Fermi), and a constant potential (constant density).
 Using (\ref{eqChiUequivalence}) we find the following equivalences
@@
 \end{eqnarray}
-Choosing parameters where the mode shape of the multimode system doesn't change much during the evolution allows us to compare the effect of mode shape on the best squeezing that can be attained relative to the two mode system.
-An example in shown in Figure~\ref{figModalSqueezingEffects1D}.
-Clearly the closer the mode shape is to constant density, the better.
+\textbf{Choosing parameters where the mode shape of the multimode system doesn't change much during the evolution allows us to compare the effect of mode shape on the best squeezing that can be attained relative to the two mode system.
+Joe: We need to be more specific here.}
+
+The results of this simulation are shown in Figure~\ref{figModalSqueezingEffects1D}, which clearly confirms the hypothesis that . Clearly the closer the mode shape is to constant density, the better.
 
@@
 The plot shows the normalized number variance in mode $\hat{a}$ as a function of the time the final coupling pulse is applied.
-Thick red line, analytic two-mode solution; solid blue line, constant density mode; dashed line, Thomas-Fermi mode; dash-dotted line, gaussian mode.
+Thick red line, analytic two-mode solution; solid blue line, constant density mode; dashed line, Thomas-Fermi mode; dash-dotted line, Gaussian mode.
 The nonlinearities of the various 1D simulations have been adjusted for their mode shape so that they are equivalent to that of the 0D, two-mode mode nonlinearity.
@@
     \centering
-    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing.eps}
+    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_1.eps}
     \label{figDimensionalSqueezingEffects:sub1}
@@
     \centering
-    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing.eps}
+    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_2.eps}
     \label{figDimensionalSqueezingEffects:sub2}

### 2012-11-28 large_number_squeezing_with_atoms.tex
Revert "Merge branch 'master' of https://github.com/anu-atom-optics/larger-number-squeezing-paper"
This reverts commit ed82b71f17ab3e4470a083b7cdbdeeeab4a2ed27, reversing
changes made to 835afe083c5c2e478c35600a99e183d36028b417.
Add figure to paper

@@
 Thick red line, analytic two-mode solution; solid blue line, 1D; dashed line, 2D; dash-dotted line, 3D.
-(a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.
+(a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=6.2$.
 (b) $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{bb}=\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.}
   \label{figDimensionalSqueezingEffects}
+\end{figure}
+
+\begin{figure}
+  \centering
+   \includegraphics[width=7cm]{figures/volume_effect_on_squeezing.eps}
+\caption{(most of this caption should be moved to main text and expanded) Effects of volume on squeezing.
+2D simulation of how the size of the box affects squeezing.
+Horizontal axis shows the linear size of the edges of the box in microns.
+The larger the box, the more low-momentum modes are included, which means worse squeezing (see Bog analysis).
+Dashed red line, analytic zero-D solution; solid blue line, relative number variance in mode $|a\rangle$.
+All simulations are 128x128 grid, 10,000 paths.
+Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=6.2$.}
+ \label{figVolumeSqueezingEffects}
 \end{figure}

### 2012-11-28 large_number_squeezing_with_atoms.tex
The edge of what I can write/claim without results.

@@
 \usepackage{caption}
+\captionsetup{compatibility=false}
 \usepackage{subcaption}
@@
 
-\section{Multimode analysis of squeezing}
+\section{Multimode analysis of squeezing} \label{sec:MMdescription}
 
@@
 \end{eqnarray}
-\textbf{Choosing parameters where the mode shape of the multimode system doesn't change much during the evolution allows us to compare the effect of mode shape on the best squeezing that can be attained relative to the two mode system.
-Joe: We need to be more specific here.}
-
-The results of this simulation are shown in Figure~\ref{figModalSqueezingEffects1D}, which clearly confirms the hypothesis that . Clearly the closer the mode shape is to constant density, the better.
+%
+Choosing parameters such that the evolution is sufficiently fast that the mode shape is reasonably stable over the simulation gives the best squeezing, and allows us to make a fair comparison between the squeezing achievable assuming that mode-matching is near optimal.
+The results of this simulation are shown in Figure~\ref{figModalSqueezingEffects1D}, which clearly confirms the hypothesis that maximum squeezing is obtained when the mode shape is closest to constant density.
+In practice, this isThis is largely due to the fact that the squeezing rates become non-uniform, but slower dynamics will also lead to issues with mode-matching.
 
@@
 Parameters: $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \hbar, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$.
-}
+Clearly, the maximum squeezing is achieved when the atomic density profile is flattest.}
     \label{figModalSqueezingEffects1D}
@@
 
-As it can be difficult to engineer specific mode shapes, another alternative is to post-select a mode.
-For example, rather than considering the variance in the total number of atoms in a mode, it may be possible to only count, say the atoms in the central portion of the atomic cloud.
-For example, suppose one had a gaussian atomic distribution, but one considers a mode that only includes the atoms within $p$ standard deviations of the centre.
-For low $p$, this mode shape is considerably closer to the constant density case than a standard gaussian.
+One way to avoid the requirement to engineer and maintain specific atomic modes is to post-select a mode.
+Rather than considering the variance in the total number of atoms in a mode, it may be possible to consider only the statistics of a spatially filtered area of the atomic cloud.
+For example, suppose the atoms were in a Gaussian atomic distribution, but we consider a mode that only includes the atoms within $p$ standard deviations of the centre.
+For low $p$, this mode shape is considerably closer to the constant density case than a standard Gaussian.
 The equivalence between the single mode case and this mode is given by
@@
 \end{equation}
-And example comparing the best attainable squeezing for this mode compared to a single gaussian is shown in Figure~\ref{XXX}.
+A simulations comparing the best attainable squeezing for this mode compared to an unfiltered Gaussian is shown in Figure~\ref{XXX}.
+We can see that the squeezing is higher.
+One obvious disadvantage to post-selecting modes is that it can severely reduce the effective number of atoms.
+This is a serious issue for most applications of squeezed sources, such as interferometry, where the signal to noise scales with the square root of the flux.
 
-\section{Using a $\pi$ pulse to improve mode-matching}
+\section{Using a $\pi$-pulse to improve mode-matching}
+Modal mismatch is generated by the different atomic potentials seen by the two internal states.
+We propose a straightforward method of minimising this effect: using a $\pi$-pulse to swap the internal states after half of the evolution.
+In the limit where each state stayed in a single mode (although that mode might change), and where the nonlinear terms did not change with time, this method would produce identical mode shapes for both pulses after the hold time.
+In practice, neither of these conditions will hold, particularly when the squeezing is completely removed by multimode effects, but the $\pi$-pulse should produce some improvement in regimes where the squeezing is only moderately reduced.
+
 \begin{itemize}
-  \item Say we can use a pi pulse to swap the populations half way through, so if the states have different nonlinearities we still end up with the same mode shapes when we recombine.
-The fact that this works will be clear from the SM equations for a(t), b(t) when the beam splitter is on
   \item show some plots of squeezing with and without the pi pulse
@@
 
+In Fig.~\ref{piPulseFig} we show the effects of a $\pi$-pulse on the squeezing generated \textbf{by some specific system}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{figures/piPulseFig.eps}
+    \caption{This is a picture of a $\pi$-pulse working.}
+    \label{piPulseFig}
+\end{figure}
+
 \section{Squeezing in 2D and 3D}
+The simulations in the previous two sections were one dimensional, and show how having spatially constant density and using a $\pi$-pulse will improve the resultant squeezing.
+While these conclusions remain valid for simulations in two or three dimensions, adding each additional dimension results in a degradation of the squeezing.
+This is true even when the potential (and the resulting atomic density) is constant across the bulk of the trap.
+Furthermore, the expected squeezing goes down as the volume of the trap increases.
+This can be seen in Fig.~\ref{1D2D3DworsererFig}, where we show the squeezing generated \textbf{by 1D, 2D and 3D BECs in a box with some parameters}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{figures/1D2D3DBox.eps}
+    \caption{Even BECs in a box get worse with both dimension and volume.}
+    \label{1D2D3DworsererFig}
+\end{figure}
+
+This dependence on dimensionality and volume are not explained by our arguments in section \ref{sec:MMdescription}, which focus on the effects of inhomogeneity within the trap.
+Fundamentally, this degradation of the squeezing is due to nonlinearity-induced coupling between different spatial modes.
+As the dimension and volume increases, the spectral density of available excited modes increases, leading to a reduction in squeezing.
+
+Excitations of a BEC are described by the Bogoliubov modes...
+
   \begin{itemize}
-  \item Two issues. 1) as we add dimensions results get worse (maybe point out how cool we are for being able to do multimode 3D simulations). 2) As our volume increases results get worse.
-  \item To gain insight into this we do a bog analysis
+  \item Introduce Bog analysis
   \item Do a comparision plot between SM analytic model and Bog to show everything agrees
@@
   \begin{itemize}
-  \item One way to approximate constant density is to look at core of BEC, since if we only look at number in the central 1 sigma portion of gaussian it's close to flat; show results compared to single mode equivalent
-
-  \item Another improvement is to put the BEC in a box since then we get constant density, discuss feasibility
   \item 3D still sucks even with box, so point out we can win by freezing out dimensions
@@
 
-The criteria we need to accomplish this depends on the number of dimensions we want to freeze out, so investigate the cases where we reduce a 3D system to an effective 0D, 1D, and 2D system seperately.
+The criteria we need to accomplish this depends on the number of dimensions we want to freeze out, so investigate the cases where we reduce a 3D system to an effective 0D, 1D, and 2D system separately.
 In all cases we assume we have a box potential with length $L$ in the dimension(s) that aren't frozen out, and tight harmonic trapping potential of frequency $\omega$ in the frozen dimensions.
@@
 \ack
-Acknowledgements in here; funding sources etc
+Acknowledgements in here; funding sources etc.  Joe's Future Fellowship.
 
@@
 \label{appendixTwoModeDerivation}
-In this Appendix we derive the equations for the number variances which were earlier stated without proof.
+In this Appendix we derive the equations for the number variances.
 We use the model described in Section \ref{secTwoModeAnalytic}, where we consider a two mode system with $N$ particles, initially prepared as a coherent state with all the population in mode $|a\rangle$ at time $t=0$.

### 2012-11-30 large_number_squeezing_with_atoms.tex
Inserted pi pulse figure into paper

@@
     \centering
-    \includegraphics[width=8cm]{figures/piPulseFig.eps}
-    \caption{This is a picture of a $\pi$-pulse working.}
+    \includegraphics[width=8cm]{figures/pi_pulse.eps}
+    \caption{Effect of a $\pi$-pulse halfway through the hold time.
+Common parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{ab}=\chi_{bb}=0$.
+Solid red line: Analytic two-mode solution, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (10.9dB squeezing); Solid blue line: no $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (4.9dB squeezing);  Dashed blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.05$ (5.8dB squeezing); Dash-dotted blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=5\times 10^{-4}$s, $\phi=3.87$ (7.2dB squeezing).}
     \label{piPulseFig}
@@
 Furthermore, the expected squeezing goes down as the volume of the trap increases.
-This can be seen in Fig.~\ref{1D2D3DworsererFig}, where we show the squeezing generated \textbf{by 1D, 2D and 3D BECs in a box with some parameters}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=8cm]{figures/1D2D3DBox.eps}
-    \caption{Even BECs in a box get worse with both dimension and volume.}
-    \label{1D2D3DworsererFig}
-\end{figure}
+This can be seen in Fig.~\ref{figDimensionalSqueezingEffects}, where we show the squeezing generated \textbf{by 1D, 2D and 3D BECs in a box with some parameters}.
 

### 2012-12-07 large_number_squeezing_with_atoms.tex
Add realistic squeezing figure to paper

@@
 
+Taking all the lessons learned, we can get good squeezing for realistic paramters in 3D.
+Figure \ref{figXXX} shows an example of a full 3D simulation, 10000 paths, box potential in 2d, tight trapping (1500Hz) in the 3rd dimension.
+\begin{figure}
+  \centering
+   \includegraphics[width=7cm]{figures/realistic_3d_squeezing_params.eps}
+\caption{Full 3D simulation of relative number squeezing in state $|a\rangle$ in a system with realistic parameters.
+Box potential in two dimensions with area 120$\mu$m $\times$ 120$\mu$m, tight harmonic trapping in 3rd dimension with $\omega=1500$Hz.
+Solid red line, analytic 0D solution; dashed blue line, full 3D multimode simulation.
+Parameters: Rb$^{85}$ atoms, $n_a = n_b =2 \times 10^5$, $U_{aa}=2.1\times 10^{-50}$ (corresponding to an s-wave scattering length of $-400 a_0$), $U_{bb}=U_{ab}=0$, $\tau_{\mathrm{hold}}=1\times 10^{-3}$s, $\phi=5.3$.
+Even with all multimode effects present, the number squeezing in the 3D system (13.9dB) approaches that of the ideal 0D case (15.1dB).}
+ \label{figRealistic3DParameters}
+\end{figure}
+
 \section{Conclusion}

### 2012-12-12 large_number_squeezing_with_atoms.tex
Yet more paper progress

@@
 In the absence of multimode effects, this squeezing can be arbitrarily large, but a range of processes limit the measurable squeezing in realistic trap configurations.
-We conclude that high levels of squeezing with large numbers of atoms can only be achieved in one-dimensional traps with very even density profiles.
-We introduce a method of maximising the measurable squeezing by using a pi-pulse during the process to improve spatial mode-matching.
+We examine these processes, and suggest methods to mitigate them.
+We conclude that high levels of squeezing with large numbers of atoms is feasible, but can realistically only be achieved in traps with even density profiles in at least two dimensions.
+We also introduce a method of maximising the measurable squeezing by using a pi-pulse during the process to improve spatial mode-matching.
 \end{abstract}
@@
 Our goal is to describe the number squeezing possibilities of a BEC comprising two relevant internal states denoted by $|a\rangle$ and $|b\rangle$.
-The Hamiltonian for a pair of coupled atomic internal modes is the spatial integral of the Hamiltonian density, which is given by:
+The Hamiltonian for a pair of coupled atomic internal modes is the spatial integral of the Hamiltonian density, which is given by
 \begin{equation}
@@
 \end{equation}
-where $\hat{\psi}_j(\mathbf{x})$ is the annihilation field operator for a particle at position $\mathbf{x}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{x})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-mode nonlinearities, and $\kappa$ describes the induced coupling between the two modes.
+where $\hat{\psi}_j(\mathbf{x})$ is the annihilation field operator for a particle at position $\mathbf{x}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{x})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-state nonlinearities, and $\kappa$ describes the coupling between the two states.
 
 We begin by considering a simplified version of the problem, assuming that only one of these spatial modes is relevant for each of the condensate's components.
-This is done by writing the multimode atomic field operators in terms of a set of basis functions $\hat{\psi}_a({\mathbf{r}},t) = \sum_n u_{n j}({\mathbf{r}}) \hat{a}_{n}(t)$ and $\hat{\psi}_b({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{b}_{n}(t)$, where $\hat{a}_{n}$ and $\hat{b}_{n}$ annihilate a particle with normalised spatial mode $u_n$ and internal state $|a\rangle$ or $|b\rangle$ respectively.
+This is done by writing the multimode atomic field operators in terms of a set of basis functions $\hat{\psi}_a({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{a}_{n}(t)$ and $\hat{\psi}_b({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{b}_{n}(t)$, where $\hat{a}_{n}$ and $\hat{b}_{n}$ annihilate a particle with normalised spatial mode $u_n$ and internal state $|a\rangle$ or $|b\rangle$ respectively.
 With this expansion, the first nonlinear term in the multimode Hamiltonian becomes
@@
 \end{equation}
-When the system remains predominantly in one mode, we can approximate this sum as the single term $\hat{H} = \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}$ where
+When the system remains predominantly in one mode, we can approximate this sum as the single term $\hat{H} = \chi_{aa} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}/2$ where
 \begin{equation}
@@
 This state is consistent with BEC coherence experiments \cite{Hadzibabic2004}.
-%@article{Hadzibabic2004,
-%  title = {Interference of an Array of Independent Bose-Einstein Condensates},
-%  author = {Hadzibabic, Zoran and Stock, Sabine and Battelier, Baptiste and Bretin, Vincent and Dalibard, Jean},
-%  journal = {Phys. Rev. Lett.},
-%  volume = {93},
-%  issue = {18},
-%  pages = {180403},
-%  numpages = {4},
-%  year = {2004},
-%  month = {Oct},
-%  doi = {10.1103/PhysRevLett.93.180403},
-%  url = {http://link.aps.org/doi/10.1103/PhysRevLett.93.180403},
-%  publisher = {American Physical Society}
-%}
-At time $t=0$ the coupling $\Omega$ is switched on and applied until time $t=t_1$, at which point it is turned off, resulting in a portion of the population being transferred into mode $|a\rangle$.
+
+At time $t=0$ the coupling $\Omega$ is switched on and applied until time $t=t_1$, at which point it is turned off, resulting in a portion of the population being transferred into mode $|b\rangle$.
 As the initial state was a coherent state in both modes, after the population transfer both $|a\rangle$ and $|b\rangle$ are described by coherent states, with mean number $\langle \hat{a}^{\dagger} \hat{a} \rangle = n_a$ and $\langle \hat{b}^{\dagger} \hat{b} \rangle = n_b$ respectively.
@@
 During this period the atoms interact solely through the nonlinear terms, with no population being transferred.
-Due to the form of the Hamiltonian, the number variance remains unchanged, but the quadrature variances in the two modes change over time, which can lead to quadrature squeezing \cite{xxx}.
+Due to the form of the Hamiltonian, the number variance remains unchanged, but the quadrature variances in the two modes change over time, which can lead to quadrature squeezing \cite{johnssonET2007}.
 
-After this hold time $\tau_{\mathrm{hold}} = t_2 - t_1 $, the coupling $\Omega$ is switched back on until time $t_3$, with a phase shift $\phi$ compared to it's first application.
+After this hold time $\tau_{\mathrm{hold}} = t_2 - t_1 $, the coupling $\Omega$ is switched back on until time $t_3$, with a phase shift $\phi$ compared to its first application.
 During this last stage, the two modes exchange population and the quadrature variance is converted into number variance, in the same way homodyne measurements are used in quantum optics to convert quadrature squeezing into number squeezing, which can be directly measured.
@@
     \caption{Schematic of the pulse sequence and atomic populations as a function of time.
-Typically the coupling sequences at the beginning and end of the sequence are much shorter than the hold time, $\tau_{\mathrm{hold}} = t_2 - t_1 $, which is where the majority of the squeezing occurs.
-\textbf{Unless I just lied}.}
+Typically the coupling sequences at the beginning and end of the sequence are much shorter than the hold time, $\tau_{\mathrm{hold}} = t_2 - t_1 $, which is where the majority of the squeezing occurs.}
     \label{figPulseScheme}
@@
 \end{eqnarray}
-then the expectation values for number, number variance, and number difference variance after appling the final coupling field for a time $\tau = t_3-t_2$ are given by
+then the expectation values for number, number variance, and number difference variance after appling the final coupling field are given by
 \begin{eqnarray}
@@
 ${\mathrm{Var}}[N_a]=N_a$, and no number squeezing is possible.
-
-While there are three independent nonlinearities $\chi_{ij}$ in the Hamiltonian (\ref{eqTwoModeHamiltonian}), the expressions for the number variances only depend on differences.
+It should also be noted that, although there are three independent nonlinearities $\chi_{ij}$ in the Hamiltonian (\ref{eqTwoModeHamiltonian}), the expressions for the number variances only depend on differences.
 This means that number squeezing is parameterised by only two independent quantities, for example $\chi_{aa}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$.
 
-{\bf{There seems to be this bogus rule of thumb in the literature that the amount of squeezing you can get is related to $\chi_{11}+\chi_{22}-2\chi_{12}$.
-Is it worth pointing out that this is wrong?
-Joe:  Absolutely, but this might be true in some kind of weak squeezing limit, so we should check for some reference to this in the literature.}}
-%Can Uaa + Ubb - 2Uab be zero and still get squeezing? Yep, Uaa=0.03, Ubb=0.01, Uab=0.2 gives squeezing for all quantities of interest
+%Can Uaa + Ubb - 2Uab be zero and still get squeezing? Yep, Uaa=0.03, Ubb=0.01, Uab=0.02 gives squeezing for all quantities of interest
 
@@
 As an example, in Figure \ref{figTwoModeAnalyticExamples} we plot the solution for the normalized number variance in mode $|a\rangle$ for a particular set of nonlinearities, choosing a number of different values of $\phi$.
-The plots illustrate the basic features of the problem: During the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=0.5 \Omega$, and has a series of minima.
+The plots illustrate the basic features of the problem: During the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=\pi / \Omega$, and has a series of minima.
 It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
+
+This two-mode analytic model allows arbitrarily good number squeezing, provided there are no limits to the number of particles in the system or to the hold time.
+Unfortunately, in any realistic system, there are more than two modes present, and multimode effects conspire to limit the level of squeezing that can be achieved.
+It is to this that we now turn.
 
@@
     \includegraphics[width=10cm]{figures/analytic_two_mode_examples.eps}
-    \caption{Normalized number variance Var$[N_a]/N_a$ in mode $\hat{a}$ as a function of final recombination time plotted on a logarithmic scale, using four different mixing angles $\phi$.
+    \caption{Normalized number variance Var$[N_a]/N_a$ in mode $|a \rangle$ as a function of final recombination time plotted on a logarithmic scale, using four different mixing angles $\phi$.
 Parameters: $n_a = n_b =5 \times 10^5$, $\chi_{aa}=0.04 \hbar, \chi_{ab}=0, \chi_{bb}=0.01\hbar, \tau_{\mathrm{hold}}=4\times 10^{-4}$s.
@@
 
-\section{Multimode analysis of squeezing} \label{sec:MMdescription}
+\section{Multimode number squeezing analysis} \label{sec:MMdescription}
 
@@
 The two-mode model shows that the parameters required to obtain best squeezing are dependent on the strength of the nonlinearity.
-This nonlinearity is a function of the mode shape and the number of atoms in that mode.
-The optimum hold time $\tau_{\mathrm{hold}}$ depends on this product, as does the recombination phase $\phi$.
-Different modes in a multimode environment will typically have varying products of nonlinearity and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.
-This means that the best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across the range of densities, result in an overall lowest possible total number variance.
+In the multimode case, the effective nonlinearity of a mode is a function of the mode shape, as is shown by Eq.~(\ref{eqChiUequivalence}).
+The optimum hold time $\tau_{\mathrm{hold}}$ depends on the product of the effective mode nonlinearity and number number of particles in that mode, as does the recombination phase $\phi$.
+Different modes in a multimode environment will typically have varying effective nonlinearity-particle number products and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.
+This means that the best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across all the spacial modes present, result in an overall lowest possible total number variance.
 
@@
 Consequently the mode shape of the atomic field changes depending on the nonlinearity-density product, which means that unless $U_{aa}|\psi_a({\mathbf{r}})|^2 = U_{bb}|\psi_b({\mathbf{r}})|^2$ for all ${\mathbf{r}}$, the mode shapes will not overlap perfectly during the final coupling pulse.
-This degraded mode-matching will result in lower efficiency in converting the quadrature squeezing to number squeezing.
-We show an example of this effect in Figure~\ref{figXXX}.
-
-\textbf{give a plot of the mode in $|a\rangle$ and $|b\rangle$ at the beginning and end of the hold time to show mismatch}
+This degraded mode-matching will result in lower efficiency when converting the quadrature squeezing to number squeezing.
 
@@
 This should enable higher squeezing.
-We cannot make strong conclusions without also considering the effects of coupling between spatial modes, however.
-The third potential problem is that coupling between modes is inevitable in the presence of a nonlinearity.
-Coupling between modes will disturb mode-matching, and mix fields with different phase evolution, which can rapidly destroy squeezing.
+We cannot make strong conclusions without also considering the effects of coupling between spatial modes, however, which leads to the third potential problem: Coupling between modes is inevitable in the presence of a nonlinearity, meaning that the modes cannot be analyzed independently.
+Coupling between modes will disturb mode-matching, as well as mixing fields with different phase evolution, which can rapidly destroy squeezing.
 
-An analytic solution to the higher dimensional problem is intractable, so we must use a numerical solution.
+An analytic solution to the multimode, higher-dimensional problem is intractable, so we must use a numerical solution.
 We use stochastic methods based on phase space representations to simulate the dynamics of the multimode quantum fields.
@@
 In our case, the only tractable phase space representation is the functional Wigner representation, which is well behaved when the quantum field is approximately Gaussian, as it is for coherent and squeezed states.
-Stochastic methods based on the functional Wigner representation have been used to model atom laser behaviour \cite{atomLaserWigner}, example2 \cite{example2Wigner}, and example3 \cite{example3Wigner}.
+Stochastic methods based on the functional Wigner representation have been used to model the behaviour or BECs and atom lasers \cite{johnssonET2007,dallET2009,dennisET2010}, as the Wigner representation allows the mapping of the system to a Fokker-Plank equation (unlike the $P$-representation), and does not suffer exponential path-weighting problems that would constrain it to very short simulation intervals (as would the positive-$P$ representation).
 
@@
 This uncontrolled approximation is called the Truncated Wigner Approximation (TWA), and while it has been used widely on ultracold gases \cite{atomlaserWigner,orOtherUCGWigner}, care must be taken that the simulation remains valid.
-\textbf{Here we insert "all the checks I've done to make sure Wigner holds over the parameters we've simulated"}
+For our simulations we checked the results of our simulations against both the analytic two-mode solution, as well as checking for typical indications of TWA breakdown such a the appearance of negative densities in lightly populated modes.
+In all the regimes we investigated the TWA remained valid, apart from the most demanding simulations which took place in 3D with high nonlinearities and long hold times.
+The difficulty with these 3D simulations was that due to the large number of modes they included, the number of particles per mode could drop to ten or less in the densest regions where the squeezing was generated.
+Over very long time scales, this low mode occupation began to result in TWA breakdown.
+As we could characterise the regime in which breakdown began to occur, however, we were able stay away from it and to ensure our simulations remained valid.
 
@@
 
-These stochastic equations take the form:
+These stochastic equations take the form
 \begin{eqnarray}
-\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} \left|\phi_{a}\right|^2 + \frac{U_{a b}}{2} \left|\phi_b\right|^2\right)\phi_a   + \kappa \phi_b  \nonumber \\
-\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} \left|\phi_{b}\right|^2 + \frac{U_{a b}}{2} \left|\phi_a\right|^2\right)\phi_b   + \kappa^* \phi_a
+\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} \left( \left|\phi_{a}\right|^2 -\frac{1}{dV}\right) + U_{a b} \left( \left|\phi_b\right|^2 -\frac{1}{2dV}\right)  \right)\phi_a   + \kappa \phi_b  \nonumber \\
+\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} \left( \left|\phi_{b}\right|^2 - \frac{1}{dV} \right) + U_{a b} \left( \left|\phi_a\right|^2 -\frac{1}{2dV} \right)   \right)\phi_b   + \kappa^* \phi_a
   \label{eqStochasticEquations}
 \end{eqnarray}
-\textbf{I believe we have some rotation terms to include and explain.  Explanation can link to paragraph above}.
+where $dV$ is the volume element of the grid on which the simulation is carried out, and the terms proportional to $1/dV$ correspond to a vacuum corrections.
+It is a peculiarity of the Wigner representation that, up to the vacuum correction terms, Eqs.~(\ref{eqStochasticEquations}) look identical to those of the coupled semi-classical nonlinear Schr{\"{o}}dinger equation.
+The quantum statistical nature of the Wigner equations enters due to the fact that Eqs.~(\ref{eqStochasticEquations}) are run many times and stochastically averaged, with each run using different set of random initial conditions describing (in this case) the noise on a coherent state.
 
@@
 
-We integrate equation (\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS \cite{Dennis2012}.
-The dimensional reduction is achieved by estimating the mode shape in three dimensions, and using dimensionally reduced values for $U_{ij}$ that match the zero-dimensional reduction given by Eq.~(\ref{eqChiUequivalence}).
+We integrate equations (\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS2 \cite{dennis2013}.
+For the one- and two-dimensional simulations, the dimensional reduction is achieved by estimating the mode shape in three dimensions, and using dimensionally reduced values for $U_{ij}$ that match the zero-dimensional reduction given by Eq.~(\ref{eqChiUequivalence}).
 An equivalent process is to match the chemical potentials of the initial states.
@@
 We compare squeezing for fields starting in the ground states of a harmonic trap with negligible nonlinearity (a Gaussian), a harmonic trap with strong nonlinearity (Thomas-Fermi), and a constant potential (constant density).
-Using (\ref{eqChiUequivalence}) we find the following equivalences
+Using Eq.~(\ref{eqChiUequivalence}) we find the following equivalences
 \begin{eqnarray}
@@
 %
-Choosing parameters such that the evolution is sufficiently fast that the mode shape is reasonably stable over the simulation gives the best squeezing, and allows us to make a fair comparison between the squeezing achievable assuming that mode-matching is near optimal.
+where $\omega_x$, $\omega_y$, $\omega_z$ are the angular frequencies of the harmonic trap, and $N$ is the number of particles in the mode of interest.
+
+Choosing parameters such that the squeezing occurs fast enough that the mode shape is reasonably stable over the simulation gives the best squeezing, and allows us to make a fair comparison between the squeezing achievable by different mode shapes without being concerned that different modes will change shapes at different rates, ruining mode-matching.
 The results of this simulation are shown in Figure~\ref{figModalSqueezingEffects1D}, which clearly confirms the hypothesis that maximum squeezing is obtained when the mode shape is closest to constant density.
-In practice, this isThis is largely due to the fact that the squeezing rates become non-uniform, but slower dynamics will also lead to issues with mode-matching.
 
@@
     \caption{Effects of mode shape on squeezing.
-The plot shows the normalized number variance in mode $\hat{a}$ as a function of the time the final coupling pulse is applied.
+The plot shows the normalized number variance in mode $|a \rangle$ as a function of the time the final coupling pulse is applied.
 Thick red line, analytic two-mode solution; solid blue line, constant density mode; dashed line, Thomas-Fermi mode; dash-dotted line, Gaussian mode.
@@
 Parameters: $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \hbar, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$.
-Clearly, the maximum squeezing is achieved when the atomic density profile is flattest.}
+Maximum squeezing is achieved when the atomic density profile is the most uniform.}
     \label{figModalSqueezingEffects1D}
@@
 We can see that the squeezing is higher.
-One obvious disadvantage to post-selecting modes is that it can severely reduce the effective number of atoms.
+One obvious disadvantage to post-selecting modes is that it  reduce the effective number of atoms.
 This is a serious issue for most applications of squeezed sources, such as interferometry, where the signal to noise scales with the square root of the flux.
@@
 Modal mismatch is generated by the different atomic potentials seen by the two internal states.
-We propose a straightforward method of minimising this effect: using a $\pi$-pulse to swap the internal states after half of the evolution.
-In the limit where each state stayed in a single mode (although that mode might change), and where the nonlinear terms did not change with time, this method would produce identical mode shapes for both pulses after the hold time.
-In practice, neither of these conditions will hold, particularly when the squeezing is completely removed by multimode effects, but the $\pi$-pulse should produce some improvement in regimes where the squeezing is only moderately reduced.
+We propose a straightforward method of minimising this effect: using a $\pi$-pulse to swap the populations of the internal states after half of the evolution.
+As can be seen by Eqs~(\ref{eqa1},\ref{eqb1}, such a pulse also leaves the quantum statistics of each state unchanged.
+In the limit where each state stayed in a single mode (although that mode might change), and exactly half of the atoms are in each state, this method would produce identical mode shapes for both modes after the hold time.
+In practice this will never work perfectly as the mode shapes will change, resulting in non-identical environments for the two modes.
+Nonetheless, in many regimes, such a $\pi$-pulse can produce some improvement.
+More importantly, as this scheme accepts that the two modes will change shape, and attempts to make them change in a similar fashion, it allows the possibility of much longer hold times.
+That is, if we are no longer constrained to hold times short enough such that the mode shapes of the atoms in states $|a \rangle$ and $|b \rangle$ do not change significantly, much more squeezing can be extracted from the system.
 
-\begin{itemize}
-  \item show some plots of squeezing with and without the pi pulse
-\end{itemize}
-
-In Fig.~\ref{piPulseFig} we show the effects of a $\pi$-pulse on the squeezing generated \textbf{by some specific system}.
-
+These points are illustrated in Fig.~\ref{piPulseFig}, which shows the best squeezing obtainable from a specific system with and without the $\pi$-pulse, as well as the extra squeezing that can be obtained by allowing a longer hold time as well as the $\pi$-pulse.
 \begin{figure}
@@
 Common parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{ab}=\chi_{bb}=0$.
-Solid red line: Analytic two-mode solution, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (10.9dB squeezing); Solid blue line: no $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (4.9dB squeezing);  Dashed blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.05$ (5.8dB squeezing); Dash-dotted blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=5\times 10^{-4}$s, $\phi=3.87$ (7.2dB squeezing).}
+Solid red line: Analytic two-mode solution, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (10.9dB squeezing); Solid blue line: no $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (4.9dB squeezing);  Dashed blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.05$ (5.8dB squeezing); Dash-dotted blue line: with $\pi$-pulse and the longer hold time $\tau_{\mathrm{hold}}=5\times 10^{-4}$s allowed by the application of the pulse, $\phi=3.87$ (7.2dB squeezing).}
     \label{piPulseFig}
@@
 Furthermore, the expected squeezing goes down as the volume of the trap increases.
-This can be seen in Fig.~\ref{figDimensionalSqueezingEffects}, where we show the squeezing generated \textbf{by 1D, 2D and 3D BECs in a box with some parameters}.
+This can be seen in Fig.~\ref{figDimensionalSqueezingEffects}, where we show the squeezing generated in 1D, 2D and 3D BECs for two different parameter regimes.
+In both cases we take the BEC to be trapped in a box potential, which in the Thomas-Fermi limit leads to a constant density profile across the BEC.
+In Figure~\ref{figDimensionalSqueezingEffects}a we consider the case where $U_{aa}=U_{bb}\neq 0$, $U_{ab}=0$ $U_{aa}\neq 0$, while in Figure~\ref{figDimensionalSqueezingEffects}b we consider the case $U_{aa}\neq 0$, $U_{ab}=U_{bb}=0$.
+In both cases it is clear that number squeezing is best in lower dimensions, and becomes degraded or non-existent as we move to a three dimensional system; the effect is worse if more than one nonlinearity is non-zero.
 
@@
 \begin{thebibliography}{10}
+\bibitem{Hadzibabic2004} Hadzibabic Z, Stock S, Battelier B, Bretin V and Dalibard J 2004 Interference of an array of independent Bose-Einstein Condensates {\it Phys. Rev. Lett.} {\bf 93} 180403
+\bibitem{johnssonET2007} Johnsson M T and Haine S A 2007 Generating squeezing in an atom laser through self-interaction {\it Phys. Rev. Lett.} {\bf 99} 010401
+
+\bibitem{dallET2009} Dall R G, Byron L J, Truscott A G, Dennis G R, Johnsson M T and Hope J J 2009 Paired-atom laser beams created via four-wave mixing {\it Phys. Rev. A} {\bf 79} 011601(R)
+
+\bibitem{dennisET2010} Dennis G R and Johnsson M T 2010 Generation of directional coherent matter beams through dynamical instabilities in Bose-Einstein condensates {\it Phys. Rev. A} {\bf 82} 033615
+
 \bibitem{gross2010} Gross C, Zibold T, Est{\`{e}}ve J and Oberthaler M K 2010 Nonlinear atom interferometer surpasses classical

### 2012-12-13 large_number_squeezing_with_atoms.tex
More work on paper

@@
 \label{secTwoModeAnalytic}
-\begin{itemize}
-  \item Lay out the scheme with diagram, define all variable etc
-  \item Present multimode analytic solution, details go into an appendix
-  \item Plot a few solutions to give an idea of what things look like, plausible times and so on
-  \item Maybe point out it agrees with Oberthaler, or save that for discussion on MM effects?
-  \item Point out that number *difference* squeezing can often be considerably higher than absolute number squeezing in a single state
-\end{itemize}
 Our goal is to describe the number squeezing possibilities of a BEC comprising two relevant internal states denoted by $|a\rangle$ and $|b\rangle$.
@@
 \end{figure}
+
+\begin{itemize}
+  \item Maybe point out it agrees with Oberthaler, or save that for discussion on MM effects?
+  \item Point out that number *difference* squeezing can often be considerably higher than absolute number squeezing in a single state
+\end{itemize}
 
@@
 
-In order to reduce the number of modes available, we can attempt to make our system single mode in one or more dimensions.
-To do this we need to ensure that that any excitation along that dimension is energetically impossible, so we always stay in the ground state.
-This in turn means we need the harmonic oscillator energy per particle to be greater than the nonlinear energy per particle.
-{\bf{*** Not happy with this explanation.
-Make it better ***}}.
+As discussed in the previous section, the existence of a large number of vacuum modes results in the degradation of squeezing due to nonlinear coupling between the modes, and this occurs regards of whether the modes are populated or not.
+The modes that play a role in this process are those with energies below the chemical potential of the system.
+In order to mitigate this effect, it is necessary to have as few of these modes as possible.
 
-The criteria we need to accomplish this depends on the number of dimensions we want to freeze out, so investigate the cases where we reduce a 3D system to an effective 0D, 1D, and 2D system separately.
-In all cases we assume we have a box potential with length $L$ in the dimension(s) that aren't frozen out, and tight harmonic trapping potential of frequency $\omega$ in the frozen dimensions.
+In any confined system, momentum modes will be discretized, with a spacing in a particular dimension given by $\delta k \sim 2\pi/L$.
 
-Reduction from 3D to 0D: Here we have tight trapping along all three dimensions, so we have a harmonic oscillator wavefunction along all dimensions.
+{\bf *** It occurs to me that I'm about to repeat what Graham will say in previous section, so leave it for now ***}
+
+Consequently, the number of low-lying momentum modes that result in degraded squeezing scales with the power of the dimension of the system.
+As shown in Figure~\ref{figDimensionalSqueezingEffects}, obtaining large number squeezing from a three-dimensional, multimode system is likely to be untenable.
+This is unfortunate, as the universe in which we live is unpleasantly three-dimensional.
+
+All is not lost, however, as it is possible to create systems that are effectively lower-dimensional.
+This is accomplished by having trapping potentials that are tight enough in one or more dimensions such that any dynamics in that dimension are effectively frozen out.
+Put another way, if the trapping potential is tight enough, then the energy required to excite any mode beyond that of the ground state is unavailable.
+Conditions required to reach this limit vary on the form of the potential, but, provided we consider a harmonic potential, a simple criterion for ensuring single mode behaviour will be
+\begin{equation}
+\hbar \omega \gg \mu
+\label{eqTightTrappingCriterionMu}
+\end{equation}
+that is, the spacing between energy levels is much greater that that of the chemical potential.
+
+Given that we have shown that constant density profiles result are required for good squeezing, we consider a 3D potential that has tight harmonic trapping in one, two or three dimensions, and a box-like potential in the other dimensions.
+Our goal is to make this system effectively two-, one- or zero-dimensional respectively.
+
+We assume a trapping frequency $\omega$ in the harmonically trapped dimensions and a confining region of length $L$ in the box potential dimensions.
+We further assume that we are working in the Thomas-Fermi limit, as this level of nonlinearity is required to obtain good squeezing.
+This means that the criterion (\ref{eqTightTrappingCriterionMu}) becomes
+\begin{equation}
+\hbar \omega \gg \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz
+\label{eqTightTrappingCriterionU}
+\end{equation}
+where the factor of $N$ is present because we require the energy per particle, not the total energy.
+
+Due to the Thomas-Fermi limit, the density profile $|\psi({\mathbf{r}})|^2$ will be constant in the box-potential directions, and due to the assumption of tight trapping, close to the harmonic oscillator ground state in the harmonically trapped directions.
+Evaluating Eq.~(\ref{eqTightTrappingCriterionU}) with these density distributions and using $U=4\pi \hbar^2 a /m$ we obtain
+\begin{eqnarray}
+\sqrt{\frac{\hbar}{m \omega}}  \gg \frac{aN}{\sqrt{2 \pi}} && \,\,\,\, ({\textrm{0D}}) \label{eq0Dreduction} \\
+L  \gg  a N / \pi && \,\,\,\, ({\textrm{1D}}) \label{eq1Dreduction} \\
+\omega  \gg  \frac{8 \pi a^2 N^2 \hbar}{m L^4}  && \,\,\,\, ({\textrm{2D}}) \label{eq2Dreduction}
+\end{eqnarray}
+as conditions we require to reduce a 3D system to an effective 0D, 1D or 2D system.
+
+To determine how feasible this dimensional reduction is, we assume Rubidium atoms, typical scattering lengths ($a\sim$10nm), large particle numbers ($N\sim 10^6$) and a confining length scale of $L\sim 200\mu$m in the box dimensions.
+
+In order to reduce the system to 0D, Eq.~(\ref{eq0Dreduction}) indicates we require a harmonic oscillator length of $\sqrt{\hbar/ m \omega}\gg 4$mm; a BEC several centimeters in diameter is not particularly feasible.
+Using Eq.~(\ref{eq1Dreduction}) demonstrates reduction to 1D requires a BEC that is $\gg3.2$mm in length; this would also be challenging.
+
+Reduction to 2D appears more promising, however. For the parameters given above, we see that we need $\omega \gg 2\pi\times 190$Hz, which is certainly feasible.
+
+We note that the reason our dimensional reduction criteria to obtain effective 0D and 1D systems are so stringent, is entirely due to our requirement of high particle number.
+For low particle number, the criteria become much more forgiving.
+An example is the experiment of Gross {\it et al.} which measured a number difference squeezing of 6.9dB in a spherical trap with a trapping frequency of $\omega=2 \pi \times 425$Hz.
+Due to the low number of atoms in their trap ($\sim 400$), they were able to achieve an effectively 0D trap with single mode behaviour.
+
+%Reduction from 3D to 0D: Here we have tight trapping along all three dimensions, so we have a harmonic oscillator wavefunction along all dimensions.
 With $N$ particles in the system, the density is given by
-\begin{equation}
-|\psi|^2 = N \left( \frac{m \omega}{\pi \hbar} \right)^{3/2} \exp \left[ -\frac{m \omega} {\hbar} (x^2 +y^2+z^2) \right] .
-\end{equation}
-In this system, the nonlinear energy per particle is
-\begin{eqnarray}
-E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
-       &=& \frac{U N}{4 \sqrt{2}} \left( \frac{m \omega}{\pi \hbar} \right)^{3/2}
-\end{eqnarray}
-where $U = \pi \hbar^2 a/m$ and $a$ is the s-wave scattering length of the particles. The criterion $\hbar \omega > E_{NL}$ thus becomes
-\begin{equation}
-%N < \frac{1}{a} \sqrt{2 \pi \hbar}{m \omega}.
-\sqrt{\frac{\hbar}{m \omega}} > \frac{aN}{\sqrt{2 \pi}}.
-\end{equation}
-Assuming Rubidium atoms, typical scattering lengths ($a\sim$10nm) and large particle numbers (say $N\sim 10^6$), we require a harmonic oscillator length $\sqrt{\hbar/ m \omega}>4$mm.
+%\begin{equation}
+%|\psi|^2 = N \left( \frac{m \omega}{\pi \hbar} \right)^{3/2} \exp \left[ -\frac{m \omega} {\hbar} (x^2 +y^2+z^2) \right] .
+%\end{equation}
+%In this system, the nonlinear energy per particle is
+%\begin{eqnarray}
+%E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
+%       &=& \frac{U N}{4 \sqrt{2}} \left( \frac{m \omega}{\pi \hbar} \right)^{3/2}
+%\end{eqnarray}
+%where $U = \pi \hbar^2 a/m$ and $a$ is the s-wave scattering length of the particles. The criterion $\hbar \omega > E_{NL}$ thus becomes
+%\begin{equation}
+%%N < \frac{1}{a} \sqrt{2 \pi \hbar}{m \omega}.
+%\sqrt{\frac{\hbar}{m \omega}} > \frac{aN}{\sqrt{2 \pi}}.
+%\end{equation}
+%Assuming Rubidium atoms, typical scattering lengths ($a\sim$10nm) and large particle numbers (say $N\sim 10^6$), we require a harmonic oscillator %length $\sqrt{\hbar/ m \omega}>4$mm.
 A BEC that's more than a centimetre across isn't particularly feasible.
+%
+%Reduction from 3D to 1D: Here we have tight trapping along $x$ and $y$, with a box potential along $z$. With $N$ particles in the system, the density is given by
+%\begin{equation}
+%|\psi|^2 = \frac{N}{L} \frac{m \omega}{\pi \hbar} \exp \left[ -\frac{m \omega} {\hbar}  (x^2 +y^2) \right] .
+%\end{equation}
+%In this system, the nonlinear energy per particle is
+%\begin{eqnarray}
+%E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
+%       &=& \frac{U N m \omega}{4 \pi \hbar L}
+%\end{eqnarray}
+%The criterion $\hbar \omega > E_{NL}$ thus becomes
+%\begin{equation}
+%L > a N / \pi.
+%\end{equation}
+%Assuming Rubidium atoms, typical scattering lengths ($a\sim$10nm) and large particle numbers (say $N\sim 10^6$) we need a BEC that's a few millimeters long, which isn't particularly feasible.
+%
+%Reduction from 3D to 2D: Here we have tight trapping along $z$, with a box potential along $x$ and $y$. With $N$ particles in the system, the density is given by
+%\begin{equation}
+%|\psi|^2 = \frac{N}{L^2} \left( {\frac{m \omega}{\pi \hbar}} \right) ^{1/2} \exp \left[ -\frac{m \omega} {\hbar} z^2  \right]
+%\end{equation}
+%In this system, the nonlinear energy per particle is
+%\begin{eqnarray}
+%E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
+%       &=& \frac{m \omega U N}{4 \pi \hbar L^2}
+%\end{eqnarray}
+%The criterion $\hbar \omega > E_{NL}$ thus becomes
+%\begin{equation}
+%\omega > \frac{8 \pi a^2 N^2 \hbar}{m L^4}.
+%\end{equation}
+%Once again taking a Rubidium atom and considering typical parameters, say $a\sim$10nm, $N\sim 10^6$, $L\sim 200\mu$m, we see that we need $\omega > 2\pi\times 190$Hz, which is feasible.
 
-Reduction from 3D to 1D: Here we have tight trapping along $x$ and $y$, with a box potential along $z$. With $N$ particles in the system, the density is given by
-\begin{equation}
-|\psi|^2 = \frac{N}{L} \frac{m \omega}{\pi \hbar} \exp \left[ -\frac{m \omega} {\hbar}  (x^2 +y^2) \right] .
-\end{equation}
-In this system, the nonlinear energy per particle is
-\begin{eqnarray}
-E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
-       &=& \frac{U N m \omega}{4 \pi \hbar L}
-\end{eqnarray}
-The criterion $\hbar \omega > E_{NL}$ thus becomes
-\begin{equation}
-L > a N / \pi.
-\end{equation}
-Assuming Rubidium atoms, typical scattering lengths ($a\sim$10nm) and large particle numbers (say $N\sim 10^6$) we need a BEC that's a few millimeters long, which isn't particularly feasible.
+We finish this section by considering an example of the sort of number squeezing that is achiveable in a realistic physical system with large particle number.
+We take a Rb$^{85}$ condensate, which possesses a number of Feschbach resonances that allow us to modify one of the $U_{ij}$ scattering lengths without affecting the others.
+We assume a box potential in two dimensions with a confinement region $L=120\mu$m$\times 120\mu$m, and a tightly trapped third dimension with a trapping frequency $\omega=2 \pi \times 1500$Hz, resulting in a near 2D system.
+We take scattering lengths of $U_{aa}=-400 a_0$, $U_{ab}=U_{bb}=0$, a total atom number of $4\times 10^5$, and a hold time of $\tau=1$ms.
+The results of a full 3D numerical simulation of this system are shown in Figure \ref{figRealistic3DParameters}, where we have ensured all of the potentially troubling Bogoliubov modes are included.
+This system achieves relative number squeezing of 13.9dB, which compares favourably to the 15.1dB generated in the ideal 0D, two-mode case.
+The main reason for not approaching the ideal case more closely was that even with a $1500$Hz harmonic trapping potential the system does remain entirely single mode in that dimension; a small admixture of higher-order modes are still excited.
 
-Reduction from 3D to 2D: Here we have tight trapping along $z$, with a box potential along $x$ and $y$. With $N$ particles in the system, the density is given by
-\begin{equation}
-|\psi|^2 = \frac{N}{L^2} \left( {\frac{m \omega}{\pi \hbar}} \right) ^{1/2} \exp \left[ -\frac{m \omega} {\hbar} z^2  \right]
-\end{equation}
-In this system, the nonlinear energy per particle is
-\begin{eqnarray}
-E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
-       &=& \frac{m \omega U N}{4 \pi \hbar L^2}
-\end{eqnarray}
-The criterion $\hbar \omega > E_{NL}$ thus becomes
-\begin{equation}
-\omega > \frac{8 \pi a^2 N^2 \hbar}{m L^4}.
-\end{equation}
-Once again taking a Rubidium atom and considering typical parameters, say $a\sim$10nm, $N\sim 10^6$, $L\sim 200\mu$m, we see that we need $\omega > 2\pi\times 190$Hz, which is feasible.
-
-The takeaway point is that dimensional reduction to 0D and 1D is possible in the case where we have low particle numbers, which is how Oberthaler got his 6dB (or 8dB depending how you count) of squeezing.
-He only had 400 atoms, so could reduce his problem to 0D and freeze out any multimode dynamics.
-But if we're interested in large number squeezing, the best we can really do is reduction from 3D to 2D.
-
-Taking all the lessons learned, we can get good squeezing for realistic paramters in 3D.
-Figure \ref{figXXX} shows an example of a full 3D simulation, 10000 paths, box potential in 2d, tight trapping (1500Hz) in the 3rd dimension.
+We also note that configurations with higher squeezing are certainly possible, but on these time scales a full 3D simulation begins to suffer breakdown of the TWA, making numerical results unreliable.
 \begin{figure}
@@
 
-\bibitem{gross2010} Gross C, Zibold T, Est{\`{e}}ve J and Oberthaler M K 2010 Nonlinear atom interferometer surpasses classical
-precision limit {\it Nature} {\bf 464} 1165--1169
 \bibitem{dennis2013} Dennis G R, Hope J J and Johnsson M T 2013 XMDS2: Fast, scalable simulation of coupled stochastic partial
@@
 \bibitem{louisell} Louisell Quantum Statistical Properties of Radiation
+
+\bibitem{gross2010} Gross C, Zibold T, Est{\`{e}}ve J and Oberthaler M K 2010 Nonlinear atom interferometer surpasses classical
+precision limit {\it Nature} {\bf 464} 1165--1169
+
 \end{thebibliography}

### 2012-12-17 large_number_squeezing_with_atoms.tex
Starting work on the Bog section

@@
 \usepackage{graphicx}
+\usepackage{amsmath}
 \usepackage{ulem}
@@
 \usepackage{subcaption}
+
+\DeclareMathOperator{\Var}{Var}
+\DeclareMathOperator{\sinc}{sinc}
+\newcommand{\abs}[1]{\ensuremath{\left|#1\right|}}
+\newcommand{\expect}[1]{\ensuremath{\left<#1\right>}}
+\newcommand{\ket}[1]{\ensuremath{\left|#1\right>}}
 
@@
 In both cases we take the BEC to be trapped in a box potential, which in the Thomas-Fermi limit leads to a constant density profile across the BEC.
-In Figure~\ref{figDimensionalSqueezingEffects}a we consider the case where $U_{aa}=U_{bb}\neq 0$, $U_{ab}=0$ $U_{aa}\neq 0$, while in Figure~\ref{figDimensionalSqueezingEffects}b we consider the case $U_{aa}\neq 0$, $U_{ab}=U_{bb}=0$.
+In Figure~\ref{figDimensionalSqueezingEffects}a we consider the case where $U_{aa}=U_{bb}\neq 0$, $U_{ab}=0$, while in Figure~\ref{figDimensionalSqueezingEffects}b we consider the case $U_{aa}\neq 0$, $U_{ab}=U_{bb}=0$.
 In both cases it is clear that number squeezing is best in lower dimensions, and becomes degraded or non-existent as we move to a three dimensional system; the effect is worse if more than one nonlinearity is non-zero.
@@
 
-Excitations of a BEC are described by the Bogoliubov modes...
+We can understand the origin and scaling of this degradation of squeezing by considering the response of the condensate to small fluctuations about the mean field.
+We do this using Bogoliubov theory \cite{PethickSmith}, which considers the first order quantum-mechanical fluctuations about the mean field.
+To simplify the analysis we consider the case in which $U_{aa} \neq 0$, $U_{bb}=U_{ab} = 0$ and the condensate has a spatially-constant density in a box of side lengths $L_x$, $L_y$ and $L_z$.
+In this limit the squeezing of the $a$ mode at $t_3$ is
+\begin{align}
+  \left.\frac{\Var[N_a](t_3)}{N_a(t_3)}\right|_\text{Bog} &= \frac{\Var[N_a](t_3) + \sum_\mathbf{k} \frac{1}{2} n_a(\mathbf{k}, t_3) \left\{3 + \cos\left[\Omega(t_3-t_2)\right] + 4 n_a(\mathbf{k}, t_3)\right\}}{N_a(t_3) + \sum_\mathbf{k} n_a(\mathbf{k}, t_3)},\\
+  n_a(\mathbf{k}, t_3) &= \left\{n_a \lambda_{aa} \cos\left[\Omega\left(t_3-t_2\right)\right] \sinc\left(\omega_\mathbf{k} \tau_\text{hold}\right) \right\}^2,\\
+  \omega_\mathbf{k} &= \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \lambda_{aa} n_a)}, \\
+  \omega^0_\mathbf{k} &= \frac{\hbar \mathbf{k}^2}{2 M},
+\end{align}
+where $n_a(\mathbf{k}, t) = \expect{\delta \hat{a}^\dagger(t) \delta\hat{a}(t)}$ is the occupation of momentum mode $\mathbf{k}$, $\Var[N_a](t)$ and $N_a(t)$ are the number variance and mean number of the $\ket{a}$ mode from the single-mode derivation, the sum over $\mathbf{k}$ is a discrete sum over the available momentum modes $\mathbf{k} = n_x/2\pi L_x \hat{\mathbf{x}} + n_y/2\pi L_y\hat{\mathbf{y}} + n_z/2\pi L_z\hat{\mathbf{z}}$, $\omega_\mathbf{k}$ is the frequency of the Bogoliubov mode $\mathbf{k}$ and $\omega^0_\mathbf{k}$ is the frequency of a free particle.
 
@@
 
+\section{Derivation of Bogoliubov squeezing formula}
+\label{appendixBogDerivation}
+
+Here we put some details of the Bog analysis.
+
+\begin{align}
+  S &= \frac{\Var(n_0) + \sum_k \frac{1}{2} n_k (3 + \cos 2\theta + 4 n_k)}{n_0 + \sum_k n_k} \\
+  n_k &= \left[\lambda \cos\theta \sinc(\omega_k \tau)\right]^2 \\
+  \hat{a}(t_3) &= \left[\hat{a}(0) - i \chi \alpha \tau_h\left(\alpha^* \hat{a}(0) + \alpha \hat{a}^\dagger(0) - 2\abs{\alpha}^2\right) \right]\cos(\theta) - i\hat{b}(0) \sin\theta e^{i\phi} \\
+  n_0 &= \left(\alpha ^2+\alpha ^4 \lambda ^2\right) \cos^2\theta + \left(\beta^2 \sin\theta  + 2\alpha\beta\cos\theta\sin\phi\right)\sin\theta\\
+  \begin{split}
+  \Var(n_0) &= \frac{1}{2} \left(\alpha ^2+\beta ^2\right)+\frac{1}{8}\alpha ^4 \left(7+ 2\beta ^2\right) \lambda ^2+ \frac{3}{4} \alpha ^8 \lambda ^4 \\
+  &\phantom{\mathrel{=}}+\frac{1}{2}\left( \alpha ^2-\beta ^2+2 \alpha ^4 \lambda ^2+2 \alpha ^8 \lambda ^4\right) \text{Cos}[2 \theta ] \\
+  &\phantom{\mathrel{=}}+\frac{1}{8}\left(\alpha ^4 \lambda ^2 \left(1-2 \beta ^2+2 \alpha ^4 \lambda ^2\right) \text{Cos}[4 \theta ] \right) \\
+  &\phantom{\mathrel{=}}+\frac{1}{8}\left(4 \alpha  \beta  \left(8 \alpha ^2 \lambda  \text{Cos}[\theta ]^3 \text{Cos}[\phi ] \text{Sin}[\theta ]+\text{Sin}[2 \theta ] \left(2 \text{Sin}[\phi ]+\alpha  \beta  \lambda  \text{Sin}[2 \theta ] \left(\alpha ^2 \lambda  \text{Cos}[2 \phi ]+\text{Sin}[2 \phi ]\right)\right)\right)\right)
+  \end{split}
+\end{align}
+
 \section*{References}
@@
 
+\bibitem{PethickSmith} C.J.~Pethick and H.~Smith, \emph{Bose-Einstein Condensation in Dilute Gases} (Cambridge University Press, Cambridge, 2002).
+
 \end{thebibliography}

### 2012-12-17 large_number_squeezing_with_atoms.tex
Extending the Bog analysis
I've included a description of the first figure that I want to include.

@@
 To simplify the analysis we consider the case in which $U_{aa} \neq 0$, $U_{bb}=U_{ab} = 0$ and the condensate has a spatially-constant density in a box of side lengths $L_x$, $L_y$ and $L_z$.
-In this limit the squeezing of the $a$ mode at $t_3$ is
+In this limit the squeezing $S_a = \Var(N_a)/N_a$ of the $a$ mode at the end of the pulse sequence is (see \ref{appendixBogDerivation})
 \begin{align}
-  \left.\frac{\Var[N_a](t_3)}{N_a(t_3)}\right|_\text{Bog} &= \frac{\Var[N_a](t_3) + \sum_\mathbf{k} \frac{1}{2} n_a(\mathbf{k}, t_3) \left\{3 + \cos\left[\Omega(t_3-t_2)\right] + 4 n_a(\mathbf{k}, t_3)\right\}}{N_a(t_3) + \sum_\mathbf{k} n_a(\mathbf{k}, t_3)},\\
-  n_a(\mathbf{k}, t_3) &= \left\{n_a \lambda_{aa} \cos\left[\Omega\left(t_3-t_2\right)\right] \sinc\left(\omega_\mathbf{k} \tau_\text{hold}\right) \right\}^2,\\
-  \omega_\mathbf{k} &= \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \lambda_{aa} n_a)}, \\
+  S_a(t_3) &= \frac{\Var[N_a](t_3) + \sum_{\mathbf{k}\neq 0} \frac{1}{2} n_a(\mathbf{k}, t_3) \left\{3 + \cos\left[\Omega(t_3-t_2)\right] + 4 n_a(\mathbf{k}, t_3)\right\}}{N_a(t_3) + \sum_{\mathbf{k}\neq 0} n_a(\mathbf{k}, t_3)}, \label{eqBogoliubovSqueezing}\\
+  n_a(\mathbf{k}, t_3) &= \left\{n_a \chi_{aa} \tau_\text{hold} \cos\left[\Omega\left(t_3-t_2\right)\right] \sinc\left(\omega_\mathbf{k} \tau_\text{hold}\right) \right\}^2,\\
+  \omega_\mathbf{k} &= \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}, \\
   \omega^0_\mathbf{k} &= \frac{\hbar \mathbf{k}^2}{2 M},
 \end{align}
-where $n_a(\mathbf{k}, t) = \expect{\delta \hat{a}^\dagger(t) \delta\hat{a}(t)}$ is the occupation of momentum mode $\mathbf{k}$, $\Var[N_a](t)$ and $N_a(t)$ are the number variance and mean number of the $\ket{a}$ mode from the single-mode derivation, the sum over $\mathbf{k}$ is a discrete sum over the available momentum modes $\mathbf{k} = n_x/2\pi L_x \hat{\mathbf{x}} + n_y/2\pi L_y\hat{\mathbf{y}} + n_z/2\pi L_z\hat{\mathbf{z}}$, $\omega_\mathbf{k}$ is the frequency of the Bogoliubov mode $\mathbf{k}$ and $\omega^0_\mathbf{k}$ is the frequency of a free particle.
+where $n_a(\mathbf{k}, t) = \expect{\hat{a}^\dagger(\mathbf{k}, t) \hat{a}(\mathbf{k}, t)}$ is the occupation of momentum mode $\mathbf{k}$, $\Var[N_a](t)$ and $N_a(t)$ are the number variance and mean number of the $a$ mode from the two-mode model in Section~\ref{secTwoModeAnalytic}, the sum over $\mathbf{k}$ is a discrete sum over the available non-zero momentum modes $\mathbf{k} = (2\pi/L_x) n_x \hat{\mathbf{x}} + (2\pi/L_y) n_y \hat{\mathbf{y}} + (2\pi/L_z) n_z \hat{\mathbf{z}}$ with $n_x$, $n_y$ and $n_z$ arbitrary integers, $\omega_\mathbf{k}$ is the frequency of the Bogoliubov mode $\mathbf{k}$ and $\omega^0_\mathbf{k}$ is the frequency of a free particle.
+
+The squeezing of mode $a$ in \eqref{eqBogoliubovSqueezing} includes the first-order multimode quantum mechanical fluctuations about the mean field and differs from the single-mode squeezing of $\Var[N_a](t_3)/N_a(t_3)$ only by terms involving the occupation of the non-zero momentum modes $n_a(\mathbf{k}, t)$.
+The contribution of these terms is always detrimental to squeezing as the contribution to the numerator of \eqref{eqBogoliubovSqueezing} is always at least as large as the contribution to the denominator.
+When the occupation of the non-zero momentum modes is a non-negligible fraction of the condensate, the observed squeezing will differ from that predicted by the two-mode model in Section~\ref{secTwoModeAnalytic} (see Fig.~\ref{figBogSqueezingValidation}).
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{.45\textwidth}
+    \centering
+    Plot of the number occupation of the condensate and non-zero momentum modes as computed by the Bog theory and a simulation.
+The simulation agrees with the Bog theory for the occupation of the non-zero momentum modes, but there is significant ~10-20\% depletion of the condensate.
+The horizontal axis is the hold time.
+  \end{subfigure}
+  \begin{subfigure}{.45\textwidth}
+    \centering
+Squeezing curve showing the two-mode model, Bogoliubov squeezing \eqref{eqBogoliubovSqueezing} and simulation.
+All three disagree, but both the Bog squeezing and simulation are worse than the two-mode model.
+The same parameters are used as in the first part of this figure
+  \end{subfigure}%
+\caption{This is not a caption.}
+  \label{figBogSqueezingValidation}
+\end{figure}
+
+The Bogoliubov theory used in deriving \eqref{eqBogoliubovSqueezing} is not number-conserving because the depletion of the condensate caused by the occupation of the non-zero momentum modes is neglected.
+Therefore when these non-zero momentum modes become sufficiently occupied for $S_a$ to differ from the two-mode model, the Bogoliubov approximation becomes invalid and $S_a$ will not agree with the full quantum-statistical multimode result.
+Although the Bogoliubov approximation cannot predict the amount by which squeezing will be degraded compared to the two-mode model, it can predict the regime in which squeezing is reduced and suggest how this problem can be avoided.
+Figure~\ref{figBogSqueezingValidation}(a) illustrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the total number of atoms in the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ for parameters where these modes reach ~20\% of the total number of atoms in the system.
+Figure~\ref{figBogSqueezingValidation}(b) shows that in this case the squeezing predicted by the Bogoliubov theory in \eqref{eqBogoliubovSqueezing} does not agree with either the two-mode model or the truncated Wigner simulations, but both the Bogoliubov theory and the truncated Wigner simulations show worse squeezing than predicted by the two-mode model.
 

### 2012-12-17 large_number_squeezing_with_atoms.tex
Minor work on Bog section

@@
 \section{Squeezing in 2D and 3D}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_1.eps}
+    \label{figDimensionalSqueezingEffects:sub1}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_2.eps}
+    \label{figDimensionalSqueezingEffects:sub2}
+  \end{subfigure}
+\caption{Effects of dimension on squeezing in two different nonlinear regimes.
+If two nonlinearities are non-zero, the deleterious effects of higher dimensions on squeezing are exacerbated.
+Thick red line, analytic two-mode solution; solid blue line, 1D; dashed line, 2D; dash-dotted line, 3D.
+(a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=6.2$.
+(b) $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{bb}=\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.}
+  \label{figDimensionalSqueezingEffects}
+\end{figure}
+
 The simulations in the previous two sections were one dimensional, and show how having spatially constant density and using a $\pi$-pulse will improve the resultant squeezing.
@@
 
+To minimise the damaging effects of the non-zero momentum modes, it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small.
+
   \begin{itemize}
@@
   \end{itemize}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_1.eps}
-    \label{figDimensionalSqueezingEffects:sub1}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_2.eps}
-    \label{figDimensionalSqueezingEffects:sub2}
-  \end{subfigure}
-\caption{Effects of dimension on squeezing in two different nonlinear regimes.
-If two nonlinearities are non-zero, the deleterious effects of higher dimensions on squeezing are exacerbated.
-Thick red line, analytic two-mode solution; solid blue line, 1D; dashed line, 2D; dash-dotted line, 3D.
-(a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=6.2$.
-(b) $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{bb}=\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.}
-  \label{figDimensionalSqueezingEffects}
-\end{figure}
 

### 2013-02-25 large_number_squeezing_with_atoms.tex
Minor updates to Bog section
Fix broken IOP/amsmath compatibility.
Add figures showing experimentalist's scaling in 1D, 2D and 3D

@@
 \documentclass{iopart}
+  \expandafter\let\csname equation*\endcsname\relax
+  \expandafter\let\csname endequation*\endcsname\relax
 %\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
@@
 To simplify the analysis we consider the case in which $U_{aa} \neq 0$, $U_{bb}=U_{ab} = 0$ and the condensate has a spatially-constant density in a box of side lengths $L_x$, $L_y$ and $L_z$.
-In this limit the squeezing $S_a = \Var(N_a)/N_a$ of the $a$ mode at the end of the pulse sequence is (see \ref{appendixBogDerivation})
+In this limit the squeezing $S_a = \Var(N_a)/N_a$ of the $a$ mode at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation})
 \begin{align}
@@
   \end{itemize}
+
+Some notes on the Bog scaling plots below: Experimentalist's scaling works as follows: V $\rightarrow$ r V; hold time $\rightarrow$ r*(hold time).
+In order to get enough lattice points to cover the momentum modes up to energy mu, the scaling goes as: 1D lattice$\rightarrow$lattice*$\sqrt{r}$ (1D); lattice$\rightarrow$lattice (2D); lattice$\rightarrow$lattice*$r^{-1/6}$.
+
+The plots show that the experimentalist's scaling increases Bog occupation (i.e.
+condensate depletion) with volume in 1D, is constant in 2D, and decreases it in 3D.
+This means squeezing gets worse with volume in 1D, is constant with volume in 2D, and gets better with volume in 3D.
+We can't actually show squeezing getting better in 3D due to TWA breakdown, but we do show bog mode occupation decreasing.
+
+The jaggedness of the line in 3D is due to the fact that we're only dealing with 10-20 lattice points in each dimension, and the lattice cutoff is often part way between two integer numbers, so we have to round up or down as we change the scaling factor.
+
+Jaggedness of squeezing line in 2D is due to stochastic nature of simulations.
+
+\begin{figure}
+  \centering
+   \includegraphics[width=7cm]{figures/1D_squeezing_scaling.eps}
+\caption{1D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
+ \label{fig1D_squeezing_scaling}
+\end{figure}
+
+\begin{figure}
+  \centering
+   \includegraphics[width=7cm]{figures/1D_Bog_mode_occupation_scaling.eps}
+\caption{1D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
+ \label{fig1D_Bog_mode_occupation_scaling}
+\end{figure}
+
+\begin{figure}
+  \centering
+   \includegraphics[width=7cm]{figures/2D_squeezing_scaling.eps}
+\caption{2D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
+ \label{fig2D_squeezing_scaling}
+\end{figure}
+
+\begin{figure}
+  \centering
+   \includegraphics[width=7cm]{figures/2D_Bog_mode_occupation_scaling.eps}
+\caption{2D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
+ \label{fig2D_Bog_mode_occupation_scaling}
+\end{figure}
+
+\begin{figure}
+  \centering
+   \includegraphics[width=7cm]{figures/3D_Bog_mode_occupation_scaling.eps}
+\caption{3D experimentalist's scaling. $N=1\times 10^4$, $L=20\mu$m.}
+ \label{fig3D_Bog_mode_occupation_scaling}
+\end{figure}
 
@@
 %
-&& + c^2 s^2 [ \hat{a}^{\dagger}_2 \hat{a}_2 + \hat{b}^{\dagger}_2 \hat{b}_2 + 4 \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 -e^{-2 i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 -e^{-2 i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 ] \nonumber \\
+&& + c^2 s^2 [ \hat{a}^{\dagger}_2 \hat{a}_2 + \hat{b}^{\dagger}_2 \hat{b}_2 + 4 \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 -e^{2 i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 -e^{-2 i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 ] \nonumber \\
 %
@@
 %
-&& + c^2 s^2 [ \hat{b}^{\dagger}_2 \hat{b}_2 + \hat{a}^{\dagger}_2 \hat{a}_2 + 4 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{a}^{\dagger}_2 \hat{a}_2 - e^{2 i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 -e^{2 i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 ] \nonumber \\
+&& + c^2 s^2 [ \hat{b}^{\dagger}_2 \hat{b}_2 + \hat{a}^{\dagger}_2 \hat{a}_2 + 4 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{a}^{\dagger}_2 \hat{a}_2 - e^{-2 i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 -e^{2 i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 ] \nonumber \\
 %

### 2013-05-12 large_number_squeezing_with_atoms.tex
Major rewrite of text for new conclusions
Appendix unexamined, and I may have gone off the rails at several
points.  I put bold statements near the most concerning, but there may
be others.  I think most of Mattias' Bog/TWA figures should be
condensed into two, so the text assumes this.

Will polish after basic truthiness is checked.

@@
 
-At time $t=0$ the coupling $\Omega$ is switched on and applied until time $t=t_1$, at which point it is turned off, resulting in a portion of the population being transferred into mode $|b\rangle$.
+At time $t=0$, the coupling $\Omega$ is switched on and applied until time $t=t_1$, at which point it is turned off, resulting in a portion of the population being transferred into mode $|b\rangle$.
 As the initial state was a coherent state in both modes, after the population transfer both $|a\rangle$ and $|b\rangle$ are described by coherent states, with mean number $\langle \hat{a}^{\dagger} \hat{a} \rangle = n_a$ and $\langle \hat{b}^{\dagger} \hat{b} \rangle = n_b$ respectively.
@@
 After this hold time $\tau_{\mathrm{hold}} = t_2 - t_1 $, the coupling $\Omega$ is switched back on until time $t_3$, with a phase shift $\phi$ compared to its first application.
-During this last stage, the two modes exchange population and the quadrature variance is converted into number variance, in the same way homodyne measurements are used in quantum optics to convert quadrature squeezing into number squeezing, which can be directly measured.
+During this last stage, the two modes exchange population and the quadrature variance is converted into number variance, in the same way that homodyne measurements are used in quantum optics to convert quadrature squeezing into number squeezing, which can be directly measured.
 In this interpretation, $\phi$ is the relative phase of the strong local oscillator, which allows specific phase angles of quadrature squeezing to be examined.
@@
 \end{eqnarray}
-then the expectation values for number, number variance, and number difference variance after appling the final coupling field are given by
+then the expectation values for number, number variance, and number difference variance after applying the final coupling field are given by
 \begin{eqnarray}
@@
 
-While the analytic expressions for the number variances are not particularly edifying, it is easy to evaluate and plot them for any given set of parameters.
-As an example, in Figure \ref{figTwoModeAnalyticExamples} we plot the solution for the normalized number variance in mode $|a\rangle$ for a particular set of nonlinearities, choosing a number of different values of $\phi$.
-The plots illustrate the basic features of the problem: During the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=\pi / \Omega$, and has a series of minima.
+In Figure \ref{figTwoModeAnalyticExamples} we plot a solution for the normalized number variance in mode $|a\rangle$ as a function of $\phi$, choosing a typical set of nonlinearities.
+This example illustrates the basic features of the system: During the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=\pi / \Omega$, and has a series of minima.
 It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
@@
 This two-mode analytic model allows arbitrarily good number squeezing, provided there are no limits to the number of particles in the system or to the hold time.
-Unfortunately, in any realistic system, there are more than two modes present, and multimode effects conspire to limit the level of squeezing that can be achieved.
-It is to this that we now turn.
+The two-mode approximation can be valid in tight traps with very low particle number, and this result correctly predicts the squeezing demonstrated by Gross \textit{et al.} \cite{Gross2010}.
+Unfortunately, for larger numbers of atoms in realistic traps there are more than two modes present, and multimode effects limit the level of squeezing that can be achieved.
 
@@
 \end{figure}
-
-\begin{itemize}
-  \item Maybe point out it agrees with Oberthaler, or save that for discussion on MM effects?
-  \item Point out that number *difference* squeezing can often be considerably higher than absolute number squeezing in a single state
-\end{itemize}
 
@@
 In our case, the only tractable phase space representation is the functional Wigner representation, which is well behaved when the quantum field is approximately Gaussian, as it is for coherent and squeezed states.
-Stochastic methods based on the functional Wigner representation have been used to model the behaviour or BECs and atom lasers \cite{johnssonET2007,dallET2009,dennisET2010}, as the Wigner representation allows the mapping of the system to a Fokker-Plank equation (unlike the $P$-representation), and does not suffer exponential path-weighting problems that would constrain it to very short simulation intervals (as would the positive-$P$ representation).
+Stochastic methods based on the functional Wigner representation have been used to model the behaviour of BECs and atom lasers \cite{johnssonET2007,dallET2009,dennisET2010}, as the Wigner representation allows the mapping of the system to a Fokker-Planck equation with positive semi-definite diffusion matrix (unlike the $P$-representation), and does not suffer exponential path-weighting problems that would constrain it to very short simulation intervals (as would the positive-$P$ representation).
 
@@
 This uncontrolled approximation is called the Truncated Wigner Approximation (TWA), and while it has been used widely on ultracold gases \cite{atomlaserWigner,orOtherUCGWigner}, care must be taken that the simulation remains valid.
-For our simulations we checked the results of our simulations against both the analytic two-mode solution, as well as checking for typical indications of TWA breakdown such a the appearance of negative densities in lightly populated modes.
-In all the regimes we investigated the TWA remained valid, apart from the most demanding simulations which took place in 3D with high nonlinearities and long hold times.
-The difficulty with these 3D simulations was that due to the large number of modes they included, the number of particles per mode could drop to ten or less in the densest regions where the squeezing was generated.
+For our simulations we checked the results of our simulations against both the analytic two-mode solution, as well as checking for typical indications of TWA breakdown such as the appearance of negative densities in lightly populated modes.
+We could also use the Bogoliubov analysis described in section \ref{sec:2D3Dsqueezing} as an independent estimate of the validity of the TWA.
+We found that the TWA remained valid except in some 3D simulations with high nonlinearities, where there were a large number of modes, and the number of particles per mode could drop to ten or less in the densest regions where the squeezing was generated.
 Over very long time scales, this low mode occupation began to result in TWA breakdown.
-As we could characterise the regime in which breakdown began to occur, however, we were able stay away from it and to ensure our simulations remained valid.
+Fortunately, the best squeezing was typically found in regions where the TWA was valid.
 
@@
 When approximating these equations on a discrete grid, the magnitude of initial noise, the relationship between stochastic averages and the expectation values, and the equations of motion all become grid-dependent.
-This is less concerning as it may at first appear, as the Hamiltonian in Eq.~(\ref{eqFieldHamiltonian}) assumes a contact potential between the atoms, and this is not correct below a length scale that can probe the details of the true potentials.
-Furthermore, well before that regime is reached, the different elements of grid dependence cancel out to give grid-independent predictions for physical observables.
+This is not a signature of a fundamental problem, as the Hamiltonian in Eq.~(\ref{eqFieldHamiltonian}) assumes a contact potential between the atoms, which is not correct below a length scale that can probe the details of the true potentials.
+Furthermore, it is not a problem in practice, as the calculation gives grid-independent predictions for physical observables well before that regime is reached.
 
@@
 
-\textbf{Note the initial state we use, and the equations.}
-
 We integrate equations (\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS2 \cite{dennis2013}.
 For the one- and two-dimensional simulations, the dimensional reduction is achieved by estimating the mode shape in three dimensions, and using dimensionally reduced values for $U_{ij}$ that match the zero-dimensional reduction given by Eq.~(\ref{eqChiUequivalence}).
-An equivalent process is to match the chemical potentials of the initial states.
+To make relevant comparisons between equivalent situations with different numbers of spatial dimensions, this mode shape is typically chosen to match the chemical potentials of the initial states.
 
@@
 The plot shows the normalized number variance in mode $|a \rangle$ as a function of the time the final coupling pulse is applied.
-Thick red line, analytic two-mode solution; solid blue line, constant density mode; dashed line, Thomas-Fermi mode; dash-dotted line, Gaussian mode.
+We compare the analytic two-mode solution (thick red), the constant density mode (solid blue), the Thomas-Fermi mode (dashed) and the Gaussian mode (dash-dotted).
 The nonlinearities of the various 1D simulations have been adjusted for their mode shape so that they are equivalent to that of the 0D, two-mode mode nonlinearity.
-Parameters: $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \hbar, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$.
+Parameters used were $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \hbar, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$.
 Maximum squeezing is achieved when the atomic density profile is the most uniform.}
@@
 
-One way to avoid the requirement to engineer and maintain specific atomic modes is to post-select a mode.
+One way to avoid any requirement to engineer and maintain specific atomic modes is to post-select a spatial mode.
 Rather than considering the variance in the total number of atoms in a mode, it may be possible to consider only the statistics of a spatially filtered area of the atomic cloud.
@@
 \end{equation}
-A simulations comparing the best attainable squeezing for this mode compared to an unfiltered Gaussian is shown in Figure~\ref{XXX}.
-We can see that the squeezing is higher.
-One obvious disadvantage to post-selecting modes is that it  reduce the effective number of atoms.
+Spatial filtering can easily produce higher squeezing, but there is the obvious disadvantage that it  significantly reduces the effective number of atoms.
 This is a serious issue for most applications of squeezed sources, such as interferometry, where the signal to noise scales with the square root of the flux.
+We will focus on methods for producing squeezing with large atomic number.
 
@@
 Modal mismatch is generated by the different atomic potentials seen by the two internal states.
-We propose a straightforward method of minimising this effect: using a $\pi$-pulse to swap the populations of the internal states after half of the evolution.
-As can be seen by Eqs~(\ref{eqa1},\ref{eqb1}, such a pulse also leaves the quantum statistics of each state unchanged.
-In the limit where each state stayed in a single mode (although that mode might change), and exactly half of the atoms are in each state, this method would produce identical mode shapes for both modes after the hold time.
+A straightforward method of minimising this effect is to use a $\pi$-pulse to swap the populations of the internal states after half of the evolution.
+As can be seen by Eqs~(\ref{eqa1}) and (\ref{eqb1}), such a pulse also leaves the quantum statistics of each state unchanged.
+In the limit where each state remains in a single mode (although that mode might change), and exactly half of the atoms are in each state, this technique ensures that both modes have identical mode shapes after the hold time.
 In practice this will never work perfectly as the mode shapes will change, resulting in non-identical environments for the two modes.
@@
     \caption{Effect of a $\pi$-pulse halfway through the hold time.
-Common parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{ab}=\chi_{bb}=0$.
+Common parameters used were $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{ab}=\chi_{bb}=0$.
 Solid red line: Analytic two-mode solution, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (10.9dB squeezing); Solid blue line: no $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (4.9dB squeezing);  Dashed blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.05$ (5.8dB squeezing); Dash-dotted blue line: with $\pi$-pulse and the longer hold time $\tau_{\mathrm{hold}}=5\times 10^{-4}$s allowed by the application of the pulse, $\phi=3.87$ (7.2dB squeezing).}
@@
 
-\section{Squeezing in 2D and 3D}
+\section{Squeezing in 2D and 3D} \label{sec:2D3Dsqueezing}
 
@@
   \end{subfigure}
-\caption{Effects of dimension on squeezing in two different nonlinear regimes.
+\caption{Effects of dimension on squeezing in two different nonlinear regimes, assuming that the chemical potential is constant for all simulations.
 If two nonlinearities are non-zero, the deleterious effects of higher dimensions on squeezing are exacerbated.
@@
 The simulations in the previous two sections were one dimensional, and show how having spatially constant density and using a $\pi$-pulse will improve the resultant squeezing.
-While these conclusions remain valid for simulations in two or three dimensions, adding each additional dimension results in a degradation of the squeezing.
-This is true even when the potential (and the resulting atomic density) is constant across the bulk of the trap.
-Furthermore, the expected squeezing goes down as the volume of the trap increases.
-This can be seen in Fig.~\ref{figDimensionalSqueezingEffects}, where we show the squeezing generated in 1D, 2D and 3D BECs for two different parameter regimes.
+While these conclusions remain valid for simulations in two or three dimensions, adding each additional dimension increases the number of spatial modes in a given energy range.
+For BECs of different dimension at the same energy per particle, this leads to a degradation of squeezing, even when the potential (and the resulting atomic density) is constant across the bulk of the trap.
+This is demonstrated in Fig.~\ref{figDimensionalSqueezingEffects}, where we show the squeezing generated in 1D, 2D and 3D BECs for two different parameter regimes.
 In both cases we take the BEC to be trapped in a box potential, which in the Thomas-Fermi limit leads to a constant density profile across the BEC.
@@
 
-This dependence on dimensionality and volume are not explained by our arguments in section \ref{sec:MMdescription}, which focus on the effects of inhomogeneity within the trap.
+To find the trap geometries that produce the best squeezing, it is important to understand the root cause of this loss of squeezing in higher dimensional traps.
+The dependence on dimensionality is not explained by the processes described in section \ref{sec:MMdescription}, which focus on the effects of inhomogeneity within the trap.
 Fundamentally, this degradation of the squeezing is due to nonlinearity-induced coupling between different spatial modes.
-As the dimension and volume increases, the spectral density of available excited modes increases, leading to a reduction in squeezing.
 
@@
 To simplify the analysis we consider the case in which $U_{aa} \neq 0$, $U_{bb}=U_{ab} = 0$ and the condensate has a spatially-constant density in a box of side lengths $L_x$, $L_y$ and $L_z$.
-In this limit the squeezing $S_a = \Var(N_a)/N_a$ of the $a$ mode at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation})
+In this limit the squeezing $S_a = \Var(N_a)/N_a$ of the $|a\rangle$ mode at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation}):
 \begin{align}
@@
 \end{align}
-where $n_a(\mathbf{k}, t) = \expect{\hat{a}^\dagger(\mathbf{k}, t) \hat{a}(\mathbf{k}, t)}$ is the occupation of momentum mode $\mathbf{k}$, $\Var[N_a](t)$ and $N_a(t)$ are the number variance and mean number of the $a$ mode from the two-mode model in Section~\ref{secTwoModeAnalytic}, the sum over $\mathbf{k}$ is a discrete sum over the available non-zero momentum modes $\mathbf{k} = (2\pi/L_x) n_x \hat{\mathbf{x}} + (2\pi/L_y) n_y \hat{\mathbf{y}} + (2\pi/L_z) n_z \hat{\mathbf{z}}$ with $n_x$, $n_y$ and $n_z$ arbitrary integers, $\omega_\mathbf{k}$ is the frequency of the Bogoliubov mode $\mathbf{k}$ and $\omega^0_\mathbf{k}$ is the frequency of a free particle.
+where $n_a(\mathbf{k}, t) = \expect{\hat{a}^\dagger(\mathbf{k}, t) \hat{a}(\mathbf{k}, t)}$ is the occupation of momentum mode $\mathbf{k}$, $\Var[N_a](t)$ and $N_a(t)$ are the number variance and mean number of the $|a\rangle$ mode from the two-mode model in Section~\ref{secTwoModeAnalytic}, the sum over $\mathbf{k}$ is a discrete sum over the available non-zero momentum modes $\mathbf{k} = (2\pi/L_x) n_x \hat{\mathbf{x}} + (2\pi/L_y) n_y \hat{\mathbf{y}} + (2\pi/L_z) n_z \hat{\mathbf{z}}$ with $n_x$, $n_y$ and $n_z$ arbitrary integers, $\omega_\mathbf{k}$ is the frequency of the Bogoliubov mode $\mathbf{k}$ and $\omega^0_\mathbf{k}$ is the frequency of a free particle.
 
-The squeezing of mode $a$ in \eqref{eqBogoliubovSqueezing} includes the first-order multimode quantum mechanical fluctuations about the mean field and differs from the single-mode squeezing of $\Var[N_a](t_3)/N_a(t_3)$ only by terms involving the occupation of the non-zero momentum modes $n_a(\mathbf{k}, t)$.
+The squeezing of mode $|a\rangle$ in \eqref{eqBogoliubovSqueezing} includes the first-order multimode quantum mechanical fluctuations about the mean field and differs from the single-mode squeezing of $\Var[N_a](t_3)/N_a(t_3)$ only by terms involving the occupation of the non-zero momentum modes $n_a(\mathbf{k}, t)$.
 The contribution of these terms is always detrimental to squeezing as the contribution to the numerator of \eqref{eqBogoliubovSqueezing} is always at least as large as the contribution to the denominator.
@@
 Therefore when these non-zero momentum modes become sufficiently occupied for $S_a$ to differ from the two-mode model, the Bogoliubov approximation becomes invalid and $S_a$ will not agree with the full quantum-statistical multimode result.
-Although the Bogoliubov approximation cannot predict the amount by which squeezing will be degraded compared to the two-mode model, it can predict the regime in which squeezing is reduced and suggest how this problem can be avoided.
-Figure~\ref{figBogSqueezingValidation}(a) illustrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the total number of atoms in the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ for parameters where these modes reach ~20\% of the total number of atoms in the system.
-Figure~\ref{figBogSqueezingValidation}(b) shows that in this case the squeezing predicted by the Bogoliubov theory in \eqref{eqBogoliubovSqueezing} does not agree with either the two-mode model or the truncated Wigner simulations, but both the Bogoliubov theory and the truncated Wigner simulations show worse squeezing than predicted by the two-mode model.
+While number-conserving variants of the Bogoliubov theory exist \cite{NumberConservingBog}, they are still perturbative approximations about the mean field, and they are not required for our purposes.
+Although the Bogoliubov approximation cannot predict the exact amount by which squeezing will be degraded compared to the two-mode model, it does predict the regime in which squeezing is reduced by coupling between spatial modes, it can thereby suggest how this problem can be avoided.
+Figure~\ref{figBogSqueezingValidation}(a) illustrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the total number of atoms in the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ for parameters where these modes approach 20\% of the total number of atoms in the system.
+Figure~\ref{figBogSqueezingValidation}(b) shows that in this case the squeezing predicted by the Bogoliubov theory in Eq.~\eqref{eqBogoliubovSqueezing} is a lower bound for the squeezing predicted by the truncated Wigner simulations.
+The two-mode analytic solution, which ignores multimode effects entirely, provides an upper bound for the squeezing.
+\textbf{Reconsider these last sentences when we have the figure.
+Maybe qualify `lower bound' for constant density models only?}  This means that we can use the Bogoliubov analysis to guide our search for parameter regimes with higher squeezing.
 
-To minimise the damaging effects of the non-zero momentum modes, it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small.
+To minimise the damaging effects of the non-zero momentum modes, it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation.
+The number of excited modes is almost entirely determined by the density of modes energetically accessible by the system.
+While this gets larger with the dimensionality and the volume of the system at a given chemical potential, in practice the dimensionality and the volume of the BEC are controlled by changing the trap parameters, which has a dramatic effect on the attainable chemical potential.
+So the true figure of merit controlling the degradation of the squeezing due to coupling between the modes is the number of the number of states energetically reachable from the chemical potential.
+This scales differently with trap size in each of 1D, 2D and 3D.
+For one-dimensional traps, the number of accessible states increases with volume, whereas for 2D traps it is constant with volume.
 
-  \begin{itemize}
-  \item Introduce Bog analysis
-  \item Do a comparision plot between SM analytic model and Bog to show everything agrees
-  \item show that the number of mode available degrades squeezing
-  \item Plot some bog results v MM sims for increasing box sizes, and increasing dimensions
-  \item Can we put some numbers on how bad 3D is compared to 2D, in order to motivate tight trapping?
-  \end{itemize}
+Fig.~\ref{fig:Combined1Dand2D} shows the non-condensed fraction predicted by the Bogoliubov analysis for one and two dimensions as a function of $r$, which is a scale factor for the size of the trap.
+\textbf{Consider not making it a function of r, but simply the size of the trap?}  The traps are assumed to have a constant potential, which minimises the effects of the multimode issues described in section \ref{sec:MMdescription}.
+The squeezing in each trap is calculated using the TWA method, and we see that the squeezing does indeed degrade at high $r$ in 1D, but remains constant in $2D$.
+For low $r$, the squeezing in 1D is limited by the other multimode effects.
+The non-smooth nature of the squeezing prediction in 2D is due to the stochastic nature of the simulations.
 
-Some notes on the Bog scaling plots below: Experimentalist's scaling works as follows: V $\rightarrow$ r V; hold time $\rightarrow$ r*(hold time).
+In Fig.~\ref{fig:3DBogScaling} we see that the number of accessible states actually decreases with volume.
+This is because the energy of the ground state of the BEC is reducing faster than the density of states at low energies is increasing.
+\textbf{Create these figures (by combining old figures) and check both for latest updates.
+Confirm that 'simulation' (blue line) means TWA, as opposed to Graham's prediction.}  The non-smooth nature of the Bogoliubov calculation is due to the coarse grid, which leads to granularity as the scale factor is changed.
+
+%Some notes on the Bog scaling plots below: Experimentalist's scaling works as follows: V $\rightarrow$ r V; hold time $\rightarrow$ r*(hold time).
 In order to get enough lattice points to cover the momentum modes up to energy mu, the scaling goes as: 1D lattice$\rightarrow$lattice*$\sqrt{r}$ (1D); lattice$\rightarrow$lattice (2D); lattice$\rightarrow$lattice*$r^{-1/6}$.
-
-The plots show that the experimentalist's scaling increases Bog occupation (i.e.
-condensate depletion) with volume in 1D, is constant in 2D, and decreases it in 3D.
-This means squeezing gets worse with volume in 1D, is constant with volume in 2D, and gets better with volume in 3D.
-We can't actually show squeezing getting better in 3D due to TWA breakdown, but we do show bog mode occupation decreasing.
-
-The jaggedness of the line in 3D is due to the fact that we're only dealing with 10-20 lattice points in each dimension, and the lattice cutoff is often part way between two integer numbers, so we have to round up or down as we change the scaling factor.
-
-Jaggedness of squeezing line in 2D is due to stochastic nature of simulations.
 
@@
 
+\textbf{Obviously 3D TWA simulation can be done in some limit - the one that produced the 15db result.  Do we say something about this?}
+
+This scaling of the estimated Bogoliubov fraction in 3D suggests that strong squeezing may actually be found in very large, weakly trapped condensates, provided the trapping potentials are box-like.
+This can be confirmed with a TWA simulation, although systematic analysis of squeezing across a wide range of large 3D traps and interaction strengths is impossible due to the breakdown of the TWA method.
+Fig.~\ref{figVolumeSqueezingEffects} shows a result of 15dB of squeezing achievable in \textbf{some parameter regime}.
+This system produces an extremely low non-condensed fraction of $0.0004$, as expected from the Bogoliubov calculations, and squeezing that is 150\% better than Bogoliubov lower bound.
+
 \begin{figure}
@@
    \includegraphics[width=7cm]{figures/volume_effect_on_squeezing.eps}
-\caption{(most of this caption should be moved to main text and expanded) Effects of volume on squeezing.
-2D simulation of how the size of the box affects squeezing.
-Horizontal axis shows the linear size of the edges of the box in microns.
-The larger the box, the more low-momentum modes are included, which means worse squeezing (see Bog analysis).
-Dashed red line, analytic zero-D solution; solid blue line, relative number variance in mode $|a\rangle$.
-All simulations are 128x128 grid, 10,000 paths.
-Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=6.2$.}
+\caption{\textbf{This figure is not yet the 'winner winner chicken dinner' result. } Demonstration of high squeezing from a large volume, 3D trap.  \textbf{Simulation parameters.}}
  \label{figVolumeSqueezingEffects}
@@
 
-\section{Mitigating multimode quantum statistical effects}
-  \begin{itemize}
-  \item 3D still sucks even with box, so point out we can win by freezing out dimensions
-  \item Show derivation for what physical situations are needed to freeze out 1, 2 and 3 dimensions.
-In general we need small number, which makes it hard since we want large number squeezing.
-Large number requirement leads to implausible geometry in 1D (i.e.
-a tube with a ridiculous aspect ratio), but going from 3D down to 2D is probably do-able
-  \end{itemize}
+%Simulating a 3D BEC in a box with the same parameters as the experimental paper works well.
+Bog mode fraction is around 4e-4, and we get squeezing of around 0.026.
+This is assuming we can set two scattering lengths to zero, and have the third equal to that of Rb87.
+%Interestingly, Graham's bog theory predicts squeezing of 0.0398 (assuming I'm doing it correctly), but both the single-mode analytic answer and the simulation give 0.026.
 
-As discussed in the previous section, the existence of a large number of vacuum modes results in the degradation of squeezing due to nonlinear coupling between the modes, and this occurs regards of whether the modes are populated or not.
-The modes that play a role in this process are those with energies below the chemical potential of the system.
-In order to mitigate this effect, it is necessary to have as few of these modes as possible.
+%Just in case this was the real version of the best data:
 
-In any confined system, momentum modes will be discretized, with a spacing in a particular dimension given by $\delta k \sim 2\pi/L$.
-
-{\bf *** It occurs to me that I'm about to repeat what Graham will say in previous section, so leave it for now ***}
-
-Consequently, the number of low-lying momentum modes that result in degraded squeezing scales with the power of the dimension of the system.
-As shown in Figure~\ref{figDimensionalSqueezingEffects}, obtaining large number squeezing from a three-dimensional, multimode system is likely to be untenable.
-This is unfortunate, as the universe in which we live is unpleasantly three-dimensional.
-
-All is not lost, however, as it is possible to create systems that are effectively lower-dimensional.
-This is accomplished by having trapping potentials that are tight enough in one or more dimensions such that any dynamics in that dimension are effectively frozen out.
-Put another way, if the trapping potential is tight enough, then the energy required to excite any mode beyond that of the ground state is unavailable.
-Conditions required to reach this limit vary on the form of the potential, but, provided we consider a harmonic potential, a simple criterion for ensuring single mode behaviour will be
-\begin{equation}
-\hbar \omega \gg \mu
-\label{eqTightTrappingCriterionMu}
-\end{equation}
-that is, the spacing between energy levels is much greater that that of the chemical potential.
-
-Given that we have shown that constant density profiles result are required for good squeezing, we consider a 3D potential that has tight harmonic trapping in one, two or three dimensions, and a box-like potential in the other dimensions.
-Our goal is to make this system effectively two-, one- or zero-dimensional respectively.
-
-We assume a trapping frequency $\omega$ in the harmonically trapped dimensions and a confining region of length $L$ in the box potential dimensions.
-We further assume that we are working in the Thomas-Fermi limit, as this level of nonlinearity is required to obtain good squeezing.
-This means that the criterion (\ref{eqTightTrappingCriterionMu}) becomes
-\begin{equation}
-\hbar \omega \gg \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz
-\label{eqTightTrappingCriterionU}
-\end{equation}
-where the factor of $N$ is present because we require the energy per particle, not the total energy.
-
-Due to the Thomas-Fermi limit, the density profile $|\psi({\mathbf{r}})|^2$ will be constant in the box-potential directions, and due to the assumption of tight trapping, close to the harmonic oscillator ground state in the harmonically trapped directions.
-Evaluating Eq.~(\ref{eqTightTrappingCriterionU}) with these density distributions and using $U=4\pi \hbar^2 a /m$ we obtain
-\begin{eqnarray}
-\sqrt{\frac{\hbar}{m \omega}}  \gg \frac{aN}{\sqrt{2 \pi}} && \,\,\,\, ({\textrm{0D}}) \label{eq0Dreduction} \\
-L  \gg  a N / \pi && \,\,\,\, ({\textrm{1D}}) \label{eq1Dreduction} \\
-\omega  \gg  \frac{8 \pi a^2 N^2 \hbar}{m L^4}  && \,\,\,\, ({\textrm{2D}}) \label{eq2Dreduction}
-\end{eqnarray}
-as conditions we require to reduce a 3D system to an effective 0D, 1D or 2D system.
-
-To determine how feasible this dimensional reduction is, we assume Rubidium atoms, typical scattering lengths ($a\sim$10nm), large particle numbers ($N\sim 10^6$) and a confining length scale of $L\sim 200\mu$m in the box dimensions.
-
-In order to reduce the system to 0D, Eq.~(\ref{eq0Dreduction}) indicates we require a harmonic oscillator length of $\sqrt{\hbar/ m \omega}\gg 4$mm; a BEC several centimeters in diameter is not particularly feasible.
-Using Eq.~(\ref{eq1Dreduction}) demonstrates reduction to 1D requires a BEC that is $\gg3.2$mm in length; this would also be challenging.
-
-Reduction to 2D appears more promising, however. For the parameters given above, we see that we need $\omega \gg 2\pi\times 190$Hz, which is certainly feasible.
-
-We note that the reason our dimensional reduction criteria to obtain effective 0D and 1D systems are so stringent, is entirely due to our requirement of high particle number.
-For low particle number, the criteria become much more forgiving.
-An example is the experiment of Gross {\it et al.} which measured a number difference squeezing of 6.9dB in a spherical trap with a trapping frequency of $\omega=2 \pi \times 425$Hz.
-Due to the low number of atoms in their trap ($\sim 400$), they were able to achieve an effectively 0D trap with single mode behaviour.
-
-%Reduction from 3D to 0D: Here we have tight trapping along all three dimensions, so we have a harmonic oscillator wavefunction along all dimensions.
-With $N$ particles in the system, the density is given by
-%\begin{equation}
-%|\psi|^2 = N \left( \frac{m \omega}{\pi \hbar} \right)^{3/2} \exp \left[ -\frac{m \omega} {\hbar} (x^2 +y^2+z^2) \right] .
-%\end{equation}
-%In this system, the nonlinear energy per particle is
-%\begin{eqnarray}
-%E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
-%       &=& \frac{U N}{4 \sqrt{2}} \left( \frac{m \omega}{\pi \hbar} \right)^{3/2}
-%\end{eqnarray}
-%where $U = \pi \hbar^2 a/m$ and $a$ is the s-wave scattering length of the particles. The criterion $\hbar \omega > E_{NL}$ thus becomes
-%\begin{equation}
-%%N < \frac{1}{a} \sqrt{2 \pi \hbar}{m \omega}.
-%\sqrt{\frac{\hbar}{m \omega}} > \frac{aN}{\sqrt{2 \pi}}.
-%\end{equation}
-%Assuming Rubidium atoms, typical scattering lengths ($a\sim$10nm) and large particle numbers (say $N\sim 10^6$), we require a harmonic oscillator %length $\sqrt{\hbar/ m \omega}>4$mm.
-A BEC that's more than a centimetre across isn't particularly feasible.
-%
-%Reduction from 3D to 1D: Here we have tight trapping along $x$ and $y$, with a box potential along $z$. With $N$ particles in the system, the density is given by
-%\begin{equation}
-%|\psi|^2 = \frac{N}{L} \frac{m \omega}{\pi \hbar} \exp \left[ -\frac{m \omega} {\hbar}  (x^2 +y^2) \right] .
-%\end{equation}
-%In this system, the nonlinear energy per particle is
-%\begin{eqnarray}
-%E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
-%       &=& \frac{U N m \omega}{4 \pi \hbar L}
-%\end{eqnarray}
-%The criterion $\hbar \omega > E_{NL}$ thus becomes
-%\begin{equation}
-%L > a N / \pi.
-%\end{equation}
-%Assuming Rubidium atoms, typical scattering lengths ($a\sim$10nm) and large particle numbers (say $N\sim 10^6$) we need a BEC that's a few millimeters long, which isn't particularly feasible.
-%
-%Reduction from 3D to 2D: Here we have tight trapping along $z$, with a box potential along $x$ and $y$. With $N$ particles in the system, the density is given by
-%\begin{equation}
-%|\psi|^2 = \frac{N}{L^2} \left( {\frac{m \omega}{\pi \hbar}} \right) ^{1/2} \exp \left[ -\frac{m \omega} {\hbar} z^2  \right]
-%\end{equation}
-%In this system, the nonlinear energy per particle is
-%\begin{eqnarray}
-%E_{NL} &=& \frac{U}{2N} \int |\psi({\mathbf{r}})|^4 dx \, dy \, dz \\
-%       &=& \frac{m \omega U N}{4 \pi \hbar L^2}
-%\end{eqnarray}
-%The criterion $\hbar \omega > E_{NL}$ thus becomes
-%\begin{equation}
-%\omega > \frac{8 \pi a^2 N^2 \hbar}{m L^4}.
-%\end{equation}
-%Once again taking a Rubidium atom and considering typical parameters, say $a\sim$10nm, $N\sim 10^6$, $L\sim 200\mu$m, we see that we need $\omega > 2\pi\times 190$Hz, which is feasible.
-
-We finish this section by considering an example of the sort of number squeezing that is achiveable in a realistic physical system with large particle number.
+%We finish this section by considering an example of the sort of number squeezing that is achieveable in a realistic physical system with large particle number.
 We take a Rb$^{85}$ condensate, which possesses a number of Feschbach resonances that allow us to modify one of the $U_{ij}$ scattering lengths without affecting the others.
@@
 
-We also note that configurations with higher squeezing are certainly possible, but on these time scales a full 3D simulation begins to suffer breakdown of the TWA, making numerical results unreliable.
-\begin{figure}
-  \centering
-   \includegraphics[width=7cm]{figures/realistic_3d_squeezing_params.eps}
-\caption{Full 3D simulation of relative number squeezing in state $|a\rangle$ in a system with realistic parameters.
-Box potential in two dimensions with area 120$\mu$m $\times$ 120$\mu$m, tight harmonic trapping in 3rd dimension with $\omega=1500$Hz.
-Solid red line, analytic 0D solution; dashed blue line, full 3D multimode simulation.
-Parameters: Rb$^{85}$ atoms, $n_a = n_b =2 \times 10^5$, $U_{aa}=2.1\times 10^{-50}$ (corresponding to an s-wave scattering length of $-400 a_0$), $U_{bb}=U_{ab}=0$, $\tau_{\mathrm{hold}}=1\times 10^{-3}$s, $\phi=5.3$.
-Even with all multimode effects present, the number squeezing in the 3D system (13.9dB) approaches that of the ideal 0D case (15.1dB).}
- \label{figRealistic3DParameters}
-\end{figure}
-
 \section{Conclusion}
 \label{sectionConclusion}
-\begin{itemize}
-  \item Kerr nonlinearities can give arbitrarily good squeezing in BECs with high number
-  \item In a straightforward approach we can't get good squeezing with high number due to MM effects
-  \item Mode matching issues can be solved with equal scattering lengths or a pi pulse or a box
-  \item Another MM effect scales with dimension and physical size
-  \item Can solve that problem with tight trapping and box
-\end{itemize}
+
+We have presented an analytic solution to the two-mode model of atomic fields with a strong Kerr non-linearity.
+The three nonlinearities between and within the modes are not yet known for many atomic species, and often have sensitive dependence on external magnetic fields due to Feshbach resonances \cite{FeshbachOrReview}, so this solution allows for a rapid estimate of the atomic number difference squeezing that can be gained from these systems as this data is obtained.
+In general, it predicts arbitrarily good squeezing in the two-mode limit.
+\textbf{Revise that.}
+
+Operating in a parameter regime where there is truly only one spatial mode, which is essential for the analytic solution to be valid, requires extremely tight trapping potentials.
+Furthermore, due to the nonlinearities, it also comes with a strong limit on the number of atoms that can be trapped.
+Interferometric applications of squeezing become valuable at high atomic flux, so we investigate methods of mitigating multimode spatial effects that degrade the squeezing.
+
+Using a constant potential in the trap (a `box' trap) allows the BEC density to be spatially uniform throughout the process.
+This is shown to alleviate the difficulties with choosing correct hold times and phases for the coupling pulses.
+Along with a pi-pulse midway during the holding period, this also alleviates the majority of the mode-matching issues for the final recombination of the two modes.
+The remaining issue with the existence of multiple spatial modes was the possibility of scattering into them by the nonlinearities.
+
+By a Bogoliubov analysis, it was shown that the coupling between different modes could be minimised in two regimes.
+The first was the tight trapping regime, where the system becomes effectively zero-dimensional, and the second was the weak trapping limit for 3D traps.
+This latter regime is in fact ideally compatible with large atomic numbers, and shows the prospect of extremely high squeezing.
+A TWA simulation demonstrated a BEC trap with $15dB$ of squeezing.
+In practice, the only limit to the number of atoms in such traps would be the stability of the trap, as the BEC would become increasingly sensitive to fluctuations as the energy per particle is reduced as the volume increases.
 
 \ack
-Acknowledgements in here; funding sources etc.  Joe's Future Fellowship.
+Acknowledgements in here; funding sources etc.  Joe's Future Fellowship.  NCI?
 

### 2013-05-24 large_number_squeezing_with_atoms.tex
Minor edits.

@@
 This is done by writing the multimode atomic field operators in terms of a set of basis functions $\hat{\psi}_a({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{a}_{n}(t)$ and $\hat{\psi}_b({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{b}_{n}(t)$, where $\hat{a}_{n}$ and $\hat{b}_{n}$ annihilate a particle with normalised spatial mode $u_n$ and internal state $|a\rangle$ or $|b\rangle$ respectively.
-With this expansion, the first nonlinear term in the multimode Hamiltonian becomes
+With this expansion, the nonlinear term in the multimode Hamiltonian for the self-interaction of the $|a\rangle$ internal state becomes
 \begin{equation}
@@
 Different modes in a multimode environment will typically have varying effective nonlinearity-particle number products and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.
-This means that the best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across all the spacial modes present, result in an overall lowest possible total number variance.
+This means that the best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across all the spatial modes present, result in an overall lowest possible total number variance.
 
@@
 One way to avoid any requirement to engineer and maintain specific atomic modes is to post-select a spatial mode.
-Rather than considering the variance in the total number of atoms in a mode, it may be possible to consider only the statistics of a spatially filtered area of the atomic cloud.
+Rather than considering the variance in the total number of atoms in each internal state, it may be possible to consider only the statistics in a spatially filtered area of the atomic cloud.
 For example, suppose the atoms were in a Gaussian atomic distribution, but we consider a mode that only includes the atoms within $p$ standard deviations of the centre.
@@
 While number-conserving variants of the Bogoliubov theory exist \cite{NumberConservingBog}, they are still perturbative approximations about the mean field, and they are not required for our purposes.
-Although the Bogoliubov approximation cannot predict the exact amount by which squeezing will be degraded compared to the two-mode model, it does predict the regime in which squeezing is reduced by coupling between spatial modes, it can thereby suggest how this problem can be avoided.
+Although the Bogoliubov approximation cannot predict the exact amount by which squeezing will be degraded compared to the two-mode model, it does predict the regime in which squeezing is reduced by coupling between spatial modes, and it can thereby suggest how this problem can be avoided.
 Figure~\ref{figBogSqueezingValidation}(a) illustrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the total number of atoms in the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ for parameters where these modes approach 20\% of the total number of atoms in the system.
@@
 While this gets larger with the dimensionality and the volume of the system at a given chemical potential, in practice the dimensionality and the volume of the BEC are controlled by changing the trap parameters, which has a dramatic effect on the attainable chemical potential.
-So the true figure of merit controlling the degradation of the squeezing due to coupling between the modes is the number of the number of states energetically reachable from the chemical potential.
+So the true figure of merit controlling the degradation of the squeezing due to coupling between the modes is the number of states energetically reachable from the chemical potential.
 This scales differently with trap size in each of 1D, 2D and 3D.
 For one-dimensional traps, the number of accessible states increases with volume, whereas for 2D traps it is constant with volume.
+For three-dimensional traps, the number of accessible states \emph{decreases} with volume.
 
@@
 
-In Fig.~\ref{fig:3DBogScaling} we see that the number of accessible states actually decreases with volume.
+In Fig.~\ref{fig:3DBogScaling} we see that the number of accessible states for three-dimensional traps actually decreases with volume.
 This is because the energy of the ground state of the BEC is reducing faster than the density of states at low energies is increasing.
 \textbf{Create these figures (by combining old figures) and check both for latest updates.
-Confirm that 'simulation' (blue line) means TWA, as opposed to Graham's prediction.}  The non-smooth nature of the Bogoliubov calculation is due to the coarse grid, which leads to granularity as the scale factor is changed.
+Confirm that `simulation' (blue line) means TWA, as opposed to Graham's prediction.}  The non-smooth nature of the Bogoliubov calculation is due to the coarse grid, which leads to granularity as the scale factor is changed.
 

### 2013-06-04 large_number_squeezing_with_atoms.tex
Fix figures, plus minor typo fixes

@@
 
-\begin{figure}
-  \centering
-   \includegraphics[width=7cm]{figures/volume_effect_on_squeezing.eps}
-\caption{\textbf{This figure is not yet the 'winner winner chicken dinner' result. } Demonstration of high squeezing from a large volume, 3D trap.  \textbf{Simulation parameters.}}
- \label{figVolumeSqueezingEffects}
-\end{figure}
-
 %Simulating a 3D BEC in a box with the same parameters as the experimental paper works well.
@@
 \begin{equation}
-\hat{a}_1 = e^{ i H_1 t/ \hbar} \, \hat{a}_0 \, e^{-i H_1 t/ \hbar}.
+\hat{a}_1 = e^{ i \hat{H}_1 t/ \hbar} \, \hat{a}_0 \, e^{-i \hat{H}_1 t/ \hbar}.
 \end{equation}
@@
 \begin{equation}
-e^X Y e^{-X} = Y + [X,Y] + \frac{1}{2!}[X,[X,Y]] + \frac{1}{3!}[X,[X,[X,Y]]] + \ldots
+e^{\hat{X}} \hat{Y} e^{-\hat{X}} = \hat{Y} + [\hat{X},\hat{Y}] + \frac{1}{2!}[\hat{X},[\hat{X},\hat{Y}]] + \frac{1}{3!}[\hat{X},[\hat{X},[\hat{X},\hat{Y}]]] + \ldots
 \label{eqHadamard}
@@
 we obtain 
-\begin{eqnarray}
-\hat{a}_1 &=& \hat{a}_0 - i \Omega t_1 \hat{b}_0 + \frac{(i \Omega t_1)^2}{2!} \hat{a}_0 - \frac{(i \Omega t_1)^3}{3!} \hat{b}_0 + \frac{(i \Omega t_1)^4}{4!} \hat{a}_2 - \ldots \nonumber \\
-          &=& \cos (\Omega t_1) \hat{a}_0 -i \sin (\Omega t_1) \hat{b}_0.
+\begin{equation}
+\hat{a}_1 =  \cos (\Omega t_1) \hat{a}_0 -i \sin (\Omega t_1) \hat{b}_0.
 \label{eqa1}
-\end{eqnarray}
+\end{equation}
 Similarly, we find
@@
 
-As $\hat{a}^{\dagger}_1 \hat{a}_1$ and $\hat{b}^{\dagger}_1 \hat{b}_1$ commute with the Hamiltonian $H_2$, they are constants of motion so we have
+As $\hat{a}^{\dagger}_1 \hat{a}_1$ and $\hat{b}^{\dagger}_1 \hat{b}_1$ commute with the Hamiltonian $\hat{H}_2$, they are constants of motion so we have
 \begin{equation}
@@
 \end{eqnarray}
-When substituted into (\ref{eqNavariance}) and (\ref{eqNbvariance}) these expecation values give the absolute number variances stated in Section \ref{secTwoModeAnalytic}.
+When substituted into (\ref{eqNavariance}) and (\ref{eqNbvariance}) these expectation values give the absolute number variances stated in Section \ref{secTwoModeAnalytic}.
 To calculate the number difference variance, we note

### 2013-06-04 large_number_squeezing_with_atoms.tex
Handle Graham's comments, more figure and caption fixes, plus random explanation improvements here and there

@@
 At time $t=0$, the coupling $\Omega$ is switched on and applied until time $t=t_1$, at which point it is turned off, resulting in a portion of the population being transferred into mode $|b\rangle$.
-As the initial state was a coherent state in both modes, after the population transfer both $|a\rangle$ and $|b\rangle$ are described by coherent states, with mean number $\langle \hat{a}^{\dagger} \hat{a} \rangle = n_a$ and $\langle \hat{b}^{\dagger} \hat{b} \rangle = n_b$ respectively.
+We assume this transfer process fast relative to the nonlinear energy time scale, ensuring that, as the initial states were coherent states, after the population transfer both $|a\rangle$ and $|b\rangle$ are still described by coherent states with mean number $\langle \hat{a}^{\dagger} \hat{a} \rangle = n_a$ and $\langle \hat{b}^{\dagger} \hat{b} \rangle = n_b$ respectively.
 
@@
 Different modes in a multimode environment will typically have varying effective nonlinearity-particle number products and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.
-This means that the best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across all the spatial modes present, result in an overall lowest possible total number variance.
+This means that the best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across all the spatial modes present, result in the overall lowest total number variance.
+Clearly, however, this averaged number variance will be still be higher than the specific mode that exhibits the best squeezing.
 
@@
 Due to the nonlinear term in the Hamiltonian, the local phase evolution of the atomic field is density dependent.
-Consequently the mode shape of the atomic field changes depending on the nonlinearity-density product, which means that unless $U_{aa}|\psi_a({\mathbf{r}})|^2 = U_{bb}|\psi_b({\mathbf{r}})|^2$ for all ${\mathbf{r}}$, the mode shapes will not overlap perfectly during the final coupling pulse.
+Consequently the mode shape of the atomic field changes depending on the nonlinearity-density product, which means that unless $U_{aa}|\psi_a({\mathbf{r}})|^2 + U_{ab}|\psi_b({\mathbf{r}})|^2 = U_{bb}|\psi_b({\mathbf{r}})|^2 + U_{ab}|\psi_a({\mathbf{r}})|^2$ for all ${\mathbf{r}}$, the mode shapes will not overlap perfectly during the final coupling pulse.
 This degraded mode-matching will result in lower efficiency when converting the quadrature squeezing to number squeezing.
@@
 As can be seen by Eqs~(\ref{eqa1}) and (\ref{eqb1}), such a pulse also leaves the quantum statistics of each state unchanged.
-In the limit where each state remains in a single mode (although that mode might change), and exactly half of the atoms are in each state, this technique ensures that both modes have identical mode shapes after the hold time.
-In practice this will never work perfectly as the mode shapes will change, resulting in non-identical environments for the two modes.
+In the limit where each state remains in a single mode, and exactly half of the atoms are in each state, this technique ensures that both modes have identical mode shapes after the hold time.
+In practice this will never work perfectly --- due to the density distributions of $\psi_a(\mathbf{r})$ and $\psi_a(\mathbf{r})$ not reflecting the ground state, resulting in non-identical environments for the two modes, some modal dynamics will occur.
 Nonetheless, in many regimes, such a $\pi$-pulse can produce some improvement.
@@
 
-Fig.~\ref{fig:Combined1Dand2D} shows the non-condensed fraction predicted by the Bogoliubov analysis for one and two dimensions as a function of $r$, which is a scale factor for the size of the trap.
-\textbf{Consider not making it a function of r, but simply the size of the trap?}  The traps are assumed to have a constant potential, which minimises the effects of the multimode issues described in section \ref{sec:MMdescription}.
-The squeezing in each trap is calculated using the TWA method, and we see that the squeezing does indeed degrade at high $r$ in 1D, but remains constant in $2D$.
-For low $r$, the squeezing in 1D is limited by the other multimode effects.
-The non-smooth nature of the squeezing prediction in 2D is due to the stochastic nature of the simulations.
+Figure \ref{fig1D_squeezing_and_bog_scaling} and Figure \ref{fig2D_squeezing_and_bog_scaling} show the non-condensed fraction predicted by the Bogoliubov analysis compared to that predicted by a full numerical simulation, in one, two and three dimensions, as a function of the trap size.
+Also shown in the 1D and 2D cases is the maximum amount of number squeezing obtainable, as predicted by the numerical simulation.
+The full squeezing scaling curve is not shown in 3D, as for the smaller trap sizes the TWA method breaks down, and results are unreliable; for the larger trap sizes with negligible non-condensate fractions, however, the simulation was carried out and a realtive number squeezing of 12.6dB was found.
+The traps are assumed to have a constant potential, which minimises the effects of the multimode issues described in section \ref{sec:MMdescription}.
+We see that the squeezing does indeed degrade at large trap sizes in 1D, but remains constant in $2D$.
+For small trap sizes, the squeezing in 1D aymptotically approaches that of the ideal 0D system.
 
-In Fig.~\ref{fig:3DBogScaling} we see that the number of accessible states for three-dimensional traps actually decreases with volume.
+We note that the non-smooth nature of the squeezing prediction in 2D is due to the stochastic nature of the simulations, and the necessity of averaging over fewer quantum trajectories due to the very high number grid points in this regime.
+The non-smooth nature of the non-condensate mode fraction in 3D, on the other hand, is due to the number of modes being both discrete and low in this regime, and counting only the non-condensed fraction in modes with an energy less than than $\mu$, the chemical potential.
+
+In Fig.~\ref{fig:3DBogScaling} we see that the number of accessible states for three-dimensional traps actually decreases with volume in 3D.
 This is because the energy of the ground state of the BEC is reducing faster than the density of states at low energies is increasing.
@@
   \centering
-   \includegraphics[width=7cm]{figures/1D_squeezing_scaling.eps}
-\caption{1D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
- \label{fig1D_squeezing_scaling}
+  \begin{subfigure}{.48\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/1D_Bog_mode_occupation_scaling.eps}
+  \end{subfigure}
+  \begin{subfigure}{.48\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/1D_squeezing_scaling.eps}
+  \end{subfigure}
+\caption{1D scaling as function of trap size, $L$.
+(Left) The proportion the non-condensed fraction in mode $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+(Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
+Parameters: $N=1\times 10^5$, $U_{aa} = 8.9\times 10^{-40}$Jm, $U_{ab}=U_{bb}=0$, $\tau_{\mathrm{hold}} = 10L$\,s.}
+  \label{fig1D_squeezing_and_bog_scaling}
 \end{figure}
@@
   \centering
-   \includegraphics[width=7cm]{figures/1D_Bog_mode_occupation_scaling.eps}
-\caption{1D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
- \label{fig1D_Bog_mode_occupation_scaling}
+  \begin{subfigure}{.48\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2D_Bog_mode_occupation_scaling.eps}
+  \end{subfigure}
+  \begin{subfigure}{.48\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2D_squeezing_scaling.eps}
+  \end{subfigure}
+\caption{2D scaling as function of trap size, L.
+(Left) The proportion the non-condensed fraction in mode $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+(Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
+Parameters: $N=1\times 10^5$, $U_{aa} = 8.9\times 10^{-44}$Jm, $U_{ab}=U_{bb}=0$, $\tau_{\mathrm{hold}} = 10^{5} L^2$\,s.}
+  \label{fig2D_squeezing_and_bog_scaling}
 \end{figure}
 
-\begin{figure}
-  \centering
-   \includegraphics[width=7cm]{figures/2D_squeezing_scaling.eps}
-\caption{2D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
- \label{fig2D_squeezing_scaling}
-\end{figure}
+%\begin{figure}
+%  \centering
+%   \includegraphics[width=7cm]{figures/1D_squeezing_scaling.eps}
+%\caption{1D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
+% \label{fig1D_squeezing_scaling}
+%\end{figure}
 
-\begin{figure}
-  \centering
-   \includegraphics[width=7cm]{figures/2D_Bog_mode_occupation_scaling.eps}
-\caption{2D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
- \label{fig2D_Bog_mode_occupation_scaling}
-\end{figure}
+%\begin{figure}
+%  \centering
+%   \includegraphics[width=7cm]{figures/1D_Bog_mode_occupation_scaling.eps}
+%\caption{1D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
+% \label{fig1D_Bog_mode_occupation_scaling}
+%\end{figure}
+
+%\begin{figure}
+%  \centering
+%   \includegraphics[width=7cm]{figures/2D_squeezing_scaling.eps}
+%\caption{2D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
+% \label{fig2D_squeezing_scaling}
+%\end{figure}
+
+%\begin{figure}
+%  \centering
+%   \includegraphics[width=7cm]{figures/2D_Bog_mode_occupation_scaling.eps}
+%\caption{2D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
+% \label{fig2D_Bog_mode_occupation_scaling}
+%\end{figure}
 
@@
    \includegraphics[width=7cm]{figures/3D_Bog_mode_occupation_scaling.eps}
-\caption{3D experimentalist's scaling. $N=1\times 10^4$, $L=20\mu$m.}
+\caption{3D scaling as function of trap size, $L$.
+(Left) The proportion the non-condensed fraction in mode $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+(Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
+Parameters: $N=1\times 10^4$, $U_{aa} = 4.7\times 10^{-49}$Jm, $U_{ab}=U_{bb}=0$, $\tau_{\mathrm{hold}} = 1.25 \times 10^{11} L^3$\,s.}
  \label{fig3D_Bog_mode_occupation_scaling}

### 2013-06-04 large_number_squeezing_with_atoms.tex
Added references, plus more tweaks and rewording

@@
 \begin{equation}
-\chi_{aa} = U_{aa} \int |u_i|^4 \, dV.
+\chi_{aa} = U_{aa} \int |u_{i}({\mathbf{r}})|^4 \, dV.
 \label{eqChiUequivalence}
@@
        && + 4 c^2 s^2 (2 n_a n_b +D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} A_2^* B_2 - e^{-i(\lambda_{aa} - \lambda_{bb})} A_2 B_2^* ) \nonumber \\
-       && + 4 i c s (c^2 - s^2) (n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B 
-                     + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^*) \label{eqNumDiffVariance}
+       && + 4 i c s (c^2 - s^2) \left( n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B \right. \nonumber \\
+       && \,\,\,\,\, \left. + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^* \right) \label{eqNumDiffVariance}
 \end{eqnarray}
@@
 Solid line, $\phi=0.10$; dash-dotted line, $\phi=1.42$; dashed line, $\phi=3.24$; dotted line, $\phi=4.71$.
-These particular parameters yield 21dB of number squeezing for the $\phi=0.1$ and $\phi=1.42$ cases.}
+These particular parameters yield 21\,dB of number squeezing for the $\phi=0.1$ and $\phi=1.42$ cases.}
     \label{figTwoModeAnalyticExamples}
@@
 We use stochastic methods based on phase space representations to simulate the dynamics of the multimode quantum fields.
-These methods alleviate the problem of directly simulating elements of the Hilbert space, the size of which increases exponentially with the number of spatial modes \cite{stochasticRefs}.
+These methods alleviate the problem of directly simulating elements of the Hilbert space, the size of which increases exponentially with the number of spatial modes \cite{gardiner1991, steelET1998}.
 Stochastic methods achieve this by finding a sufficiently well-behaved quasi-probability representation for the density matrix, which can then be simulated as the average behaviour of a number of low-dimensional samples.
@@
 
-Under the TWA, the equation of motion for the functional Wigner distribution is a Fokker-Planck equation, and can therefore be sampled by a set of stochastic equations, as described in \cite{GardinerQuantumNoise}.
+Under the TWA, the equation of motion for the functional Wigner distribution is a Fokker-Planck equation, and can therefore be sampled by a set of stochastic equations, as described in \cite{gardiner1991}.
 These stochastic fields $\phi(\mathbf{x})$ are related to field operator expectation values by
@@
 \begin{equation}
-\chi = U \frac{{\mathrm{Erf}} (\sqrt{2p})^3}{{\mathrm{Erf}}(p)^6} \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}}
+\chi = U \frac{{\mathrm{Erf}} (\sqrt{2p})^3}{{\mathrm{Erf}}(p)^6} \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}}.
 \end{equation}
-Spatial filtering can easily produce higher squeezing, but there is the obvious disadvantage that it  significantly reduces the effective number of atoms.
+Such spatial filtering can easily produce higher squeezing, but there is the obvious disadvantage that it  significantly reduces the effective number of atoms.
 This is a serious issue for most applications of squeezed sources, such as interferometry, where the signal to noise scales with the square root of the flux.
@@
 To minimise the damaging effects of the non-zero momentum modes, it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation.
+
 The number of excited modes is almost entirely determined by the density of modes energetically accessible by the system.
@@
 Also shown in the 1D and 2D cases is the maximum amount of number squeezing obtainable, as predicted by the numerical simulation.
-The full squeezing scaling curve is not shown in 3D, as for the smaller trap sizes the TWA method breaks down, and results are unreliable; for the larger trap sizes with negligible non-condensate fractions, however, the simulation was carried out and a realtive number squeezing of 12.6dB was found.
+The full squeezing scaling curve is not shown in 3D, as for the smaller trap sizes the TWA method breaks down, and results are unreliable; for the larger trap sizes with negligible non-condensate fractions, however, the simulation was carried out and a relative number squeezing of 12.6dB was found.
 The traps are assumed to have a constant potential, which minimises the effects of the multimode issues described in section \ref{sec:MMdescription}.
@@
 
+\bibitem{haineET2009} S. A. Haine and M. T. Johnsson 2009 Dynamic scheme for generating number squeezing in Bose-Einstein condensates
+through nonlinear interactions \emph{Phys. Rev. A} {\bf 80} 023611.
+
 \bibitem{dennisET2010} Dennis G R and Johnsson M T 2010 Generation of directional coherent matter beams through dynamical instabilities in Bose-Einstein condensates {\it Phys. Rev. A} {\bf 82} 033615
+
+\bibitem{gardiner1991} C. W. Gardiner, \emph{Quantum Noise}, (Springer-Verlag, Berlin 1991).
+
+\bibitem{steelET1998} M.
+J.
+Steel, M.
+K.
+Olsen, L.
+I.
+Plimak, P.
+D.
+Drummond, S.
+M.
+Tan, M.
+J.
+Collett, D.
+F.
+Walls, and R.
+Graham 2010 Dynamical quantum noise in trapped Bose-Einstein condensates {\it Phys.
+Rev.
+A} {\bf 58}, 4824.
 
@@
 differential equations {\it Comput. Phys. Commun.} {\bf 184} 201--208
-\bibitem{louisell} Louisell Quantum Statistical Properties of Radiation
+
+\bibitem{louisell} W. H. Louisell \emph{Quantum Statistical Properties of Radiation}, (Wiley 1990).
 

### 2013-06-05 large_number_squeezing_with_atoms.tex
Redefine \chi so that it has units of per second, fix Hamiltonians so that \Omega terms are multiplied by \hbar.

@@
 \end{equation}
-When the system remains predominantly in one mode, we can approximate this sum as the single term $\hat{H} = \chi_{aa} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}/2$ where
+When the system remains predominantly in one mode, we can approximate this sum as the single term $\hat{H} = \hbar\chi_{aa} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}/2$ where
 \begin{equation}
-\chi_{aa} = U_{aa} \int |u_{i}({\mathbf{r}})|^4 \, dV.
+\chi_{aa} = \frac{1}{\hbar} U_{aa} \int |u_{i}({\mathbf{r}})|^4 \, dV.
 \label{eqChiUequivalence}
@@
 \begin{equation}
-\hat{H} = \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
-          + \chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
-          + \frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}
-          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger}  \hat{a} )
+\hat{H} = \hbar\frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
+          + \hbar\chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
+          + \hbar\frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}
+          + \hbar\Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger}  \hat{a} )
 \label{eqTwoModeHamiltonian}
@@
 c &=& \cos(\Omega (t_3 - t_2)) \\
-\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} / \hbar \\
+\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} \\
 A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
@@
     \caption{Normalized number variance Var$[N_a]/N_a$ in mode $|a \rangle$ as a function of final recombination time plotted on a logarithmic scale, using four different mixing angles $\phi$.
-Parameters: $n_a = n_b =5 \times 10^5$, $\chi_{aa}=0.04 \hbar, \chi_{ab}=0, \chi_{bb}=0.01\hbar, \tau_{\mathrm{hold}}=4\times 10^{-4}$s.
+Parameters: $n_a = n_b =5 \times 10^5$, $\chi_{aa}=0.04\, \text{s}^{-1}, \chi_{ab}=0, \chi_{bb}=0.01\,\text{s}^{-1}, \tau_{\mathrm{hold}}=4\times 10^{-4}$s.
 Solid line, $\phi=0.10$; dash-dotted line, $\phi=1.42$; dashed line, $\phi=3.24$; dotted line, $\phi=4.71$.
@@
 \begin{eqnarray}
-\chi &=& U \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}} \,\,\,\, ({\textrm{Gaussian}}) \\
+\hbar \chi &=& U \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}} \,\,\,\, ({\textrm{Gaussian}}) \\
 %
-\chi &=& \frac{4}{7} \left( \frac{15 U \omega_x \omega_y \omega_z}{16 \pi \sqrt{2}} \right)^{2/5} \left( \frac{m}{N} \right)^{3/5} \,\,\,\, ({\textrm{Thomas-Fermi}}) \\
+\hbar \chi &=& \frac{4}{7} \left( \frac{15 U \omega_x \omega_y \omega_z}{16 \pi \sqrt{2}} \right)^{2/5} \left( \frac{m}{N} \right)^{3/5} \,\,\,\, ({\textrm{Thomas-Fermi}}) \\
 %
-\chi &=& \frac{U}{V} \,\,\,\, {\textrm{(constant density)}}
+\hbar \chi &=& \frac{U}{V} \,\,\,\, {\textrm{(constant density)}}
 \end{eqnarray}
@@
 The nonlinearities of the various 1D simulations have been adjusted for their mode shape so that they are equivalent to that of the 0D, two-mode mode nonlinearity.
-Parameters used were $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \hbar, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$.
+Parameters used were $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \text{s}^{-1}, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$.
 Maximum squeezing is achieved when the atomic density profile is the most uniform.}
@@
 \begin{equation}
-\chi = U \frac{{\mathrm{Erf}} (\sqrt{2p})^3}{{\mathrm{Erf}}(p)^6} \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}}.
+\hbar\chi = U \frac{{\mathrm{Erf}} (\sqrt{2p})^3}{{\mathrm{Erf}}(p)^6} \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}}.
 \end{equation}
@@
     \caption{Effect of a $\pi$-pulse halfway through the hold time.
-Common parameters used were $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{ab}=\chi_{bb}=0$.
+Common parameters used were $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\,\text{s}^{-1}$, $\chi_{ab}=\chi_{bb}=0$.
 Solid red line: Analytic two-mode solution, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (10.9dB squeezing); Solid blue line: no $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (4.9dB squeezing);  Dashed blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.05$ (5.8dB squeezing); Dash-dotted blue line: with $\pi$-pulse and the longer hold time $\tau_{\mathrm{hold}}=5\times 10^{-4}$s allowed by the application of the pulse, $\phi=3.87$ (7.2dB squeezing).}
@@
 Thick red line, analytic two-mode solution; solid blue line, 1D; dashed line, 2D; dash-dotted line, 3D.
-(a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=6.2$.
-(b) $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{bb}=\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.}
+(a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\,\text{s}^{-1}$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=6.2$.
+(b) $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\,\text{s}^{-1}$, $\chi_{bb}=\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.}
   \label{figDimensionalSqueezingEffects}
@@
 To simplify the analysis we consider the case in which $U_{aa} \neq 0$, $U_{bb}=U_{ab} = 0$ and the condensate has a spatially-constant density in a box of side lengths $L_x$, $L_y$ and $L_z$.
-In this limit the squeezing $S_a = \Var(N_a)/N_a$ of the $|a\rangle$ mode at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation}):
+In this limit the occupation of the non-zero momentum modes in the $\ket{a}$ internal state at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation}):
 \begin{align}
-  S_a(t_3) &= \frac{\Var[N_a](t_3) + \sum_{\mathbf{k}\neq 0} \frac{1}{2} n_a(\mathbf{k}, t_3) \left\{3 + \cos\left[\Omega(t_3-t_2)\right] + 4 n_a(\mathbf{k}, t_3)\right\}}{N_a(t_3) + \sum_{\mathbf{k}\neq 0} n_a(\mathbf{k}, t_3)}, \label{eqBogoliubovSqueezing}\\
-  n_a(\mathbf{k}, t_3) &= \left\{n_a \chi_{aa} \tau_\text{hold} \cos\left[\Omega\left(t_3-t_2\right)\right] \sinc\left(\omega_\mathbf{k} \tau_\text{hold}\right) \right\}^2,\\
-  \omega_\mathbf{k} &= \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}, \\
-  \omega^0_\mathbf{k} &= \frac{\hbar \mathbf{k}^2}{2 M},
+  n_a(\mathbf{k}, t_3) &= \left\{\frac{1}{\hbar} n_a \chi_{aa} \tau_hold \cos\left[\Omega\left(t_3-t_2\right)\right] \sinc\left(\omega_\mathbf{k} \tau_\text{hold}\right) \right\}^2,
 \end{align}
+where the Bogoliubov mode frequency $\omega_\mathbf{k} = \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}$, and the free particle frequency is 
+$  \omega^0_\mathbf{k} = \frac{\hbar \mathbf{k}^2}{2 M}, $
 where $n_a(\mathbf{k}, t) = \expect{\hat{a}^\dagger(\mathbf{k}, t) \hat{a}(\mathbf{k}, t)}$ is the occupation of momentum mode $\mathbf{k}$, $\Var[N_a](t)$ and $N_a(t)$ are the number variance and mean number of the $|a\rangle$ mode from the two-mode model in Section~\ref{secTwoModeAnalytic}, the sum over $\mathbf{k}$ is a discrete sum over the available non-zero momentum modes $\mathbf{k} = (2\pi/L_x) n_x \hat{\mathbf{x}} + (2\pi/L_y) n_y \hat{\mathbf{y}} + (2\pi/L_z) n_z \hat{\mathbf{z}}$ with $n_x$, $n_y$ and $n_z$ arbitrary integers, $\omega_\mathbf{k}$ is the frequency of the Bogoliubov mode $\mathbf{k}$ and $\omega^0_\mathbf{k}$ is the frequency of a free particle.
@@
 \hat{H_1} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b} 
-          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger} \hat{a} )
+          + \hbar \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger} \hat{a} )
 \end{equation}
@@
 \begin{equation}
-\hat{H_2} = \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
-          + \chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
-          + \frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}.
+\hat{H_2} = \hbar\frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
+          + \hbar\chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
+          + \hbar\frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}.
 \end{equation}
@@
 \hat{H_3} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b}
-          + \Omega (e^{i\phi} \hat{a}^{\dagger} \hat{b} + e^{-i\phi} \hat{b}^{\dagger} \hat{a} ).
+          + \hbar\Omega (e^{i\phi} \hat{a}^{\dagger} \hat{b} + e^{-i\phi} \hat{b}^{\dagger} \hat{a} ).
 \end{equation}
@@
 \end{equation}
-where we have defined $\lambda_{ij} = \chi_{ij} (t_2-t_1)/\hbar$. Utilizing (\ref{eqHadamard}) to move $\hat{a}_1$ through the exponential, with some algebra one can show that
+where we have defined $\lambda_{ij} = \chi_{ij} (t_2-t_1)$. Utilizing (\ref{eqHadamard}) to move $\hat{a}_1$ through the exponential, with some algebra one can show that
 \begin{eqnarray}

### 2013-06-05 large_number_squeezing_with_atoms.tex
Added text beginning on the Bog scaling argument.

@@
 \begin{eqnarray}
-s &=& \sin(\Omega (t_3 - t_2)) \label{eqsDef} \\
-c &=& \cos(\Omega (t_3 - t_2)) \\
+s &=& \sin(\theta) \label{eqsDef} \\
+c &=& \cos(\theta) \\
+\theta &=& \Omega(t_3 - t_2) \\
 \lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} \\
@@
 To simplify the analysis we consider the case in which $U_{aa} \neq 0$, $U_{bb}=U_{ab} = 0$ and the condensate has a spatially-constant density in a box of side lengths $L_x$, $L_y$ and $L_z$.
-In this limit the occupation of the non-zero momentum modes in the $\ket{a}$ internal state at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation}):
+In this limit the occupation of the non-zero momentum modes in the $\ket{a}$ internal state at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation})
 \begin{align}
-  n_a(\mathbf{k}, t_3) &= \left\{\frac{1}{\hbar} n_a \chi_{aa} \tau_hold \cos\left[\Omega\left(t_3-t_2\right)\right] \sinc\left(\omega_\mathbf{k} \tau_\text{hold}\right) \right\}^2,
+  n_a(\mathbf{k}, t_3) &= \expect{\hat{a}^\dagger(\mathbf{k},t) \hat{a}(\mathbf{k},t)} = \left[n_a \chi_{aa} \tau_\text{hold} \cos(\theta) \sinc\left(\omega_\mathbf{k} \tau_\text{hold}\right) \right]^2, \label{eqBogOccupation}
 \end{align}
-where the Bogoliubov mode frequency $\omega_\mathbf{k} = \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}$, and the free particle frequency is 
-$  \omega^0_\mathbf{k} = \frac{\hbar \mathbf{k}^2}{2 M}, $
-where $n_a(\mathbf{k}, t) = \expect{\hat{a}^\dagger(\mathbf{k}, t) \hat{a}(\mathbf{k}, t)}$ is the occupation of momentum mode $\mathbf{k}$, $\Var[N_a](t)$ and $N_a(t)$ are the number variance and mean number of the $|a\rangle$ mode from the two-mode model in Section~\ref{secTwoModeAnalytic}, the sum over $\mathbf{k}$ is a discrete sum over the available non-zero momentum modes $\mathbf{k} = (2\pi/L_x) n_x \hat{\mathbf{x}} + (2\pi/L_y) n_y \hat{\mathbf{y}} + (2\pi/L_z) n_z \hat{\mathbf{z}}$ with $n_x$, $n_y$ and $n_z$ arbitrary integers, $\omega_\mathbf{k}$ is the frequency of the Bogoliubov mode $\mathbf{k}$ and $\omega^0_\mathbf{k}$ is the frequency of a free particle.
+where the Bogoliubov mode frequency is $\omega_\mathbf{k} = \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}$, and the free particle frequency is $\omega^0_\mathbf{k} = \hbar \mathbf{k}^2/2 M$.
+The Bogoliubov modes can be expected to have an impact on the squeezing of the system when their occupation is a non-negligible fraction of the total number of atoms.
+The proportional occupation of the Bogoliubov modes (i.e.\ the condensate depletion) is
+\begin{align}
+  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &= \frac{1}{N}\sum_{\mathbf{k}\neq 0} n_a(\mathbf{k}, t), \label{eqCondensateDepletion}
+\end{align}
+where the sum over $\mathbf{k} \neq 0$ is taken over the available non-zero momentum modes $\mathbf{k} = (2\pi/L_x) n_x \hat{\mathbf{x}} + (2\pi/L_y) n_y \hat{\mathbf{y}} + (2\pi/L_z) n_z \hat{\mathbf{z}}$ with $n_x$, $n_y$ and $n_z$ arbitrary integers (not all zero).
+
+[Insert motivation that the impact of Bogoliubov modes is detrimental here.
+I could wave my hands and say that more modes is bad, as seen in the previous parts of the article, or I could refer to the squeezing of the Bogoliubov modes:
+\begin{align}
+  S_{a,\mathbf{k}\neq 0}(t_3) &= \frac{\frac{1}{2}\sum_{\mathbf{k}\neq0}n_a(\mathbf{k},t_3)\left[3 + \cos\theta + n_a(\mathbf{k}, t_3)\right]}{\sum_{\mathbf{k}\neq0} n_a(\mathbf{k}, t_3)},
+\end{align}
+which is always greater than 1, but this isn't convincing either as the total squeezing includes the number variance of the $\mathbf{k=0}$ mode as well as the correlations between the $\mathbf{k}=0$ mode and the $\mathbf{k}\neq 0$ modes.
+If you do it naively, you find that when the Bog modes have a significant impact on the squeezing, they have also significantly changed it at $t=t_2$, which must be the same as $t=t_1$.
+The problem is that the Bogoliubov theory used isn't number-conserving.
+The best argument I have at the moment for Bog theory being a good indicator for when things go bad is that a significant Bog fraction implies a depleted condensate, and that will almost certainly make squeezing bad as it will invalidate the single-mode theory.
+More freedom == bad?
+In this case, the best evidence for the utility of Bog theory would seem to be Figure~\ref{fig1D_squeezing_and_bog_scaling}.]
+
+Good squeezing requires that the Bogoliubov fraction or condensate depletion be small.
+To better understand the scaling of $N_{\mathbf{k}\neq0}/N$, we will approximate the sum in \eqref{eqCondensateDepletion} with an integral.
+In 1D \eqref{eqCondensateDepletion} becomes
+\begin{align}
+  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &= \frac{(n_a \lambda_{aa} \cos\theta)^2}{N} \sum_{k \neq 0} \sinc^2\left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right).
+\end{align}
+In the limit that $\omega_{k_\text{min}} \tau_\text{hold} \ll 1$, where $k_\text{min} = 2\pi/L$ this discrete sum may be approximated by the integral
+\begin{align}
+  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \lambda_{aa} \cos\theta)^2}{N} \int_{-\infty}^{+\infty} \sinc^2 \left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right) \frac{d k}{\Delta k},
+\end{align}
+where $\Delta k = 2\pi/L$.  Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$, the integral can be written as
+\begin{align}
+  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \lambda_{aa} \cos\theta)^2}{N} \frac{L}{\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \lambda_{aa} n_a)}\, d\Phi
+\end{align}
+Define $\Lambda_{a} = \lambda_{aa} n_a$ to simplify the expression and we get
+\begin{align}
+  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 L \cos^2\theta } {\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} f_\text{1D}(\Lambda_a),
+\end{align}
+where
+\begin{align}
+  f_\text{1D}(\Lambda) &= \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
+\end{align}
+
+In 2D and 3D you get
+\begin{align}
+  \frac{N^{\text{2D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 A \cos^2\theta } {4\frac{\hbar}{M}\tau_\text{hold}} f_\text{2D}(\Lambda_a), \\
+  \frac{N^{\text{3D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 V \cos^2\theta } {\sqrt{2}\pi^2\left(\frac{\hbar}{M}\tau_\text{hold}\right)^{3/2}} f_\text{3D}(\Lambda_a),
+\end{align}
+where
+\begin{align}
+  f_\text{2D}(\Lambda) &= \int_0^{\infty} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi\\
+  f_\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
+\end{align}
 

### 2013-06-11 large_number_squeezing_with_atoms.tex
Minor changes to bog section.
Move the labels to the correct place in an extended eqnarray block

@@
 %
-{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \\
+{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \nonumber\\
        && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2) \nonumber \\
@@
        && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
-       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ) \nonumber\\
+       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ) \\
 %
@@
 \begin{align}
-  S_{a,\mathbf{k}\neq 0}(t_3) &= \frac{\frac{1}{2}\sum_{\mathbf{k}\neq0}n_a(\mathbf{k},t_3)\left[3 + \cos\theta + n_a(\mathbf{k}, t_3)\right]}{\sum_{\mathbf{k}\neq0} n_a(\mathbf{k}, t_3)},
+  S_{a,\mathbf{k}\neq 0}(t_3) &= \frac{\frac{1}{2}\sum_{\mathbf{k}\neq0}n_a(\mathbf{k},t_3)\left[3 + \cos 2\theta + n_a(\mathbf{k}, t_3)\right]}{\sum_{\mathbf{k}\neq0} n_a(\mathbf{k}, t_3)},
 \end{align}
@@
 \begin{align}
-  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &= \frac{(n_a \lambda_{aa} \cos\theta)^2}{N} \sum_{k \neq 0} \sinc^2\left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right).
+  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &= \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \sum_{k \neq 0} \sinc^2\left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right).
 \end{align}
@@
 \begin{align}
-  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \lambda_{aa} \cos\theta)^2}{N} \int_{-\infty}^{+\infty} \sinc^2 \left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right) \frac{d k}{\Delta k},
+  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \int_{-\infty}^{+\infty} \sinc^2 \left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right) \frac{d k}{\Delta k},
 \end{align}
@@
 \begin{align}
-  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \lambda_{aa} \cos\theta)^2}{N} \frac{L}{\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \lambda_{aa} n_a)}\, d\Phi
+  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \frac{L}{\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \lambda_{aa} n_a)}\, d\Phi
 \end{align}
-Define $\Lambda_{a} = \lambda_{aa} n_a$ to simplify the expression and we get
+Define $\Lambda_{a} = n_a \chi_{aa} \tau_\text{hold}$ to simplify the expression and we get
 \begin{align}

### 2013-06-11 large_number_squeezing_with_atoms.tex
Complete the description of the scaling laws.

@@
 \begin{align}
-  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 L \cos^2\theta } {\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} f_\text{1D}(\Lambda_a),
+  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 L \cos^2\theta } {\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} f^\text{1D}(\Lambda_a),
 \end{align}
@@
 \begin{align}
-  f_\text{1D}(\Lambda) &= \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
+  f^\text{1D}(\Lambda) &= \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
 \end{align}
@@
 \begin{align}
-  \frac{N^{\text{2D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 A \cos^2\theta } {4\frac{\hbar}{M}\tau_\text{hold}} f_\text{2D}(\Lambda_a), \\
-  \frac{N^{\text{3D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 V \cos^2\theta } {\sqrt{2}\pi^2\left(\frac{\hbar}{M}\tau_\text{hold}\right)^{3/2}} f_\text{3D}(\Lambda_a),
+  \frac{N^{\text{2D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 A \cos^2\theta } {4\frac{\hbar}{M}\tau_\text{hold}} f^\text{2D}(\Lambda_a), \\
+  \frac{N^{\text{3D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 V \cos^2\theta } {\sqrt{2}\pi^2\left(\frac{\hbar}{M}\tau_\text{hold}\right)^{3/2}} f^\text{3D}(\Lambda_a),
 \end{align}
@@
 \begin{align}
-  f_\text{2D}(\Lambda) &= \int_0^{\infty} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi\\
-  f_\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
+  f^\text{2D}(\Lambda) &= \int_0^{\infty} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi\\
+  f^\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
 \end{align}
+
+The squeezing predicted by the two-mode model will only be valid in higher dimensions when $N_{\mathbf{k}\neq0}/N \ll 1$, when this condition is violated, the squeezing will be detrimentally affected.
+We therefore want to examine the scaling of $N_{\mathbf{k}\neq 0}/N$ with system size in 1D, 2D and 3D.
+
+For constant total number $N$ and occupations of the $\ket{a}$ and $\ket{b}$ states at time $t_1$, the optimum squeezing as determined by the two-mode model will have a particular value of $\tau_\text{hold}$, $\theta$ and $\phi$.
+As the dimensions of the system are scaled, $\hbar \chi_{ij} = U^\text{1D}_{ij}/L = U^\text{2D}_{ij}/A = U^\text{3D}_{ij}/V$ will change, but to keep the same squeezing (and the same equivalent two-mode model) $\lambda_{ij} = \chi_{ij} \tau_\text{hold}$ must be kept constant (see Eq.~\eqref{eqsDef}--\eqref{eqNumDiffVariance}).
+Therefore $\Lambda_a = n_a\chi_{aa}\tau_\text{hold}$ must also be kept constant.
+With this in mind, we can write $\tau_\text{hold}$ in terms of the system size and the nonlinearity.
+We obtain:
+\begin{align}
+  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx \sqrt{L} \cdot \frac{1}{N} \sqrt{\frac{M U^\text{1D}_{aa} n_a}{2 \hbar^2 \pi^2}} \Lambda_a^{3/2} \cos^2\theta f^\text{1D}(\Lambda_a) &&\propto \sqrt{L} \\
+  \frac{N^\text{2D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx 1 \cdot \frac{1}{N} \frac{M U^\text{2D}_{aa} n_a }{4 \hbar^2} \Lambda_a \cos^2\theta f^\text{2D}(\Lambda_a) &&\propto \text{constant} \\
+  \frac{N^\text{3D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx \frac{1}{\sqrt{V}} \cdot \frac{1}{N} \left(\frac{M U^\text{3D}_{aa} n_a}{\hbar^2}\right)^{3/2} \frac{\Lambda_a^{1/2} \cos^2\theta f^\text{3D}(\Lambda_a)}{\sqrt{2} \pi^2} &&\propto \frac{1}{\sqrt{V}}
+\end{align}
+
+In 1D the condensate depletion $N_{\mathbf{k}\neq 0}/N$ increases with increasing system size, in 2D the condensate depletion remains constant, and in 3D the condensate depletion decreases as the system size increases.
+This suggests that the best system to use will be a box potential, and the system size can be increased until the condensate depletion becomes negligibly small.
 

### 2013-06-11 large_number_squeezing_with_atoms.tex
Going back and cleaning up the Bog section.

@@
 The Bogoliubov modes can be expected to have an impact on the squeezing of the system when their occupation is a non-negligible fraction of the total number of atoms.
-The proportional occupation of the Bogoliubov modes (i.e.\ the condensate depletion) is
+The fractional occupation of the Bogoliubov modes is
 \begin{align}
@@
 \end{align}
-where the sum over $\mathbf{k} \neq 0$ is taken over the available non-zero momentum modes $\mathbf{k} = (2\pi/L_x) n_x \hat{\mathbf{x}} + (2\pi/L_y) n_y \hat{\mathbf{y}} + (2\pi/L_z) n_z \hat{\mathbf{z}}$ with $n_x$, $n_y$ and $n_z$ arbitrary integers (not all zero).
+where the sum over $\mathbf{k} \neq 0$ is taken over the available non-zero momentum modes $\mathbf{k} = (2\pi/L_x) n_x \hat{\mathbf{x}} + (2\pi/L_y) n_y \hat{\mathbf{y}} + (2\pi/L_z) n_z \hat{\mathbf{z}}$ with $n_x$, $n_y$ and $n_z$ arbitrary integers that are not all zero.
 
-[Insert motivation that the impact of Bogoliubov modes is detrimental here.
-I could wave my hands and say that more modes is bad, as seen in the previous parts of the article, or I could refer to the squeezing of the Bogoliubov modes:
-\begin{align}
-  S_{a,\mathbf{k}\neq 0}(t_3) &= \frac{\frac{1}{2}\sum_{\mathbf{k}\neq0}n_a(\mathbf{k},t_3)\left[3 + \cos 2\theta + n_a(\mathbf{k}, t_3)\right]}{\sum_{\mathbf{k}\neq0} n_a(\mathbf{k}, t_3)},
-\end{align}
-which is always greater than 1, but this isn't convincing either as the total squeezing includes the number variance of the $\mathbf{k=0}$ mode as well as the correlations between the $\mathbf{k}=0$ mode and the $\mathbf{k}\neq 0$ modes.
-If you do it naively, you find that when the Bog modes have a significant impact on the squeezing, they have also significantly changed it at $t=t_2$, which must be the same as $t=t_1$.
-The problem is that the Bogoliubov theory used isn't number-conserving.
-The best argument I have at the moment for Bog theory being a good indicator for when things go bad is that a significant Bog fraction implies a depleted condensate, and that will almost certainly make squeezing bad as it will invalidate the single-mode theory.
-More freedom == bad?
-In this case, the best evidence for the utility of Bog theory would seem to be Figure~\ref{fig1D_squeezing_and_bog_scaling}.]
+Figure~\ref{figBogSqueezingValidation} demonstrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the fractional occupation of the Bogoliubov modes for parameters where the occupation of these modes approach 20\%\footnote{FIXME} of the total occupation of the system.
+The disagreement for larger fractional occupations occurs because the Bogoliubov theory used in deriving \eqref{eqCondensateDepletion} is not number-conserving as it neglects the effect of the Bogoliubov modes on the condensate.
 
-Good squeezing requires that the Bogoliubov fraction or condensate depletion be small.
-To better understand the scaling of $N_{\mathbf{k}\neq0}/N$, we will approximate the sum in \eqref{eqCondensateDepletion} with an integral.
-In 1D \eqref{eqCondensateDepletion} becomes
+To minimise the damaging effects of the non-zero momentum modes it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation.
+To better understand when $N_{\mathbf{k}\neq 0}/N$ can be expected to be small, we approximate the sum in \eqref{eqCondensateDepletion} with an integral.
+In 1D Eq.~\eqref{eqCondensateDepletion} is
 \begin{align}
@@
 \end{align}
-In the limit that $\omega_{k_\text{min}} \tau_\text{hold} \ll 1$, where $k_\text{min} = 2\pi/L$ this discrete sum may be approximated by the integral
+We wish to approximate this sum by the integral
 \begin{align}
-  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \int_{-\infty}^{+\infty} \sinc^2 \left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right) \frac{d k}{\Delta k},
+  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \int_{-\infty}^{+\infty} \sinc^2 \left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right) \frac{d k}{\Delta k}, \label{eqCondensateDepletionFirstIntegralApproximation}
 \end{align}
-where $\Delta k = 2\pi/L$.  Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$, the integral can be written as
-\begin{align}
-  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \frac{L}{\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \lambda_{aa} n_a)}\, d\Phi
-\end{align}
-Define $\Lambda_{a} = n_a \chi_{aa} \tau_\text{hold}$ to simplify the expression and we get
+where $\Delta k = 2\pi/L$.
+This approximation will be valid in the limit that $\omega_{k_\text{min}} \tau_\text{hold} \ll 1$, where $k_\text{min} = 2\pi/L$ as for small $k$ the allowed values $k = n_x 2\pi /L$ will closely sample the region $\xi = \sqrt{\omega_k^0(\omega_k^0 + 2 \chi_{aa} n_a)} \tau_\text{hold} < \pi$ where $\sinc^2(\xi)$ is significant.
+
+Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$ and defining $\Lambda_a = n_a \chi_{aa} \tau_\text{hold}$ the expression \eqref{eqCondensateDepletionFirstIntegralApproximation} can be written as
 \begin{align}
@@
 \begin{align}
-  f^\text{1D}(\Lambda) &= \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
+  f^\text{1D}(\Lambda) &= \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi.
 \end{align}
-
-In 2D and 3D you get
+In 2D and 3D the equivalent expressions are
 \begin{align}
@@
 \begin{align}
-  f^\text{2D}(\Lambda) &= \int_0^{\infty} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi\\
-  f^\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
+  f^\text{2D}(\Lambda) &= \int_0^{\infty} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi,\\
+  f^\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi.
 \end{align}
@@
 
+Qualitatively, the occupation of the non-zero momentum modes is almost entirely determined by the number of Bogoliubov modes energetically accessible to the system.  While this gets larger
+
+The Bogoliubov theory used in deriving \eqref{eqCondensateDepletion} is not number-conserving
+
+To minimise the damaging effects of the non-zero momentum modes, it is better to
+
+The Bogoliubov theory used in deriving \eqref{eqBogOccupation} is not number-conserving because the effect of the Bogoliubov modes on the condensate is neglected.
+
+To minimise the
+
+Good squeezing will only be observed when the fractional occupation of the Bogoliubov modes $N_{\mathbf{k}\neq 0}/N$ is small.
+
+% I'm commenting this out and will essentially gloss over the matter.  We will illustrate that Bog modes are detrimental, rather than arguing the fact.
+% [Insert motivation that the impact of Bogoliubov modes is detrimental here.
+I could wave my hands and say that more modes is bad, as seen in the previous parts of the article, or I could refer to the squeezing of the Bogoliubov modes:
+% \begin{align}
+%   S_{a,\mathbf{k}\neq 0}(t_3) &= \frac{\frac{1}{2}\sum_{\mathbf{k}\neq0}n_a(\mathbf{k},t_3)\left[3 + \cos 2\theta + n_a(\mathbf{k}, t_3)\right]}{\sum_{\mathbf{k}\neq0} n_a(\mathbf{k}, t_3)},
+% \end{align}
+% which is always greater than 1, but this isn't convincing either as the total squeezing includes the number variance of the $\mathbf{k=0}$ mode as well as the correlations between the $\mathbf{k}=0$ mode and the $\mathbf{k}\neq 0$ modes.
+If you do it naively, you find that when the Bog modes have a significant impact on the squeezing, they have also significantly changed it at $t=t_2$, which must be the same as $t=t_1$.
+The problem is that the Bogoliubov theory used isn't number-conserving.
+The best argument I have at the moment for Bog theory being a good indicator for when things go bad is that a significant Bog fraction implies a depleted condensate, and that will almost certainly make squeezing bad as it will invalidate the single-mode theory.
+More freedom == bad?
+In this case, the best evidence for the utility of Bog theory would seem to be Figure~\ref{fig1D_squeezing_and_bog_scaling}.]
+
 The squeezing of mode $|a\rangle$ in \eqref{eqBogoliubovSqueezing} includes the first-order multimode quantum mechanical fluctuations about the mean field and differs from the single-mode squeezing of $\Var[N_a](t_3)/N_a(t_3)$ only by terms involving the occupation of the non-zero momentum modes $n_a(\mathbf{k}, t)$.
@@
     \centering
-  \begin{subfigure}{.45\textwidth}
-    \centering
-    Plot of the number occupation of the condensate and non-zero momentum modes as computed by the Bog theory and a simulation.
+    \framebox{
+    \parbox{0.8\textwidth}{
+    Plot of the number occupation of the condensate and non-zero momentum modes as computed by the Bog theory (direct summation and integral approximation) and a simulation.
 The simulation agrees with the Bog theory for the occupation of the non-zero momentum modes, but there is significant ~10-20\% depletion of the condensate.
-The horizontal axis is the hold time.
-  \end{subfigure}
-  \begin{subfigure}{.45\textwidth}
-    \centering
-Squeezing curve showing the two-mode model, Bogoliubov squeezing \eqref{eqBogoliubovSqueezing} and simulation.
-All three disagree, but both the Bog squeezing and simulation are worse than the two-mode model.
-The same parameters are used as in the first part of this figure
-  \end{subfigure}%
+The horizontal axis is the hold time.}
+    }
 \caption{This is not a caption.}

### 2013-06-11 large_number_squeezing_with_atoms.tex
Minor changes.

@@
 To minimise the damaging effects of the non-zero momentum modes it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation.
-To better understand when $N_{\mathbf{k}\neq 0}/N$ can be expected to be small, we approximate the sum in \eqref{eqCondensateDepletion} with an integral.
+To better understand when $N_{\mathbf{k}\neq 0}/N$ will be small, we approximate the sum in \eqref{eqCondensateDepletion} with an integral.
 In 1D Eq.~\eqref{eqCondensateDepletion} is
@@
 \end{align}
-where $\Delta k = 2\pi/L$.
+where $\Delta k = 2\pi/L$ and $L$ is the length of the 1D box condensate.
 This approximation will be valid in the limit that $\omega_{k_\text{min}} \tau_\text{hold} \ll 1$, where $k_\text{min} = 2\pi/L$ as for small $k$ the allowed values $k = n_x 2\pi /L$ will closely sample the region $\xi = \sqrt{\omega_k^0(\omega_k^0 + 2 \chi_{aa} n_a)} \tau_\text{hold} < \pi$ where $\sinc^2(\xi)$ is significant.
@@
   f^\text{2D}(\Lambda) &= \int_0^{\infty} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi,\\
-  f^\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi.
+  f^\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi,
 \end{align}
+and $A$ is the area of the two-dimensional system and $V$ is the volume of the three-dimensional system.
 

### 2013-06-12 large_number_squeezing_with_atoms.tex
Revised the Bog section.
It just needs a figure now.

@@
 \usepackage{amsmath}
-\usepackage{ulem}
+% \usepackage{ulem}
 \usepackage{amssymb}
@@
 where the Bogoliubov mode frequency is $\omega_\mathbf{k} = \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}$, and the free particle frequency is $\omega^0_\mathbf{k} = \hbar \mathbf{k}^2/2 M$.
-The Bogoliubov modes can be expected to have an impact on the squeezing of the system when their occupation is a non-negligible fraction of the total number of atoms.
-The fractional occupation of the Bogoliubov modes is
+The non-zero momentum modes can be expected to have an impact on the squeezing of the system when their occupation is a non-negligible fraction of the total number of atoms.
+The fractional occupation of these modes is
 \begin{align}
@@
 
-Figure~\ref{figBogSqueezingValidation} demonstrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the fractional occupation of the Bogoliubov modes for parameters where the occupation of these modes approach 20\%\footnote{FIXME} of the total occupation of the system.
-The disagreement for larger fractional occupations occurs because the Bogoliubov theory used in deriving \eqref{eqCondensateDepletion} is not number-conserving as it neglects the effect of the Bogoliubov modes on the condensate.
+\begin{figure}
+    \centering
+    \framebox{
+    \parbox{0.8\textwidth}{
+    Plot of the number occupation of the condensate and non-zero momentum modes as computed by the Bog theory (direct summation and integral approximation) and a simulation.
+The simulation agrees with the Bog theory for the occupation of the non-zero momentum modes, but there is significant ~10-20\% depletion of the condensate.
+The horizontal axis is the hold time.}
+    }
+\caption{This is not a caption.}
+  \label{figBogSqueezingValidation}
+\end{figure}
+
+Figure~\ref{figBogSqueezingValidation} demonstrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the fractional occupation of the non-zero momentum modes for parameters where the occupation of these modes approach 20\%\footnote{FIXME} of the total occupation of the system.
+The disagreement for larger fractional occupations occurs because the Bogoliubov theory used in deriving \eqref{eqCondensateDepletion} is not number-conserving as it neglects the effect of the non-zero momentum modes on the condensate.
+
+% I'm commenting this out and will essentially gloss over the matter.  We will illustrate that Bog modes are detrimental, rather than arguing the fact.
+% [Insert motivation that the impact of Bogoliubov modes is detrimental here.
+I could wave my hands and say that more modes is bad, as seen in the previous parts of the article, or I could refer to the squeezing of the Bogoliubov modes:
+% \begin{align}
+%   S_{a,\mathbf{k}\neq 0}(t_3) &= \frac{\frac{1}{2}\sum_{\mathbf{k}\neq0}n_a(\mathbf{k},t_3)\left[3 + \cos 2\theta + n_a(\mathbf{k}, t_3)\right]}{\sum_{\mathbf{k}\neq0} n_a(\mathbf{k}, t_3)},
+% \end{align}
+% which is always greater than 1, but this isn't convincing either as the total squeezing includes the number variance of the $\mathbf{k=0}$ mode as well as the correlations between the $\mathbf{k}=0$ mode and the $\mathbf{k}\neq 0$ modes.
+If you do it naively, you find that when the Bog modes have a significant impact on the squeezing, they have also significantly changed it at $t=t_2$, which must be the same as $t=t_1$.
+The problem is that the Bogoliubov theory used isn't number-conserving.
+The best argument I have at the moment for Bog theory being a good indicator for when things go bad is that a significant Bog fraction implies a depleted condensate, and that will almost certainly make squeezing bad as it will invalidate the single-mode theory.
+More freedom == bad?
+In this case, the best evidence for the utility of Bog theory would seem to be Figure~\ref{fig1D_squeezing_and_bog_scaling}.]
 
@@
 \begin{align}
-  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 L \cos^2\theta } {\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} f^\text{1D}(\Lambda_a),
+  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 L \cos^2\theta } {\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} f^\text{1D}(\Lambda_a), \label{eqBogFraction1D}
 \end{align}
@@
 \begin{align}
-  \frac{N^{\text{2D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 A \cos^2\theta } {4\frac{\hbar}{M}\tau_\text{hold}} f^\text{2D}(\Lambda_a), \\
-  \frac{N^{\text{3D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 V \cos^2\theta } {\sqrt{2}\pi^2\left(\frac{\hbar}{M}\tau_\text{hold}\right)^{3/2}} f^\text{3D}(\Lambda_a),
+  \frac{N^{\text{2D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 A \cos^2\theta } {4\frac{\hbar}{M}\tau_\text{hold}} f^\text{2D}(\Lambda_a),  \label{eqBogFraction2D}\\
+  \frac{N^{\text{3D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 V \cos^2\theta } {\sqrt{2}\pi^2\left(\frac{\hbar}{M}\tau_\text{hold}\right)^{3/2}} f^\text{3D}(\Lambda_a), \label{eqBogFraction3D}
 \end{align}
@@
 and $A$ is the area of the two-dimensional system and $V$ is the volume of the three-dimensional system.
+Figure~\ref{figBogSqueezingValidation} demonstrates the agreement in 1D between the integral approximation to $N_{\mathbf{k}\neq 0}/N$ in \eqref{eqBogFraction1D}, the summation expression in \eqref{eqCondensateDepletion} and a truncated Wigner simulation.
 
-The squeezing predicted by the two-mode model will only be valid in higher dimensions when $N_{\mathbf{k}\neq0}/N \ll 1$, when this condition is violated, the squeezing will be detrimentally affected.
-We therefore want to examine the scaling of $N_{\mathbf{k}\neq 0}/N$ with system size in 1D, 2D and 3D.
-
-For constant total number $N$ and occupations of the $\ket{a}$ and $\ket{b}$ states at time $t_1$, the optimum squeezing as determined by the two-mode model will have a particular value of $\tau_\text{hold}$, $\theta$ and $\phi$.
-As the dimensions of the system are scaled, $\hbar \chi_{ij} = U^\text{1D}_{ij}/L = U^\text{2D}_{ij}/A = U^\text{3D}_{ij}/V$ will change, but to keep the same squeezing (and the same equivalent two-mode model) $\lambda_{ij} = \chi_{ij} \tau_\text{hold}$ must be kept constant (see Eq.~\eqref{eqsDef}--\eqref{eqNumDiffVariance}).
-Therefore $\Lambda_a = n_a\chi_{aa}\tau_\text{hold}$ must also be kept constant.
-With this in mind, we can write $\tau_\text{hold}$ in terms of the system size and the nonlinearity.
-We obtain:
+We wish to examine the scaling of the fractional occupation of the non-zero momentum modes $N_{\mathbf{k}\neq0}/N$ with the physical size of the condensate.
+Other system parameters such as the total number $N$ and the occupation of the $\ket{a}$ and $\ket{b}$ states at $t_1$ ($n_a$ and $n_b$, respectively) are kept constant and chosen to maximise the squeezing of the equivalent two-mode model presented in Section~\ref{secTwoModeAnalytic}.
+In particular this fixes the values of $\theta$, $\phi$ and $\lambda_{ij} = \chi_{ij}\tau_\text{hold}$.
+As the size of the condensate is scaled, $\hbar \chi_{ij} = U^\text{1D}_{ij}/L = U^\text{2D}_{ij}/A = U^\text{3D}_{ij}/V$ will change, but $\tau_\text{hold}$ will vary to keep $\lambda_{ij}$ constant.
+Therefore $\Lambda_a = n_a\chi_{aa}\tau_\text{hold}$ will also be constant.
+With this in mind, we can write $\tau_\text{hold}$ in terms of the system size and the nonlinearity $U_{aa}$.
+The fractional occupation of the non-zero momentum modes can then be shown to scale with the physical size of the condensate as:
 \begin{align}
-  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx \sqrt{L} \cdot \frac{1}{N} \sqrt{\frac{M U^\text{1D}_{aa} n_a}{2 \hbar^2 \pi^2}} \Lambda_a^{3/2} \cos^2\theta f^\text{1D}(\Lambda_a) &&\propto \sqrt{L} \\
-  \frac{N^\text{2D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx 1 \cdot \frac{1}{N} \frac{M U^\text{2D}_{aa} n_a }{4 \hbar^2} \Lambda_a \cos^2\theta f^\text{2D}(\Lambda_a) &&\propto \text{constant} \\
-  \frac{N^\text{3D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx \frac{1}{\sqrt{V}} \cdot \frac{1}{N} \left(\frac{M U^\text{3D}_{aa} n_a}{\hbar^2}\right)^{3/2} \frac{\Lambda_a^{1/2} \cos^2\theta f^\text{3D}(\Lambda_a)}{\sqrt{2} \pi^2} &&\propto \frac{1}{\sqrt{V}}
+  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx \sqrt{L} \cdot \frac{1}{N} \sqrt{\frac{M U^\text{1D}_{aa} n_a}{2 \hbar^2 \pi^2}} \Lambda_a^{3/2} \cos^2\theta f^\text{1D}(\Lambda_a) &&\propto \sqrt{L}, \\
+  \frac{N^\text{2D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx 1 \cdot \frac{1}{N} \frac{M U^\text{2D}_{aa} n_a }{4 \hbar^2} \Lambda_a \cos^2\theta f^\text{2D}(\Lambda_a) &&\propto \text{constant}, \\
+  \frac{N^\text{3D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx \frac{1}{\sqrt{V}} \cdot \frac{1}{N} \left(\frac{M U^\text{3D}_{aa} n_a}{\hbar^2}\right)^{3/2} \frac{\Lambda_a^{1/2} \cos^2\theta f^\text{3D}(\Lambda_a)}{\sqrt{2} \pi^2} &&\propto \frac{1}{\sqrt{V}}.
 \end{align}
-
-In 1D the condensate depletion $N_{\mathbf{k}\neq 0}/N$ increases with increasing system size, in 2D the condensate depletion remains constant, and in 3D the condensate depletion decreases as the system size increases.
+The fractional occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}/N$ thus scales differently with trap size in each of 1D, 2D and 3D.
+For one-dimensional traps, the fractional occupation of these modes increases with system size, whereas for 2D traps it is constant.
+For three-dimensional traps, the fractional occupation of the non-zero momentum modes \emph{decreases} with volume.
 This suggests that the best system to use will be a box potential, and the system size can be increased until the condensate depletion becomes negligibly small.
 
-Qualitatively, the occupation of the non-zero momentum modes is almost entirely determined by the number of Bogoliubov modes energetically accessible to the system.  While this gets larger
-
-The Bogoliubov theory used in deriving \eqref{eqCondensateDepletion} is not number-conserving
-
-To minimise the damaging effects of the non-zero momentum modes, it is better to
-
-The Bogoliubov theory used in deriving \eqref{eqBogOccupation} is not number-conserving because the effect of the Bogoliubov modes on the condensate is neglected.
-
-To minimise the
-
-Good squeezing will only be observed when the fractional occupation of the Bogoliubov modes $N_{\mathbf{k}\neq 0}/N$ is small.
-
-% I'm commenting this out and will essentially gloss over the matter.  We will illustrate that Bog modes are detrimental, rather than arguing the fact.
-% [Insert motivation that the impact of Bogoliubov modes is detrimental here.
-I could wave my hands and say that more modes is bad, as seen in the previous parts of the article, or I could refer to the squeezing of the Bogoliubov modes:
-% \begin{align}
-%   S_{a,\mathbf{k}\neq 0}(t_3) &= \frac{\frac{1}{2}\sum_{\mathbf{k}\neq0}n_a(\mathbf{k},t_3)\left[3 + \cos 2\theta + n_a(\mathbf{k}, t_3)\right]}{\sum_{\mathbf{k}\neq0} n_a(\mathbf{k}, t_3)},
-% \end{align}
-% which is always greater than 1, but this isn't convincing either as the total squeezing includes the number variance of the $\mathbf{k=0}$ mode as well as the correlations between the $\mathbf{k}=0$ mode and the $\mathbf{k}\neq 0$ modes.
-If you do it naively, you find that when the Bog modes have a significant impact on the squeezing, they have also significantly changed it at $t=t_2$, which must be the same as $t=t_1$.
-The problem is that the Bogoliubov theory used isn't number-conserving.
-The best argument I have at the moment for Bog theory being a good indicator for when things go bad is that a significant Bog fraction implies a depleted condensate, and that will almost certainly make squeezing bad as it will invalidate the single-mode theory.
-More freedom == bad?
-In this case, the best evidence for the utility of Bog theory would seem to be Figure~\ref{fig1D_squeezing_and_bog_scaling}.]
-
-The squeezing of mode $|a\rangle$ in \eqref{eqBogoliubovSqueezing} includes the first-order multimode quantum mechanical fluctuations about the mean field and differs from the single-mode squeezing of $\Var[N_a](t_3)/N_a(t_3)$ only by terms involving the occupation of the non-zero momentum modes $n_a(\mathbf{k}, t)$.
-The contribution of these terms is always detrimental to squeezing as the contribution to the numerator of \eqref{eqBogoliubovSqueezing} is always at least as large as the contribution to the denominator.
-When the occupation of the non-zero momentum modes is a non-negligible fraction of the condensate, the observed squeezing will differ from that predicted by the two-mode model in Section~\ref{secTwoModeAnalytic} (see Fig.~\ref{figBogSqueezingValidation}).
-
-\begin{figure}
-    \centering
-    \framebox{
-    \parbox{0.8\textwidth}{
-    Plot of the number occupation of the condensate and non-zero momentum modes as computed by the Bog theory (direct summation and integral approximation) and a simulation.
-The simulation agrees with the Bog theory for the occupation of the non-zero momentum modes, but there is significant ~10-20\% depletion of the condensate.
-The horizontal axis is the hold time.}
-    }
-\caption{This is not a caption.}
-  \label{figBogSqueezingValidation}
-\end{figure}
-
-The Bogoliubov theory used in deriving \eqref{eqBogoliubovSqueezing} is not number-conserving because the depletion of the condensate caused by the occupation of the non-zero momentum modes is neglected.
-Therefore when these non-zero momentum modes become sufficiently occupied for $S_a$ to differ from the two-mode model, the Bogoliubov approximation becomes invalid and $S_a$ will not agree with the full quantum-statistical multimode result.
-While number-conserving variants of the Bogoliubov theory exist \cite{NumberConservingBog}, they are still perturbative approximations about the mean field, and they are not required for our purposes.
-Although the Bogoliubov approximation cannot predict the exact amount by which squeezing will be degraded compared to the two-mode model, it does predict the regime in which squeezing is reduced by coupling between spatial modes, and it can thereby suggest how this problem can be avoided.
-Figure~\ref{figBogSqueezingValidation}(a) illustrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the total number of atoms in the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ for parameters where these modes approach 20\% of the total number of atoms in the system.
-Figure~\ref{figBogSqueezingValidation}(b) shows that in this case the squeezing predicted by the Bogoliubov theory in Eq.~\eqref{eqBogoliubovSqueezing} is a lower bound for the squeezing predicted by the truncated Wigner simulations.
-The two-mode analytic solution, which ignores multimode effects entirely, provides an upper bound for the squeezing.
-\textbf{Reconsider these last sentences when we have the figure.
-Maybe qualify `lower bound' for constant density models only?}  This means that we can use the Bogoliubov analysis to guide our search for parameter regimes with higher squeezing.
-
-To minimise the damaging effects of the non-zero momentum modes, it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation.
-
-The number of excited modes is almost entirely determined by the density of modes energetically accessible by the system.
-While this gets larger with the dimensionality and the volume of the system at a given chemical potential, in practice the dimensionality and the volume of the BEC are controlled by changing the trap parameters, which has a dramatic effect on the attainable chemical potential.
-So the true figure of merit controlling the degradation of the squeezing due to coupling between the modes is the number of states energetically reachable from the chemical potential.
-This scales differently with trap size in each of 1D, 2D and 3D.
-For one-dimensional traps, the number of accessible states increases with volume, whereas for 2D traps it is constant with volume.
-For three-dimensional traps, the number of accessible states \emph{decreases} with volume.
+% End of Graham's content
 

### 2013-06-12 large_number_squeezing_with_atoms.tex
Added the Bog validation figure.

@@
     \centering
-    \framebox{
-    \parbox{0.8\textwidth}{
-    Plot of the number occupation of the condensate and non-zero momentum modes as computed by the Bog theory (direct summation and integral approximation) and a simulation.
-The simulation agrees with the Bog theory for the occupation of the non-zero momentum modes, but there is significant ~10-20\% depletion of the condensate.
-The horizontal axis is the hold time.}
-    }
-\caption{This is not a caption.}
+    \includegraphics[width=10cm]{figures/Bog_validation}
+\caption{Plot of the fractional occupation of the non-zero momentum modes as a function of the hold time $\tau_\text{hold}$.
+The blue dashed line is the Bogoliubov integral approximation given in Eq.~\ref{eqBogFraction1D}, the purple dot-dashed line is the Bogoliubov summation expression given in Eq.~\ref{eqCondensateDepletion}, and the solid black line is the result of a truncated Wigner simulation.
+Parameters: $L=600\,\mu\text{m}$, $N=4 \times 10^5$, $n_a=n_b=2\times 10^5$, $\chi_{aa} = 2.67 \times 10^{-2}\, \text{s}^{-1}$, $\chi_{ab}=\chi_{bb}=0$, $\theta=0$.}
   \label{figBogSqueezingValidation}
@@
 
-Figure~\ref{figBogSqueezingValidation} demonstrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the fractional occupation of the non-zero momentum modes for parameters where the occupation of these modes approach 20\%\footnote{FIXME} of the total occupation of the system.
+Figure~\ref{figBogSqueezingValidation} demonstrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the fractional occupation of the non-zero momentum modes for parameters where the occupation of these modes approach 15\% of the total occupation of the system.
 The disagreement for larger fractional occupations occurs because the Bogoliubov theory used in deriving \eqref{eqCondensateDepletion} is not number-conserving as it neglects the effect of the non-zero momentum modes on the condensate.

### 2013-06-12 large_number_squeezing_with_atoms.tex
Normalise terminology.
I was using "fractional occupation of the non-zero momentum modes" in my bog section when the shorter phrase "non-condensed fraction" was used elsewhere.  I updated my section to use the shorter phrase.

@@
 To simplify the analysis we consider the case in which $U_{aa} \neq 0$, $U_{bb}=U_{ab} = 0$ and the condensate has a spatially-constant density in a box of side lengths $L_x$, $L_y$ and $L_z$.
-In this limit the occupation of the non-zero momentum modes in the $\ket{a}$ internal state at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation})
+In this limit the occupation of the non-zero momentum modes (the non-condensed fraction) in the $\ket{a}$ internal state at the end of the pulse sequence is (a derivation is given in \ref{appendixBogDerivation})
 \begin{align}
@@
     \includegraphics[width=10cm]{figures/Bog_validation}
-\caption{Plot of the fractional occupation of the non-zero momentum modes as a function of the hold time $\tau_\text{hold}$.
+\caption{Plot of the non-condensed fraction $N_{\mathbf{k}\neq0}/N$ as a function of the hold time $\tau_\text{hold}$.
 The blue dashed line is the Bogoliubov integral approximation given in Eq.~\ref{eqBogFraction1D}, the purple dot-dashed line is the Bogoliubov summation expression given in Eq.~\ref{eqCondensateDepletion}, and the solid black line is the result of a truncated Wigner simulation.
@@
 
-Figure~\ref{figBogSqueezingValidation} demonstrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the fractional occupation of the non-zero momentum modes for parameters where the occupation of these modes approach 15\% of the total occupation of the system.
-The disagreement for larger fractional occupations occurs because the Bogoliubov theory used in deriving \eqref{eqCondensateDepletion} is not number-conserving as it neglects the effect of the non-zero momentum modes on the condensate.
+Figure~\ref{figBogSqueezingValidation} demonstrates the agreement between truncated Wigner simulations and the Bogoliubov theory for the non-condensed fraction for parameters where the occupation of these modes approach 15\% of the total occupation of the system.
+The disagreement for larger non-condensed fractions occurs because the Bogoliubov theory used in deriving \eqref{eqCondensateDepletion} is not number-conserving as it neglects the effect of the non-zero momentum modes on the condensate.
 
@@
 
-To minimise the damaging effects of the non-zero momentum modes it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation.
+To minimise the damaging effects of the non-condensed fraction it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation.
 To better understand when $N_{\mathbf{k}\neq 0}/N$ will be small, we approximate the sum in \eqref{eqCondensateDepletion} with an integral.
@@
 
-Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$ and defining $\Lambda_a = n_a \chi_{aa} \tau_\text{hold}$ the expression \eqref{eqCondensateDepletionFirstIntegralApproximation} can be written as
+Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$ and defining $\Lambda_a = n_a \chi_{aa} \tau_\text{hold}$ the non-condensed fraction \eqref{eqCondensateDepletionFirstIntegralApproximation} can be written as
 \begin{align}
@@
 
-We wish to examine the scaling of the fractional occupation of the non-zero momentum modes $N_{\mathbf{k}\neq0}/N$ with the physical size of the condensate.
+We wish to examine the scaling of the non-condensed fraction $N_{\mathbf{k}\neq0}/N$ with the physical size of the condensate.
 Other system parameters such as the total number $N$ and the occupation of the $\ket{a}$ and $\ket{b}$ states at $t_1$ ($n_a$ and $n_b$, respectively) are kept constant and chosen to maximise the squeezing of the equivalent two-mode model presented in Section~\ref{secTwoModeAnalytic}.
@@
 With this in mind, we can write $\tau_\text{hold}$ in terms of the system size and the nonlinearity $U_{aa}$.
-The fractional occupation of the non-zero momentum modes can then be shown to scale with the physical size of the condensate as:
+The non-condensed fraction can then be shown to scale with the physical size of the condensate as:
 \begin{align}
@@
 \end{align}
-The fractional occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}/N$ thus scales differently with trap size in each of 1D, 2D and 3D.
-For one-dimensional traps, the fractional occupation of these modes increases with system size, whereas for 2D traps it is constant.
-For three-dimensional traps, the fractional occupation of the non-zero momentum modes \emph{decreases} with volume.
-This suggests that the best system to use will be a box potential, and the system size can be increased until the condensate depletion becomes negligibly small.
+The non-condensed fraction $N_{\mathbf{k}\neq 0}/N$ thus scales differently with trap size in each of 1D, 2D and 3D.
+For one-dimensional traps, the non-condensed fraction increases with system size, whereas for 2D traps it is constant.
+For three-dimensional traps, the non-condensed fraction \emph{decreases} with volume.
+This suggests that the best system to use will be a three-dimensional box potential, and the system size can be increased until the condensate depletion becomes negligibly small.
 
@@
 
-Figure \ref{fig1D_squeezing_and_bog_scaling} and Figure \ref{fig2D_squeezing_and_bog_scaling} show the non-condensed fraction predicted by the Bogoliubov analysis compared to that predicted by a full numerical simulation, in one, two and three dimensions, as a function of the trap size.
+Figures~\ref{fig1D_squeezing_and_bog_scaling}--\ref{fig3D_Bog_mode_occupation_scaling} show the non-condensed fraction predicted by the Bogoliubov analysis compared to that predicted by a full numerical simulation, in one, two and three dimensions, as a function of the trap size.
 Also shown in the 1D and 2D cases is the maximum amount of number squeezing obtainable, as predicted by the numerical simulation.

### 2013-06-12 large_number_squeezing_with_atoms.tex
coupling between spatial modes -> momentum modes
What was said previously was wrong.  It should be nonlinearity-induced coupling between momentum modes.
Added a brief summary of the Bog derivation.

@@
 \newcommand{\ket}[1]{\ensuremath{\left|#1\right>}}
+\renewcommand{\Re}{\operatorname{Re}}
+\renewcommand{\Im}{\operatorname{Im}}
 
@@
 The dependence on dimensionality is not explained by the processes described in section \ref{sec:MMdescription}, which focus on the effects of inhomogeneity within the trap.
-Fundamentally, this degradation of the squeezing is due to nonlinearity-induced coupling between different spatial modes.
+Fundamentally, this degradation of the squeezing is due to nonlinearity-induced coupling between different momentum modes.
 
@@
 
-Here we put some details of the Bog analysis.
+The Bogoliubov model presented in Section~\ref{sec:2D3Dsqueezing} is taken in the limit of a homogeneous condensate with $\chi_{aa}\neq 0$, $\chi_{ab}=\chi_{bb}=0$.
+In this limit, the Hamiltonian for the internal state $\ket{a}$ during the hold time is
+\begin{align}
+  \hat{H}_{a} &= \int \left[ \hat{\psi}_a^\dagger \left(-\frac{\hbar^2}{2 m}\nabla^2\right) \hat{\psi}_a + \frac{U_{aa}}{2} \hat{\psi}_a^\dagger \hat{\psi}_a^\dagger \hat{\psi}_a \hat{\psi}_a \right] \,dV \label{eqBogOriginalHamiltonian}
+\end{align}
+To find the response of the condensate to small fluctuations we define fluctuation operators for the non-zero momentum modes $\delta\hat{a}_\mathbf{k} = \hat{a}_\mathbf{k} - \alpha_\mathbf{k}$ where $\alpha_\mathbf{k} = \expect{\hat{a}_\mathbf{k}}$ is the condensate mean field ($\alpha_\mathbf{k}=0$ for $\mathbf{k} \neq 0$ as the condensate is homogenous).
+The Hamiltonian \eqref{eqBogOriginalHamiltonian} is then expanded in terms of $\delta\hat{a}_\mathbf{k}$ retaining only the lowest order contributions.
+The lowest order contribution is second order:
+\begin{align}
+  \hat{H}_{a}^{(2)} &= \sum_{\mathbf{k}\neq 0} \left[ \frac{\hbar^2 \mathbf{k}^2}{2M} \delta\hat{a}_\mathbf{k}^\dagger \delta\hat{a}_\mathbf{k} + \frac{1}{2} \hbar \chi_{aa} n_a \left(\delta\hat{a}_\mathbf{k}^\dagger \delta\hat{a}_{-\mathbf{k}}^\dagger + \delta\hat{a}_{-\mathbf{k}} \delta\hat{a}_{\mathbf{k}} \right) \right].
+\end{align}
+This Hamiltonian is diagonalised with the standard Bogoliubov transformation \cite{PethickSmith}
+\begin{align}
+  \delta \hat{a}_\mathbf{k}(t) &= u_\mathbf{k} \hat{b}_\mathbf{k}(t) - v_k \hat{b}_{-\mathbf{k}}^\dagger(t),\\
+  u_k^2 &= v_k^2+1 = \frac{1}{2} \left(\frac{\omega^0_\mathbf{k} + 2 \chi_{aa}n_a}{\omega_\mathbf{k}} +1\right),
+\end{align}
+where $\omega^0_\mathbf{k} = \hbar \mathbf{k}^2/2M$ is the frequency of a free particle, and $\omega_\mathbf{k} = \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}$ is the frequency of the Bogoliubov modes:
+\begin{align}
+  \delta\hat{b}_\mathbf{k}(t) &= \exp\left(-i \omega_\mathbf{k} t\right)\delta\hat{b}_\mathbf{k}(0).
+\end{align}
 
+The expectation value of the number of atoms in momentum mode $\hbar\mathbf{k}$ after the hold time is then found to be
 \begin{align}
-  S &= \frac{\Var(n_0) + \sum_k \frac{1}{2} n_k (3 + \cos 2\theta + 4 n_k)}{n_0 + \sum_k n_k} \\
-  n_k &= \left[\lambda \cos\theta \sinc(\omega_k \tau)\right]^2 \\
-  \hat{a}(t_3) &= \left[\hat{a}(0) - i \chi \alpha \tau_h\left(\alpha^* \hat{a}(0) + \alpha \hat{a}^\dagger(0) - 2\abs{\alpha}^2\right) \right]\cos(\theta) - i\hat{b}(0) \sin\theta e^{i\phi} \\
-  n_0 &= \left(\alpha ^2+\alpha ^4 \lambda ^2\right) \cos^2\theta + \left(\beta^2 \sin\theta  + 2\alpha\beta\cos\theta\sin\phi\right)\sin\theta\\
+  \expect{\hat{n}_\mathbf{k}(t_2)} &= \expect{\delta\hat{a}^\dagger_\mathbf{k} \delta\hat{a}_\mathbf{k}} \\
   \begin{split}
-  \Var(n_0) &= \frac{1}{2} \left(\alpha ^2+\beta ^2\right)+\frac{1}{8}\alpha ^4 \left(7+ 2\beta ^2\right) \lambda ^2+ \frac{3}{4} \alpha ^8 \lambda ^4 \\
-  &\phantom{\mathrel{=}}+\frac{1}{2}\left( \alpha ^2-\beta ^2+2 \alpha ^4 \lambda ^2+2 \alpha ^8 \lambda ^4\right) \text{Cos}[2 \theta ] \\
-  &\phantom{\mathrel{=}}+\frac{1}{8}\left(\alpha ^4 \lambda ^2 \left(1-2 \beta ^2+2 \alpha ^4 \lambda ^2\right) \text{Cos}[4 \theta ] \right) \\
-  &\phantom{\mathrel{=}}+\frac{1}{8}\left(4 \alpha  \beta  \left(8 \alpha ^2 \lambda  \text{Cos}[\theta ]^3 \text{Cos}[\phi ] \text{Sin}[\theta ]+\text{Sin}[2 \theta ] \left(2 \text{Sin}[\phi ]+\alpha  \beta  \lambda  \text{Sin}[2 \theta ] \left(\alpha ^2 \lambda  \text{Cos}[2 \phi ]+\text{Sin}[2 \phi ]\right)\right)\right)\right)
+  &=\left[1 + 4 u_\mathbf{k}^2(2 u_\mathbf{k}^2 - 1)\sin^2(\omega_\mathbf{k} \tau_\text{hold})\right] \expect{\delta \hat{a}^\dagger_\mathbf{k} \hat{a}_\mathbf{k}} \\
+  &\phantom{\mathrel{=}}- 4 u_\mathbf{k} v_\mathbf{k} \sin(\omega_\mathbf{k}\tau_\text{hold}) \Im\left\{\left[1-i2 u_\mathbf{k}^2 \sin(\omega_k \tau_\text{hold})\right] \expect{\hat{a}_{-\mathbf{k}} \hat{a}_\mathbf{k}}\right\}\\
+  &\phantom{\mathrel{=}} + 4 u_\mathbf{k}^2 v_\mathbf{k}^2 \sin^2(\omega_\mathbf{k}\tau_\text{hold})
   \end{split}
 \end{align}
+If the non-zero momentum modes are initially in a vacuum state then this reduces to
+\begin{align}
+  \expect{\hat{n}_\mathbf{k}(t_2)} &= 4 u_\mathbf{k}^2 v_\mathbf{k}^2 \sin^2(\omega_\mathbf{k}\tau_\text{hold}), \\
+  &= \left[n_a \chi_{aa} \tau_\text{hold} \sinc(\omega_\mathbf{k} \tau_\text{hold}) \right]^2.
+\end{align}
+After the final Rabi coupling pulse the occupation of the non-zero momentum mode $\hbar\mathbf{k}$ is
+\begin{align}
+  n_a(\mathbf{k}, t_3) &= \left[n_a \chi_{aa} \tau_\text{hold} \cos(\theta)\sinc(\omega_\mathbf{k} \tau_\text{hold}) \right]^2,
+\end{align}
+which is the expression given earlier in Eq.~\eqref{eqBogOccupation}.
 

### 2013-06-13 large_number_squeezing_with_atoms.tex
Fix some missing \delta's.
Add a few introductory references

--- 2013-06-12 large_number_squeezing_with_atoms.tex
@@
 \end{itemize}
+
+History: Quantum optics led to creating nonclassical states after the laser was
+invented. The advent of the BEC led to the ability to start recreating these types
+of states with massive particles. One of the first things that was considered was
+quadrature squeezing, as this can be done with intrinsic nonlinearities of the atoms,
+which photons don’t have. Quadrature squeezing using nonlinearities has been around
+a while (cite Drummond) due to being useful for EPR. Also been shown you can use
+it for atom lasers. Possible to use atom nonlinearities to continuously quadrature
+squeeze an atom laser [?] To get number squeezing, however, need to mix quadrature
+squeezed states or mix a quadrature squeezed state with a coherent local oscillator.
+Can be hard to mode match beams of atoms, so doing everything in trap can useful.
+That’s why number squeezing (aka spin squeezing) often tends to use a Ramsey-type
+interferometry scheme. [4].
+These sorts of things have led to experimentally detectable amounts of number
+squeezing [10]
 

### 2013-06-13 large_number_squeezing_with_atoms.tex

--- 2013-06-12 large_number_squeezing_with_atoms.tex
@@
   &=\left[1 + 4 u_\mathbf{k}^2(2 u_\mathbf{k}^2 - 1)\sin^2(\omega_\mathbf{k} \tau_\text{hold})\right] \expect{\delta \hat{a}^\dagger_\mathbf{k} \hat{a}_\mathbf{k}} \\
-  &\phantom{\mathrel{=}}- 4 u_\mathbf{k} v_\mathbf{k} \sin(\omega_\mathbf{k}\tau_\text{hold}) \Im\left\{\left[1-i2 u_\mathbf{k}^2 \sin(\omega_k \tau_\text{hold})\right] \expect{\hat{a}_{-\mathbf{k}} \hat{a}_\mathbf{k}}\right\}\\
+  &\phantom{\mathrel{=}}- 4 u_\mathbf{k} v_\mathbf{k} \sin(\omega_\mathbf{k}\tau_\text{hold}) \Im\left\{\left[1-i2 u_\mathbf{k}^2 \sin(\omega_k \tau_\text{hold})\right] \expect{\delta\hat{a}_{-\mathbf{k}} \delta\hat{a}_\mathbf{k}}\right\}\\
   &\phantom{\mathrel{=}} + 4 u_\mathbf{k}^2 v_\mathbf{k}^2 \sin^2(\omega_\mathbf{k}\tau_\text{hold})

### 2013-06-14 large_number_squeezing_with_atoms.tex
Fix lots of errors in my Appendix found by Mattias.
Fix another error in my Appendix.
Add references plus minor tweaks

--- 2013-06-12 large_number_squeezing_with_atoms.tex
@@
 \end{itemize}
+
+Quantum optics led to creating nonclassical states after the laser was invented \cite{???}.
+The experimental realization of the BEC \cite{???} led to the ability to start recreating these types of states with massive particles.
+
+One of the advantages of atoms over photons is that they interact with each other, unlike photons, which allows nonlinear interaction effects without all the fluffery that optics requires, like crystals etc.
+
+It has been recognized that the intrinsic nonlinearities of atoms could lead to effects like quadrature squeezing \cite{kheruntsyanET2005, johnssonET2007, haineET2009, doeringET2010}.
+Can also get number squeezing using spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissocation of BECs \cite{kheruntsyanET2002, davisET2008, savage2007, ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, or double-well potentials \cite{esteve2008, maussang2010} or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}.
+
+{\bf Note that the FWM stuff is using the s-wave scatter sphere from condensate collisions, so pretty useless for interferometry}
+
+The creation of these non-classical atomic states has relevence for fundamental tests of quantum mechanics such as the EPR paradox \cite{pu2000, sorensen2001, haine2005,zhao2007, ogren2010}, as well as atom inteferometry \cite{sorensenET1999, kuzmichET2000, meyerET2001, liebfriedET2004, roosET2006, doeringET2010, gross2010}
+
+While these techniques have resulted in impressive levels of experimentally detectable number squeezing \cite{gross2010}, they have been limited to small numbers of atoms, typically a few hundred.
+
+To really win in the interferometry game, you need good squeezing for large numbers of atoms.
+The reason people have used small numbers of atoms is that it allows you to move into the single-mode regime, where the trap is tight enough such that all higher modes are frozen out \cite{LiET2009}
+
+If this is not the case multimode stuff needs to be considered, and that's what we do in this paper.
+
+Most previous simulations of this sort of stuff tend to be in 1D or 2D for simplicity \cite{kheruntsyanET2002, kheruntsyanET2005, johnssonET2007, haineET2009}.
+They reduce the problem from 3D to 1D by integrating out one or two dimensions \cite{???}.
+Semi-classically this isn't a problem, because you can assume the modal dynamics have symmetry.
+However as we show in this paper, when dealing with QFT problems this naive approach can give you the wrong answer.
+We can do full 3D simulations because we're cool and we have XMDS2 \cite{dennis2013}.
 
@@
 The traps are assumed to have a constant potential, which minimises the effects of the multimode issues described in section \ref{sec:MMdescription}.
-We see that the squeezing does indeed degrade at large trap sizes in 1D, but remains constant in $2D$.
+We see that the squeezing does indeed degrade at large trap sizes in 1D, but remains constant in 2D.
 For small trap sizes, the squeezing in 1D aymptotically approaches that of the ideal 0D system.
@@
 \caption{1D scaling as function of trap size, $L$.
-(Left) The proportion the non-condensed fraction in mode $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+(Left) The proportion the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
 (Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
@@
 \caption{2D scaling as function of trap size, L.
-(Left) The proportion the non-condensed fraction in mode $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+(Left) The proportion the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
 (Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
-Parameters: $N=1\times 10^5$, $U_{aa} = 8.9\times 10^{-44}$Jm, $U_{ab}=U_{bb}=0$, $\tau_{\mathrm{hold}} = 10^{5} L^2$\,s.}
+Parameters: $N=1\times 10^5$, $U_{aa} = 8.9\times 10^{-44}$Jm$^2$, $U_{ab}=U_{bb}=0$, $\tau_{\mathrm{hold}} = 10^{5} L^2$\,s.}
   \label{fig2D_squeezing_and_bog_scaling}
@@
 \caption{3D scaling as function of trap size, $L$.
-(Left) The proportion the non-condensed fraction in mode $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+(Left) The proportion the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
 (Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
-Parameters: $N=1\times 10^4$, $U_{aa} = 4.7\times 10^{-49}$Jm, $U_{ab}=U_{bb}=0$, $\tau_{\mathrm{hold}} = 1.25 \times 10^{11} L^3$\,s.}
+Parameters: $N=1\times 10^4$, $U_{aa} = 4.7\times 10^{-49}$Jm$^3$, $U_{ab}=U_{bb}=0$, $\tau_{\mathrm{hold}} = 1.25 \times 10^{11} L^3$\,s.}
  \label{fig3D_Bog_mode_occupation_scaling}
@@
 
-By a Bogoliubov analysis, it was shown that the coupling between different modes could be minimised in two regimes.
+Using a Bogoliubov analysis, it was shown that the coupling between different modes could be minimised in two regimes.
 The first was the tight trapping regime, where the system becomes effectively zero-dimensional, and the second was the weak trapping limit for 3D traps.
 This latter regime is in fact ideally compatible with large atomic numbers, and shows the prospect of extremely high squeezing.
-A TWA simulation demonstrated a BEC trap with $15dB$ of squeezing.
+A TWA simulation demonstrated a BEC trap with $15$\,dB of squeezing.
 In practice, the only limit to the number of atoms in such traps would be the stability of the trap, as the BEC would become increasingly sensitive to fluctuations as the energy per particle is reduced as the volume increases.
@@
 
-The Bogoliubov model presented in Section~\ref{sec:2D3Dsqueezing} is taken in the limit of a homogeneous condensate with $\chi_{aa}\neq 0$, $\chi_{ab}=\chi_{bb}=0$.
+We consider the Bogoliubov model presented in Section~\ref{sec:2D3Dsqueezing} taken in the limit of a homogeneous condensate with $\chi_{aa}\neq 0$, $\chi_{ab}=\chi_{bb}=0$.
 In this limit, the Hamiltonian for the internal state $\ket{a}$ during the hold time is
 \begin{align}
-  \hat{H}_{a} &= \int \left[ \hat{\psi}_a^\dagger \left(-\frac{\hbar^2}{2 m}\nabla^2\right) \hat{\psi}_a + \frac{U_{aa}}{2} \hat{\psi}_a^\dagger \hat{\psi}_a^\dagger \hat{\psi}_a \hat{\psi}_a \right] \,dV \label{eqBogOriginalHamiltonian}
+  \hat{H}_{a} &= \int \left[ \hat{\psi}_a^\dagger \left(-\frac{\hbar^2}{2 m}\nabla^2\right) \hat{\psi}_a + \frac{U_{aa}}{2} \hat{\psi}_a^\dagger \hat{\psi}_a^\dagger \hat{\psi}_a \hat{\psi}_a \right] \,dV.
+\label{eqBogOriginalHamiltonian}
 \end{align}
@@
 \begin{align}
-  \delta \hat{a}_\mathbf{k}(t) &= u_\mathbf{k} \hat{b}_\mathbf{k}(t) - v_k \hat{b}_{-\mathbf{k}}^\dagger(t),\\
-  u_k^2 &= v_k^2+1 = \frac{1}{2} \left(\frac{\omega^0_\mathbf{k} + 2 \chi_{aa}n_a}{\omega_\mathbf{k}} +1\right),
+  \delta \hat{a}_\mathbf{k}(t) &= u_\mathbf{k} \delta\hat{b}_\mathbf{k}(t) - v_k \delta\hat{b}_{-\mathbf{k}}^\dagger(t),\\
+  u_k^2 &= v_k^2+1 = \frac{1}{2} \left(\frac{\omega^0_\mathbf{k} + \chi_{aa}n_a}{\omega_\mathbf{k}} +1\right),
 \end{align}
@@
 \begin{align}
-  \expect{\hat{n}_\mathbf{k}(t_2)} &= \expect{\delta\hat{a}^\dagger_\mathbf{k} \delta\hat{a}_\mathbf{k}} \\
+  \expect{\hat{n}_\mathbf{k}(t_2)} &= \expect{\delta\hat{a}^\dagger_\mathbf{k}(t_2) \delta\hat{a}_\mathbf{k}(t_2)} \\
   \begin{split}
-  &=\left[1 + 4 u_\mathbf{k}^2(2 u_\mathbf{k}^2 - 1)\sin^2(\omega_\mathbf{k} \tau_\text{hold})\right] \expect{\delta \hat{a}^\dagger_\mathbf{k} \hat{a}_\mathbf{k}} \\
-  &\phantom{\mathrel{=}}- 4 u_\mathbf{k} v_\mathbf{k} \sin(\omega_\mathbf{k}\tau_\text{hold}) \Im\left\{\left[1-i2 u_\mathbf{k}^2 \sin(\omega_k \tau_\text{hold})\right] \expect{\hat{a}_{-\mathbf{k}} \hat{a}_\mathbf{k}}\right\}\\
-  &\phantom{\mathrel{=}} + 4 u_\mathbf{k}^2 v_\mathbf{k}^2 \sin^2(\omega_\mathbf{k}\tau_\text{hold})
+  &=\left[1 + 4 u_\mathbf{k}^2(2 u_\mathbf{k}^2 - 1)\sin^2(\omega_\mathbf{k} \tau_\text{hold})\right] \expect{\delta \hat{a}^\dagger_\mathbf{k}(t_1) \delta\hat{a}_\mathbf{k}(t_1)} \\
+  &\phantom{\mathrel{=}}- 4 u_\mathbf{k} v_\mathbf{k} \sin(\omega_\mathbf{k}\tau_\text{hold}) \Im\left\{\left[1-i2 u_\mathbf{k}^2 \sin(\omega_k \tau_\text{hold})\right] \expect{\delta\hat{a}_{-\mathbf{k}}(t_1) \delta\hat{a}_\mathbf{k}(t_1)}\right\}\\
+  &\phantom{\mathrel{=}} + 4 u_\mathbf{k}^2 v_\mathbf{k}^2 \sin^2(\omega_\mathbf{k}\tau_\text{hold}).
   \end{split}
 \end{align}
-If the non-zero momentum modes are initially in a vacuum state then this reduces to
+If the non-zero momentum modes are in a vacuum state at $t=t_1$ then this reduces to
 \begin{align}
-  \expect{\hat{n}_\mathbf{k}(t_2)} &= 4 u_\mathbf{k}^2 v_\mathbf{k}^2 \sin^2(\omega_\mathbf{k}\tau_\text{hold}), \\
+  \expect{\hat{n}_\mathbf{k}(t_2)} &= 4 u_\mathbf{k}^2 v_\mathbf{k}^2 \sin^2(\omega_\mathbf{k}\tau_\text{hold}) \\
   &= \left[n_a \chi_{aa} \tau_\text{hold} \sinc(\omega_\mathbf{k} \tau_\text{hold}) \right]^2.
@@
 \begin{align}
-  n_a(\mathbf{k}, t_3) &= \left[n_a \chi_{aa} \tau_\text{hold} \cos(\theta)\sinc(\omega_\mathbf{k} \tau_\text{hold}) \right]^2,
+  n_a(\mathbf{k}, t_3) &= \left[n_a \chi_{aa} \tau_\text{hold} \cos(\theta)\sinc(\omega_\mathbf{k} \tau_\text{hold}) \right]^2
 \end{align}
@@
 \begin{thebibliography}{10}
+\bibitem{kheruntsyanET2002} Kheruntsyan K V and Drummond P D 2002 Quantum correlated twin atomic beams via photodissociation of a molecular Bose-Einstein condensate \emph {Phys.
+Rev.} A {\bf 66} 031602(R)
+
+\bibitem{kheruntsyanET2005} Kheruntsyan K V, Olsen M K and Drummond P D 2005 Einstein-Podolsky-Rosen correlations via dissociation of a molecular Bose-Einstein condensate \emph{Phys.
+Rev.} A {\bf 66} 031602(R)
+
+\bibitem{kitagawaET1993} Kitagawa M and Ueda M 1995 Spin squeezed states \emph{Phys. Rev.} A {\bf 47} 5138
+
+\bibitem{sorensenET1999} Hald J, S{\"{o}}rensen J L, Schori C, and Polzik E S 1999 Spin squeezed atoms: a macroscopic entangled ensemble created by light \emph{Phys. Rev. Lett.} {\bf 83} 1319
+
+\bibitem{kuzmichET2000} Kuzmich A, Mandel L, and Bigelow N P 2000 Generation of spin squeezing via
+continuous quantum nondemolition measurement \emph{Phys. Rev. Lett.} {\bf 85} 1594
+
+\bibitem{meyerET2001} Meyer V, Rowe M A, Kielpinski D, Sackett C A, Itano W M, Monroe C, and Wineland D J 2001 Experimental demonstration of entanglement-enhanced rotation angle estimation using trapped ions \emph{Phys.
+Rev.
+Lett.} {\bf 86} 5870
+
+\bibitem{liebfriedET2004} Leibfried D, Barrett M D, Schaetz T, Britton J, Chiaverini J, Itano W M, Jost J D, Langer D, Wineland D J 2004 Toward Heisenberg-limited spectroscopy with multiparticle entangled states \emph{Science} {\bf 304} 1476
+
+\bibitem {roosET2006} Roos C F, Chwalla M, Kim K, Riebe M, and Blatt R Designer atoms for
+quantum metrology 2006 \emph{Nature} {\bf 443}, 316
+
+\bibitem{LiET2009} Li Y, Treutlein P, Reichel J, and Sinatra A 2009 Spin squeezing in a bimodal condensate: spatial dynamics and particle losses \emph{Eur. Phys. J.} B {\bf 68} 365
+
+\bibitem{jascula2010} Jaskula J C, Bonneau M, Partridge G B, Krachmalnicoff V, Deuar P, Kheruntsyan K V, Aspect A, Boiron D, and Westbrook C I 2010 Sub-Poissonian Number Differences in Four-Wave Mixing of Matter Waves \emph{Phys.
+Rev.
+Lett.} {\bf 105} 190402
+
+\bibitem{doeringET2010} D{\"{o}}ring D, McDonald G, Debs J E, Figl C, Altin P A, Bachor H-A, Robins N P, and Close J D 2010 Quantum-projection-noise-limited interferometry with coherent atoms in a Ramsey-type setup \emph{Phys.
+Rev.} A {\bf 81} 043633
+
+\bibitem{davisET2008} Davis M J, Thwaite S J, Olsen M K, and Kheruntsyan K V 2008 Pairing mean-field theory for the dynamics of dissociation of molecular Bose-Einstein condensates \emph{Phys.
+Rev.} A {\bf 77} 023617
+
+\bibitem{savage2007} Savage C M and Kheruntsyan K V 2007 Spatial Pair Correlations of Atoms in Molecular Dissociation \emph{Phys. Rev. Lett.} {\bf 99} 220404
+
+\bibitem{pu2000} Pu H and Meystre P 2000 Creating Macroscopic Atomic Einstein-Podolsky-Rosen States from Bose-Einstein Condensates \emph{Phys. Rev. Lett.} {\bf 85} 3987
+
+\bibitem{sorensen2001} S{\"{o}}rensen A, Duan L-M, Cirac J I, Zoller P 2001 Many-particle entanglement with Bose-Einstein condensates \emph{Nature} {\bf 409} 63
+
+\bibitem{haine2005} Haine S A and Hope J J 2005 Outcoupling from a Bose-Einstein condensate with squeezed light to produce entangled atom laser beams, \emph{Phys. Rev.} A {\bf 72} 033601
+
+\bibitem{zhao2007} Zhao B, Chen Z, Pan J, Schmiedmayer J, Recati A, Astrakharchik G E, Calarco T 2007 Deterministic Entanglement via Molecular Dissociation in Integrated Atom Optics \emph{Phys.
+Rev.} A {\bf 75} 042312
+
+\bibitem{esteve2008} J. Est{\`{e}}eve, C. Gross, A. Weller, S. Giovanazzi, and M. Oberthaler 2008 Squeezing and entanglement in a Bose-Einstein condensate \emph{Nature} {\bf 455} 1216 (2008)
+
+\bibitem{maussang2010} Maussang, G.E.
+Marti, T.
+Schneider, P.
+Treutlein, Y.
+Li, A.
+Sinatra, R.
+Long, J.
+Est{\`{e}}ve, and J.
+Reichel 2010 Enhanced and reduced atom number fluctuations in a BEC splitter \emph{Phys.
+Rev.
+Lett.} {\bf 105} 080403
+
+\bibitem{sinatra2012} A. Sinatra, J.-C. Dornstetter, Y. Castin 2012 Spin squeezing in Bose-Einstein condensates: Limits imposed by decoherence and non-zero temperature \emph{Front. Phys.} {\bf 7} 86
+
+\bibitem{perrin2008} A.
+Perrin, C.
+M.
+Savage, D.
+Boiron, V.
+Krachmalnicoff, C.
+I.
+Westbrook and Kheruntsyan K V 2008 Atomic four-wave mixing via condensate collisions \emph{New J.
+Phys.} {\bf 10} 045021
+
+\bibitem{ogren2010} M. {\"{O}}gren and Kheruntsyan K V 2010 Role of spatial inhomogeneity in dissociation of trapped molecular condensates \emph{Phys. Rev. A} {\bf 82} 013641
+
+\bibitem{ogren2009} M. {\"{O}}gren and Kheruntsyan K V 2009 Atom-atom correlations in colliding Bose-Einstein condensates \emph{Phys. Rev. A} {\bf 79} 021606
+
 \bibitem{Hadzibabic2004} Hadzibabic Z, Stock S, Battelier B, Bretin V and Dalibard J 2004 Interference of an array of independent Bose-Einstein Condensates {\it Phys. Rev. Lett.} {\bf 93} 180403
-\bibitem{johnssonET2007} Johnsson M T and Haine S A 2007 Generating squeezing in an atom laser through self-interaction {\it Phys. Rev. Lett.} {\bf 99} 010401
 
-\bibitem{dallET2009} Dall R G, Byron L J, Truscott A G, Dennis G R, Johnsson M T and Hope J J 2009 Paired-atom laser beams created via four-wave mixing {\it Phys. Rev. A} {\bf 79} 011601(R)
+\bibitem{johnssonET2007} Johnsson M T and Haine S A 2007 Generating squeezing in an atom laser through self-interaction \emph{Phys. Rev. Lett.} {\bf 99} 010401
 
-\bibitem{haineET2009} S. A. Haine and M. T. Johnsson 2009 Dynamic scheme for generating number squeezing in Bose-Einstein condensates
-through nonlinear interactions \emph{Phys. Rev. A} {\bf 80} 023611.
+\bibitem{dallET2009} Dall R G, Byron L J, Truscott A G, Dennis G R, Johnsson M T and Hope J J 2009 Paired-atom laser beams created via four-wave mixing \emph{Phys. Rev.} A {\bf 79} 011601(R)
 
-\bibitem{dennisET2010} Dennis G R and Johnsson M T 2010 Generation of directional coherent matter beams through dynamical instabilities in Bose-Einstein condensates {\it Phys. Rev. A} {\bf 82} 033615
+\bibitem{haineET2009} S. A. Haine and M. T. Johnsson 2009 Dynamic scheme for generating number squeezing in Bose-Einstein condensates through nonlinear interactions \emph{Phys. Rev.} A {\bf 80} 023611
+
+\bibitem{dennisET2010} G.
+R.
+Dennis and M.
+T.
+Johnsson 2010 Generation of directional coherent matter beams through dynamical instabilities in Bose-Einstein condensates \emph{Phys.
+Rev.} A {\bf 82} 033615
 
@@
 Walls, and R.
-Graham 2010 Dynamical quantum noise in trapped Bose-Einstein condensates {\it Phys.
-Rev.
-A} {\bf 58}, 4824.
+Graham 2010 Dynamical quantum noise in trapped Bose-Einstein condensates \emph{Phys.
+Rev.} A {\bf 58} 4824
 
-\bibitem{dennis2013} Dennis G R, Hope J J and Johnsson M T 2013 XMDS2: Fast, scalable simulation of coupled stochastic partial
-differential equations {\it Comput. Phys. Commun.} {\bf 184} 201--208
+\bibitem{dennis2013} G.
+R.
+Dennis, J.
+J.
+Hope and M.
+T.
+Johnsson 2013 XMDS2: Fast, scalable simulation of coupled stochastic partial differential equations {\it Comput.
+Phys.
+Commun.} {\bf 184} 201--208
 
-\bibitem{louisell} W. H. Louisell \emph{Quantum Statistical Properties of Radiation}, (Wiley 1990).
+\bibitem{louisell} W. H. Louisell \emph{Quantum Statistical Properties of Radiation}, (Wiley 1990)
 
-\bibitem{gross2010} Gross C, Zibold T, Est{\`{e}}ve J and Oberthaler M K 2010 Nonlinear atom interferometer surpasses classical
-precision limit {\it Nature} {\bf 464} 1165--1169
+\bibitem{gross2010} Gross C, Zibold T, Est{\`{e}}ve J and Oberthaler M K 2010 Nonlinear atom interferometer surpasses classical precision limit \emph{Nature} {\bf 464} 1165--1169
 
-\bibitem{PethickSmith} C.J.~Pethick and H.~Smith, \emph{Bose-Einstein Condensation in Dilute Gases} (Cambridge University Press, Cambridge, 2002).
+\bibitem{PethickSmith} Pethick C J and Smith H, \emph{Bose-Einstein Condensation in Dilute Gases} (Cambridge University Press, Cambridge, 2002).
 

### 2013-07-09 large_number_squeezing_with_atoms.tex
Results conclusions viable draft

@@
 
-\title{Large squeezing in multicomponent trapped Bose-Einstein condensates}
+\title{Large squeezing in Bose-Einstein condensates with large numbers of atoms}
 
@@
 We examine these processes, and suggest methods to mitigate them.
-We conclude that high levels of squeezing with large numbers of atoms is feasible, but can realistically only be achieved in traps with even density profiles in at least two dimensions.
+We conclude that high levels of squeezing with large numbers of atoms is feasible, but can realistically only be achieved in traps with particular trap geometries.
 We also introduce a method of maximising the measurable squeezing by using a pi-pulse during the process to improve spatial mode-matching.
@@
 \label{sectionIntroduction}
-\begin{itemize}
-  \item General non-classical state generation with atoms
-  \item Talk about number squeezing with atoms
-  \item Link to interferometry
-  \item Motivate why large number squeezing
-  \item Point out no one has done it before, only low number squeeze
-  \item Lay out what sections we have and what we've done
-\end{itemize}
-
-Quantum optics led to creating nonclassical states after the laser was invented \cite{???}.
-The experimental realization of the BEC \cite{???} led to the ability to start recreating these types of states with massive particles.
-
-One of the advantages of atoms over photons is that they interact with each other, unlike photons, which allows nonlinear interaction effects without all the fluffery that optics requires, like crystals etc.
-
-It has been recognized that the intrinsic nonlinearities of atoms could lead to effects like quadrature squeezing \cite{kheruntsyanET2005, johnssonET2007, haineET2009, doeringET2010}.
-Can also get number squeezing using spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissocation of BECs \cite{kheruntsyanET2002, davisET2008, savage2007, ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, or double-well potentials \cite{esteve2008, maussang2010} or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}.
-
-{\bf Note that the FWM stuff is using the s-wave scatter sphere from condensate collisions, so pretty useless for interferometry}
-
-The creation of these non-classical atomic states has relevence for fundamental tests of quantum mechanics such as the EPR paradox \cite{pu2000, sorensen2001, haine2005,zhao2007, ogren2010}, as well as atom inteferometry \cite{sorensenET1999, kuzmichET2000, meyerET2001, liebfriedET2004, roosET2006, doeringET2010, gross2010}
-
-While these techniques have resulted in impressive levels of experimentally detectable number squeezing \cite{gross2010}, they have been limited to small numbers of atoms, typically a few hundred.
-
-To really win in the interferometry game, you need good squeezing for large numbers of atoms.
-The reason people have used small numbers of atoms is that it allows you to move into the single-mode regime, where the trap is tight enough such that all higher modes are frozen out \cite{LiET2009}
-
-If this is not the case multimode stuff needs to be considered, and that's what we do in this paper.
-
-Most previous simulations of this sort of stuff tend to be in 1D or 2D for simplicity \cite{kheruntsyanET2002, kheruntsyanET2005, johnssonET2007, haineET2009}.
-They reduce the problem from 3D to 1D by integrating out one or two dimensions \cite{???}.
-Semi-classically this isn't a problem, because you can assume the modal dynamics have symmetry.
-However as we show in this paper, when dealing with QFT problems this naive approach can give you the wrong answer.
-We can do full 3D simulations because we're cool and we have XMDS2 \cite{dennis2013}.
+
+Quantum degenerate ultracold atomic sources have been used to study fundamental processes such as quantum phase transitions \cite{Leggett2001,Buluta2009}, quantum non-equilibrium thermodynamics \cite{Kinoshita:2006,Barnett:2011} and entanglement of massive particles \cite{Kheruntsyan:2005}.
+Their ability to be highly isolated from their environment and controlled precisely by optical, magnetic and rf fields makes them an ideal platform for engineering and manipulating non-trivial quantum states.
+In quantum optics, the most common way to produce entangled states has been through squeezing, and it has been recognised that the intrinsic nonlinearities of atoms naturally lead to quadrature squeezing \cite{kheruntsyanET2005, johnssonET2007, haineET2009, doeringET2010}.
+Number squeezing can be achieved with spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissociation of Bose-Einstein Condensates (BECs) \cite{kheruntsyanET2002, davisET2008, savage2007, ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, double-well potentials \cite{esteve2008, maussang2010}, or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}.
+The creation of these non-classical atomic states has relevance for fundamental tests of quantum mechanics such as the EPR paradox \cite{pu2000, sorensen2001, haine2005,zhao2007, ogren2010}, as well as atom inteferometry \cite{sorensenET1999, kuzmichET2000, meyerET2001, liebfriedET2004, roosET2006, doeringET2010, gross2010}.
+While these techniques have resulted in significant experimentally detectable number squeezing \cite{gross2010}, they have been limited to small numbers of atoms, typically a few hundred.
+
+Ultracold atomic interferometers are also reaching state of the art sensitivities in precision measurements \cite{Robins2012}.
+While all current atom interferometry is limited by technical noise sources, it appears that BEC-based coherent sources offer the best chance of overcoming some of these issues, for the same reasons that optical lasers are typically preferred for optical interferometry \cite{Szigeti2012}.
+If these technical noise sources can be controlled, atom interferometers will be limited by atomic shot noise, which leads to a sensitivity that scales as the square root of the atomic flux.
+From that point, sensitivity will only be improved by increasing the atomic flux, and/or using non-classical quantum states such as squeezed states to go below the shot noise limit.
+Producing squeezed states for metrological purposes is therefore only relevant when it can be achieved in a context of large numbers of atoms.
+This paper examines the generation of atomic squeezing using the inherent Kerr-like nonlinearities of ultracold atoms, and demonstrates a trap geometry in which metrologically relevant levels of squeezing are present in trap geometries that are compatible with large atomic number.
+
+The creation, detection and application of squeezing requires precise mode-matching, and is therefore very sensitive to uncontrolled coupling between spatial modes of the atomic field.
+Atomic squeezing experiments have typically attempted to operate in the single-mode regime, where the trap is tight enough such that all higher modes are frozen out \cite{LiET2009}.
+However, tight traps limit the total number of atoms, as the 3-body recombination rate places an effective maximum atomic density in the gas, and well before that limit, the interatomic interactions break the single-mode operation.
+Both of these effects limit the number of atoms that can be effectively squeezed, so we will examine a broader range of trap geometries, and calculate the dynamics in the presence of multiple spatial modes.
+Many similar previous simulations have been performed in one or two spatial dimensions, by integrating out some of the remaining dimensions \cite{kheruntsyanET2002, kheruntsyanET2005, johnssonET2007, haineET2009}.
+For simulations of semiclassical fields this can be a reasonable approximation, but it can ignore spontaneous scattering processes in full quantum field theory calculations, so we take care to include full 3D simulations.
+
+In section \ref{secTwoModeAnalytic} we examine the ideal behaviour when the system is reduced to a single spatial mode, extracting a full analytic solution to the squeezing for arbitrary \textbf{whatever it is we extended, along with reference of what was covered previously.}  In section \ref{sec:MMdescription} we explore the idea of how multimode behaviour will become important in physical systems, and introduce our model for simulating higher dimensional trap geometries.
+Sec.~\ref{sec:1Dsim} analyses one-dimensional models, highlighting some of the issues regarding optimising squeezing parameters and mode-matching in squeezing experiments.
+Sec.~ref{sec:pipulse} shows that the application of a $\pi$-pulse during the process can dramatically improve mode-matching issues.
+Sec.~\ref{sec:2D3Dsqueezing} investigates measurable squeezing in 2D and 3D, and uses a Bogoliubov analysis to explain the  scaling with dimensionality.
+Sec.~\ref{sectionConclusion} describes the conclusion that large volume, flat-bottomed, three-dimensional traps, which are obviously compatible with large atom number, can produce extremely high degrees of squeezing.
 
@@
 Our goal is to describe the number squeezing possibilities of a BEC comprising two relevant internal states denoted by $|a\rangle$ and $|b\rangle$.
-The Hamiltonian for a pair of coupled atomic internal modes is the spatial integral of the Hamiltonian density, which is given by
-\begin{equation}
-\hat{\mathcal{H}} = \sum_{j} \hat{\psi}_j^{\dagger}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_j\right)\hat{\psi}_j 
-          + \sum_{i j}\frac{U_{i j}}{2} \hat{\psi}_i^{\dagger} \hat{\psi}_j^{\dagger} \hat{\psi}_j \hat{\psi}_i
-          + \kappa \hat{\psi}_a^{\dagger} \hat{\psi}_b + \kappa^* \hat{\psi}_b^{\dagger}  \hat{\psi}_a
+The Hamiltonian for a pair of coupled atomic internal modes is the spatial integral of the Hamiltonian density:
+\begin{equation}
+\hat{\mathcal{H}} = \sum_{j\in \{a,b\}} \hat{\psi}_j^{\dagger}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_j\right)\hat{\psi}_j 
+          + \sum_{i,j}\frac{U_{i j}}{2} \hat{\psi}_i^{\dagger} \hat{\psi}_j^{\dagger} \hat{\psi}_j \hat{\psi}_i
+          + \kappa \hat{\psi}_a^{\dagger} \hat{\psi}_b + \kappa^* \hat{\psi}_b^{\dagger}  \hat{\psi}_a,
 \label{eqFieldHamiltonian}
@@
 We begin by considering a simplified version of the problem, assuming that only one of these spatial modes is relevant for each of the condensate's components.
-This is done by writing the multimode atomic field operators in terms of a set of basis functions $\hat{\psi}_a({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{a}_{n}(t)$ and $\hat{\psi}_b({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{b}_{n}(t)$, where $\hat{a}_{n}$ and $\hat{b}_{n}$ annihilate a particle with normalised spatial mode $u_n$ and internal state $|a\rangle$ or $|b\rangle$ respectively.
+This is done by writing the multimode atomic field operators in terms of a set of basis functions $\hat{\psi}_a({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{a}_{n}(t)$ and $\hat{\psi}_b({\mathbf{r}},t) = \sum_n u_{n}({\mathbf{r}}) \hat{b}_{n}(t)$, where $\hat{a}_{n}$ and $\hat{b}_{n}$ annihilate a particle with normalised spatial mode $u_n({\mathbf{r}})$ and internal state $|a\rangle$ or $|b\rangle$ respectively.
 With this expansion, the nonlinear term in the multimode Hamiltonian for the self-interaction of the $|a\rangle$ internal state becomes
@@
 The system is prepared with all the population in $|a\rangle$, and vacuum in $|b\rangle$.
-We choose state $|a\rangle$ to be occupied as a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number in our system.
-As our simulation and experiment are insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, which is also equivalent to mixture of Poissonian-distributed number states.
+We choose state $|a\rangle$ to be occupied as a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number.
+As our simulation and experiment are insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, or a mixture of Poissonian-distributed number states.
 This state is consistent with BEC coherence experiments \cite{Hadzibabic2004}.
 
-At time $t=0$, the coupling $\Omega$ is switched on and applied until time $t=t_1$, at which point it is turned off, resulting in a portion of the population being transferred into mode $|b\rangle$.
-We assume this transfer process fast relative to the nonlinear energy time scale, ensuring that, as the initial states were coherent states, after the population transfer both $|a\rangle$ and $|b\rangle$ are still described by coherent states with mean number $\langle \hat{a}^{\dagger} \hat{a} \rangle = n_a$ and $\langle \hat{b}^{\dagger} \hat{b} \rangle = n_b$ respectively.
-
-The coupling $\Omega$ is then switched off until time $t_2$.
-During this period the atoms interact solely through the nonlinear terms, with no population being transferred.
-Due to the form of the Hamiltonian, the number variance remains unchanged, but the quadrature variances in the two modes change over time, which can lead to quadrature squeezing \cite{johnssonET2007}.
-
+At time $t=0$, the coupling $\Omega$ is applied until time $t=t_1$, resulting in a portion of the population being transferred into mode $|b\rangle$.
+We assume this transfer process is fast relative to the nonlinear energy time scale, ensuring that, as the initial states were coherent states, after the population transfer both $|a\rangle$ and $|b\rangle$ are still described by coherent states with mean number $\langle \hat{a}^{\dagger} \hat{a} \rangle = n_a$ and $\langle \hat{b}^{\dagger} \hat{b} \rangle = n_b$ respectively.
+The coupling $\Omega$ is then switched off until time $t_2$, while the atoms interact solely through the nonlinear terms.
 After this hold time $\tau_{\mathrm{hold}} = t_2 - t_1 $, the coupling $\Omega$ is switched back on until time $t_3$, with a phase shift $\phi$ compared to its first application.
@@
 This experiment can also be interpreted as a Ramsey interferometer with a final beam splitter phase of $\phi$.
-
 The entire sequence of pulses and the resulting populations are shown schematically in Figure~\ref{figPulseScheme}.
@@
 This should enable higher squeezing.
-We cannot make strong conclusions without also considering the effects of coupling between spatial modes, however, which leads to the third potential problem: Coupling between modes is inevitable in the presence of a nonlinearity, meaning that the modes cannot be analyzed independently.
+We cannot make strong conclusions without also considering the effects of coupling between spatial modes, however, which leads to the third potential problem: coupling between modes is inevitable in the presence of a nonlinearity, meaning that the modes cannot be analyzed independently.
 Coupling between modes will disturb mode-matching, as well as mixing fields with different phase evolution, which can rapidly destroy squeezing.
@@
 \begin{eqnarray}
-\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} \left( \left|\phi_{a}\right|^2 -\frac{1}{dV}\right) + U_{a b} \left( \left|\phi_b\right|^2 -\frac{1}{2dV}\right)  \right)\phi_a   + \kappa \phi_b  \nonumber \\
-\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} \left( \left|\phi_{b}\right|^2 - \frac{1}{dV} \right) + U_{a b} \left( \left|\phi_a\right|^2 -\frac{1}{2dV} \right)   \right)\phi_b   + \kappa^* \phi_a
+\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} n_a + U_{a b} \tilde{n}_b  \right)\phi_a   + \kappa \phi_b  \nonumber \\
+\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} n_b + U_{a b} \tilde{n}_a   \right)\phi_b   + \kappa^* \phi_a
+%\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} \left( \left|\phi_{a}\right|^2 -\frac{1}{dV}\right) + U_{a b} \left( \left|\phi_b\right|^2 -\frac{1}{2dV}\right)  \right)\phi_a   + \kappa \phi_b  \nonumber \\
+%\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} \left( \left|\phi_{b}\right|^2 - \frac{1}{dV} \right) + U_{a b} \left( \left|\phi_a\right|^2 -\frac{1}{2dV} \right)   \right)\phi_b   + \kappa^* \phi_a
   \label{eqStochasticEquations}
 \end{eqnarray}
-where $dV$ is the volume element of the grid on which the simulation is carried out, and the terms proportional to $1/dV$ correspond to a vacuum corrections.
+where $n_j=\left|\phi_{j}\right|^2 -\frac{1}{dV}$, $\tilde{n}_j=\left|\phi_{j}\right|^2 -\frac{1}{2 dV}$, $dV$ is the volume element of the grid on which the simulation is carried out, and the terms proportional to $1/dV$ correspond to vacuum corrections.
 It is a peculiarity of the Wigner representation that, up to the vacuum correction terms, Eqs.~(\ref{eqStochasticEquations}) look identical to those of the coupled semi-classical nonlinear Schr{\"{o}}dinger equation.
@@
 
-\section{Effects of trap geometries on squeezing in 1D}
+\section{Effects of trap geometries on squeezing in 1D} \label{sec:1Dsim}
 
@@
 
-\section{Using a $\pi$-pulse to improve mode-matching}
+\section{Using a $\pi$-pulse to improve mode-matching} \label{sec:pipulse}
 Modal mismatch is generated by the different atomic potentials seen by the two internal states.
@@
 
-These points are illustrated in Fig.~\ref{piPulseFig}, which shows the best squeezing obtainable from a specific system with and without the $\pi$-pulse, as well as the extra squeezing that can be obtained by allowing a longer hold time as well as the $\pi$-pulse.
+These effects are illustrated in Fig.~\ref{piPulseFig}, which shows the best squeezing obtainable from a specific system with and without the $\pi$-pulse, as well as the extra squeezing that can be obtained by allowing a longer hold time as well as the $\pi$-pulse.
 \begin{figure}
@@
 
-To minimise the damaging effects of the non-condensed fraction it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation.
-To better understand when $N_{\mathbf{k}\neq 0}/N$ will be small, we approximate the sum in \eqref{eqCondensateDepletion} with an integral.
+To minimise the damaging effects of the non-condensed fraction it is better to operate in a regime where the total occupation of the non-zero momentum modes $N_{\mathbf{k}\neq 0}$ is small, which is also fortuitously the regime in which the Bogoliubov theory is a good approximation, so we can investigate in more detail.
 In 1D Eq.~\eqref{eqCondensateDepletion} is
@@
 \end{align}
-We wish to approximate this sum by the integral
+We approximate this sum by the integral
 \begin{align}
@@
 
-Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$ and defining $\Lambda_a = n_a \chi_{aa} \tau_\text{hold}$ the non-condensed fraction \eqref{eqCondensateDepletionFirstIntegralApproximation} can be written as
+Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$ and defining $\Lambda_a = n_a \chi_{aa} \tau_\text{hold}$, the non-condensed fraction \eqref{eqCondensateDepletionFirstIntegralApproximation} can be written as
 \begin{align}
@@
 
-Figures~\ref{fig1D_squeezing_and_bog_scaling}--\ref{fig3D_Bog_mode_occupation_scaling} show the non-condensed fraction predicted by the Bogoliubov analysis compared to that predicted by a full numerical simulation, in one, two and three dimensions, as a function of the trap size.
+Figures~\ref{fig1D_squeezing_and_bog_scaling}--\ref{fig3D_Bog_mode_occupation_scaling} show the non-condensed fraction predicted by the Bogoliubov analysis compared to that predicted by a full numerical simulation in one, two and three dimensions, as a function of the trap size.
 Also shown in the 1D and 2D cases is the maximum amount of number squeezing obtainable, as predicted by the numerical simulation.
-The full squeezing scaling curve is not shown in 3D, as for the smaller trap sizes the TWA method breaks down, and results are unreliable; for the larger trap sizes with negligible non-condensate fractions, however, the simulation was carried out and a relative number squeezing of 12.6dB was found.
+The full squeezing scaling curve is not shown in 3D, as for the smaller trap sizes the TWA method breaks down, and results are unreliable; for the larger trap sizes with negligible non-condensate fractions, however, the TWA simulation was possible in the large trap volume limit, where the non-condensed fraction was low.
 The traps are assumed to have a constant potential, which minimises the effects of the multimode issues described in section \ref{sec:MMdescription}.
 We see that the squeezing does indeed degrade at large trap sizes in 1D, but remains constant in 2D.
-For small trap sizes, the squeezing in 1D aymptotically approaches that of the ideal 0D system.
+For small trap sizes, the squeezing in 1D asymptotically approaches that of the ideal 0D system.
 
@@
 
-In Fig.~\ref{fig:3DBogScaling} we see that the number of accessible states for three-dimensional traps actually decreases with volume in 3D.
+The simulation shown in Fig.~\ref{fig:3DBogScaling} verified that the number of accessible states for three-dimensional traps decreases with volume in 3D.
 This is because the energy of the ground state of the BEC is reducing faster than the density of states at low energies is increasing.
-\textbf{Create these figures (by combining old figures) and check both for latest updates.
-Confirm that `simulation' (blue line) means TWA, as opposed to Graham's prediction.}  The non-smooth nature of the Bogoliubov calculation is due to the coarse grid, which leads to granularity as the scale factor is changed.
-
-%Some notes on the Bog scaling plots below: Experimentalist's scaling works as follows: V $\rightarrow$ r V; hold time $\rightarrow$ r*(hold time).
-In order to get enough lattice points to cover the momentum modes up to energy mu, the scaling goes as: 1D lattice$\rightarrow$lattice*$\sqrt{r}$ (1D); lattice$\rightarrow$lattice (2D); lattice$\rightarrow$lattice*$r^{-1/6}$.
+This scaling of the estimated Bogoliubov fraction in 3D suggests that strong squeezing may actually be found in very large, weakly trapped condensates, provided the trapping potentials are box-like.
+Although systematic analysis of squeezing across a wide range of large 3D traps and interaction strengths is impossible due to the breakdown of the TWA method, fortunately the method is accurate when the highest squeezing is produced.
+While inter-state scattering lengths of most atomic species and states are not well characterised, extremely high squeezing can be found in systems with large, three-dimensional box-like potentials, and asymmetry in the three scattering lengths.
+A sample 3D simulation with this geometry \textbf{with some lengths and some atom number}, and only a single non-zero scattering length, gives a result of 15dB of squeezing.
+This system produces an extremely low non-condensed fraction of $0.0004$, in agreement with the Bogoliubov calculations.
+%From the "winner winner chicken dinner" email:  Simulating a 3D BEC in a box with the same parameters as the experimental paper works well.
+Bog mode fraction is around 4e-4, and we get squeezing of around 0.026.
+This is assuming we can set two scattering lengths to zero, and have the third equal to that of Rb87.
 
@@
   \end{subfigure}
-\caption{1D scaling as function of trap size, $L$.
-(Left) The proportion the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+\caption{Non-condensed fraction and squeezing as function of trap size $L$ in a 1D trap.
+(Left) The proportion of the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
 (Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
@@
   \end{subfigure}
-\caption{2D scaling as function of trap size, L.
+\caption{Non-condensed fraction and squeezing as function of trap size $L$ in a 2D trap.
 (Left) The proportion the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
@@
    \includegraphics[width=7cm]{figures/3D_Bog_mode_occupation_scaling.eps}
-\caption{3D scaling as function of trap size, $L$.
+\caption{Non-condensed fraction and squeezing as function of trap size $L$ in a 3D trap.
 (Left) The proportion the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
@@
 
-\textbf{Obviously 3D TWA simulation can be done in some limit - the one that produced the 15db result.  Do we say something about this?}
-
-This scaling of the estimated Bogoliubov fraction in 3D suggests that strong squeezing may actually be found in very large, weakly trapped condensates, provided the trapping potentials are box-like.
-This can be confirmed with a TWA simulation, although systematic analysis of squeezing across a wide range of large 3D traps and interaction strengths is impossible due to the breakdown of the TWA method.
-Fig.~\ref{figVolumeSqueezingEffects} shows a result of 15dB of squeezing achievable in \textbf{some parameter regime}.
-This system produces an extremely low non-condensed fraction of $0.0004$, as expected from the Bogoliubov calculations, and squeezing that is 150\% better than Bogoliubov lower bound.
-
-%Simulating a 3D BEC in a box with the same parameters as the experimental paper works well.
-Bog mode fraction is around 4e-4, and we get squeezing of around 0.026.
-This is assuming we can set two scattering lengths to zero, and have the third equal to that of Rb87.
-%Interestingly, Graham's bog theory predicts squeezing of 0.0398 (assuming I'm doing it correctly), but both the single-mode analytic answer and the simulation give 0.026.
-
-%Just in case this was the real version of the best data:
+%Old best attempt:
 
 %We finish this section by considering an example of the sort of number squeezing that is achieveable in a realistic physical system with large particle number.
-We take a Rb$^{85}$ condensate, which possesses a number of Feschbach resonances that allow us to modify one of the $U_{ij}$ scattering lengths without affecting the others.
+We take a Rb$^{85}$ condensate, which possesses a number of Feshbach resonances that allow us to modify one of the $U_{ij}$ scattering lengths without affecting the others.
 We assume a box potential in two dimensions with a confinement region $L=120\mu$m$\times 120\mu$m, and a tightly trapped third dimension with a trapping frequency $\omega=2 \pi \times 1500$Hz, resulting in a near 2D system.
@@
 
-We have presented an analytic solution to the two-mode model of atomic fields with a strong Kerr non-linearity.
-The three nonlinearities between and within the modes are not yet known for many atomic species, and often have sensitive dependence on external magnetic fields due to Feshbach resonances \cite{FeshbachOrReview}, so this solution allows for a rapid estimate of the atomic number difference squeezing that can be gained from these systems as this data is obtained.
+The large Kerr-type nonlinearities present between neutral atoms in BEC present the possibility of significant squeezing, but they also cause potentially complicated multimode behaviour.
+Furthermore, the nonlinearities between multiple internal atomic states are not yet known for many atomic species, and often have sensitive dependence on external magnetic fields due to Feshbach resonances \cite{FeshbachOrReview}.
+This paper analyses the requirements for producing useful squeezing for large atomic numbers.
+
+We have presented an analytic solution to the two-mode model of atomic fields with a strong Kerr non-linearity, allowing for a rapid estimate of the atomic number difference squeezing that can be gained from these systems as scattering length data is obtained.
+\textbf{Reiterate what is new in this analytic solution}.
 In general, it predicts arbitrarily good squeezing in the two-mode limit.
-\textbf{Revise that.}
-
-Operating in a parameter regime where there is truly only one spatial mode, which is essential for the analytic solution to be valid, requires extremely tight trapping potentials.
-Furthermore, due to the nonlinearities, it also comes with a strong limit on the number of atoms that can be trapped.
-Interferometric applications of squeezing become valuable at high atomic flux, so we investigate methods of mitigating multimode spatial effects that degrade the squeezing.
+We then examined trap geometries in which this limit can be approached.
 
@@
 
-Using a Bogoliubov analysis, it was shown that the coupling between different modes could be minimised in two regimes.
-The first was the tight trapping regime, where the system becomes effectively zero-dimensional, and the second was the weak trapping limit for 3D traps.
+Excluding multiple spatial modes on the basis of energy makes the analytic solution valid, but requires extremely tight trapping potentials.
+This also strongly limits the number of atoms that can be trapped, due to inelastic scattering rates.
+Using a Bogoliubov analysis, it was shown that the coupling between different modes could be minimised in a second regime, namely the \textit{weak} trapping limit for 3D traps.
 This latter regime is in fact ideally compatible with large atomic numbers, and shows the prospect of extremely high squeezing.
@@
 \ack
-Acknowledgements in here; funding sources etc.  Joe's Future Fellowship.  NCI?
+JH acknowledges the support of an ARC Future Fellowship.  \textbf{NCI?}
 
@@
 As this system is self-contained and global phase is unmeasurable, all phase measurements can only rely on the difference in phase between the two modes.
-This phase rotation is therefore irrelevent and will be ignored in the following analysis.
+This phase rotation is therefore irrelevant and will be ignored in the following analysis.
 
@@
 \end{align}
-where $\omega^0_\mathbf{k} = \hbar \mathbf{k}^2/2M$ is the frequency of a free particle, and $\omega_\mathbf{k} = \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}$ is the frequency of the Bogoliubov modes:
+where $\omega^0_\mathbf{k} = \hbar \mathbf{k}^2/2M$ is the frequency of a free particle, and $\omega_\mathbf{k} = \sqrt{\omega^0_\mathbf{k}(\omega^0_\mathbf{k} + 2 \chi_{aa} n_a)}$ are the frequencies of the Bogoliubov modes:
 \begin{align}

### 2013-07-10 large_number_squeezing_with_atoms.tex
Fixing minor typos.

@@
 We examine the feasibility of creating and measuring large relative number squeezing in multicomponent trapped Bose-Einstein condensates.
-In the absence of multimode effects, this squeezing can be arbitrarily large, but a range of processes limit the measurable squeezing in realistic trap configurations.
+In the absence of multimode effects, this squeezing can be arbitrarily large for arbitrarily large condensates, but a range of processes limit the measurable squeezing in realistic trap configurations.
 We examine these processes, and suggest methods to mitigate them.
@@
 In section \ref{secTwoModeAnalytic} we examine the ideal behaviour when the system is reduced to a single spatial mode, extracting a full analytic solution to the squeezing for arbitrary \textbf{whatever it is we extended, along with reference of what was covered previously.}  In section \ref{sec:MMdescription} we explore the idea of how multimode behaviour will become important in physical systems, and introduce our model for simulating higher dimensional trap geometries.
-Sec.~\ref{sec:1Dsim} analyses one-dimensional models, highlighting some of the issues regarding optimising squeezing parameters and mode-matching in squeezing experiments.
-Sec.~ref{sec:pipulse} shows that the application of a $\pi$-pulse during the process can dramatically improve mode-matching issues.
-Sec.~\ref{sec:2D3Dsqueezing} investigates measurable squeezing in 2D and 3D, and uses a Bogoliubov analysis to explain the  scaling with dimensionality.
-Sec.~\ref{sectionConclusion} describes the conclusion that large volume, flat-bottomed, three-dimensional traps, which are obviously compatible with large atom number, can produce extremely high degrees of squeezing.
+Section~\ref{sec:1Dsim} analyses one-dimensional models, highlighting some of the issues regarding optimising squeezing parameters and mode-matching in squeezing experiments.
+Section~\ref{sec:pipulse} shows that the application of a $\pi$-pulse during the process can dramatically improve mode-matching issues.
+Section~\ref{sec:2D3Dsqueezing} investigates measurable squeezing in 2D and 3D, and uses a Bogoliubov analysis to explain the  scaling with dimensionality.
+Section~\ref{sectionConclusion} describes the conclusion that large volume, flat-bottomed, three-dimensional traps, which are obviously compatible with large atom number, can produce extremely high degrees of squeezing.
 
@@
 \end{equation}
-When the system remains predominantly in one mode, we can approximate this sum as the single term $\hat{H} = \hbar\chi_{aa} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}/2$ where
-\begin{equation}
-\chi_{aa} = \frac{1}{\hbar} U_{aa} \int |u_{i}({\mathbf{r}})|^4 \, dV.
+When the system remains predominantly in spatial mode $u_0(\mathbf{r})$, we can approximate this sum as the single term $\hat{H} = \hbar\chi_{aa} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}/2$ where
+\begin{equation}
+\chi_{aa} = \frac{1}{\hbar} U_{aa} \int |u_{0}({\mathbf{r}})|^4 \, dV.
 \label{eqChiUequivalence}
@@
 The system is prepared with all the population in $|a\rangle$, and vacuum in $|b\rangle$.
-We choose state $|a\rangle$ to be occupied as a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number.
-As our simulation and experiment are insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, or a mixture of Poissonian-distributed number states.
+We choose state $|a\rangle$ to be initially a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number.
+As the system is insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, or a mixture of Poissonian-distributed number states.
 This state is consistent with BEC coherence experiments \cite{Hadzibabic2004}.
@@
 In Figure \ref{figTwoModeAnalyticExamples} we plot a solution for the normalized number variance in mode $|a\rangle$ as a function of $\phi$, choosing a typical set of nonlinearities.
-This example illustrates the basic features of the system: During the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=\pi / \Omega$, and has a series of minima.
+This example illustrates the basic features of the system: during the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=\pi / \Omega$, and has a series of minima.
 It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
@@
 In the multimode case, the effective nonlinearity of a mode is a function of the mode shape, as is shown by Eq.~(\ref{eqChiUequivalence}).
-The optimum hold time $\tau_{\mathrm{hold}}$ depends on the product of the effective mode nonlinearity and number number of particles in that mode, as does the recombination phase $\phi$.
+The optimum hold time $\tau_{\mathrm{hold}}$ depends on the product of the effective mode nonlinearity and the number of particles in that mode, as does the recombination phase $\phi$.
 Different modes in a multimode environment will typically have varying effective nonlinearity-particle number products and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.
@@
 These sample objects are the same size as the equivalent classical field.
-In our case, the only tractable phase space representation is the functional Wigner representation, which is well behaved when the quantum field is approximately Gaussian, as it is for coherent and squeezed states.
-Stochastic methods based on the functional Wigner representation have been used to model the behaviour of BECs and atom lasers \cite{johnssonET2007,dallET2009,dennisET2010}, as the Wigner representation allows the mapping of the system to a Fokker-Planck equation with positive semi-definite diffusion matrix (unlike the $P$-representation), and does not suffer exponential path-weighting problems that would constrain it to very short simulation intervals (as would the positive-$P$ representation).
+In our case, the only tractable phase space representation is the functional Wigner representation \cite{steelET1998}, which is well behaved when the quantum field is approximately Gaussian, as it is for coherent and squeezed states.
+Stochastic methods based on the functional Wigner representation have been used to model the behaviour of BECs and atom lasers \cite{johnssonET2007,dallET2009,dennisET2010}, as the Wigner representation allows the mapping of the system to a Fokker-Planck equation with positive semi-definite diffusion matrix (unlike the $P$-representation \cite{gardiner1991}), and does not suffer exponential path-weighting problems that would constrain it to very short simulation intervals (as would the positive-$P$ representation \cite{gardiner1991}).
 
@@
 A sample 3D simulation with this geometry \textbf{with some lengths and some atom number}, and only a single non-zero scattering length, gives a result of 15dB of squeezing.
-This system produces an extremely low non-condensed fraction of $0.0004$, in agreement with the Bogoliubov calculations.
+This system produces an extremely low non-condensed fraction of $4 \times 10^{-4}$, in agreement with the Bogoliubov calculations.
 %From the "winner winner chicken dinner" email:  Simulating a 3D BEC in a box with the same parameters as the experimental paper works well.

### 2013-07-11 large_number_squeezing_with_atoms.tex
Fix figure axis labels and font sizes

@@
     \centering
-    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_1.eps}
+    \includegraphics[width=6cm]{figures/dimensional_effects_on_squeezing_1.eps}
     \label{figDimensionalSqueezingEffects:sub1}
@@
     \centering
-    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_2.eps}
+    \includegraphics[width=6cm]{figures/dimensional_effects_on_squeezing_2.eps}
     \label{figDimensionalSqueezingEffects:sub2}

### 2013-07-11 large_number_squeezing_with_atoms.tex
Rewrite part of conclusion, fix/add a lot of references
Also add discussion of 15dB squeezing and how box traps have been experimentally demonstrated

@@
 Their ability to be highly isolated from their environment and controlled precisely by optical, magnetic and rf fields makes them an ideal platform for engineering and manipulating non-trivial quantum states.
-In quantum optics, the most common way to produce entangled states has been through squeezing, and it has been recognised that the intrinsic nonlinearities of atoms naturally lead to quadrature squeezing \cite{kheruntsyanET2005, johnssonET2007, haineET2009, doeringET2010}.
+In quantum optics, the most common way to produce entangled states has been through squeezing, and it has been recognised that the intrinsic nonlinearities of atoms naturally lead to quadrature squeezing \cite{Kheruntsyan:2005, johnssonET2007, haineET2009, doeringET2010}.
 Number squeezing can be achieved with spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissociation of Bose-Einstein Condensates (BECs) \cite{kheruntsyanET2002, davisET2008, savage2007, ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, double-well potentials \cite{esteve2008, maussang2010}, or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}.
@@
 
-Ultracold atomic interferometers are also reaching state of the art sensitivities in precision measurements \cite{Robins2012}.
-While all current atom interferometry is limited by technical noise sources, it appears that BEC-based coherent sources offer the best chance of overcoming some of these issues, for the same reasons that optical lasers are typically preferred for optical interferometry \cite{Szigeti2012}.
+Ultracold atomic interferometers are also reaching state of the art sensitivities in precision measurements \cite{robins:2013}.
+While all current atom interferometry is limited by technical noise sources, it appears that BEC-based coherent sources offer the best chance of overcoming some of these issues, for the same reasons that optical lasers are typically preferred for optical interferometry \cite{szigeti:2012}.
 If these technical noise sources can be controlled, atom interferometers will be limited by atomic shot noise, which leads to a sensitivity that scales as the square root of the atomic flux.
@@
 Both of these effects limit the number of atoms that can be effectively squeezed, so we will examine a broader range of trap geometries, and calculate the dynamics in the presence of multiple spatial modes.
-Many similar previous simulations have been performed in one or two spatial dimensions, by integrating out some of the remaining dimensions \cite{kheruntsyanET2002, kheruntsyanET2005, johnssonET2007, haineET2009}.
+Many similar previous simulations have been performed in one or two spatial dimensions, by integrating out some of the remaining dimensions \cite{kheruntsyanET2002, Kheruntsyan:2005, johnssonET2007, haineET2009}.
 For simulations of semiclassical fields this can be a reasonable approximation, but it can ignore spontaneous scattering processes in full quantum field theory calculations, so we take care to include full 3D simulations.
@@
 This two-mode analytic model allows arbitrarily good number squeezing, provided there are no limits to the number of particles in the system or to the hold time.
-The two-mode approximation can be valid in tight traps with very low particle number, and this result correctly predicts the squeezing demonstrated by Gross \textit{et al.} \cite{Gross2010}.
+The two-mode approximation can be valid in tight traps with very low particle number, and this result correctly predicts the squeezing demonstrated by Gross \textit{et al.} \cite{gross2010}.
 Unfortunately, for larger numbers of atoms in realistic traps there are more than two modes present, and multimode effects limit the level of squeezing that can be achieved.
@@
 The equation of motion for the functional Wigner distribution contains derivatives of third order which we assume to be negligible.
-This uncontrolled approximation is called the Truncated Wigner Approximation (TWA), and while it has been used widely on ultracold gases \cite{atomlaserWigner,orOtherUCGWigner}, care must be taken that the simulation remains valid.
+This uncontrolled approximation is called the Truncated Wigner Approximation (TWA), and while it has been used widely on ultracold gases \cite{johnssonET2007,johnssonHope:2007,haineET2009,dennisET2010,steelET1998}, care must be taken that the simulation remains valid.
 For our simulations we checked the results of our simulations against both the analytic two-mode solution, as well as checking for typical indications of TWA breakdown such as the appearance of negative densities in lightly populated modes.
@@
 \frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} n_b + U_{a b} \tilde{n}_a   \right)\phi_b   + \kappa^* \phi_a
-%\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} \left( \left|\phi_{a}\right|^2 -\frac{1}{dV}\right) + U_{a b} \left( \left|\phi_b\right|^2 -\frac{1}{2dV}\right)  \right)\phi_a   + \kappa \phi_b  \nonumber \\
-%\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} \left( \left|\phi_{b}\right|^2 - \frac{1}{dV} \right) + U_{a b} \left( \left|\phi_a\right|^2 -\frac{1}{2dV} \right)   \right)\phi_b   + \kappa^* \phi_a
   \label{eqStochasticEquations}
@@
 
-% End of Graham's content
-
 Figures~\ref{fig1D_squeezing_and_bog_scaling}--\ref{fig3D_Bog_mode_occupation_scaling} show the non-condensed fraction predicted by the Bogoliubov analysis compared to that predicted by a full numerical simulation in one, two and three dimensions, as a function of the trap size.
@@
 
-The simulation shown in Fig.~\ref{fig:3DBogScaling} verified that the number of accessible states for three-dimensional traps decreases with volume in 3D.
+The simulation shown in Fig.~\ref{fig3D_Bog_mode_occupation_scaling} verified that the number of accessible states for three-dimensional traps decreases with volume in 3D.
 This is because the energy of the ground state of the BEC is reducing faster than the density of states at low energies is increasing.
 This scaling of the estimated Bogoliubov fraction in 3D suggests that strong squeezing may actually be found in very large, weakly trapped condensates, provided the trapping potentials are box-like.
+Such box-like traps have already been demonstrated and shown to be capable of producing BECs \cite{gaunt2013}.
+
 Although systematic analysis of squeezing across a wide range of large 3D traps and interaction strengths is impossible due to the breakdown of the TWA method, fortunately the method is accurate when the highest squeezing is produced.
-While inter-state scattering lengths of most atomic species and states are not well characterised, extremely high squeezing can be found in systems with large, three-dimensional box-like potentials, and asymmetry in the three scattering lengths.
-A sample 3D simulation with this geometry \textbf{with some lengths and some atom number}, and only a single non-zero scattering length, gives a result of 15dB of squeezing.
-This system produces an extremely low non-condensed fraction of $4 \times 10^{-4}$, in agreement with the Bogoliubov calculations.
-%From the "winner winner chicken dinner" email:  Simulating a 3D BEC in a box with the same parameters as the experimental paper works well.
-Bog mode fraction is around 4e-4, and we get squeezing of around 0.026.
-This is assuming we can set two scattering lengths to zero, and have the third equal to that of Rb87.
+While inter-state scattering lengths of most atomic species and states are not well characterised, extremely high squeezing can be found in systems with large, three-dimensional box potentials, and asymmetry in the three scattering lengths.
+
+As a demonstration of the plausibility of such a scheme, we carried out a full 3D simulation using the geometry of the optical box potential described in Ref.~\cite{gaunt2013}, using the same potentials (a cylindrical ``box'' with length 63\,$\mu$m and diameter 30\,$\mu$m) and condensate number ($N=10^5$ $^{87}$Rb atoms).
+We set one scattering length to $5$\,nm and the others to zero: $a_{00}=5$\,nm, $a_{10}=a_{11}=0$.
+This system produced number squeezing of 15dB, and an extremely low non-condensed fraction of $4 \times 10^{-4}$.
+Both of these numbers are in agreement with the Bogoliubov calculations.
 
@@
 
-%\begin{figure}
-%  \centering
-%   \includegraphics[width=7cm]{figures/1D_squeezing_scaling.eps}
-%\caption{1D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
-% \label{fig1D_squeezing_scaling}
-%\end{figure}
-
-%\begin{figure}
-%  \centering
-%   \includegraphics[width=7cm]{figures/1D_Bog_mode_occupation_scaling.eps}
-%\caption{1D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
-% \label{fig1D_Bog_mode_occupation_scaling}
-%\end{figure}
-
-%\begin{figure}
-%  \centering
-%   \includegraphics[width=7cm]{figures/2D_squeezing_scaling.eps}
-%\caption{2D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
-% \label{fig2D_squeezing_scaling}
-%\end{figure}
-
-%\begin{figure}
-%  \centering
-%   \includegraphics[width=7cm]{figures/2D_Bog_mode_occupation_scaling.eps}
-%\caption{2D experimentalist's scaling. $N=1\times 10^5$, $L=100\mu$m.}
-% \label{fig2D_Bog_mode_occupation_scaling}
-%\end{figure}
-
 \begin{figure}
@@
 
-The large Kerr-type nonlinearities present between neutral atoms in BEC present the possibility of significant squeezing, but they also cause potentially complicated multimode behaviour.
-Furthermore, the nonlinearities between multiple internal atomic states are not yet known for many atomic species, and often have sensitive dependence on external magnetic fields due to Feshbach resonances \cite{FeshbachOrReview}.
-This paper analyses the requirements for producing useful squeezing for large atomic numbers.
-
-We have presented an analytic solution to the two-mode model of atomic fields with a strong Kerr non-linearity, allowing for a rapid estimate of the atomic number difference squeezing that can be gained from these systems as scattering length data is obtained.
+The large Kerr-type nonlinearities present between neutral atoms in BEC present the possibility of significant squeezing, but they also cause potentially complicated multimode behaviour when large BECs are used.
+Furthermore, the nonlinearities between multiple internal atomic states are not yet known for many atomic species, and often have sensitive dependence on external magnetic fields due to Feshbach resonances \cite{cornish:2000,altin:2010}.
+This paper analyses the requirements for producing useful squeezing for systems with large numbers of atoms.
+
+We have presented an analytic solution to the two-mode model of atomic fields with a strong Kerr non-linearity, allowing for a rapid estimate of the atomic number difference squeezing that can be obtained from these systems as scattering length data is obtained.
 \textbf{Reiterate what is new in this analytic solution}.
@@
 Using a constant potential in the trap (a `box' trap) allows the BEC density to be spatially uniform throughout the process.
-This is shown to alleviate the difficulties with choosing correct hold times and phases for the coupling pulses.
-Along with a pi-pulse midway during the holding period, this also alleviates the majority of the mode-matching issues for the final recombination of the two modes.
-The remaining issue with the existence of multiple spatial modes was the possibility of scattering into them by the nonlinearities.
-
-Excluding multiple spatial modes on the basis of energy makes the analytic solution valid, but requires extremely tight trapping potentials.
-This also strongly limits the number of atoms that can be trapped, due to inelastic scattering rates.
-Using a Bogoliubov analysis, it was shown that the coupling between different modes could be minimised in a second regime, namely the \textit{weak} trapping limit for 3D traps.
+This is shown to alleviate the difficulties with choosing correct hold times and phases for the coupling pulses, and also alleviates the majority of the mode-matching issues for the final recombination of the two modes.
+If such a box potential is not available, the application of a pi-pulse part way through the hold period can also improve mode matching by imparting the same modal dynamics to both states.
+
+Even in the best case, where a box potential is used, the existence of multiple possible momentum modes allows the possibility of cross-mode scattering via nonlinear processes, which can degrade or totally remove any number squeezing.
+This degradation is visible only in full 3D simulations, and not lower dimensional approximations, due to its dependence on the number of available modes.
+It is worth noting that despite their difficulty, in multimode systems with large atom number, full 3D simulations need to be carried out to accurately determine the presence of number or spin squeezing.
+
+Excluding these modes by making them energetically unfavourable can reduce the system to a single mode problem, making our analytic solution valid, but requires extremely tight trapping potentials.
+Such a regime also strongly limits the number of atoms that can be trapped, partly due to the rate at which the required trapping frequencies grow with atom number, and partly due to inelastic scattering rates.
+
+Using a Bogoliubov analysis, however, it was shown that the coupling between different modes could be minimised in a second regime, namely the \textit{weak} trapping limit for 3D traps.
 This latter regime is in fact ideally compatible with large atomic numbers, and shows the prospect of extremely high squeezing.
-A TWA simulation demonstrated a BEC trap with $15$\,dB of squeezing.
+A 3D TWA simulation using an experimentally demonstrated trapping potential yielded $15$\,dB of squeezing with $10^5$ particles.
 In practice, the only limit to the number of atoms in such traps would be the stability of the trap, as the BEC would become increasingly sensitive to fluctuations as the energy per particle is reduced as the volume increases.
@@
 \ack
-JH acknowledges the support of an ARC Future Fellowship.  \textbf{NCI?}
+JH acknowledges the support of an ARC Future Fellowship.
 
@@
 \begin{thebibliography}{10}
+
+\bibitem{Buluta2009} Buluta I and Nori F 2009 Quantum Simulators \emph{Science} {\bf 326} 108
+
 \bibitem{kheruntsyanET2002} Kheruntsyan K V and Drummond P D 2002 Quantum correlated twin atomic beams via photodissociation of a molecular Bose-Einstein condensate \emph {Phys.
@@
 
-\bibitem{kheruntsyanET2005} Kheruntsyan K V, Olsen M K and Drummond P D 2005 Einstein-Podolsky-Rosen correlations via dissociation of a molecular Bose-Einstein condensate \emph{Phys.
+\bibitem{Kheruntsyan:2005} Kheruntsyan K V, Olsen M K and Drummond P D 2005 Einstein-Podolsky-Rosen correlations via dissociation of a molecular Bose-Einstein condensate \emph{Phys.
 Rev.} A {\bf 66} 031602(R)
@@
 
-\bibitem{kuzmichET2000} Kuzmich A, Mandel L, and Bigelow N P 2000 Generation of spin squeezing via
-continuous quantum nondemolition measurement \emph{Phys. Rev. Lett.} {\bf 85} 1594
+\bibitem{kuzmichET2000} Kuzmich A, Mandel L, and Bigelow N P 2000 Generation of spin squeezing via continuous quantum nondemolition measurement \emph{Phys. Rev. Lett.} {\bf 85} 1594
 
@@
 
+\bibitem{robins:2013} Robins N P, Altin P A, Debs J E and Close J D 2013 Atom lasers: Production, properties and prospects for precision inertial measurement \emph{Phys. Reports} {\bf 529} 265
+
+\bibitem{szigeti:2012} Szigeti S S, Debs J E, Hope J J, Robins N P and Close J D 2012 Why momentum width matters for atom interferometry with Bragg pulses \emph{N. J. Phys.} {\bf 14} 023009
+
 \bibitem{doeringET2010} D{\"{o}}ring D, McDonald G, Debs J E, Figl C, Altin P A, Bachor H-A, Robins N P, and Close J D 2010 Quantum-projection-noise-limited interferometry with coherent atoms in a Ramsey-type setup \emph{Phys.
@@
 
-\bibitem{esteve2008} J. Est{\`{e}}eve, C. Gross, A. Weller, S. Giovanazzi, and M. Oberthaler 2008 Squeezing and entanglement in a Bose-Einstein condensate \emph{Nature} {\bf 455} 1216 (2008)
-
-\bibitem{maussang2010} Maussang, G.E.
-Marti, T.
-Schneider, P.
-Treutlein, Y.
-Li, A.
-Sinatra, R.
-Long, J.
-Est{\`{e}}ve, and J.
-Reichel 2010 Enhanced and reduced atom number fluctuations in a BEC splitter \emph{Phys.
+\bibitem{esteve2008} Est{\`{e}}eve J, Gross C, Weller A, Giovanazzi S, and Oberthaler M 2008 Squeezing and entanglement in a Bose-Einstein condensate \emph{Nature} {\bf 455} 1216
+
+\bibitem{maussang2010} Maussang K, Marti G E, Schneider T, Treutlein P, Li Y, Sinatra A, Long R, Est{\`{e}}ve J, and Reichel J 2010 Enhanced and reduced atom number fluctuations in a BEC splitter \emph{Phys.
 Rev.
@@
 
-\bibitem{sinatra2012} A. Sinatra, J.-C. Dornstetter, Y. Castin 2012 Spin squeezing in Bose-Einstein condensates: Limits imposed by decoherence and non-zero temperature \emph{Front. Phys.} {\bf 7} 86
-
-\bibitem{perrin2008} A.
-Perrin, C.
-M.
-Savage, D.
-Boiron, V.
-Krachmalnicoff, C.
-I.
-Westbrook and Kheruntsyan K V 2008 Atomic four-wave mixing via condensate collisions \emph{New J.
-Phys.} {\bf 10} 045021
-
-\bibitem{ogren2010} M. {\"{O}}gren and Kheruntsyan K V 2010 Role of spatial inhomogeneity in dissociation of trapped molecular condensates \emph{Phys. Rev. A} {\bf 82} 013641
-
-\bibitem{ogren2009} M. {\"{O}}gren and Kheruntsyan K V 2009 Atom-atom correlations in colliding Bose-Einstein condensates \emph{Phys. Rev. A} {\bf 79} 021606
-
-\bibitem{Hadzibabic2004} Hadzibabic Z, Stock S, Battelier B, Bretin V and Dalibard J 2004 Interference of an array of independent Bose-Einstein Condensates {\it Phys. Rev. Lett.} {\bf 93} 180403
+\bibitem{sinatra2012} Sinatra A, Dornstetter J-C, Castin Y 2012 Spin squeezing in Bose-Einstein condensates: Limits imposed by decoherence and non-zero temperature \emph{Front. Phys.} {\bf 7} 86
+
+\bibitem{perrin2008} Perrin A, Savage C M, Boiron D, Krachmalnicoff V, Westbrook C I and Kheruntsyan K V 2008 Atomic four-wave mixing via condensate collisions \emph{New J. Phys.} {\bf 10} 045021
+
+\bibitem{ogren2010} {\"{O}}gren M and Kheruntsyan K V 2010 Role of spatial inhomogeneity in dissociation of trapped molecular condensates \emph{Phys. Rev. A} {\bf 82} 013641
+
+\bibitem{ogren2009} {\"{O}}gren M and Kheruntsyan K V 2009 Atom-atom correlations in colliding Bose-Einstein condensates \emph{Phys. Rev. A} {\bf 79} 021606
+
+\bibitem{Hadzibabic2004} Hadzibabic Z, Stock S, Battelier B, Bretin V and Dalibard J 2004 Interference of an array of independent Bose-Einstein Condensates \emph{Phys. Rev. Lett.} {\bf 93} 180403
 
@@
 
-\bibitem{haineET2009} S. A. Haine and M. T. Johnsson 2009 Dynamic scheme for generating number squeezing in Bose-Einstein condensates through nonlinear interactions \emph{Phys. Rev.} A {\bf 80} 023611
-
-\bibitem{dennisET2010} G.
-R.
-Dennis and M.
-T.
-Johnsson 2010 Generation of directional coherent matter beams through dynamical instabilities in Bose-Einstein condensates \emph{Phys.
-Rev.} A {\bf 82} 033615
-
-\bibitem{gardiner1991} C. W. Gardiner, \emph{Quantum Noise}, (Springer-Verlag, Berlin 1991).
-
-\bibitem{steelET1998} M.
-J.
-Steel, M.
-K.
-Olsen, L.
-I.
-Plimak, P.
-D.
-Drummond, S.
-M.
-Tan, M.
-J.
-Collett, D.
-F.
-Walls, and R.
-Graham 2010 Dynamical quantum noise in trapped Bose-Einstein condensates \emph{Phys.
+\bibitem{johnssonHope:2007} Johnsson M T and Hope J J 2007 Multimode quantum limits to the linewidth of an atom laser \emph{Phys. Rev A} {\bf 75} 0432619
+
+\bibitem{haineET2009} Haine S A and Johnsson M T 2009 Dynamic scheme for generating number squeezing in Bose-Einstein condensates through nonlinear interactions \emph{Phys. Rev.} A {\bf 80} 023611
+
+\bibitem{dennisET2010} Dennis G R and Johnsson M T 2010 Generation of directional coherent matter beams through dynamical instabilities in Bose-Einstein condensates \emph{Phys. Rev.} A {\bf 82} 033615
+
+\bibitem{gardiner1991} Gardiner C W, \emph{Quantum Noise}, (Springer-Verlag, Berlin 1991).
+
+\bibitem{steelET1998} Steel M J, Olsen M K, Plimak L I, Drummond P D, Tan S M, Collett M J, Walls D F, and Graham R 2010 Dynamical quantum noise in trapped Bose-Einstein condensates \emph{Phys.
 Rev.} A {\bf 58} 4824
 
-\bibitem{dennis2013} G.
-R.
-Dennis, J.
-J.
-Hope and M.
-T.
-Johnsson 2013 XMDS2: Fast, scalable simulation of coupled stochastic partial differential equations {\it Comput.
-Phys.
-Commun.} {\bf 184} 201--208
-
-\bibitem{louisell} W. H. Louisell \emph{Quantum Statistical Properties of Radiation}, (Wiley 1990)
+\bibitem{dennis2013} Dennis G R, Hope J J and Johnsson M T 2013 XMDS2: Fast, scalable simulation of coupled stochastic partial differential equations \emph{Comput. Phys. Commun.} {\bf 184} 201--208
+
+\bibitem{louisell} Louisell W H \emph{Quantum Statistical Properties of Radiation}, (Wiley 1990)
 
@@
 
+\bibitem{gaunt2013} Gaunt A L, Schmidutz T F, Gotlibovych I, Smith R P, and Hadzibabic Z 2013 Bose-Einstein Condensation of Atoms in a Uniform Potential \emph{Phys. Rev. Lett.} {\bf 110} 200406
+
+\bibitem{cornish:2000} Cornish S, Claussen N, Roberts J, Cornell E, and Wieman C 2000 Stable $^{85}$Rb Bose-Einstein Condensates with Widely Tunable Interactions \emph{Phys. Rev. Lett.} {\bf 85} 1795
+
+\bibitem{altin:2010} Altin P A, Robins N P, D{\"{o}}ring D, Debs J E, Poldy R, Figl C and Close J D 2010 $^{85}$Rb tunable-interaction Bose-Einstein condensate machine \emph{Rev.
+Sci.
+Instr.} {\bf 81} 063103
+
 \end{thebibliography}

### 2013-07-12 large_number_squeezing_with_atoms.tex
Add and order refs, discuss the point of our analytic solution
Also various other scattered edits

@@
 We conclude that high levels of squeezing with large numbers of atoms is feasible, but can realistically only be achieved in traps with particular trap geometries.
-We also introduce a method of maximising the measurable squeezing by using a pi-pulse during the process to improve spatial mode-matching.
+We also introduce a method of maximising the measurable squeezing by using a $\pi$-pulse during the process to improve spatial mode-matching.
 \end{abstract}
@@
 In quantum optics, the most common way to produce entangled states has been through squeezing, and it has been recognised that the intrinsic nonlinearities of atoms naturally lead to quadrature squeezing \cite{Kheruntsyan:2005, johnssonET2007, haineET2009, doeringET2010}.
-Number squeezing can be achieved with spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissociation of Bose-Einstein Condensates (BECs) \cite{kheruntsyanET2002, davisET2008, savage2007, ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, double-well potentials \cite{esteve2008, maussang2010}, or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}.
-The creation of these non-classical atomic states has relevance for fundamental tests of quantum mechanics such as the EPR paradox \cite{pu2000, sorensen2001, haine2005,zhao2007, ogren2010}, as well as atom inteferometry \cite{sorensenET1999, kuzmichET2000, meyerET2001, liebfriedET2004, roosET2006, doeringET2010, gross2010}.
-While these techniques have resulted in significant experimentally detectable number squeezing \cite{gross2010}, they have been limited to small numbers of atoms, typically a few hundred.
+Number squeezing can be achieved with spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissociation of Bose-Einstein Condensates (BECs) \cite{kheruntsyanET2002,davisET2008,savage2007,ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, double-well potentials \cite{esteve2008, maussang2010}, or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}.
+The creation of these non-classical atomic states has relevance for fundamental tests of quantum mechanics such as the EPR paradox \cite{ogren2010, pu2000, sorensen2001, haine2005,zhao2007}, as well as atom interferometry \cite{doeringET2010, gross2010, sorensenET1999, kuzmichET2000, meyerET2001, liebfriedET2004, roosET2006}.
+While these techniques have resulted in significant experimentally detectable number squeezing \cite{esteve2008,gross2010}, they have been limited to small numbers of atoms, typically a few hundred.
 
@@
 Both of these effects limit the number of atoms that can be effectively squeezed, so we will examine a broader range of trap geometries, and calculate the dynamics in the presence of multiple spatial modes.
-Many similar previous simulations have been performed in one or two spatial dimensions, by integrating out some of the remaining dimensions \cite{kheruntsyanET2002, Kheruntsyan:2005, johnssonET2007, haineET2009}.
+Many similar previous simulations have been performed in one or two spatial dimensions, by integrating out some of the remaining dimensions \cite{Kheruntsyan:2005, johnssonET2007, haineET2009, kheruntsyanET2002}.
 For simulations of semiclassical fields this can be a reasonable approximation, but it can ignore spontaneous scattering processes in full quantum field theory calculations, so we take care to include full 3D simulations.
 
-In section \ref{secTwoModeAnalytic} we examine the ideal behaviour when the system is reduced to a single spatial mode, extracting a full analytic solution to the squeezing for arbitrary \textbf{whatever it is we extended, along with reference of what was covered previously.}  In section \ref{sec:MMdescription} we explore the idea of how multimode behaviour will become important in physical systems, and introduce our model for simulating higher dimensional trap geometries.
+In Section \ref{secTwoModeAnalytic} we examine the ideal behaviour of a nonlinear bosonic two-state system when that system possesses only a single spatial mode, extracting full analytic solutions for the number squeezing in each of the two states, as well as the number difference squeezing between the two states.
+These solutions of the ideal case provide a check on the multimode cases we consider later in the paper, as well as providing a bound on the best possible squeezing for the system.
+
+In Section \ref{sec:MMdescription} we show how multimode behaviour becomes important in physical systems, and introduce our model for simulating higher dimensional trap geometries.
 Section~\ref{sec:1Dsim} analyses one-dimensional models, highlighting some of the issues regarding optimising squeezing parameters and mode-matching in squeezing experiments.
-Section~\ref{sec:pipulse} shows that the application of a $\pi$-pulse during the process can dramatically improve mode-matching issues.
+Section~\ref{sec:pipulse} shows that the application of a $\pi$-pulse during the process can significantly improve mode-matching issues.
 Section~\ref{sec:2D3Dsqueezing} investigates measurable squeezing in 2D and 3D, and uses a Bogoliubov analysis to explain the scaling with dimensionality.
@@
 
-The system is prepared with all the population in $|a\rangle$, and vacuum in $|b\rangle$.
-We choose state $|a\rangle$ to be initially a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number.
+When considering squeezing, the Hamiltonian \ref{eqTwoModeHamiltonian} has typically been handled within the framework of spin squeezing, which has a long history \cite{kitagawaET1993, esteve2008,sorensenET1999, liebfriedET2004, wineland1994, steel+collett}.
+This approach involves exploiting the equivalence between the algebra of two harmonic oscillators and that of angular momentum, and writing
+\begin{eqnarray}
+\hat{S}_+ &=& \hat{a}^{\dagger} \hat{b} \\
+\hat{S}_- &=& \hat{a} \hat{b}^{\dagger} \\
+\hat{S}_x &=& \frac{1}{2} \left( \hat{S}_+ + \hat{S}_-\right) \\
+\hat{S}_y &=& \frac{1}{2i} \left( \hat{S}_+ - \hat{S}_-\right) \\
+\hat{S}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
+\end{eqnarray}
+Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) is transformed into the Josephson Hamiltonian \cite{steel+collett}
+\begin{equation}
+\Delta \omega \, \hat{S}_z + \chi \hat{J}_z^2 + \Omega \hat{J}_x
+\label{eqJosephson}
+\end{equation}
+where the $\Delta \omega$ is a function of the various nonlinearities and $\chi=\chi_{aa}+\chi_{bb}-2\chi{ab}$ are functions of the various nonlinearities, which is then solved using angular momentum algebra.
+
+We take a different approach and solve Eq.~(\ref{eqTwoModeHamiltonian}) directly for two reasons.
+First, we want access to the number variances of each state, not just the variance of the number difference between the states.
+Second, Eq.~(\ref{eqJosephson}) has led to the common misconception that no squeezing can be generated in this system if $\chi_{aa}+\chi_{bb}-2\chi{ab}=0$, as the effects of the nonlinearities on $\Delta \omega$ are dismissed as an irrelevent detuning.
+This is only the case if a spin-echo pulse is used half way through the application of (\ref{eqTwoModeHamiltonian}) (which is the protocol used in most spin squeezing experiments) or extra detunings are added to (\ref{eqJosephson}).
+Directly solving the Hamiltonian (\ref{eqTwoModeHamiltonian}) makes this clearer.
+
+The system is prepared with all the population initially in $|a\rangle$, and vacuum in $|b\rangle$.
+We choose state $|a\rangle$ to initially be a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number.
 As the system is insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, or a mixture of Poissonian-distributed number states.
-This state is consistent with BEC coherence experiments \cite{Hadzibabic2004}.
+Such a state is consistent with BEC coherence experiments \cite{Hadzibabic2004}.
 
@@
 
-This system can be solved analytically, and expressions for quadrature, absolute and relative number squeezing can be derived.
-We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}.
+We solve this system analytically, and derive expressions for absolute and relative number squeezing.
+We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation} as it is somewhat laborious.
 If we define
@@
 The equation of motion for the functional Wigner distribution contains derivatives of third order which we assume to be negligible.
-This uncontrolled approximation is called the Truncated Wigner Approximation (TWA), and while it has been used widely on ultracold gases \cite{johnssonET2007,johnssonHope:2007,haineET2009,dennisET2010,steelET1998}, care must be taken that the simulation remains valid.
+This uncontrolled approximation is called the Truncated Wigner Approximation (TWA), and while it has been used widely on ultracold gases \cite{johnssonET2007,haineET2009,steelET1998,dennisET2010,johnssonHope:2007}, care must be taken that the simulation remains valid.
 For our simulations we checked the results of our simulations against both the analytic two-mode solution, as well as checking for typical indications of TWA breakdown such as the appearance of negative densities in lightly populated modes.
@@
 We have presented an analytic solution to the two-mode model of atomic fields with a strong Kerr non-linearity, allowing for a rapid estimate of the atomic number difference squeezing that can be obtained from these systems as scattering length data is obtained.
-\textbf{Reiterate what is new in this analytic solution}.
 In general, it predicts arbitrarily good squeezing in the two-mode limit.
@@
 This is shown to alleviate the difficulties with choosing correct hold times and phases for the coupling pulses, and also alleviates the majority of the mode-matching issues for the final recombination of the two modes.
-If such a box potential is not available, the application of a pi-pulse part way through the hold period can also improve mode matching by imparting the same modal dynamics to both states.
+If such a box potential is not available, the application of a $\pi$-pulse part way through the hold period can also improve mode matching by imparting the same modal dynamics to both states.
 
@@
 
-\bibitem{Buluta2009} Buluta I and Nori F 2009 Quantum Simulators \emph{Science} {\bf 326} 108
+\bibitem{Leggett2001} Leggett A J 2001 Bose-Einstein condensation in the alkali gases: Some fundamental concepts \emph{Rev. Mod. Phys.} {\bf 73} 307
+
+\bibitem{Buluta2009} Buluta I and Nori F 2009 Quantum simulators \emph{Science} {\bf 326} 108
+
+\bibitem{Kinoshita:2006} Kinoshita T, Wenger T R and Weiss D S 2006 A quantum Newton's cradle \emph{Nature} {\bf 440} 900
+
+\bibitem{Barnett:2011} Barnett R, Polkovnikov A and Vengalattore M 2011 Prethermalization in quenched spinor condensates \emph{Phys. Rev.} A {\bf 84} 023606
+
+\bibitem{Kheruntsyan:2005} Kheruntsyan K V, Olsen M K and Drummond P D 2005 Einstein-Podolsky-Rosen correlations via dissociation of a molecular Bose-Einstein condensate \emph{Phys.
+Rev.} A {\bf 66} 031602(R)
+
+\bibitem{johnssonET2007} Johnsson M T and Haine S A 2007 Generating squeezing in an atom laser through self-interaction \emph{Phys. Rev. Lett.} {\bf 99} 010401
+
+\bibitem{haineET2009} Haine S A and Johnsson M T 2009 Dynamic scheme for generating number squeezing in Bose-Einstein condensates through nonlinear interactions \emph{Phys. Rev.} A {\bf 80} 023611
+
+\bibitem{doeringET2010} D{\"{o}}ring D, McDonald G, Debs J E, Figl C, Altin P A, Bachor H-A, Robins N P, and Close J D 2010 Quantum-projection-noise-limited interferometry with coherent atoms in a Ramsey-type setup \emph{Phys.
+Rev.} A {\bf 81} 043633
+
+\bibitem{kitagawaET1993} Kitagawa M and Ueda M 1995 Spin squeezed states \emph{Phys. Rev.} A {\bf 47} 5138
+
+\bibitem{gross2010} Gross C, Zibold T, Est{\`{e}}ve J and Oberthaler M K 2010 Nonlinear atom interferometer surpasses classical precision limit \emph{Nature} {\bf 464} 1165--1169
+
+\bibitem{sinatra2012} Sinatra A, Dornstetter J-C, Castin Y 2012 Spin squeezing in Bose-Einstein condensates: Limits imposed by decoherence and non-zero temperature \emph{Front. Phys.} {\bf 7} 86
 
@@
 
-\bibitem{Kheruntsyan:2005} Kheruntsyan K V, Olsen M K and Drummond P D 2005 Einstein-Podolsky-Rosen correlations via dissociation of a molecular Bose-Einstein condensate \emph{Phys.
-Rev.} A {\bf 66} 031602(R)
-
-\bibitem{kitagawaET1993} Kitagawa M and Ueda M 1995 Spin squeezed states \emph{Phys. Rev.} A {\bf 47} 5138
+\bibitem{davisET2008} Davis M J, Thwaite S J, Olsen M K, and Kheruntsyan K V 2008 Pairing mean-field theory for the dynamics of dissociation of molecular Bose-Einstein condensates \emph{Phys.
+Rev.} A {\bf 77} 023617
+
+\bibitem{savage2007} Savage C M and Kheruntsyan K V 2007 Spatial pair correlations of atoms in molecular dissociation \emph{Phys. Rev. Lett.} {\bf 99} 220404
+
+\bibitem{ogren2010} {\"{O}}gren M and Kheruntsyan K V 2010 Role of spatial inhomogeneity in dissociation of trapped molecular condensates \emph{Phys. Rev. A} {\bf 82} 013641
+
+\bibitem{perrin2008} Perrin A, Savage C M, Boiron D, Krachmalnicoff V, Westbrook C I and Kheruntsyan K V 2008 Atomic four-wave mixing via condensate collisions \emph{New J. Phys.} {\bf 10} 045021
+
+\bibitem{ogren2009} {\"{O}}gren M and Kheruntsyan K V 2009 Atom-atom correlations in colliding Bose-Einstein condensates \emph{Phys. Rev. A} {\bf 79} 021606
+
+\bibitem{jascula2010} Jaskula J C, Bonneau M, Partridge G B, Krachmalnicoff V, Deuar P, Kheruntsyan K V, Aspect A, Boiron D, and Westbrook C I 2010 Sub-Poissonian number differences in four-wave mixing of matter waves \emph{Phys.
+Rev.
+Lett.} {\bf 105} 190402
+
+\bibitem{esteve2008} Est{\`{e}}ve J, Gross C, Weller A, Giovanazzi S, and Oberthaler M 2008 Squeezing and entanglement in a Bose-Einstein condensate \emph{Nature} {\bf 455} 1216
+
+\bibitem{maussang2010} Maussang K, Marti G E, Schneider T, Treutlein P, Li Y, Sinatra A, Long R, Est{\`{e}}ve J, and Reichel J 2010 Enhanced and reduced atom number fluctuations in a BEC splitter \emph{Phys.
+Rev.
+Lett.} {\bf 105} 080403
+
+\bibitem{pu2000} Pu H and Meystre P 2000 Creating macroscopic atomic Einstein-Podolsky-Rosen states from Bose-Einstein condensates \emph{Phys. Rev. Lett.} {\bf 85} 3987
+
+\bibitem{sorensen2001} S{\"{o}}rensen A, Duan L-M, Cirac J I, Zoller P 2001 Many-particle entanglement with Bose-Einstein condensates \emph{Nature} {\bf 409} 63
+
+\bibitem{haine2005} Haine S A and Hope J J 2005 Outcoupling from a Bose-Einstein condensate with squeezed light to produce entangled atom laser beams, \emph{Phys. Rev.} A {\bf 72} 033601
+
+\bibitem{zhao2007} Zhao B, Chen Z, Pan J, Schmiedmayer J, Recati A, Astrakharchik G E, Calarco T 2007 Deterministic Entanglement via Molecular Dissociation in Integrated Atom Optics \emph{Phys.
+Rev.} A {\bf 75} 042312
 
@@
 
+\bibitem{robins:2013} Robins N P, Altin P A, Debs J E and Close J D 2013 Atom lasers: Production, properties and prospects for precision inertial measurement \emph{Phys. Reports} {\bf 529} 265
+
+\bibitem{szigeti:2012} Szigeti S S, Debs J E, Hope J J, Robins N P and Close J D 2012 Why momentum width matters for atom interferometry with Bragg pulses \emph{N. J. Phys.} {\bf 14} 023009
+
 \bibitem{LiET2009} Li Y, Treutlein P, Reichel J, and Sinatra A 2009 Spin squeezing in a bimodal condensate: spatial dynamics and particle losses \emph{Eur. Phys. J.} B {\bf 68} 365
 
-\bibitem{jascula2010} Jaskula J C, Bonneau M, Partridge G B, Krachmalnicoff V, Deuar P, Kheruntsyan K V, Aspect A, Boiron D, and Westbrook C I 2010 Sub-Poissonian Number Differences in Four-Wave Mixing of Matter Waves \emph{Phys.
-Rev.
-Lett.} {\bf 105} 190402
-
-\bibitem{robins:2013} Robins N P, Altin P A, Debs J E and Close J D 2013 Atom lasers: Production, properties and prospects for precision inertial measurement \emph{Phys. Reports} {\bf 529} 265
-
-\bibitem{szigeti:2012} Szigeti S S, Debs J E, Hope J J, Robins N P and Close J D 2012 Why momentum width matters for atom interferometry with Bragg pulses \emph{N. J. Phys.} {\bf 14} 023009
-
-\bibitem{doeringET2010} D{\"{o}}ring D, McDonald G, Debs J E, Figl C, Altin P A, Bachor H-A, Robins N P, and Close J D 2010 Quantum-projection-noise-limited interferometry with coherent atoms in a Ramsey-type setup \emph{Phys.
-Rev.} A {\bf 81} 043633
-
-\bibitem{davisET2008} Davis M J, Thwaite S J, Olsen M K, and Kheruntsyan K V 2008 Pairing mean-field theory for the dynamics of dissociation of molecular Bose-Einstein condensates \emph{Phys.
-Rev.} A {\bf 77} 023617
-
-\bibitem{savage2007} Savage C M and Kheruntsyan K V 2007 Spatial Pair Correlations of Atoms in Molecular Dissociation \emph{Phys. Rev. Lett.} {\bf 99} 220404
-
-\bibitem{pu2000} Pu H and Meystre P 2000 Creating Macroscopic Atomic Einstein-Podolsky-Rosen States from Bose-Einstein Condensates \emph{Phys. Rev. Lett.} {\bf 85} 3987
-
-\bibitem{sorensen2001} S{\"{o}}rensen A, Duan L-M, Cirac J I, Zoller P 2001 Many-particle entanglement with Bose-Einstein condensates \emph{Nature} {\bf 409} 63
-
-\bibitem{haine2005} Haine S A and Hope J J 2005 Outcoupling from a Bose-Einstein condensate with squeezed light to produce entangled atom laser beams, \emph{Phys. Rev.} A {\bf 72} 033601
-
-\bibitem{zhao2007} Zhao B, Chen Z, Pan J, Schmiedmayer J, Recati A, Astrakharchik G E, Calarco T 2007 Deterministic Entanglement via Molecular Dissociation in Integrated Atom Optics \emph{Phys.
-Rev.} A {\bf 75} 042312
-
-\bibitem{esteve2008} Est{\`{e}}eve J, Gross C, Weller A, Giovanazzi S, and Oberthaler M 2008 Squeezing and entanglement in a Bose-Einstein condensate \emph{Nature} {\bf 455} 1216
-
-\bibitem{maussang2010} Maussang K, Marti G E, Schneider T, Treutlein P, Li Y, Sinatra A, Long R, Est{\`{e}}ve J, and Reichel J 2010 Enhanced and reduced atom number fluctuations in a BEC splitter \emph{Phys.
-Rev.
-Lett.} {\bf 105} 080403
-
-\bibitem{sinatra2012} Sinatra A, Dornstetter J-C, Castin Y 2012 Spin squeezing in Bose-Einstein condensates: Limits imposed by decoherence and non-zero temperature \emph{Front. Phys.} {\bf 7} 86
-
-\bibitem{perrin2008} Perrin A, Savage C M, Boiron D, Krachmalnicoff V, Westbrook C I and Kheruntsyan K V 2008 Atomic four-wave mixing via condensate collisions \emph{New J. Phys.} {\bf 10} 045021
-
-\bibitem{ogren2010} {\"{O}}gren M and Kheruntsyan K V 2010 Role of spatial inhomogeneity in dissociation of trapped molecular condensates \emph{Phys. Rev. A} {\bf 82} 013641
-
-\bibitem{ogren2009} {\"{O}}gren M and Kheruntsyan K V 2009 Atom-atom correlations in colliding Bose-Einstein condensates \emph{Phys. Rev. A} {\bf 79} 021606
+\bibitem{wineland1994} Wineland D J, Bollinger J J, Itano W M and Heinzen D J 1994 Squeezed atomic states and projection noise in spectroscopy \emph{Phys. Rev.} A {\bf 50} 67
+
+\bibitem{steel+collett} Steel M and Collett M 1998 Quantum state of two trapped Bose-Einstein condensates with a Josephson coupling \emph{Phys. Rev.} A {\bf 57} 2920
 
 \bibitem{Hadzibabic2004} Hadzibabic Z, Stock S, Battelier B, Bretin V and Dalibard J 2004 Interference of an array of independent Bose-Einstein Condensates \emph{Phys. Rev. Lett.} {\bf 93} 180403
-
-\bibitem{johnssonET2007} Johnsson M T and Haine S A 2007 Generating squeezing in an atom laser through self-interaction \emph{Phys. Rev. Lett.} {\bf 99} 010401
-
-\bibitem{dallET2009} Dall R G, Byron L J, Truscott A G, Dennis G R, Johnsson M T and Hope J J 2009 Paired-atom laser beams created via four-wave mixing \emph{Phys. Rev.} A {\bf 79} 011601(R)
-
-\bibitem{johnssonHope:2007} Johnsson M T and Hope J J 2007 Multimode quantum limits to the linewidth of an atom laser \emph{Phys. Rev A} {\bf 75} 0432619
-
-\bibitem{haineET2009} Haine S A and Johnsson M T 2009 Dynamic scheme for generating number squeezing in Bose-Einstein condensates through nonlinear interactions \emph{Phys. Rev.} A {\bf 80} 023611
-
-\bibitem{dennisET2010} Dennis G R and Johnsson M T 2010 Generation of directional coherent matter beams through dynamical instabilities in Bose-Einstein condensates \emph{Phys. Rev.} A {\bf 82} 033615
 
@@
 
+\bibitem{dallET2009} Dall R G, Byron L J, Truscott A G, Dennis G R, Johnsson M T and Hope J J 2009 Paired-atom laser beams created via four-wave mixing \emph{Phys. Rev.} A {\bf 79} 011601(R)
+
+\bibitem{dennisET2010} Dennis G R and Johnsson M T 2010 Generation of directional coherent matter beams through dynamical instabilities in Bose-Einstein condensates \emph{Phys. Rev.} A {\bf 82} 033615
+
+\bibitem{johnssonHope:2007} Johnsson M T and Hope J J 2007 Multimode quantum limits to the linewidth of an atom laser \emph{Phys. Rev A} {\bf 75} 0432619
+
 \bibitem{dennis2013} Dennis G R, Hope J J and Johnsson M T 2013 XMDS2: Fast, scalable simulation of coupled stochastic partial differential equations \emph{Comput. Phys. Commun.} {\bf 184} 201--208
 
-\bibitem{louisell} Louisell W H \emph{Quantum Statistical Properties of Radiation}, (Wiley 1990)
-
-\bibitem{gross2010} Gross C, Zibold T, Est{\`{e}}ve J and Oberthaler M K 2010 Nonlinear atom interferometer surpasses classical precision limit \emph{Nature} {\bf 464} 1165--1169
-
 \bibitem{PethickSmith} Pethick C J and Smith H, \emph{Bose-Einstein Condensation in Dilute Gases} (Cambridge University Press, Cambridge, 2002).
 
-\bibitem{gaunt2013} Gaunt A L, Schmidutz T F, Gotlibovych I, Smith R P, and Hadzibabic Z 2013 Bose-Einstein Condensation of Atoms in a Uniform Potential \emph{Phys. Rev. Lett.} {\bf 110} 200406
-
-\bibitem{cornish:2000} Cornish S, Claussen N, Roberts J, Cornell E, and Wieman C 2000 Stable $^{85}$Rb Bose-Einstein Condensates with Widely Tunable Interactions \emph{Phys. Rev. Lett.} {\bf 85} 1795
+\bibitem{gaunt2013} Gaunt A L, Schmidutz T F, Gotlibovych I, Smith R P, and Hadzibabic Z 2013 Bose-Einstein condensation of atoms in a uniform potential \emph{Phys. Rev. Lett.} {\bf 110} 200406
+
+\bibitem{cornish:2000} Cornish S, Claussen N, Roberts J, Cornell E, and Wieman C 2000 Stable $^{85}$Rb Bose-Einstein condensates with widely tunable Interactions \emph{Phys. Rev. Lett.} {\bf 85} 1795
 
@@
 
+\bibitem{louisell} Louisell W H \emph{Quantum Statistical Properties of Radiation}, (Wiley 1990)
+
 \end{thebibliography}

### 2013-07-12 large_number_squeezing_with_atoms.tex
Minor changes to Mattias' appendix and a comment where the language should probably be made a little less informal.

--- 2013-07-11 large_number_squeezing_with_atoms.tex
@@
 
-Beginning with the Hamiltonian $H_1$, during the period $(0, t_1)$ mode $\hat{a}$ will evolve as
+Beginning with the Hamiltonian $\hat{H}_1$, during the period $(0, t_1)$ mode $\hat{a}$ will evolve as
 \begin{equation}
@@
 From the form of (\ref{eqa1}) and (\ref{eqb1}) it is clear that as both modes $\hat{a}$ and $\hat{b}$ began in coherent states, they will remain in coherent states, with only their amplitudes changing.
-Specifically, at time $t_1$ the system is in a product of coherent states, which we denote $|\alpha, \beta\rangle$ with $\alpha=\sqrt{n_a}$ and $\beta=-i\sqrt{n_b}$.
+Specifically, at time $t_1$ the system is in a product of coherent states, which we denote $|\alpha, \beta\rangle$ with $\alpha=\sqrt{n_a}$ and $\beta=-i\sqrt{n_b}$ with $n_a + n_b = N$.
 
-Next we consider evolution under $H_2$.
+Next we consider evolution under $\hat{H}_2$.
 Taking mode $\hat{a}$ first, we note that as it commutes with the $\hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}$ term, and since $\hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}$ commutes with $\hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}$ we have
@@
 
-Next we tackle $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$. Using (\ref{eqa2}) and (\ref{eqb2}) and the fact that $\hat{a}$ and $\hat{b}$ commute, we have
+Next we tackle (\textbf{too informal?}) $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$.
+Using (\ref{eqa2}) and (\ref{eqb2}) and the fact that $\hat{a}$ and $\hat{b}$ commute, we have
 \begin{eqnarray}
@@
   e^{i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
-  \hat{b}_1 \hat{a}^{\dagger_1}
+  \hat{b}_1 \hat{a}^{\dagger}_1
   e^{i (\lambda_{ab}-\lambda_{bb}) \hat{b}^{\dagger}_1 \hat{b}_1}
@@
 We write (\ref{eqNumDiffVarianceAppendix}) in terms of $\hat{a}_2$ and $\hat{b}_2$ by using using (\ref{eqa3}) and (\ref{eqb3}), and then use the expectation values calculated earlier in this Appendix.
-After some algebra we obtain the expression (\ref{eqNumDiffVariance}) given in Section \ref{secTwoModeAnalytic}.
+This procedure gives the expression (\ref{eqNumDiffVariance}) given in Section \ref{secTwoModeAnalytic}.
 

### 2013-07-12 large_number_squeezing_with_atoms.tex

--- 2013-07-12 large_number_squeezing_with_atoms.tex
@@
 From the form of (\ref{eqa1}) and (\ref{eqb1}) it is clear that as both modes $\hat{a}$ and $\hat{b}$ began in coherent states, they will remain in coherent states, with only their amplitudes changing.
-Specifically, at time $t_1$ the system is in a product of coherent states, which we denote $|\alpha, \beta\rangle$ with $\alpha=\sqrt{n_a}$ and $\beta=-i\sqrt{n_b}$.
-
-Next we consider evolution under $H_2$.
+Specifically, at time $t_1$ the system is in a product of coherent states, which we denote $|\alpha, \beta\rangle$ with $\alpha=\sqrt{n_a}$ and $\beta=-i\sqrt{n_b}$ with $n_a + n_b = N$.
+
+Next we consider evolution under $\hat{H}_2$.
 Taking mode $\hat{a}$ first, we note that as it commutes with the $\hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}$ term, and since $\hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}$ commutes with $\hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}$ we have

### 2013-07-12 large_number_squeezing_with_atoms.tex
Fix figure 3, minor appendix tweaks

--- 2013-07-12 large_number_squeezing_with_atoms.tex
@@
 We use the model described in Section \ref{secTwoModeAnalytic}, where we consider a two mode system with $N$ particles, initially prepared as a coherent state with all the population in mode $|a\rangle$ at time $t=0$.
-A coupling pulse is then applied, so that the system evolves under the Hamiltonian
-\begin{equation}
-\hat{H_1} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b} 
-          + \hbar \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger} \hat{a} )
+A coupling pulse is then applied, assumed to be on resonance, so that the system evolves under the Hamiltonian
+\begin{equation}
+\hat{H_1} = \hbar \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger} \hat{a} )
 \end{equation}
@@
 \begin{equation}
-\hat{H_3} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b}
-          + \hbar\Omega (e^{i\phi} \hat{a}^{\dagger} \hat{b} + e^{-i\phi} \hat{b}^{\dagger} \hat{a} ).
-\end{equation}
-It is clear that the linear self-energy terms proportional to $\omega$ result in a constant phase rotation of the fields $\hat{a}$ and $\hat{b}$.
-As this system is self-contained and global phase is unmeasurable, all phase measurements can only rely on the difference in phase between the two modes.
-This phase rotation is therefore irrelevant and will be ignored in the following analysis.
+\hat{H_3} = \hbar\Omega (e^{i\phi} \hat{a}^{\dagger} \hat{b} + e^{-i\phi} \hat{b}^{\dagger} \hat{a} ).
+\end{equation}
 
@@
 
-Next we tackle $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$. Using (\ref{eqa2}) and (\ref{eqb2}) and the fact that $\hat{a}$ and $\hat{b}$ commute, we have
+We now consider $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$. Using (\ref{eqa2}) and (\ref{eqb2}) and the fact that $\hat{a}$ and $\hat{b}$ commute, we have
 \begin{eqnarray}

### 2013-07-12 large_number_squeezing_with_atoms.tex
Redo GD's fix which I screwed up
Finish re-applying my changes that Mattias accidentally backed-out.

@@
 We will work in the Heisenberg picture, and derive expressions for $\hat{a}(t_3)$ and $\hat{b}(t_3)$ in terms of $\hat{a}(t_2)$ and $\hat{b}(t_2)$, which in turn can be expressed in terms of $\hat{a}(t_1)$ and $\hat{b}(t_1)$.
-We will show that the state of the system at time $t_1$ consists of seperate coherent states in modes $|a\rangle$ and $|b\rangle$, enabling us to calculate expectation values at time $t_3$ in terms of the known state at $t_1$.
+We will show that the state of the system at time $t_1$ consists of separate coherent states in modes $|a\rangle$ and $|b\rangle$, enabling us to calculate expectation values at time $t_3$ in terms of the known state at $t_1$.
 

### 2013-07-12 large_number_squeezing_with_atoms.tex
Fix another typo.

@@
 First, we want access to the number variances of each state, not just the variance of the number difference between the states.
-Second, Eq.~(\ref{eqJosephson}) has led to the common misconception that no squeezing can be generated in this system if $\chi_{aa}+\chi_{bb}-2\chi{ab}=0$, as the effects of the nonlinearities on $\Delta \omega$ are dismissed as an irrelevent detuning.
+Second, Eq.~(\ref{eqJosephson}) has led to the common misconception that no squeezing can be generated in this system if $\chi_{aa}+\chi_{bb}-2\chi_{ab}=0$, as the effects of the nonlinearities on $\Delta \omega$ are dismissed as an irrelevent detuning.
 This is only the case if a spin-echo pulse is used half way through the application of (\ref{eqTwoModeHamiltonian}) (which is the protocol used in most spin squeezing experiments) or extra detunings are added to (\ref{eqJosephson}).

### 2013-07-16 large_number_squeezing_with_atoms.tex
First set of proofreading edits

@@
 The creation of these non-classical atomic states has relevance for fundamental tests of quantum mechanics such as the EPR paradox \cite{ogren2010, pu2000, sorensen2001, haine2005,zhao2007}, as well as atom interferometry \cite{doeringET2010, gross2010, sorensenET1999, kuzmichET2000, meyerET2001, liebfriedET2004, roosET2006}.
-While these techniques have resulted in significant experimentally detectable number squeezing \cite{esteve2008,gross2010}, they have been limited to small numbers of atoms, typically a few hundred.
+While these techniques have resulted in significant experimentally detectable number squeezing, they have been limited to small numbers of atoms, typically a few hundred \cite{gross2010, esteve2008}.
 
@@
 If these technical noise sources can be controlled, atom interferometers will be limited by atomic shot noise, which leads to a sensitivity that scales as the square root of the atomic flux.
-From that point, sensitivity will only be improved by increasing the atomic flux, and/or using non-classical quantum states such as squeezed states to go below the shot noise limit.
+From that point, sensitivity will only be improved by increasing the atomic flux, and/or using non-classical quantum states such as squeezed states to go below the shot noise limit \cite{robins:2013}.
 Producing squeezed states for metrological purposes is therefore only relevant when it can be achieved in a context of large numbers of atoms.
@@
 Atomic squeezing experiments have typically attempted to operate in the single-mode regime, where the trap is tight enough such that all higher modes are frozen out \cite{LiET2009}.
-However, tight traps limit the total number of atoms, as the 3-body recombination rate places an effective maximum atomic density in the gas, and well before that limit, the interatomic interactions break the single-mode operation.
+However, tight traps limit the total number of atoms, as the 3-body recombination rate places an effective maximum atomic density in the gas, and well before that limit the interatomic interactions break the single-mode operation.
 Both of these effects limit the number of atoms that can be effectively squeezed, so we will examine a broader range of trap geometries, and calculate the dynamics in the presence of multiple spatial modes.
@@
 \end{equation}
-where $\hat{\psi}_j(\mathbf{x})$ is the annihilation field operator for a particle at position $\mathbf{x}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{x})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-state nonlinearities, and $\kappa$ describes the coupling between the two states.
+where $\hat{\psi}_j(\mathbf{r})$ is the annihilation field operator for a particle at position $\mathbf{r}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{r})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-state nonlinearities, and $\kappa$ describes the coupling between the two states.
 
@@
 
-When considering squeezing, the Hamiltonian \ref{eqTwoModeHamiltonian} has typically been handled within the framework of spin squeezing, which has a long history \cite{kitagawaET1993, esteve2008,sorensenET1999, liebfriedET2004, wineland1994, steel+collett}.
+When considering squeezing, the Hamiltonian (\ref{eqTwoModeHamiltonian}) has typically been handled within the framework of spin squeezing, which has a long history \cite{kitagawaET1993, esteve2008,sorensenET1999, liebfriedET2004, wineland1994, steel+collett}.
 This approach involves exploiting the equivalence between the algebra of two harmonic oscillators and that of angular momentum, and writing
@@
 \begin{equation}
-\Delta \omega \, \hat{S}_z + \chi \hat{J}_z^2 + \Omega \hat{J}_x
+\hat{H}_J = \Delta \omega \, \hat{S}_z + \chi \hat{J}_z^2 + \Omega \hat{J}_x
 \label{eqJosephson}
 \end{equation}
-where the $\Delta \omega$ is a function of the various nonlinearities and $\chi=\chi_{aa}+\chi_{bb}-2\chi{ab}$ are functions of the various nonlinearities, which is then solved using angular momentum algebra.
+where $\Delta \omega$ is a function of the various nonlinearities and $\chi=\chi_{aa}+\chi_{bb}-2\chi_{ab}$. The system described by $\hat{H}_J$ is then solved by using angular momentum algebra.
 
@@
 This is only the case if a spin-echo pulse is used half way through the application of (\ref{eqTwoModeHamiltonian}) (which is the protocol used in most spin squeezing experiments) or extra detunings are added to (\ref{eqJosephson}).
-Directly solving the Hamiltonian (\ref{eqTwoModeHamiltonian}) makes this clearer.
+Directly solving Eq.~(\ref{eqTwoModeHamiltonian}) makes this clearer.
 
 The system is prepared with all the population initially in $|a\rangle$, and vacuum in $|b\rangle$.
-We choose state $|a\rangle$ to initially be a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number.
+We choose the initial state $|a\rangle$ to be a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number.
 As the system is insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, or a mixture of Poissonian-distributed number states.
@@
 After this hold time $\tau_{\mathrm{hold}} = t_2 - t_1 $, the coupling $\Omega$ is switched back on until time $t_3$, with a phase shift $\phi$ compared to its first application.
-During this last stage, the two modes exchange population and the quadrature variance is converted into number variance, in the same way that homodyne measurements are used in quantum optics to convert quadrature squeezing into number squeezing, which can be directly measured.
-In this interpretation, $\phi$ is the relative phase of the strong local oscillator, which allows specific phase angles of quadrature squeezing to be examined.
+During this last stage, the two modes exchange population and the quadrature variances are converted into number variance, in the same way that homodyne measurements are used in quantum optics to convert quadrature squeezing into number squeezing, which can be directly measured.
+In this interpretation, $\phi$ is the relative phase of the strong local oscillator, which allows specific phase angles of the quadrature squeezing to be examined.
 This experiment can also be interpreted as a Ramsey interferometer with a final beam splitter phase of $\phi$.
@@
 \end{eqnarray}
-In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, e.g.
+In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, i.e.
 ${\mathrm{Var}}[N_a]=N_a$, and no number squeezing is possible.
@@
 The two-mode approximation can be valid in tight traps with very low particle number, and this result correctly predicts the squeezing demonstrated by Gross \textit{et al.} \cite{gross2010}.
-Unfortunately, for larger numbers of atoms in realistic traps there are more than two modes present, and multimode effects limit the level of squeezing that can be achieved.
+Unfortunately, for larger numbers of atoms in realistic traps there are more than two modes present, and these multimode effects limit the level of squeezing that can be achieved.
 
@@
 
-An analytic solution to the multimode, higher-dimensional problem is intractable, so we must use a numerical solution.
+An analytic solution to the multimode, higher-dimensional problem is intractable, so we must turn to numerical solutions.
 We use stochastic methods based on phase space representations to simulate the dynamics of the multimode quantum fields.
@@
 This uncontrolled approximation is called the Truncated Wigner Approximation (TWA), and while it has been used widely on ultracold gases \cite{johnssonET2007,haineET2009,steelET1998,dennisET2010,johnssonHope:2007}, care must be taken that the simulation remains valid.
-For our simulations we checked the results of our simulations against both the analytic two-mode solution, as well as checking for typical indications of TWA breakdown such as the appearance of negative densities in lightly populated modes.
-We could also use the Bogoliubov analysis described in section \ref{sec:2D3Dsqueezing} as an independent estimate of the validity of the TWA.
+For our simulations we checked the results of our simulations against the analytic two-mode solution, as well as checking for typical indications of TWA breakdown such as the appearance of negative densities in lightly populated modes.
+We were also able to use the Bogoliubov analysis described in Section \ref{sec:2D3Dsqueezing} as an independent estimate of the validity of the TWA.
 We found that the TWA remained valid except in some 3D simulations with high nonlinearities, where there were a large number of modes, and the number of particles per mode could drop to ten or less in the densest regions where the squeezing was generated.
@@
 \begin{eqnarray}
-\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{x}) + U_{a a} n_a + U_{a b} \tilde{n}_b  \right)\phi_a   + \kappa \phi_b  \nonumber \\
-\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{x}) + U_{b b} n_b + U_{a b} \tilde{n}_a   \right)\phi_b   + \kappa^* \phi_a
+\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{r}) + (U_{a a} n_a + U_{a b} \tilde{n}_b  \right)\phi_a   + \kappa \phi_b  \nonumber \\
+\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{r}) + U_{b b} n_b + U_{a b} \tilde{n}_a   \right)\phi_b   + \kappa^* \phi_a
   \label{eqStochasticEquations}
@@
 
-We integrate equations (\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS2 \cite{dennis2013}.
+We integrate Eqs.~(\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS2 \cite{dennis2013}.
 For the one- and two-dimensional simulations, the dimensional reduction is achieved by estimating the mode shape in three dimensions, and using dimensionally reduced values for $U_{ij}$ that match the zero-dimensional reduction given by Eq.~(\ref{eqChiUequivalence}).
@@
 Common parameters used were $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\,\text{s}^{-1}$, $\chi_{ab}=\chi_{bb}=0$.
-Solid red line: Analytic two-mode solution, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (10.9dB squeezing); Solid blue line: no $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.0$ (4.9dB squeezing);  Dashed blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.05$ (5.8dB squeezing); Dash-dotted blue line: with $\pi$-pulse and the longer hold time $\tau_{\mathrm{hold}}=5\times 10^{-4}$s allowed by the application of the pulse, $\phi=3.87$ (7.2dB squeezing).}
+Solid red line: Analytic two-mode solution, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$ (10.9dB squeezing); Solid blue line: no $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$ (4.9dB squeezing);  Dashed blue line: with $\pi$-pulse, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$ (5.8dB squeezing); Dash-dotted blue line: with $\pi$-pulse and the longer hold time $\tau_{\mathrm{hold}}=5\times 10^{-4}$s allowed by the application of the pulse, $\phi=3.87$ (7.2dB squeezing).}
     \label{piPulseFig}

### 2013-07-17 large_number_squeezing_with_atoms.tex
Second round of proofreading tweaks

@@
 
-Ultracold atomic interferometers are also reaching state of the art sensitivities in precision measurements \cite{robins:2013}.
+Ultracold atomic interferometers are also reaching state of the art sensitivities in precision measurements \cite{robins:2013, altinET2013}.
 While all current atom interferometry is limited by technical noise sources, it appears that BEC-based coherent sources offer the best chance of overcoming some of these issues, for the same reasons that optical lasers are typically preferred for optical interferometry \cite{szigeti:2012}.
@@
 \end{equation}
-where $\hat{\psi}_j(\mathbf{r})$ is the annihilation field operator for a particle at position $\mathbf{r}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{r})$ is the trapping potential for atoms in internal state $|j\rangle$, $U_{ij}$ describe the various inter- and intra-state nonlinearities, and $\kappa$ describes the coupling between the two states.
+where $\hat{\psi}_j(\mathbf{r})$ is the annihilation field operator for a particle at position $\mathbf{r}$ and in internal state $|j\rangle$, $m$ is the mass of the atoms, $V_j(\mathbf{r})$ is the trapping potential for atoms in internal state $|j\rangle$, $\kappa$ describes the coupling between the two states, and the $U_{ij}$ describe the various inter- and intra-state nonlinearities and are given by
+\begin{equation}
+U_{ij} = \frac{4 \pi \hbar^2 m} {a_{ij}}
+\label{eqUij}
+\end{equation}
+where $a_{ij}$ are the inter- and intra-species s-wave scattering lengths.
 
@@
 \begin{eqnarray}
-\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{r}) + (U_{a a} n_a + U_{a b} \tilde{n}_b  \right)\phi_a   + \kappa \phi_b  \nonumber \\
+\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{r}) + U_{a a} n_a + U_{a b} \tilde{n}_b  \right)\phi_a   + \kappa \phi_b  \nonumber \\
 \frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{r}) + U_{b b} n_b + U_{a b} \tilde{n}_a   \right)\phi_b   + \kappa^* \phi_a
@@
 \caption{Plot of the non-condensed fraction $N_{\mathbf{k}\neq0}/N$ as a function of the hold time $\tau_\text{hold}$.
-The blue dashed line is the Bogoliubov integral approximation given in Eq.~\ref{eqBogFraction1D}, the purple dot-dashed line is the Bogoliubov summation expression given in Eq.~\ref{eqCondensateDepletion}, and the solid black line is the result of a truncated Wigner simulation.
+The blue dashed line is the Bogoliubov integral approximation given in Eq.~(\ref{eqBogFraction1D}), the purple dot-dashed line is the Bogoliubov summation expression given in Eq.~(\ref{eqCondensateDepletion}), and the solid black line is the result of a truncated Wigner simulation.
 Parameters: $L=600\,\mu\text{m}$, $N=4 \times 10^5$, $n_a=n_b=2\times 10^5$, $\chi_{aa} = 2.67 \times 10^{-2}\, \text{s}^{-1}$, $\chi_{ab}=\chi_{bb}=0$, $\theta=0$.}
@@
 In particular this fixes the values of $\theta$, $\phi$ and $\lambda_{ij} = \chi_{ij}\tau_\text{hold}$.
-As the size of the condensate is scaled, $\hbar \chi_{ij} = U^\text{1D}_{ij}/L = U^\text{2D}_{ij}/A = U^\text{3D}_{ij}/V$ will change, but $\tau_\text{hold}$ will vary to keep $\lambda_{ij}$ constant.
+As the size of the condensate is scaled, $\hbar \chi_{ij} = U^\text{1D}_{ij}/L = U^\text{2D}_{ij}/A = U^\text{3D}_{ij}/V$ will change, but we will vary $\tau_\text{hold}$ to keep $\lambda_{ij}$ constant.
 Therefore $\Lambda_a = n_a\chi_{aa}\tau_\text{hold}$ will also be constant.
@@
 Also shown in the 1D and 2D cases is the maximum amount of number squeezing obtainable, as predicted by the numerical simulation.
-The full squeezing scaling curve is not shown in 3D, as for the smaller trap sizes the TWA method breaks down, and results are unreliable; for the larger trap sizes with negligible non-condensate fractions, however, the TWA simulation was possible in the large trap volume limit, where the non-condensed fraction was low.
-The traps are assumed to have a constant potential, which minimises the effects of the multimode issues described in section \ref{sec:MMdescription}.
+The full squeezing scaling curve is not shown in 3D, as for the smaller trap sizes the TWA method breaks down, and results are unreliable.
+For the larger trap sizes with small non-condensate fractions, however, the TWA simulation remained valid, and we obtained relative number squeezing of 12.7 dB.
+The traps are assumed to have a constant potential, which minimises the effects of the multimode issues described in Section \ref{sec:MMdescription}.
 We see that the squeezing does indeed degrade at large trap sizes in 1D, but remains constant in 2D.
@@
 
-The simulation shown in Fig.~\ref{fig3D_Bog_mode_occupation_scaling} verified that the number of accessible states for three-dimensional traps decreases with volume in 3D.
-This is because the energy of the ground state of the BEC is reducing faster than the density of states at low energies is increasing.
-This scaling of the estimated Bogoliubov fraction in 3D suggests that strong squeezing may actually be found in very large, weakly trapped condensates, provided the trapping potentials are box-like.
+The simulation shown in Fig.~\ref{fig3D_Bog_mode_occupation_scaling} demonstrates that provided particle number is kept fixed, the number of accessible states for three-dimensional traps decreases with volume in 3D.
+This is because the energy of the ground state of the BEC (the chemical potential, $\mu$) is reducing faster than the density of states at low energies is increasing, which in turn results in fewer states that can become dynamically occupied.
+This scaling of the estimated Bogoliubov fraction in 3D suggests that strong squeezing will be found in very large, weakly trapped condensates, provided the trapping potentials are box-like.
 Such box-like traps have already been demonstrated and shown to be capable of producing BECs \cite{gaunt2013}.
@@
 As a demonstration of the plausibility of such a scheme, we carried out a full 3D simulation using the geometry of the optical box potential described in Ref.~\cite{gaunt2013}, using the same potentials (a cylindrical ``box'' with length 63\,$\mu$m and diameter 30\,$\mu$m) and condensate number ($N=10^5$ $^{87}$Rb atoms).
-We set one scattering length to $5$\,nm and the others to zero: $a_{00}=5$\,nm, $a_{10}=a_{11}=0$.
+We set one s-wave scattering length to $5$\,nm and the others to zero: $a_{00}=5$\,nm, $a_{10}=a_{11}=0$.
+The associated nonlinearities are given by Eq.~(\ref{eqUij}).
 This system produced number squeezing of 15dB, and an extremely low non-condensed fraction of $4 \times 10^{-4}$.
@@
 Furthermore, the nonlinearities between multiple internal atomic states are not yet known for many atomic species, and often have sensitive dependence on external magnetic fields due to Feshbach resonances \cite{cornish:2000,altin:2010}.
-This paper analyses the requirements for producing useful squeezing for systems with large numbers of atoms.
-
-We have presented an analytic solution to the two-mode model of atomic fields with a strong Kerr non-linearity, allowing for a rapid estimate of the atomic number difference squeezing that can be obtained from these systems as scattering length data is obtained.
+This paper analyses the requirements for producing useful squeezing via these nonlinearities for systems with large numbers of atoms.
+
+We have presented an analytic solution to the two-mode model of atomic fields with a strong Kerr nonlinearity, allowing for a rapid estimate of the atomic number difference squeezing that can be obtained from these systems as scattering length data is obtained.
 In general, it predicts arbitrarily good squeezing in the two-mode limit.
@@
 
-\bibitem {roosET2006} Roos C F, Chwalla M, Kim K, Riebe M, and Blatt R Designer atoms for
+\bibitem{roosET2006} Roos C F, Chwalla M, Kim K, Riebe M, and Blatt R Designer atoms for
 quantum metrology 2006 \emph{Nature} {\bf 443}, 316
@@
 \bibitem{robins:2013} Robins N P, Altin P A, Debs J E and Close J D 2013 Atom lasers: Production, properties and prospects for precision inertial measurement \emph{Phys. Reports} {\bf 529} 265
+
+\bibitem{altinET2013} Altin P A \emph{et al.} Precision atomic gravimeter based on Bragg diffraction 2013 \emph{New Journal of Physics} {\bf 15} 023009
 

### 2013-07-17 large_number_squeezing_with_atoms.tex
Rewritten Josephson Hamiltonian section and some proofreading changes.

@@
 We examine these processes, and suggest methods to mitigate them.
-We conclude that high levels of squeezing with large numbers of atoms is feasible, but can realistically only be achieved in traps with particular trap geometries.
+We conclude that high levels of squeezing with large numbers of atoms is feasible, but can realistically only be achieved in particular trap geometries.
 We also introduce a method of maximising the measurable squeezing by using a $\pi$-pulse during the process to improve spatial mode-matching.
@@
 Section~\ref{sec:pipulse} shows that the application of a $\pi$-pulse during the process can significantly improve mode-matching issues.
-Section~\ref{sec:2D3Dsqueezing} investigates measurable squeezing in 2D and 3D, and uses a Bogoliubov analysis to explain the scaling with dimensionality.
+Section~\ref{sec:2D3Dsqueezing} investigates measurable squeezing in 2D and 3D, and uses a Bogoliubov analysis to explain the scaling with system size and dimensionality.
 Section~\ref{sectionConclusion} describes the conclusion that large volume, flat-bottomed, three-dimensional traps, which are obviously compatible with large atom number, can produce extremely high degrees of squeezing.
 
-\section{Model and two mode solutions}
+\section{Model and single-mode solutions}
 \label{secTwoModeAnalytic}
@@
 \begin{equation}
-U_{ij} = \frac{4 \pi \hbar^2 m} {a_{ij}}
+U_{ij} = \frac{4 \pi \hbar^2 a_{ij}} {m}
 \label{eqUij}
@@
 \begin{eqnarray}
-\hat{S}_+ &=& \hat{a}^{\dagger} \hat{b} \\
-\hat{S}_- &=& \hat{a} \hat{b}^{\dagger} \\
-\hat{S}_x &=& \frac{1}{2} \left( \hat{S}_+ + \hat{S}_-\right) \\
-\hat{S}_y &=& \frac{1}{2i} \left( \hat{S}_+ - \hat{S}_-\right) \\
-\hat{S}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
-\end{eqnarray}
-Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) is transformed into the Josephson Hamiltonian \cite{steel+collett}
-\begin{equation}
-\hat{H}_J = \Delta \omega \, \hat{S}_z + \chi \hat{J}_z^2 + \Omega \hat{J}_x
-\label{eqJosephson}
-\end{equation}
-where $\Delta \omega$ is a function of the various nonlinearities and $\chi=\chi_{aa}+\chi_{bb}-2\chi_{ab}$. The system described by $\hat{H}_J$ is then solved by using angular momentum algebra.
-
-We take a different approach and solve Eq.~(\ref{eqTwoModeHamiltonian}) directly for two reasons.
+\hat{J}_+ &=& \hat{a}^{\dagger} \hat{b} \\
+\hat{J}_- &=& \hat{a} \hat{b}^{\dagger} \\
+\hat{J}_x &=& \frac{1}{2} \left( \hat{J}_+ + \hat{J}_-\right) \\
+\hat{J}_y &=& \frac{1}{2i} \left( \hat{J}_+ - \hat{J}_-\right) \\
+\hat{J}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
+\end{eqnarray}
+Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) becomes
+\begin{equation}
+\hat{H} = \hbar\Delta \omega(\hat{N}) \, \hat{J}_z + \frac{1}{2}\hbar\chi_+ \hat{J}_z^2 + 2\hbar\Omega \hat{J}_x - \frac{1}{4}\left(\chi_{aa} + \chi_{bb}\right)\hat{N} + \frac{1}{8} \left(\chi_{aa} + \chi_{bb} + 2\chi_{ab}\right) \hat{N}^2
+\label{eqNonlinearJosephson}
+\end{equation}
+where $\hat{N}=\hat{n}_1 + \hat{n}_2$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_1 - \chi_2\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$.
+For initial states which are total number Fock states, the last two terms are physically meaningless phase shifts and the first term can be removed by an appropriate choice of detuning.
+In this case Eq.~\eqref{eqNonlinearJosephson} reduces to the Josephson Hamiltonian with zero detuning \cite{steel+collett}
+\begin{equation}
+  \hat{H}_J = \frac{1}{2} \hbar\chi_+ \hat{J}_z^2 + 2\hbar \Omega \hat{J}_x. \label{eqJosephson}
+\end{equation}
+The system described by $\hat{H}_J$ is then solved using angular momentum algebra.
+
+The reduction of Eq.~\eqref{eqTwoModeHamiltonian} to the Josephson Hamiltonian has led to the common misconception that no squeezing can be generated in this system if $\chi_+ = 0$ as the effects of the $\Delta\omega$ term are usually dismissed as an irrelevant detuning.
+This is not true in general for initial states other than a total number Fock state as the $\Delta\omega(\hat{N})\hat{J}_z$ term leads to a number-dependent detuning.
+We demonstrate in this paper that for the case of an initial coherent state this term leads to squeezing.
+A spin-echo pulse applied half-way through the experiment which exchanges the population of modes $\ket{a}$ and $\ket{b}$ (which is the protocol used in most spin squeezing experiments) also cancels the effects of this term, and in this case squeezing is only generated if $\chi_+ \neq 0$.
+In this work we directly solve Eq.~\eqref{eqTwoModeHamiltonian} demonstrating that squeezing is generated (without a spin-echo pulse) even if $\chi_+=0$.
+% We discuss the effects of a spin-echo pulse in Section~\ref{sec:pipulse}
+
+% We take a different approach and solve Eq.~(\ref{eqTwoModeHamiltonian}) directly for two reasons.
 First, we want access to the number variances of each state, not just the variance of the number difference between the states.

### 2013-07-17 large_number_squeezing_with_atoms.tex
Minor changes to Josephson paragraph.

@@
 A spin-echo pulse applied half-way through the experiment which exchanges the population of modes $\ket{a}$ and $\ket{b}$ (which is the protocol used in most spin squeezing experiments) also cancels the effects of this term, and in this case squeezing is only generated if $\chi_+ \neq 0$.
-In this work we directly solve Eq.~\eqref{eqTwoModeHamiltonian} demonstrating that squeezing is generated (without a spin-echo pulse) even if $\chi_+=0$.
+In this work we directly solve Eq.~\eqref{eqTwoModeHamiltonian} demonstrating that squeezing is generated (in the absence of a spin-echo pulse) even if $\chi_+=0$.
 % We discuss the effects of a spin-echo pulse in Section~\ref{sec:pipulse}

### 2013-07-17 large_number_squeezing_with_atoms.tex
Fix typo in Josephson paragraph.

@@
 \end{equation}
-where $\hat{N}=\hat{n}_1 + \hat{n}_2$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_1 - \chi_2\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$.
+where $\hat{N}=\hat{a}^\dagger\hat{a} + \hat{b}^\dagger\hat{b}$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_1 - \chi_2\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$.
 For initial states which are total number Fock states, the last two terms are physically meaningless phase shifts and the first term can be removed by an appropriate choice of detuning.

### 2013-07-17 large_number_squeezing_with_atoms.tex
More typos.

--- 2013-07-17 large_number_squeezing_with_atoms.tex
@@
 \end{equation}
-where $\hat{N}=\hat{n}_1 + \hat{n}_2$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_1 - \chi_2\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$.
+where $\hat{N}=\hat{a}^\dagger\hat{a} + \hat{b}^\dagger\hat{b}$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_{aa} - \chi_{bb}\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$.
 For initial states which are total number Fock states, the last two terms are physically meaningless phase shifts and the first term can be removed by an appropriate choice of detuning.

### 2013-07-17 large_number_squeezing_with_atoms.tex
Punctuating equations.

@@
 \begin{equation}
-U_{ij} = \frac{4 \pi \hbar^2 a_{ij}} {m}
+U_{ij} = \frac{4 \pi \hbar^2 a_{ij}} {m},
 \label{eqUij}
@@
 \begin{equation}
-\hat{H} = \frac{U_{aa}}{2} \int dV \sum_{nmpq} u_n^* u_m^* u_p u_q \, \hat{a}^{\dagger}_{n} \hat{a}^{\dagger}_{m} \hat{a}_{p} \hat{a}_{q} 
+\hat{H} = \frac{U_{aa}}{2} \int dV \sum_{nmpq} u_n^* u_m^* u_p u_q \, \hat{a}^{\dagger}_{n} \hat{a}^{\dagger}_{m} \hat{a}_{p} \hat{a}_{q} .
 \end{equation}
@@
           + \hbar\frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}
-          + \hbar\Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger}  \hat{a} )
+          + \hbar\Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger}  \hat{a} ),
 \label{eqTwoModeHamiltonian}
@@
 \begin{eqnarray}
-\hat{J}_+ &=& \hat{a}^{\dagger} \hat{b} \\
-\hat{J}_- &=& \hat{a} \hat{b}^{\dagger} \\
-\hat{J}_x &=& \frac{1}{2} \left( \hat{J}_+ + \hat{J}_-\right) \\
-\hat{J}_y &=& \frac{1}{2i} \left( \hat{J}_+ - \hat{J}_-\right) \\
+\hat{J}_+ &=& \hat{a}^{\dagger} \hat{b}, \\
+\hat{J}_- &=& \hat{a} \hat{b}^{\dagger}, \\
+\hat{J}_x &=& \frac{1}{2} \left( \hat{J}_+ + \hat{J}_-\right), \\
+\hat{J}_y &=& \frac{1}{2i} \left( \hat{J}_+ - \hat{J}_-\right), \\
 \hat{J}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
@@
 \begin{equation}
-\hat{H} = \hbar\Delta \omega(\hat{N}) \, \hat{J}_z + \frac{1}{2}\hbar\chi_+ \hat{J}_z^2 + 2\hbar\Omega \hat{J}_x - \frac{1}{4}\left(\chi_{aa} + \chi_{bb}\right)\hat{N} + \frac{1}{8} \left(\chi_{aa} + \chi_{bb} + 2\chi_{ab}\right) \hat{N}^2
+\hat{H} = \hbar\Delta \omega(\hat{N}) \, \hat{J}_z + \frac{1}{2}\hbar\chi_+ \hat{J}_z^2 + 2\hbar\Omega \hat{J}_x - \frac{1}{4}\left(\chi_{aa} + \chi_{bb}\right)\hat{N} + \frac{1}{8} \left(\chi_{aa} + \chi_{bb} + 2\chi_{ab}\right) \hat{N}^2,
 \label{eqNonlinearJosephson}
@@
 We solve this system analytically, and derive expressions for absolute and relative number squeezing.
-We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation} as it is somewhat laborious.
+We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}.
 If we define
 \begin{eqnarray}
-s &=& \sin(\theta) \label{eqsDef} \\
-c &=& \cos(\theta) \\
-\theta &=& \Omega(t_3 - t_2) \\
-\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} \\
-A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
-A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
-B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
-B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
-D &=& AB^* - A^*B \label{eqDDef}
+s &=& \sin(\theta), \label{eqsDef} \\
+c &=& \cos(\theta), \\
+\theta &=& \Omega(t_3 - t_2), \\
+\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}}, \\
+A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )], \\
+A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )], \\
+B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )], \\
+B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )], \\
+D &=& AB^* - A^*B, \label{eqDDef}
 \end{eqnarray}
@@
 \begin{eqnarray}
-N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D \\
-%
-N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D \\
+N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D, \\
+%
+N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D, \\
 %
@@
        && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
-       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ) \\
+       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ), \\
 %
@@
        && + 2 i c^3 s n_b (e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - e^{i(\lambda_{bb} - \lambda_{ab})} A B^* ) \nonumber \\
-       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ) \\
+       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ), \\
 %
@@
        && + 4 i c s (c^2 - s^2) \left( n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B \right. \nonumber \\
-       && \,\,\,\,\, \left. + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^* \right) \label{eqNumDiffVariance}
+       && \,\,\,\,\, \left. + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^* \right). \label{eqNumDiffVariance}
 \end{eqnarray}
@@
 \begin{equation}
-\langle :\hat{\psi}^\dagger(\mathbf{x_1})\cdots\hat{\psi}^\dagger(\mathbf{x_n})\hat{\psi}(\mathbf{y_1})\cdots\hat{\psi}(\mathbf{y_m}):_{\mbox{sym}}\rangle = \mathbb{E}\left[\phi^*(\mathbf{x_1})\cdots\phi^*(\mathbf{x_n})\phi(\mathbf{y_1})\cdots\phi(\mathbf{y_m})\right]
+\langle :\hat{\psi}^\dagger(\mathbf{x_1})\cdots\hat{\psi}^\dagger(\mathbf{x_n})\hat{\psi}(\mathbf{y_1})\cdots\hat{\psi}(\mathbf{y_m}):_{\mbox{sym}}\rangle = \mathbb{E}\left[\phi^*(\mathbf{x_1})\cdots\phi^*(\mathbf{x_n})\phi(\mathbf{y_1})\cdots\phi(\mathbf{y_m})\right],
 \label{eqExpectationRelations}
@@
 \begin{eqnarray}
-\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{r}) + U_{a a} n_a + U_{a b} \tilde{n}_b  \right)\phi_a   + \kappa \phi_b  \nonumber \\
-\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{r}) + U_{b b} n_b + U_{a b} \tilde{n}_a   \right)\phi_b   + \kappa^* \phi_a
+\frac{d \phi_a}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_a(\mathbf{r}) + U_{a a} n_a + U_{a b} \tilde{n}_b  \right)\phi_a   + \kappa \phi_b , \nonumber \\
+\frac{d \phi_b}{dt} = -\frac{i}{\hbar}\left(-\frac{\hbar^2}{2 m}\nabla^2+V_b(\mathbf{r}) + U_{b b} n_b + U_{a b} \tilde{n}_a   \right)\phi_b   + \kappa^* \phi_a,
   \label{eqStochasticEquations}
@@
 Using Eq.~(\ref{eqChiUequivalence}) we find the following equivalences
-\begin{eqnarray}
-\hbar \chi &=& U \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}} \,\,\,\, ({\textrm{Gaussian}}) \\
-%
-\hbar \chi &=& \frac{4}{7} \left( \frac{15 U \omega_x \omega_y \omega_z}{16 \pi \sqrt{2}} \right)^{2/5} \left( \frac{m}{N} \right)^{3/5} \,\,\,\, ({\textrm{Thomas-Fermi}}) \\
-%
-\hbar \chi &=& \frac{U}{V} \,\,\,\, {\textrm{(constant density)}}
-\end{eqnarray}
+\begin{align}
+\hbar \chi &= U \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}}, &&({\textrm{Gaussian}}) \\
+%
+\hbar \chi &= \frac{4}{7} \left( \frac{15 U \omega_x \omega_y \omega_z}{16 \pi \sqrt{2}} \right)^{2/5} \left( \frac{m}{N} \right)^{3/5}, &&({\textrm{Thomas-Fermi}}) \\
+%
+\hbar \chi &= \frac{U}{V}, &&{\textrm{(constant density)}}
+\end{align}
 %

### 2013-07-18 large_number_squeezing_with_atoms.tex
Simplifying language and converting some US-isms to Australian English for consistency.

@@
 The system is prepared with all the population initially in $|a\rangle$, and vacuum in $|b\rangle$.
-We choose the initial state $|a\rangle$ to be a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$, where $N$ is the total particle number.
-As the system is insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, or a mixture of Poissonian-distributed number states.
+We choose the initial state $|a\rangle$ to be a coherent state with average number $\langle \hat{a}^{\dagger} \hat{a} \rangle = N$.
+As the system is insensitive to the initial phase of the coherent state, this is equivalent to a mixture of coherent states of uncertain phase, or a mixture of Poisson-distributed number states.
 Such a state is consistent with BEC coherence experiments \cite{Hadzibabic2004}.
@@
 
-In Figure \ref{figTwoModeAnalyticExamples} we plot a solution for the normalized number variance in mode $|a\rangle$ as a function of $\phi$, choosing a typical set of nonlinearities.
+In Figure \ref{figTwoModeAnalyticExamples} we plot a solution for the normalised number variance in mode $|a\rangle$ as a function of $\phi$, choosing a typical set of nonlinearities.
 This example illustrates the basic features of the system: during the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=\pi / \Omega$, and has a series of minima.
@@
     \caption{Effects of mode shape on squeezing.
-The plot shows the normalized number variance in mode $|a \rangle$ as a function of the time the final coupling pulse is applied.
+The plot shows the normalised number variance in mode $|a \rangle$ as a function of the time the final coupling pulse is applied.
 We compare the analytic two-mode solution (thick red), the constant density mode (solid blue), the Thomas-Fermi mode (dashed) and the Gaussian mode (dash-dotted).
@@
 \end{equation}
-where we have defined $\lambda_{ij} = \chi_{ij} (t_2-t_1)$. Utilizing (\ref{eqHadamard}) to move $\hat{a}_1$ through the exponential, with some algebra one can show that
+where we have defined $\lambda_{ij} = \chi_{ij} (t_2-t_1)$. Using (\ref{eqHadamard}) to move $\hat{a}_1$ through the exponential, with some algebra one can show that
 \begin{eqnarray}
@@
 \end{equation}
-Finally, to obtain $\hat{a}_3$ and $\hat{b}_3$ we utilize (\ref{eqa1}) and (\ref{eqb1}) with the phase factor attached to the $\hat{b}$ operator to obtain
+Finally, to obtain $\hat{a}_3$ and $\hat{b}_3$ we use (\ref{eqa1}) and (\ref{eqb1}) with the phase factor attached to the $\hat{b}$ operator to obtain
 \begin{eqnarray}

### 2013-07-18 large_number_squeezing_with_atoms.tex
Add new figure showing squeezing in case where effective chi is zero

--- 2013-07-17 large_number_squeezing_with_atoms.tex
@@
 We examine these processes, and suggest methods to mitigate them.
-We conclude that high levels of squeezing with large numbers of atoms is feasible, but can realistically only be achieved in particular trap geometries.
+We conclude that high levels of squeezing with large numbers of atoms is feasible, but can realistically only be achieved in traps with particular trap geometries.
 We also introduce a method of maximising the measurable squeezing by using a $\pi$-pulse during the process to improve spatial mode-matching.
@@
 Section~\ref{sec:pipulse} shows that the application of a $\pi$-pulse during the process can significantly improve mode-matching issues.
-Section~\ref{sec:2D3Dsqueezing} investigates measurable squeezing in 2D and 3D, and uses a Bogoliubov analysis to explain the scaling with system size and dimensionality.
+Section~\ref{sec:2D3Dsqueezing} investigates measurable squeezing in 2D and 3D, and uses a Bogoliubov analysis to explain the scaling with dimensionality.
 Section~\ref{sectionConclusion} describes the conclusion that large volume, flat-bottomed, three-dimensional traps, which are obviously compatible with large atom number, can produce extremely high degrees of squeezing.
 
-\section{Model and single-mode solutions}
+\section{Model and two mode solutions}
 \label{secTwoModeAnalytic}
@@
 \begin{equation}
-U_{ij} = \frac{4 \pi \hbar^2 a_{ij}} {m}
+U_{ij} = \frac{4 \pi \hbar^2 m} {a_{ij}}
 \label{eqUij}
@@
 \begin{eqnarray}
-\hat{J}_+ &=& \hat{a}^{\dagger} \hat{b} \\
-\hat{J}_- &=& \hat{a} \hat{b}^{\dagger} \\
-\hat{J}_x &=& \frac{1}{2} \left( \hat{J}_+ + \hat{J}_-\right) \\
-\hat{J}_y &=& \frac{1}{2i} \left( \hat{J}_+ - \hat{J}_-\right) \\
-\hat{J}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
-\end{eqnarray}
-Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) becomes
-\begin{equation}
-\hat{H} = \hbar\Delta \omega(\hat{N}) \, \hat{J}_z + \frac{1}{2}\hbar\chi_+ \hat{J}_z^2 + 2\hbar\Omega \hat{J}_x - \frac{1}{4}\left(\chi_{aa} + \chi_{bb}\right)\hat{N} + \frac{1}{8} \left(\chi_{aa} + \chi_{bb} + 2\chi_{ab}\right) \hat{N}^2
-\label{eqNonlinearJosephson}
-\end{equation}
-where $\hat{N}=\hat{a}^\dagger\hat{a} + \hat{b}^\dagger\hat{b}$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_{aa} - \chi_{bb}\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$.
-For initial states which are total number Fock states, the last two terms are physically meaningless phase shifts and the first term can be removed by an appropriate choice of detuning.
-In this case Eq.~\eqref{eqNonlinearJosephson} reduces to the Josephson Hamiltonian with zero detuning \cite{steel+collett}
-\begin{equation}
-  \hat{H}_J = \frac{1}{2} \hbar\chi_+ \hat{J}_z^2 + 2\hbar \Omega \hat{J}_x. \label{eqJosephson}
-\end{equation}
-The system described by $\hat{H}_J$ is then solved using angular momentum algebra.
-
-The reduction of Eq.~\eqref{eqTwoModeHamiltonian} to the Josephson Hamiltonian has led to the common misconception that no squeezing can be generated in this system if $\chi_+ = 0$ as the effects of the $\Delta\omega$ term are usually dismissed as an irrelevant detuning.
-This is not true in general for initial states other than a total number Fock state as the $\Delta\omega(\hat{N})\hat{J}_z$ term leads to a number-dependent detuning.
-We demonstrate in this paper that for the case of an initial coherent state this term leads to squeezing.
-A spin-echo pulse applied half-way through the experiment which exchanges the population of modes $\ket{a}$ and $\ket{b}$ (which is the protocol used in most spin squeezing experiments) also cancels the effects of this term, and in this case squeezing is only generated if $\chi_+ \neq 0$.
-In this work we directly solve Eq.~\eqref{eqTwoModeHamiltonian} demonstrating that squeezing is generated (in the absence of a spin-echo pulse) even if $\chi_+=0$.
-% We discuss the effects of a spin-echo pulse in Section~\ref{sec:pipulse}
-
-% We take a different approach and solve Eq.~(\ref{eqTwoModeHamiltonian}) directly for two reasons.
+\hat{S}_+ &=& \hat{a}^{\dagger} \hat{b} \\
+\hat{S}_- &=& \hat{a} \hat{b}^{\dagger} \\
+\hat{S}_x &=& \frac{1}{2} \left( \hat{S}_+ + \hat{S}_-\right) \\
+\hat{S}_y &=& \frac{1}{2i} \left( \hat{S}_+ - \hat{S}_-\right) \\
+\hat{S}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
+\end{eqnarray}
+Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) is transformed into the Josephson Hamiltonian \cite{steel+collett}
+\begin{equation}
+\hat{H}_J = \Delta \omega \, \hat{S}_z + \chi \hat{J}_z^2 + \Omega \hat{J}_x
+\label{eqJosephson}
+\end{equation}
+where $\Delta \omega$ is a function of the various nonlinearities and $\chi=\chi_{aa}+\chi_{bb}-2\chi_{ab}$. The system described by $\hat{H}_J$ is then solved by using angular momentum algebra.
+
+We take a different approach and solve Eq.~(\ref{eqTwoModeHamiltonian}) directly for two reasons.
 First, we want access to the number variances of each state, not just the variance of the number difference between the states.
@@
 
-%Can Uaa + Ubb - 2Uab be zero and still get squeezing? Yep, Uaa=0.03, Ubb=0.01, Uab=0.02 gives squeezing for all quantities of interest
+%Can Uaa + Ubb - 2Uab be zero and still get squeezing?
+Yep, Uaa=0.03, Ubb=0.01, Uab=0.02, na=5e4, nb=5e4, thold = 2e-3, phi=1.92  gives squeezing in |a>.
+If we use a pi pulse, however, then Uaa + Ubb - 2Uab is the only parameter which matters.
+
+% Uaa=0.03, Ubb=0.01, Uab=0.02, na=2e5, nb=2e5, thold = 2e-3. max squeeze in |a> phi = 1.67. max squeeze in |b> phi = 4.81. max squeeze in |a> -  |b> phi = 1.55
+%4e-6
 
@@
     \label{figTwoModeAnalyticExamples}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=10cm]{figures/zero_effective_chi.eps}
+    \caption{Relative number squeezing in state $|a\rangle$ (solid line) and number difference squeezing between states $|a\rangle$ and $|b\rangle$ (dashed line) as a function of final recombination time in the case where $\chi_{aa} + \chi_{bb} - 2 \chi_{ab} = 0$.
+Parameters: $n_a = n_b = 2 \times 10^5$, $\chi_{aa}=0.03\, \text{s}^{-1}, \chi_{ab}=0.02, \chi_{bb}=0.01\,\text{s}^{-1}, \tau_{\mathrm{hold}}=2\times 10^{-3}$s.
+Solid line, $\phi=1.67$; dashed line, $\phi=1.55$.}
+    \label{figZeroEffectiveChi}
 \end{figure}

### 2013-07-18 large_number_squeezing_with_atoms.tex
Reworked \pi pulse section.

@@
 \DeclareMathOperator{\sinc}{sinc}
+\DeclareMathOperator{\erf}{Erf}
 \newcommand{\abs}[1]{\ensuremath{\left|#1\right|}}
@@
 
-The zero-dimensional, two-mode model described in the previous section shows that arbitrarily high squeezing can be achieved using the intrinsic nonlinearities in a BEC, but it ignores the existence of multiple spatial modes.
+The zero-dimensional, two-mode model described in the previous section shows that arbitrarily good squeezing can be achieved using the intrinsic nonlinearities in a BEC, but it ignores the existence of multiple spatial modes.
 There are three problems caused by the existence of multiple modes, and these will limit the achievable squeezing.
@@
 This should enable higher squeezing.
-We cannot make strong conclusions without also considering the effects of coupling between spatial modes, however, which leads to the third potential problem: coupling between modes is inevitable in the presence of a nonlinearity, meaning that the modes cannot be analyzed independently.
+We cannot make strong conclusions without also considering the effects of coupling between spatial modes, however, which leads to the third potential problem: coupling between modes is inevitable in the presence of a nonlinearity, meaning that the modes cannot be analysed independently.
 Coupling between modes will disturb mode-matching, as well as mixing fields with different phase evolution, which can rapidly destroy squeezing.
@@
 \begin{equation}
-\hbar\chi = U \frac{{\mathrm{Erf}} (\sqrt{2p})^3}{{\mathrm{Erf}}(p)^6} \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}}.
+\hbar\chi = U \frac{\erf (\sqrt{2p})^3}{\erf(p)^6} \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}}.
 \end{equation}
@@
 \section{Using a $\pi$-pulse to improve mode-matching} \label{sec:pipulse}
+
 Modal mismatch is generated by the different atomic potentials seen by the two internal states.
-A straightforward method of minimising this effect is to use a $\pi$-pulse to swap the populations of the internal states after half of the evolution.
-As can be seen by Eqs~(\ref{eqa1}) and (\ref{eqb1}), such a pulse also leaves the quantum statistics of each state unchanged.
-In the limit where each state remains in a single mode, and exactly half of the atoms are in each state, this technique ensures that both modes have identical mode shapes after the hold time.
-In practice this will never work perfectly --- due to the density distributions of $\psi_a(\mathbf{r})$ and $\psi_a(\mathbf{r})$ not reflecting the ground state, resulting in non-identical environments for the two modes, some modal dynamics will occur.
+A straightforward method of minimising this effect in the case that modes $\ket{a}$ and $\ket{b}$ are equally occupied after the initial coupling pulse is to apply a $\pi$-pulse to swap the populations of the internal states after half of the evolution.
+As can be seen by Eqs.~\eqref{eqa1} and \eqref{eqb1}, such a pulse also leaves the quantum statistics of each state unchanged.
+The idea of this scheme is to make the effective potential seen by the atoms in state $\ket{a}$ during the second half of the hold time very similar to that seen by the atoms that were in state $\ket{a}$ during the first half of the hold time.
+In the limit of short hold times and for equal occupations of the two modes, this technique reduces the difference in the mode shapes after the hold time.
+In practice this will never work perfectly --- due to the density distributions of $\psi_a(\mathbf{r})$ and $\psi_b(\mathbf{r})$ not being in the ground state, their mode shapes will have changed by the time the $\pi$-pulse is applied, and so the spatial dynamics in the second half of the hold time will not be quite the same as those during the first half.
 Nonetheless, in many regimes, such a $\pi$-pulse can produce some improvement.

### 2013-07-18 large_number_squeezing_with_atoms.tex
Added a reference in the text to the new figure showing that squeezing is possible even when \chi_+ = 0.

@@
 It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
+As mentioned previously, squeezing is possible in this system even when the nonlinearity in the reduced Josephson Hamiltonian $\chi_+=\chi_{aa} + \chi_{bb} - 2\chi_{ab} = 0$.
+Figure~\ref{figZeroEffectiveChi} demonstrates that significant squeezing is possible in this system if a spin-echo pulse is not used.
 
@@
     \includegraphics[width=10cm]{figures/analytic_two_mode_examples.eps}
-    \caption{Normalized number variance Var$[N_a]/N_a$ in mode $|a \rangle$ as a function of final recombination time plotted on a logarithmic scale, using four different mixing angles $\phi$.
+    \caption{Normalised number variance Var$[N_a]/N_a$ in mode $|a \rangle$ as a function of final recombination time plotted on a logarithmic scale, using four different mixing angles $\phi$.
 Parameters: $n_a = n_b =5 \times 10^5$, $\chi_{aa}=0.04\, \text{s}^{-1}, \chi_{ab}=0, \chi_{bb}=0.01\,\text{s}^{-1}, \tau_{\mathrm{hold}}=4\times 10^{-4}$s.
@@
     \includegraphics[width=10cm]{figures/zero_effective_chi.eps}
-    \caption{Relative number squeezing in state $|a\rangle$ (solid line) and number difference squeezing between states $|a\rangle$ and $|b\rangle$ (dashed line) as a function of final recombination time in the case where $\chi_{aa} + \chi_{bb} - 2 \chi_{ab} = 0$.
+    \caption{Relative number squeezing in state $|a\rangle$ (solid line) and number difference squeezing between states $|a\rangle$ and $|b\rangle$ (dashed line) as a function of final recombination time in the case where $\chi_+ = \chi_{aa} + \chi_{bb} - 2 \chi_{ab} = 0$.
 Parameters: $n_a = n_b = 2 \times 10^5$, $\chi_{aa}=0.03\, \text{s}^{-1}, \chi_{ab}=0.02, \chi_{bb}=0.01\,\text{s}^{-1}, \tau_{\mathrm{hold}}=2\times 10^{-3}$s.

### 2013-07-18 large_number_squeezing_with_atoms.tex
More changes to the \pi pulse section.

@@
 Modal mismatch is generated by the different atomic potentials seen by the two internal states.
-A straightforward method of minimising this effect in the case that modes $\ket{a}$ and $\ket{b}$ are equally occupied after the initial coupling pulse is to apply a $\pi$-pulse to swap the populations of the internal states after half of the evolution.
+A straightforward method of minimising this effect is to apply a $\pi$-pulse to swap the populations of the internal states after half of the evolution.
 As can be seen by Eqs.~\eqref{eqa1} and \eqref{eqb1}, such a pulse also leaves the quantum statistics of each state unchanged.
-The idea of this scheme is to make the effective potential seen by the atoms in state $\ket{a}$ during the second half of the hold time very similar to that seen by the atoms that were in state $\ket{a}$ during the first half of the hold time.
-In the limit of short hold times and for equal occupations of the two modes, this technique reduces the difference in the mode shapes after the hold time.
+The idea of this scheme is to make the effective potential seen by the atoms in state $\ket{a}$ during the second half of the hold time very similar to that seen by the atoms that were in state $\ket{a}$ during the first half of the hold time, and will only work if the modes $\ket{a}$ and $\ket{b}$ have an equal occupation at the start of the hold time, or if $\chi_+=\chi_{aa} + \chi_{bb} - 2\chi_{ab}=0$.
+In the latter case, this reduces to a spin-echo pulse, and as the squeezing effects of the $\Delta\omega$ term in Eq.~\eqref{eqNonlinearJosephson} are cancelled in this case, there will be no squeezing.
+However, in the limit of short hold times, for equal occupations of the two modes, and when $\chi_+\neq 0$, this technique reduces the difference in the mode shapes after the hold time.
 In practice this will never work perfectly --- due to the density distributions of $\psi_a(\mathbf{r})$ and $\psi_b(\mathbf{r})$ not being in the ground state, their mode shapes will have changed by the time the $\pi$-pulse is applied, and so the spatial dynamics in the second half of the hold time will not be quite the same as those during the first half.

### 2013-07-22 large_number_squeezing_with_atoms.tex
Punctuation.
Tweak intro and add more references

@@
 Ultracold atomic interferometers are also reaching state of the art sensitivities in precision measurements \cite{robins:2013, altinET2013}.
-While all current atom interferometry is limited by technical noise sources, it appears that BEC-based coherent sources offer the best chance of overcoming some of these issues, for the same reasons that optical lasers are typically preferred for optical interferometry \cite{szigeti:2012}.
-If these technical noise sources can be controlled, atom interferometers will be limited by atomic shot noise, which leads to a sensitivity that scales as the square root of the atomic flux.
-From that point, sensitivity will only be improved by increasing the atomic flux, and/or using non-classical quantum states such as squeezed states to go below the shot noise limit \cite{robins:2013}.
+While all current atom interferometry is limited by technical noise sources, it appears that BEC-based coherent sources offer the best chance of overcoming some of these issues, for the same reasons that optical lasers are typically preferred for optical interferometry.
+In particular, it appears that Bose-condensed sources allow for higher fidelity mirrors and beamsplitters \cite{szigeti:2012}, are less sensitive to the effects of distortions in the optical wavefronts of the beamsplitters and mirrors and the Coriolis effect \cite{Debs:2011, robins:2013}, and are robust to the adverse effects of strong outcoupling when feedback-cooled \cite{semiclassicalstabilization, Szigeti:2009, Szigeti:2010}.
+
+If these technical noise sources can be reduced and/or controlled, atom interferometers will be limited by atomic shot noise, which leads to a sensitivity that scales as the square root of the atomic flux.
+From that point, sensitivity will only be improved by increasing the atomic flux and brightness \cite{semiclassicallimit2007, outcouplercomparison, jeppesenET2008}, and/or using non-classical quantum states such as squeezed states to go below the shot noise limit \cite{robins:2013}.
+
 Producing squeezed states for metrological purposes is therefore only relevant when it can be achieved in a context of large numbers of atoms.
@@
 Other system parameters such as the total number $N$ and the occupation of the $\ket{a}$ and $\ket{b}$ states at $t_1$ ($n_a$ and $n_b$, respectively) are kept constant and chosen to maximise the squeezing of the equivalent two-mode model presented in Section~\ref{secTwoModeAnalytic}.
-In particular this fixes the values of $\theta$, $\phi$ and $\lambda_{ij} = \chi_{ij}\tau_\text{hold}$.
+In particular, this fixes the values of $\theta$, $\phi$, and $\lambda_{ij} = \chi_{ij}\tau_\text{hold}$.
 As the size of the condensate is scaled, $\hbar \chi_{ij} = U^\text{1D}_{ij}/L = U^\text{2D}_{ij}/A = U^\text{3D}_{ij}/V$ will change, but we will vary $\tau_\text{hold}$ to keep $\lambda_{ij}$ constant.
@@
 
+\bibitem{Debs:2011} Debs J E, Altin P A, Barter T H, D{\"{o}}ring D, Dennis G R, McDonald G, Anderson R P, Close J D, and
+Robins N P 2011 Cold-atom gravimetry with a Bose-Einstein condensate {\emph Phys. Rev.} A {\bf 84} 033610
+
+\bibitem{semiclassicalstabilization} Johnsson M T, Haine S A and Hope J J 2005 Stabilizing an atom laser using spatially selective pumping and feedback \emph{Phys. Rev.} A {\bf 72} 053603
+
+\bibitem{Szigeti:2009} Szigeti S S, Hush M R, Carvalho A R R, and Hope J J 2009 Continuous measurement feedback control of a Bose-Einstein condensate using phase-contrast imaging \emph{Phys.
+Rev.} A {\bf 80} 013614
+
+\bibitem{Szigeti:2010} S.
+S.
+Szigeti, M.
+R.
+Hush, A.
+R.
+R.
+Carvalho, and J.
+J.
+Hope 2010 Feedback control of an interacting Bose-Einstein condensate using phase-contrast imaging \emph{Phys.
+Rev.} A {\bf 82} 043632
+
+\bibitem{semiclassicallimit2007} Johnsson M T, Haine S A, Hope J J, Robins N P, Figl C, Jeppesen M, Dugu{\'{e}} J and Close, J 2007 Semiclassical limits to the linewidth of an atom laser \emph{Phys.
+Rev.} A {\bf 75} 043618
+
+\bibitem{outcouplercomparison} Dugu{\'{e}} J, Robins N P, Figl C, Jeppesen M, Johnsson M T, Hope J J, and Close J D 2007 Investigation and comparison of multi-state and two-state atom laser output-couplers \emph{Phys.
+Rev.} A {\bf 75} 053602
+
+\bibitem{jeppesenET2008} Jeppesen M, Dugu{\'{e}}, Dennis G R, Johnsson M T, Figl C, Robins N P and Close J D 2008 Approaching the Heisenberg limit in an atom laser \emph{Phys. Rev.} A {\bf 77} 063618
+
 \bibitem{LiET2009} Li Y, Treutlein P, Reichel J, and Sinatra A 2009 Spin squeezing in a bimodal condensate: spatial dynamics and particle losses \emph{Eur. Phys. J.} B {\bf 68} 365

### 2013-07-24 large_number_squeezing_with_atoms.tex
Go thunderbirds, go!

@@
 \usepackage{amsmath}
+\usepackage{cite}
 % \usepackage{ulem}
@@
 
-\title{Large squeezing in Bose-Einstein condensates with large numbers of atoms}
+\title{Squeezing in Bose-Einstein condensates with large numbers of atoms}
 
@@
 Atomic squeezing experiments have typically attempted to operate in the single-mode regime, where the trap is tight enough such that all higher modes are frozen out \cite{LiET2009}.
-However, tight traps limit the total number of atoms, as the 3-body recombination rate places an effective maximum atomic density in the gas, and well before that limit the interatomic interactions break the single-mode operation.
-Both of these effects limit the number of atoms that can be effectively squeezed, so we will examine a broader range of trap geometries, and calculate the dynamics in the presence of multiple spatial modes.
+However, tight traps limit the total number of atoms, as the 3-body recombination rate places an effective maximum atomic density in the gas, and well before that limit, the interatomic interactions break the single-mode operation.
+Both of these effects limit the number of atoms that can be squeezed effectively, so we will examine a broader range of trap geometries, and calculate the dynamics in the presence of multiple spatial modes.
 Many similar previous simulations have been performed in one or two spatial dimensions, by integrating out some of the remaining dimensions \cite{Kheruntsyan:2005, johnssonET2007, haineET2009, kheruntsyanET2002}.
@@
 These solutions of the ideal case provide a check on the multimode cases we consider later in the paper, as well as providing a bound on the best possible squeezing for the system.
-
 In Section \ref{sec:MMdescription} we show how multimode behaviour becomes important in physical systems, and introduce our model for simulating higher dimensional trap geometries.
@@
     \includegraphics[width=10cm]{figures/zero_effective_chi.eps}
-    \caption{Relative number squeezing in state $|a\rangle$ (solid line) and number difference squeezing between states $|a\rangle$ and $|b\rangle$ (dashed line) as a function of final recombination time in the case where $\chi_+ = \chi_{aa} + \chi_{bb} - 2 \chi_{ab} = 0$.
+    \caption{Number squeezing in state $|a\rangle$ (solid line) and number difference squeezing between states $|a\rangle$ and $|b\rangle$ (dashed line) as a function of final recombination time in the case where $\chi_+ = \chi_{aa} + \chi_{bb} - 2 \chi_{ab} = 0$.
 Parameters: $n_a = n_b = 2 \times 10^5$, $\chi_{aa}=0.03\, \text{s}^{-1}, \chi_{ab}=0.02, \chi_{bb}=0.01\,\text{s}^{-1}, \tau_{\mathrm{hold}}=2\times 10^{-3}$s.
@@
 However, in the limit of short hold times, for equal occupations of the two modes, and when $\chi_+\neq 0$, this technique reduces the difference in the mode shapes after the hold time.
-In practice this will never work perfectly --- due to the density distributions of $\psi_a(\mathbf{r})$ and $\psi_b(\mathbf{r})$ not being in the ground state, their mode shapes will have changed by the time the $\pi$-pulse is applied, and so the spatial dynamics in the second half of the hold time will not be quite the same as those during the first half.
+In practice this will never work perfectly, as $\psi_a(\mathbf{r})$ and $\psi_b(\mathbf{r})$ will typically have evolved by the time the $\pi$-pulse is applied, and so the spatial dynamics in the second half of the hold time will not be quite the same as those during the first half.
 Nonetheless, in many regimes, such a $\pi$-pulse can produce some improvement.

### 2013-10-03 large_number_squeezing_with_atoms.tex
Address comments of referee 1.
Also add the referee reports as text files plus inline comments
to the first report.

@@
 %Uncomment for PACS numbers title message
-%\pacs{00.00, 20.00, 42.10}
+\pacs{03.75.Mn, 03.75.Kk, 42.50.Dv, 03.75.Dg}
 % Keywords required only for MST, PB, PMB, PM, JOA, JOB? 
@@
 Their ability to be highly isolated from their environment and controlled precisely by optical, magnetic and rf fields makes them an ideal platform for engineering and manipulating non-trivial quantum states.
-In quantum optics, the most common way to produce entangled states has been through squeezing, and it has been recognised that the intrinsic nonlinearities of atoms naturally lead to quadrature squeezing \cite{Kheruntsyan:2005, johnssonET2007, haineET2009, doeringET2010}.
+In quantum optics, the most common way to produce entangled states has been through squeezing, and it has been recognised that the intrinsic nonlinearities of atoms naturally lead to quadrature squeezing \cite{Kheruntsyan:2005, johnssonET2007, wuester2008, haineET2009, doeringET2010}.
 Number squeezing can be achieved with spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissociation of Bose-Einstein Condensates (BECs) \cite{kheruntsyanET2002,davisET2008,savage2007,ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, double-well potentials \cite{esteve2008, maussang2010}, or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}.
@@
 where $\hat{N}=\hat{a}^\dagger\hat{a} + \hat{b}^\dagger\hat{b}$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_{aa} - \chi_{bb}\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$.
-For initial states which are total number Fock states, the last two terms are physically meaningless phase shifts and the first term can be removed by an appropriate choice of detuning.
-In this case Eq.~\eqref{eqNonlinearJosephson} reduces to the Josephson Hamiltonian with zero detuning \cite{steel+collett}
+In this form, the last two terms in Eq.~(\ref{eqNonlinearJosephson}) are proportional to powers of the number operator, and so will commute with the number variance operators for state $|a\rangle$ and $|b\rangle$.
+This means they cannot cause number squeezing of any type, and can be dropped in the search for a minimal, squeezing-generating Hamiltonian.
+Further, in the case where the system is in a total number Fock state, the first term in Eq.~(\ref{eqNonlinearJosephson}) can be removed by an appropriate choice of detuning.
+In this case, dropping the first and last two terms transforms Eq.~(\ref{eqNonlinearJosephson}) into the Josephson Hamiltonian with zero detuning \cite{steel+collett}:
 \begin{equation}
@@
 
-The reduction of Eq.~\eqref{eqTwoModeHamiltonian} to the Josephson Hamiltonian has led to the common misconception that no squeezing can be generated in this system if $\chi_+ = 0$ as the effects of the $\Delta\omega$ term are usually dismissed as an irrelevant detuning.
+The reduction of Eq.~\eqref{eqTwoModeHamiltonian} to Eq.~\eqref{eqJosephson} has led to the common misconception that no squeezing can be generated in this system if $\chi_+ = 0$ as the effects of the $\Delta\omega$ term are usually dismissed as an irrelevant detuning.
 This is not true in general for initial states other than a total number Fock state as the $\Delta\omega(\hat{N})\hat{J}_z$ term leads to a number-dependent detuning.
@@
 \begin{eqnarray}
+\tau &=& t_3 - t_2, \\
+\theta &=& \Omega \tau, \\
 s &=& \sin(\theta), \label{eqsDef} \\
 c &=& \cos(\theta), \\
-\theta &=& \Omega(t_3 - t_2), \\
 \lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}}, \\
@@
 In Figure \ref{figTwoModeAnalyticExamples} we plot a solution for the normalised number variance in mode $|a\rangle$ as a function of $\phi$, choosing a typical set of nonlinearities.
-This example illustrates the basic features of the system: during the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=\pi / \Omega$, and has a series of minima.
+This example illustrates the basic features of the system: during the final coupling period, the normalised number variance undergoes cyclic variation with a period $T=\pi / \Omega$, and has a series of minima.
 It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.
@@
 
+In the multimode case, the single mode definition for number squeezing, $\langle\mathrm{Var}[\hat{N}_i] \rangle / \langle \hat{N}_i \rangle$, where $\hat{N}_i$ is the number operator for state $|i\rangle \in \{|a\rangle, |b\rangle \}$, is no longer adequate.
+In the multimode case we define the number squeezing in states $|i\rangle$ to be
+\begin{equation}
+\mathrm{Squeezing} = \frac{\mathrm{Var} \left[ \sum_j \hat{N}_i(x_j) \right]} { \sum_j \hat{N}_i(x_j) }
+\label{eqMultimodeSqueezingDefinition}
+\end{equation}
+where the sums are over spatial modes defined as delta functions in position, and there is seperate number operator $\hat{N}_i(x_j)$ for each mode $j$ in state $|i\rangle$.
+
 \section{Effects of trap geometries on squeezing in 1D} \label{sec:1Dsim}
@@
     \caption{Effects of mode shape on squeezing.
-The plot shows the normalised number variance in mode $|a \rangle$ as a function of the time the final coupling pulse is applied.
+The plot shows the normalised number variance defined by Eq.~(\ref{eqMultimodeSqueezingDefinition}) in mode $|a \rangle$ as a function of the time the final coupling pulse is applied.
 We compare the analytic two-mode solution (thick red), the constant density mode (solid blue), the Thomas-Fermi mode (dashed) and the Gaussian mode (dash-dotted).
@@
 
+\bibitem{wuester2008} W{\"{u}}ster S, Dabrowska-W{\"{u}}ster B J, Scott S M, Close J D and Savage C M 2008
+ Quantum-field dynamics of expanding and contracting Bose-Einstein condensates \emph{Phys. Rev.} A {\bf 77} 023619
+
 \bibitem{haineET2009} Haine S A and Johnsson M T 2009 Dynamic scheme for generating number squeezing in Bose-Einstein condensates through nonlinear interactions \emph{Phys. Rev.} A {\bf 80} 023611

### 2013-10-10 Response to referees.txt
Nearly ready to resubmit

We thank the referees for their detailed reports.  We have addressed their 
concerns and clarified answers to their questions in a new version of the 
manuscript.

Response to suggestions of referee 1:

1. We have clarified the wording of the discussion at the bottom of page 4, to
better explain the derivation of Eq. (12), which is the reduced, approximate 
form of (11).

2. We had indeed overlooked defining \tau. We have now added the definition
of \tau, and renamed the period of cyclic variations from \tau to T to
avoid confusion.

3. We have also added a paragraph at the end of Section 3 clearly explaining
our definition of number squeezing in the multimode case, and refer to this
definition in the caption of Figure 4.

4. We thank the referee for bring the extra reference on atomic Kerr squeezing
to our attention, and we have added it to our paper.

Response to questions and suggestions of referee 2:

1.  We have added a discussion of the practicality of using squeezing to improve a precision 
measurement relative to the possibility of increasing the atom number.  In fact, these two methods 
should ideally be used together if possible, as they are essentially entirely compatible and 
independent improvements to the sensitivity.  We include the highly relevant reference that the 
referee quoted.

2.  We have added a comment that the large number, box-potential geometries are most obviously
adapted to precision inertial measurements rather than in-trap measurements with goals of spatial 
resolution.

3.  We have rewritten our discussion of the effect of trap geometry on squeezing in response to 
the referee's very reasonable comments that it was confusing and apparently contradictory.  The 
effect of dimensionality changes character completely depending on whether the density is kept 
constant, or the total atom number is kept constant.  We have elaborated on this point, and made
the manuscript clear on this key issue.

     These two conclusions seem contradictory, and it is not clear how to relate them. This is
    made more difficult because figure 10 is missing the part labelled "Right", which as I
    understand should show the achievable squeezing in the 3D case. The authors appear to
    resolve this seeming contradiction in the conclusion, but it is unclear from the text how
    this relates to the parameters used in the different simulations.

### 2013-10-10 large_number_squeezing_with_atoms.tex
Nearly ready to resubmit

@@
 
-Producing squeezed states for metrological purposes is therefore only relevant when it can be achieved in a context of large numbers of atoms.
+A precision measurement using interferometry is always performed at the maximum allowable flux in order to maximise sensitivity.
+While the application of squeezing can be performed to some extent independently of the technical considerations that tend to limit flux, it is clearly most metrologically relevant when it can be achieved in the context of the maximum flux.
+There are many practical limits on the usable atomic flux, typically set by the available trap parameters and loss rates.
+Squeezing has already been used interferometrically on measurements using chip-based systems, which are particularly limited in their maximum atomic number \cite{Ockeloen2013}.
+Outside very specific circumstances, however, atom interferometry is a sensitive tool on an absolute scale when the atomic flux can
+be large.
 This paper examines the generation of atomic squeezing using the inherent Kerr-like nonlinearities of ultracold atoms, and demonstrates a trap geometry in which metrologically relevant levels of squeezing are present in trap geometries that are compatible with large atomic number.
@@
 
-In the multimode case, the single mode definition for number squeezing, $\langle\mathrm{Var}[\hat{N}_i] \rangle / \langle \hat{N}_i \rangle$, where $\hat{N}_i$ is the number operator for state $|i\rangle \in \{|a\rangle, |b\rangle \}$, is no longer adequate.
-In the multimode case we define the number squeezing in states $|i\rangle$ to be
-\begin{equation}
-\mathrm{Squeezing} = \frac{\mathrm{Var} \left[ \sum_j \hat{N}_i(x_j) \right]} { \sum_j \hat{N}_i(x_j) }
+The total number squeezing in the multimode case is quantified by the normalised number variances $\mathcal{V}_i$ in internal states $|i\rangle \in \{|a\rangle, |b\rangle \}$:
+\begin{equation}
+\mathcal{V}_i = \langle\mathrm{Var}[\hat{N}_i] \rangle / \langle \hat{N}_i \rangle.
 \label{eqMultimodeSqueezingDefinition}
 \end{equation}
-where the sums are over spatial modes defined as delta functions in position, and there is seperate number operator $\hat{N}_i(x_j)$ for each mode $j$ in state $|i\rangle$.
+In the multimode case, this is calculated using the total number operator $\hat{N}_i = \sum_j \hat{N}_{i j}$, where $\hat{N}_{i j}$ is the number operator for the state with spatial mode $j$ and internal state $|i\rangle$, which is independent of the spatial mode basis.
 
@@
 
-The simulations in the previous two sections were one dimensional, and show how having spatially constant density and using a $\pi$-pulse will improve the resultant squeezing.
+The simulations in the previous two sections were one dimensional, and show how having spatially constant atomic density and using a $\pi$-pulse will improve the resultant squeezing.
 While these conclusions remain valid for simulations in two or three dimensions, adding each additional dimension increases the number of spatial modes in a given energy range.
-For BECs of different dimension at the same energy per particle, this leads to a degradation of squeezing, even when the potential (and the resulting atomic density) is constant across the bulk of the trap.
+For BECs of different dimension \textit{at the same energy per particle}, this leads to a degradation of squeezing, even when the potential (and the resulting atomic density) is constant across the bulk of the trap.
 This is demonstrated in Fig.~\ref{figDimensionalSqueezingEffects}, where we show the squeezing generated in 1D, 2D and 3D BECs for two different parameter regimes.
@@
 
-We wish to examine the scaling of the non-condensed fraction $N_{\mathbf{k}\neq0}/N$ with the physical size of the condensate.
-Other system parameters such as the total number $N$ and the occupation of the $\ket{a}$ and $\ket{b}$ states at $t_1$ ($n_a$ and $n_b$, respectively) are kept constant and chosen to maximise the squeezing of the equivalent two-mode model presented in Section~\ref{secTwoModeAnalytic}.
+We wish to examine the scaling of the non-condensed fraction $N_{\mathbf{k}\neq0}/N$ with the physical size of the condensate, \textit{where the total number $N$  is kept constant}.
+This is importantly different to the case above where the density is kept constant, as here the density and energy per particle will go down with volume.
+Other system parameters such as the occupation of the $\ket{a}$ and $\ket{b}$ states at $t_1$ ($n_a$ and $n_b$, respectively) are also kept constant and chosen to maximise the squeezing of the equivalent two-mode model presented in Section~\ref{secTwoModeAnalytic}.
 In particular, this fixes the values of $\theta$, $\phi$, and $\lambda_{ij} = \chi_{ij}\tau_\text{hold}$.
@@
 This suggests that the best system to use will be a three-dimensional box potential, and the system size can be increased until the condensate depletion becomes negligibly small.
+We have therefore determined that dimensionality of the system switches from a curse to a blessing, provided we allow the density to go down as we increase the volume.
+This does not limit the total number accessible, but it means that the volume must go up faster than the number to maintain high squeezing.
 
@@
 In practice, the only limit to the number of atoms in such traps would be the stability of the trap, as the BEC would become increasingly sensitive to fluctuations as the energy per particle is reduced as the volume increases.
+These trap geometries would most obviously be suited for precision inertial measurements rather than in-trap measurements with goals of spatial resolution.
 
@@
 
-\bibitem{jeppesenET2008} Jeppesen M, Dugu{\'{e}}, Dennis G R, Johnsson M T, Figl C, Robins N P and Close J D 2008 Approaching the Heisenberg limit in an atom laser \emph{Phys. Rev.} A {\bf 77} 063618
+\bibitem{jeppesenET2008} Jeppesen M, Dugu{\'{e}} J, Dennis G R, Johnsson M T, Figl C, Robins N P and Close J D 2008 Approaching the Heisenberg limit in an atom laser \emph{Phys.
+Rev.} A {\bf 77} 063618
+
+\bibitem{Ockeloen2013} Ockeloen C F, Schmied R, Riedel M F, and Treutlein P, \emph{Phys. Rev. Lett.} {\bf 111} 143001 (2013).
 

### 2013-10-10 Response to referees.txt
Final changes addressing referee 2's comments

@@
 
-1.  We have added a discussion of the practicality of using squeezing to improve a precision 
-measurement relative to the possibility of increasing the atom number.  In fact, these two methods 
-should ideally be used together if possible, as they are essentially entirely compatible and 
-independent improvements to the sensitivity.  We include the highly relevant reference that the 
-referee quoted.
+1.  We have added a discussion of the practicality of using squeezing to 
+improve a precision measurement relative to the possibility of increasing the 
+atom number.  In fact, these two methods should ideally be used together if 
+possible, as they are essentially entirely compatible and independent 
+improvements to the sensitivity.  We include the highly relevant reference that 
+the referee quoted.
 
-2.  We have added a comment that the large number, box-potential geometries are most obviously
-adapted to precision inertial measurements rather than in-trap measurements with goals of spatial 
-resolution.
+2.  We have added a comment that the large number, box-potential geometries are 
+most obviously adapted to precision inertial measurements rather than in-trap 
+measurements with goals of spatial resolution.
 
-3.  We have rewritten our discussion of the effect of trap geometry on squeezing in response to 
-the referee's very reasonable comments that it was confusing and apparently contradictory.  The 
-effect of dimensionality changes character completely depending on whether the density is kept 
-constant, or the total atom number is kept constant.  We have elaborated on this point, and made
+3.  We have rewritten our discussion of the effect of trap geometry on 
+squeezing in response to the referee's very reasonable comments that it was 
+confusing and apparently contradictory.  The effect of dimensionality changes 
+character completely depending on whether the density is kept constant, or the 
+total atom number is kept constant.  We have elaborated on this point, and made 
 the manuscript clear on this key issue.
 
-     These two conclusions seem contradictory, and it is not clear how to relate them. This is
-    made more difficult because figure 10 is missing the part labelled "Right", which as I
-    understand should show the achievable squeezing in the 3D case. The authors appear to
-    resolve this seeming contradiction in the conclusion, but it is unclear from the text how
-    this relates to the parameters used in the different simulations.
+4. We have removed the reference to the non-existent "Right" figure in the 
+caption to Figure 10, and re-written the caption for greater clarity.

### 2013-10-10 large_number_squeezing_with_atoms.tex
Final changes addressing referee 2's comments

@@
 \caption{Non-condensed fraction and squeezing as function of trap size $L$ in a 2D trap.
-(Left) The proportion the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+(Left) The proportion of the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
 (Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
@@
    \includegraphics[width=7cm]{figures/3D_Bog_mode_occupation_scaling.eps}
-\caption{Non-condensed fraction and squeezing as function of trap size $L$ in a 3D trap.
-(Left) The proportion the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
-(Right) Relative number squeezing in mode $|a\rangle$ as a function of trap size.
+\caption{Non-condensed fraction as function of trap size $L$ in a 3D trap.
+The Figure shows the proportion of the non-condensed fraction in state $|a\rangle$ relative to total particle number, with the Bogoliubov theory model in red and the full numerical solution in blue.
+Unlike Figures \ref{fig2D_squeezing_and_bog_scaling} and \ref{fig1D_squeezing_and_bog_scaling}, the squeezing scaling curve is not shown here because in 3D the TWA method breaks down for small trap sizes.
+For the larger trap sizes with small non-condensate fractions, however, the TWA simulation remained valid, and we obtained relative number squeezing of 12.7 dB.
 Parameters: $N=1\times 10^4$, $U_{aa} = 4.7\times 10^{-49}$Jm$^3$, $U_{ab}=U_{bb}=0$, $\tau_{\mathrm{hold}} = 1.25 \times 10^{11} L^3$\,s.}
